### 2024-10-19 paper/paper.qmd
committing 03-clean_data script and cleaned dataset

---
title: "2024 US Presidential Election Model"
subtitle: "My subtitle if needed"
author: 
  - Jimin Lee
  - Sarah Ding
  - Xiyan Chen
thanks: "Code and data are available at: [https://github.com/jamiejiminlee/2024_US_Elections.git](https://github.com/jamiejiminlee/2024_US_Elections.git)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-10-19 scripts/02-download_data.R

@@ -13,10 +13,10 @@ library(tidyverse)
 
 
 #### Download data ####
-raw_data <- read_csv("/Users/jamielee/Downloads/president_polls.csv")
+raw_poll_data <- read_csv("/Users/jamielee/Downloads/president_polls.csv")
 
 
 #### Save data ####
 # [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(raw_data, "data/01-raw_data/raw_data.csv")
+
+write_csv(raw_poll_data, "data/01-raw_data/raw_data.csv")

### 2024-10-19 scripts/03-clean_data.R

@@ -11,34 +11,12 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_poll_data <- read_csv("data/01-raw_data/raw_data.csv")
 
 cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+  raw_poll_data |>
+  
+  
 
 #### Save data ####
 write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2024-10-19 paper/paper.qmd
committ linear model script and model file

@@
 library(tidyverse)
-library(palmerpenguins)
+library(janitor)
+library(lubridate)
+library(modelsummary)
+library(rstanarm)
+library(readr)
+library(dplyr)
 ```
@@
 # Data {#sec-data}
+```{r}
+# Load the cleaned poll data
+clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
+
+# Group by state and candidate, and calculate the average percentage (pct)
+state_average <- clean_poll_data |>
+  group_by(state, candidate_name) |>
+  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
+  pivot_wider(names_from = candidate_name, values_from = avg_pct) |>
+  rename(Trump_pct = "Donald Trump", Harris_pct = "Kamala Harris")
+
+# Replace missing values only in the numeric columns (Trump_pct, Harris_pct)
+state_average <- state_average |>
+  mutate(across(c(Trump_pct, Harris_pct), ~replace_na(., 0)))
+```
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+To predict the election outcome, we will assume that the higher pct in a state between Trump and Harris determines the winner of the election.
+The difference in pct is used as the predictor for which candidate wins each state.
 
@@
 Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 

### 2024-10-19 scripts/03-clean_data.R

@@ -1,22 +1,29 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Jimin Lee, 
+# Date: 19 October 2024 
+# Contact: jamiejimin.lee@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
+library(lubridate)
+library(readr)
+library(dplyr)
 
-#### Clean data ####
-raw_poll_data <- read_csv("data/01-raw_data/raw_data.csv")
+#### Prepare Dataset ####
+clean_poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
+  clean_names()
 
-cleaned_data <-
-  raw_poll_data |>
+# Filter data to Harris estimates based on high-quality polls after she declared
+clean_poll_data <- clean_poll_data |>
+  filter(numeric_grade >= 3, transparency_score > 5, candidate_name %in% c("Donald Trump", "Kamala Harris"))
   
-  
+head(clean_poll_data)
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+write_csv(clean_poll_data, "data/02-analysis_data/analysis_data.csv")
+

### 2024-10-19 scripts/06-model_data.R

@@ -13,24 +13,25 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Predicting the election outcome:
+# We'll assume that the higher pct in a state between Trump and Harris determines the winner. 
+# We can use the difference in pct as a predictor for which candidate wins each state. 
 
+# Create a binary outcome for prediction:m 1 if Trump is predicted to win, 0 if Harris
+state_average <- state_average |>
+  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
+
+# Create a linear model using pct differences to predict the outcome
+linear_model <- lm(winner ~ Trump_pct + Harris_pct, data = state_average)
+
+summary(linear_model)
 
 #### Save model ####
 saveRDS(
-  first_model,
+  linear_model,
   file = "models/first_model.rds"
 )
 

### 2024-10-19 scripts/00-simulate_data.R
committing text changes to scripts for the preambles

@@ -1,12 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of US electoral divisions, including the 
-  #state and party that won each division.
-# Author: Jimin Lee, Sarah Ding, Xiayi Chen
+# Purpose: Generates a simulated dataset for polling data in the 2024 US Presidential Election.
+# Author: Jimin Lee, Sarah Ding, Xiyan Chen
 # Date: 19 October 2024
-# Contact: jamiejimin.lee@mail.utoronto.ca
+# Contact: jamiejimin.lee@mail.utoronto.ca, sarah.ding@mail.utoronto.ca, xiyan.chen@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
+# Any other information needed? N/A
 
 
 #### Workspace setup ####

### 2024-10-19 scripts/03-clean_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Jimin Lee, 
+# Purpose: Prepares polling data for analysis in the 2024 US Presidential Election model by filtering the raw dataset to retain only high-quality polls and focuses on estimates for Donald Trump and Kamala Harris.
+# Author: Jimin Lee, Sarah Ding, Xiyan Chen
 # Date: 19 October 2024 
-# Contact: jamiejimin.lee@mail.utoronto.ca 
+# Contact: jamiejimin.lee@mail.utoronto.ca, sarah.ding@mail.utoronto.ca, xiyan.chen@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
+# Any other information needed? N/A
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-10-19 scripts/06-model_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the predicted outcome for the 2024 US Presidential Election
+# Author: Jimin Lee, Sarah Ding, Xiyan Chen
+# Date: 19 October 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca, sarah.ding@mail.utoronto.ca, xiyan.chen@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
+# Any other information needed? N/A
 
 
 #### Workspace setup ####

### 2024-10-19 paper/paper.qmd
saving code changes to paper - data section

@@
 library(dplyr)
+library(ggplot2)
+library(usmap)
 ```
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+@fig-averagetable provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model. 
+```{r}
+#| label: fig-averagetable
+#| fig-cap: Summary Table of Polling Averages by State
+#| echo: false
+
+# Create winner column: 1 if Trump_pct is greater than Harris_pct, 0 otherwise
+state_average <- state_average |>
+  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
+
+average_table <- state_average |>
+  select(state, Trump_pct, Harris_pct, winner) |>
+  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))
+
+average_table
+```
+
+@fig-barplot illustrates the average polling support for Donald Trump and Kamala Harris in each state.
+The side-by-side comparison represents the percentage of support each candidate has received, based on aggregated poll data from pollsters with high-quality scores.
+Trump's support is shown in red, while Harris's support is depicted in red.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-barplot
+#| fig-cap: Average Poll Percentage by State
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+# Reshape data to a long format for side-by-side bars
+stateaverage_long <- state_average %>%
+  pivot_longer(cols = c(Trump_pct, Harris_pct),
+               names_to = "candidate",
+               values_to = "support_pct")
+
+ggplot(stateaverage_long, aes(x=reorder(state, -support_pct), y= support_pct, fill = candidate)) + 
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
+  labs(title = "Average Poll Percentages for Trump and Harris by State",
+       x = "State", y = "Average Percentage") +
+  scale_fill_manual(name = "Candidate", values = c("Trump_pct" = "red", "Harris_pct" = "blue")) +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(axis.text.x = element_text(angle = 90, hjust = 1))
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+@fig-map visualizes the predicted winner of the 2024 US Presidential Election across all states, based on aggregated polling data.
+Each state is color-coded according to the candidate predicted to win - blue represents Donald Trump, and red represents Kamala Harris.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-map
+#| fig-cap: Predicted Election Winner by State
 #| echo: false
@@
 
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+library(ggplot2)
+library(usmap)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Ensuring state names are in the correct format
+state_average <- state_average %>%
+  mutate(state = str_to_title(state))
+
+# Creating binary variable winner column: 1 if Trump has more support, 0 if Harris has more support. 
+state_average <- state_average %>%
+  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
+
+# Creating predicted_winner column to label Trump and Harris
+state_average <- state_average |>
+  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))
+
+# Map visualization showing predicted winner by state
+plot_usmap(data = state_average, values = "predicted_winner", regions = "state") + 
+  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") + 
+  labs(title = "Predicted Winner of 2024 US Presidential Election by state") + 
+  theme(legend.position = "right")
 ```
-
-Talk way more about it.
 

### 2024-10-19 paper/paper.qmd
rendering pdf

@@
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+#modelsummary::modelsummary(
+  #list(
+    #"First model" = first_model
+ # ),
+ # statistic = "mad",
+ # fmt = 2
+#)
 ```
@@
 
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+#pp_check(first_model) +
+  #theme_classic() +
+  #theme(legend.position = "bottom")
 
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
+#posterior_vs_prior(first_model) +
+ # theme_minimal() +
+  #scale_color_brewer(palette = "Set1") +
+ # theme(legend.position = "bottom") +
+  #coord_flip()
 ```
@@
 
-plot(first_model, "trace")
+#plot(first_model, "trace")
 
-plot(first_model, "rhat")
+#plot(first_model, "rhat")
 ```

### 2024-10-20 paper/paper.qmd
committing code and text changes

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Placeholder: This paper presents a predictive model for the 2024 United States Presidential Election, utilizing polling data from various states.
+The model aggregates high-quality poll data, focusing on support for the two main candidates, Donald Trump and Kamala Harris.
+Using state-level polling averages, we create a linear model to predict the election outcome in each state.
+The model compares the polling percentages of both candidates to identify the likely winner.
+A geographic visualization provides a state-by-state breakdown, showing which candidate is predicted to win based on the polling data.
+The results offer insights into the competitive dynamics of the election and demonstrate the power of polling data in forecasting political outcomes.
+Furthermore, we discuss the methodology used to filter for high-quality pollsters and ensure data accuracy.
+The paper concludes with a reflection on the model’s limitations and the potential impact of polling biases on electoral predictions."
 format: pdf
@@
 
-Overview paragraph
+This paper develops a predictive model for the 2024 US Presidential Election using high-quality polling data.
+By aggregating state-level poll results for Donald Trump and Kamala Harris, we forecast the likely winner in each state.
+Our analysis leverages a linear regression model, with state-wise polling percentages as predictor variables, to estimate the outcome of the election.
 
-Estimand paragraph
+The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, based on state-level polling averages.
+The binary outcome variable indicates whether Trump (1) or Harris (0) is predicted to win each state.
 
-Results paragraph
+Our model predicts the election outcome by comparing the average polling percentages for Trump and Harris in each state.
+The results show a geographic distribution of support, with some states clearly favoring one candidate.
+The map visualization highlights these state-by-state predictions, offering insight into key battleground states and strongholds for both candidates.
 
-Why it matters paragraph
+Accurate election predictions are crucial for understanding voter dynamics and campaign strategies.
+By focusing on high-quality polling data, this model provides a more reliable forecast of the election outcome, which can help inform both political analysts and the public about potential election results and voting trends.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The remainder of this paper is structured as follows.
+In @sec-data, we describe the data and the variables used in the analysis.
+@sec-model outlines the model setup and estimation strategy.
+@sec-results presents the results, including the predicted election outcomes and visualizations.
+Finally, @sec-discussion the implications and limitations of our findings.
 
 # Data {#sec-data}
-```{r}
-# Load the cleaned poll data
-clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
-
-# Group by state and candidate, and calculate the average percentage (pct)
-state_average <- clean_poll_data |>
-  group_by(state, candidate_name) |>
-  summarise(avg_pct = mean(pct, na.rm = TRUE)) |>
-  pivot_wider(names_from = candidate_name, values_from = avg_pct) |>
-  rename(Trump_pct = "Donald Trump", Harris_pct = "Kamala Harris")
-
-# Replace missing values only in the numeric columns (Trump_pct, Harris_pct)
-state_average <- state_average |>
-  mutate(across(c(Trump_pct, Harris_pct), ~replace_na(., 0)))
-```
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
-
-Overview text
+Placholder - We use the statistical programming language R [@citeR] for all data manipulation, modeling, and visualization in this study.
+Our data [@shelter] was obtained from the public polling aggregation website FiveThirtyEight, which provides comprehensive polling data for the 2024 US Presidential Election [@fivethirtyeight].
+The dataset includes national polls conducted by various polling organizations, capturing voter support for the two leading candidates, Donald Trump and Kamala Harris.
+The dataset records critical variables such as the pollster, sample size, and percentage of voter support for each candidate.
 
@@
 
+The outcome variable in our analysis is the binary variable winner, which represents the predicted winner of the 2024 US Presidential Election in each state.
+The value of winner is set to 1 if Donald Trump is predicted to win the state, and 0 if Kamala Harris is predicted to win.
+This binary outcome is determined by comparing the polling support for both candidates within each state.
+By setting up this binary variable, we aim to forecast which candidate will secure more votes in each state based on the aggregated polling data.
+
 @fig-averagetable provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model.
+
 ```{r}
@@
 #| echo: false
+
+# Load cleaned data
+clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
+
+# Aggregating polling percentages for Trump and Harris by state
+state_average <- clean_poll_data %>%
+  group_by(state) %>%
+  summarise(
+    Trump_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE),
+    Harris_pct = mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE)
+  )
 
@@
 ```
+
+## Predictor variables
+
+The predictor variables used in the model are the aggregated polling percentages for Donald Trump (Trump_pct) and Kamala Harris (Harris_pct).
+These variables are calculated as the average support for each candidate, using polling data filtered to include only high-quality pollsters with a numeric grade of 3 or higher and transparency scores above 6.
+These averages reflect the level of support for each candidate across all polls in each state.
+To further illustrate the distribution of support, a side-by-side bar graph is provided, comparing the average polling percentages for both candidates across all states.
+This visualization offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
 
@@
 
-@fig-map visualizes the predicted winner of the 2024 US Presidential Election across all states, based on aggregated polling data.
-Each state is color-coded according to the candidate predicted to win - blue represents Donald Trump, and red represents Kamala Harris.
+# Model {#sec-model}
 
-```{r}
-#| label: fig-map
-#| fig-cap: Predicted Election Winner by State
-#| echo: false
-#| warning: false
-#| message: false
-
-library(ggplot2)
-library(usmap)
-
-# Ensuring state names are in the correct format
-state_average <- state_average %>%
-  mutate(state = str_to_title(state))
-
-# Creating binary variable winner column: 1 if Trump has more support, 0 if Harris has more support. 
-state_average <- state_average %>%
-  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
-
-# Creating predicted_winner column to label Trump and Harris
-state_average <- state_average |>
-  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))
-
-# Map visualization showing predicted winner by state
-plot_usmap(data = state_average, values = "predicted_winner", regions = "state") + 
-  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") + 
-  labs(title = "Predicted Winner of 2024 US Presidential Election by state") + 
-  theme(legend.position = "right")
-```
-
-## Predictor variables
-
-Add graphs, tables and text.
-
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modeling strategy is twofold.
+Firstly, we aim to predict the winner of the 2024 US Presidential Election in each state based on aggregated polling data.
+Secondly, we seek to understand the relationship between polling support for each candidate and the likelihood of winning a state.
+Here we briefly describe the Bayesian analysis model used to investigate these relationships.
+This model allows for uncertainty in the predictions by incorporating prior distributions for the parameters, providing a probabilistic framework for estimating the election outcome.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+## Model
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+Define \( y_i \) as the binary outcome variable indicating whether Donald Trump is predicted to win state \( i \) (1 if Trump wins, 0 if Harris wins).
+The predictor variables are \( \beta_i \), the average percentage of polling support for Donald Trump in state \( i \), and \( \gamma_i \), the average percentage of polling support for Kamala Harris in state \( i \).
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+\[
+y_i|\mu_i, \sigma \sim \mbox{Bernoulli}(\mu_i)
+\]
+\[
+\mu_i = \alpha + \beta_i + \gamma_i
+\]
+\[
+\alpha \sim \mbox{Normal}(0, 2.5)
+\]
+\[
+\beta_i \sim \mbox{Normal}(0, 2.5)
+\]
+\[
+\gamma_i \sim \mbox{Normal}(0, 2.5)
+\]
 
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+We run the model in R [@citeR] using the `rstanarm` package [@rstanarm].
+The model uses default priors from `rstanarm`, and estimates the likelihood of Donald Trump winning each state based on polling averages for both candidates.
 
@@
 
+The model used to predict the 2024 US Presidential Election outcome is a logistic regression model.
+The binary outcome variable, \( y_i \), represents whether Donald Trump (1) or Kamala Harris (0) is predicted to win each state.
+The predictor variables, \( \beta_i \) and \( \gamma_i \), are the average polling percentages for Trump and Harris, respectively, in state \( i \).
+We use a logistic regression model because the outcome is binary, making it appropriate for estimating the probability of a Trump victory in each state.
+The model is implemented in R using the `glm()` function with a binomial family, providing a probabilistic prediction of the election outcome based on polling data.
+
+### Model Justification
+
+We expect a positive relationship between polling support for each candidate and the likelihood of winning a state.
+In particular, as the percentage of support for Donald Trump (\(\beta_i\)) increases, the probability of Trump winning state \(i\) also increases.
+Similarly, as Kamala Harris's support (\(\gamma_i\)) increases, the likelihood of her winning the state rises.
+The logistic regression model is suitable for capturing these dynamics, as it allows us to estimate the probability of a binary outcome (win or lose) using continuous predictor variables.
+The model outputs probabilities that are constrained between 0 and 1, which aligns with the binary nature of our outcome variable (\(y_i\)).
+In essence, we expect a higher \(\beta_i\) to correlate with a higher probability of Trump winning a state, while a higher \(\gamma_i\) suggests a greater likelihood of a Harris victory.
+
+# Results {#sec-results}
+
+@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state, based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
+States where Trump is predicted to win are shown in red, while states where Harris is predicted to win are displayed in blue.
+Several states appear in gray, indicating that the model could not generate a prediction due to incomplete or missing polling data.
+The results highlight clear regional differences, with Trump expected to perform strongly in the southern and midwestern states, while Harris is favored in parts of the Northeast and Midwest.
+The predicted outcomes underscore the importance of swing states, where polling results are closely contested and could shift the election results.
+
+### Predicted Election Winner By State
+```{r}
+#| label: fig-statemap
+#| fig-cap: Predicted Election Winner by State
+#| echo: false
+#| warning: false
+#| message: false
+
+#### Workspace setup ####
+library(tidyverse)
+library(usmap)
+library(ggplot2)
+
+# Remove any states with missing polling data for Trump or Harris
+state_average <- state_average %>%
+  filter(!is.na(Trump_pct) & !is.na(Harris_pct))
+
+# Proceed with the prediction for all states
+state_average <- state_average %>%
+  mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))
+
+# Classify the predicted winner: Trump if probability > 0.5, else Harris
+state_average <- state_average %>%
+  mutate(predicted_winner = ifelse(predicted_prob > 0.5, "Trump", "Harris"))
+
+#### Map Visualization ####
+plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
+  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
+  labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
+  theme(legend.position = "right")
+```
+
+It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.
+These uncolored states reflect the absence of reliable or sufficient polling information required for an accurate prediction.
+This underscores the limitations of the data used in the model, where predictions are only made for states with adequate polling coverage.
+As a result, certain regions lack predictions, which could be addressed by including more comprehensive polling data in future analyses.
+
+### Summary of Predicted Election Results by State
+
+@tbl-modelresults presents a summary of the predicted 2024 US Presidential Election results on a state-by-state basis.
+It includes the average polling percentages for Donald Trump and Kamala Harris, the predicted probability of Trump winning each state based on a logistic regression model, and the predicted winner in each state.
+
 ```{r}
@@
 #| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| tbl-cap: "Predicted 2024 US Presidential Election outcomes by state, based on polling percentages for Trump and Harris"
 #| warning: false
 
-#modelsummary::modelsummary(
-  #list(
-    #"First model" = first_model
- # ),
- # statistic = "mad",
- # fmt = 2
-#)
+#### Create and Display the Summary Table ####
+summary_table <- state_average %>%
+  select(state, Trump_pct, Harris_pct, predicted_prob, predicted_winner) %>%
+  rename(
+    "State" = state,
+    "Trump Polling (%)" = Trump_pct,
+    "Harris Polling (%)" = Harris_pct,
+    "Predicted Trump Win Probability" = predicted_prob,
+    "Predicted Winner" = predicted_winner
+  )
+
+knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State")
+
 ```
 
-# Discussion
+According to our model results, in **Florida**, Trump’s polling percentage (53.00%) is substantially higher than Harris’s (40.50%), leading to a near-certain predicted win for Trump with a probability of 0.9998.
+Conversely, in **Michigan**, Harris holds a slight polling lead (46.27% vs.
+44.52%), resulting in a predicted win for her with a lower probability of 0.036.
+Several states, such as **Montana** and **Missouri**, have extremely high probabilities of a Trump victory, while others, like **Nebraska CD-2** and **North Carolina**, lean toward Harris.
+The inclusion of predicted probabilities adds nuance by highlighting the confidence levels of the model’s predictions for each state, especially in swing states where polling percentages are closely contested.
 
-## First discussion point {#sec-first-point}
+# Discussion {#sec-discussion}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Interpretation of Results {#sec-results-interpretation}
 
-## Second discussion point
+The results of the logistic regression model provide insights into the predicted winner of the 2024 US Presidential Election based on polling percentages for Donald Trump and Kamala Harris.
+The side-by-side bar graph comparing polling averages across states reveals a clear geographic divide.
+Trump tends to perform better in southern and midwestern states, while Harris shows stronger support in the northeastern and western regions.
+Swing states, where the polling differences between the two candidates are narrower, emerge as crucial battlegrounds that could determine the election outcome.
+The map visualization reinforces these trends, visually highlighting where each candidate is expected to win.
+These results align with historical voting patterns, but the model also emphasizes the close competition in key states that are not traditionally battlegrounds.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+## Model Performance {#sec-model-performance}
 
-## Third discussion point
+The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state.
+The coefficients for the predictor variables (polling percentages for Trump and Harris) indicate a strong relationship between candidate support and the probability of winning.
+The model outputs probabilities for each state, which are useful for gauging the certainty of predictions, especially in closely contested states.
+The model’s performance is reinforced by the robustness of the results in states where one candidate has a clear polling lead, providing accurate predictions.
+However, states with narrower margins reflect greater uncertainty, which is critical in understanding election dynamics.
+
+### Interpretation of Polling Averages by State
+
+The bar graph displays the average polling percentages for Donald Trump (in red) and Kamala Harris (in blue) across various states.
+The comparison shows significant variability in support for each candidate.
+In some states like **Montana** and **Missouri**, Trump’s support is notably higher, with a clear margin over Harris.
+In contrast, **Colorado** stands out as having nearly equal polling support for both candidates, suggesting a closely contested race in that state.
+
+Swing states such as **Florida**, **North Carolina**, and **Pennsylvania** exhibit relatively close polling percentages, indicating that the election outcomes in these states could be pivotal and difficult to predict.
+The plot also includes states like **Virginia** and **Texas**, where the support levels are more balanced, though Trump seems to hold a slight lead.
+
+The state labeled as **NA** at the end may indicate missing or incomplete data, emphasizing the need for comprehensive polling coverage in all states for more reliable predictions.
+
+Overall, the side-by-side comparison highlights the states where each candidate holds a significant lead and where the race remains competitive, offering a detailed view of the election's geographic dynamics.
 
@@
 
-Weaknesses and next steps should also be included.
+Despite the model's predictive power, there are several weaknesses that should be noted.
+First, polling data itself is prone to biases, such as underrepresentation of certain demographic groups or inaccuracies in sampling.
+These issues can distort the predictions, particularly in states with fewer or lower-quality polls.
+Additionally, voter turnout and last-minute shifts in public opinion are not accounted for, which can significantly alter the outcome, especially in close races.
+The model also assumes that polling percentages directly translate to votes, overlooking other factors like voter mobilization efforts and campaign dynamics that may influence election outcomes.
+
+To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data.
+Factors such as voter turnout models, economic conditions, and social media sentiment analysis may provide a more holistic view of voter behavior.
+Additionally, real-time polling data integration could allow for dynamic updates to predictions as the election nears, improving the timeliness and accuracy of forecasts.
+Moreover, regional adjustments to account for differences in polling methodology or demographic shifts would help to address some of the biases inherent in polling data.
+Expanding the model to include these aspects would offer a more comprehensive understanding of the election landscape and enhance the precision of predictions.
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
 ```{r}
+``{r}
 #| echo: false
 #| eval: true
-#| message: false
+#| label: tbl-modelsummary
+#| tbl-cap: "Explanatory models of election outcome based on polling percentages for Trump and Harris"
 #| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
 
-#plot(first_model, "trace")
+#### Model Summary Table ####
+# Create and display a model summary table
+modelsummary::modelsummary(
+  list("First Model" = logistic_model),
+  statistic = c("std.error", "p.value"),  # Use standard error and p-value
+  fmt = 2,  # Format numbers to 2 decimal places
+  output = "markdown"  # Output format for Quarto or RMarkdown
+)
 
-#plot(first_model, "rhat")
 ```
+
+@tbl-modelsummary presents the results of the logistic regression model predicting the winner of the 2024 US Presidential Election in each state based on polling percentages for Donald Trump and Kamala Harris.
+The coefficients indicate that an increase in Trump’s polling percentage (\( \beta = 8.72 \)) is positively associated with the probability of winning a state, whereas an increase in Harris’s polling percentage (\( \beta = -31.63 \)) is negatively associated with Trump’s probability of winning.
+The standard errors for both predictors are quite large, suggesting high variability in the estimates and possible uncertainty in the model.
+
+The model diagnostics, including the AIC (6.0) and BIC (8.1), suggest that the model fits the data reasonably well given the small sample size of 15 observations.
+However, the low log-likelihood (0.000) and root mean squared error (RMSE = 0.00) indicate that the model may not be well-calibrated, likely due to the small dataset or the structure of the data.
+These results emphasize the importance of using a larger sample size.
 

### 2024-10-20 scripts/06-model_data.R

@@ -6,33 +6,36 @@
 # License: MIT
 # Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 # Any other information needed? N/A
-
-
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
+# Load the cleaned polling data
 clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
 
-### Model data ####
-# Predicting the election outcome:
-# We'll assume that the higher pct in a state between Trump and Harris determines the winner. 
-# We can use the difference in pct as a predictor for which candidate wins each state. 
 
-# Create a binary outcome for prediction:m 1 if Trump is predicted to win, 0 if Harris
-state_average <- state_average |>
+### Prepare Model Data ####
+# Aggregating polling percentages for Trump and Harris by state
+state_average <- clean_poll_data %>%
+  group_by(state) %>%
+  summarise(
+    Trump_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE),
+    Harris_pct = mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE)
+  )
+
+# Create a binary outcome for prediction: 1 if Trump is predicted to win, 0 if Harris
+state_average <- state_average %>%
   mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
 
-# Create a linear model using pct differences to predict the outcome
-linear_model <- lm(winner ~ Trump_pct + Harris_pct, data = state_average)
+#### Bayesian Logistic Regression Model with Regularization ####
+# Using a regularized logistic regression model from rstanarm to avoid convergence issues
+logistic_model <- stan_glm(winner ~ Trump_pct + Harris_pct, data = state_average, family = binomial, 
+                           prior = student_t(df = 7), chains = 4, iter = 2000)
 
-summary(linear_model)
+# Display the summary of the model
+summary(logistic_model)
 
-#### Save model ####
-saveRDS(
-  linear_model,
-  file = "models/first_model.rds"
-)
-
-
+#### Save Model ####
+# Save the trained logistic model for future use
+saveRDS(logistic_model, file = "models/first_model.rds")

### 2024-10-20 paper/paper.qmd
committing changes to paper.qmd and rendered file

@@
 #| echo: false
+#| warning: false
+#| message: false
 
@@
 
-average_table
+# Render the simplest possible table
+knitr::kable(average_table, caption = "Summary Table of Polling Averages by State")
+
 ```
@@
 #| echo: false
+#| warning: false
 
@@
 
-first_model <-
+logistic_model <-
   readRDS(file = here::here("models/first_model.rds"))
@@
 #| message: false
+#| cache: false
 
@@
 library(ggplot2)
+library(rstanarm)
 
-# Remove any states with missing polling data for Trump or Harris
+# Check for missing values in state_average and ensure data is clean
 state_average <- state_average %>%
-  filter(!is.na(Trump_pct) & !is.na(Harris_pct))
+  filter(!is.na(Trump_pct) & !is.na(Harris_pct)) %>%
+  ungroup()
 
-# Proceed with the prediction for all states
+# Check that state_average has the correct number of rows and no grouping
+print(nrow(state_average))  # Ensure this number matches the expected rows
+
+# Proceed with prediction after ensuring the data is clean
+predicted_prob <- predict(logistic_model, newdata = state_average, type = "response")
+
+# Ensure predicted_prob matches the number of rows in state_average
+if (length(predicted_prob) != nrow(state_average)) {
+  stop("Mismatch between predicted_prob and state_average rows")
+}
+
+# Add predicted_prob to state_average
 state_average <- state_average %>%
-  mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))
+  mutate(predicted_prob = predicted_prob)
 
@@
 
-#### Map Visualization ####
+# Map visualization showing predicted winner by state
 plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
@@
   theme(legend.position = "right")
+
+# Remove any states with missing polling data for Trump or Harris
+#state_average <- state_average %>%
+  #filter(!is.na(Trump_pct) & !is.na(Harris_pct))
+
+# Proceed with the prediction for all states
+#state_average <- state_average %>%
+  #mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))
+
+# Classify the predicted winner: Trump if probability > 0.5, else Harris
+#state_average <- state_average %>%
+  #mutate(predicted_winner = ifelse(predicted_prob > 0.5, "Trump", "Harris"))
+
+#### Map Visualization ####
+#plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
+  #scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
+  #labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
+  #theme(legend.position = "right")
 ```
@@
 
-knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State")
+knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State", escape = TRUE)
 
@@
 ```{r}
-``{r}
 #| echo: false
@@
 
-#### Model Summary Table ####
-# Create and display a model summary table
+# Ensure the model exists
+if (exists("logistic_model")) {
+  # Create and display a model summary table in LaTeX format for PDF
   modelsummary::modelsummary(
@@
     fmt = 2,  # Format numbers to 2 decimal places
-  output = "markdown"  # Output format for Quarto or RMarkdown
+    output = "latex"  # Use LaTeX for PDF output
   )
+} else {
+  print("Model not found")
+}
+
+#### Model Summary Table ####
+# Create and display a model summary table
+#modelsummary::modelsummary(
+  #list("First Model" = logistic_model),
+  #statistic = c("std.error", "p.value"),  # Use standard error and p-value
+  #fmt = 2,  # Format numbers to 2 decimal places
+  #output = "markdown"  # Output format for Quarto or RMarkdown
+#)
 

### 2024-10-20 scripts/06-model_data.R

@@ -39,3 +39,4 @@ summary(logistic_model)
 #### Save Model ####
 # Save the trained logistic model for future use
 saveRDS(logistic_model, file = "models/first_model.rds")
+

### 2024-10-20 paper/paper.qmd
committing rendered file

@@
 ## Model
-
 Define \( y_i \) as the binary outcome variable indicating whether Donald Trump is predicted to win state \( i \) (1 if Trump wins, 0 if Harris wins).
@@
 \[
-y_i|\mu_i, \sigma \sim \mbox{Bernoulli}(\mu_i)
+y_i|\mu_i \sim \text{Bernoulli}(\mu_i)
 \]
@@
 \[
-\alpha \sim \mbox{Normal}(0, 2.5)
+\alpha \sim \text{Normal}(0, 2.5)
 \]
 \[
-\beta_i \sim \mbox{Normal}(0, 2.5)
+\beta_i \sim \text{Normal}(0, 2.5)
 \]
 \[
-\gamma_i \sim \mbox{Normal}(0, 2.5)
+\gamma_i \sim \text{Normal}(0, 2.5)
 \]
@@
 
-### Model Justification
+## Model Justification
 
@@
 
-### Predicted Election Winner By State
+## Predicted Election Winner By State
 ```{r}
@@
   ungroup()
-
-# Check that state_average has the correct number of rows and no grouping
-print(nrow(state_average))  # Ensure this number matches the expected rows
 

### 2024-10-20 paper/paper.qmd
committing rendered file

@@
 ## Model set-up
-
-## Model
 Define \( y_i \) as the binary outcome variable indicating whether Donald Trump is predicted to win state \( i \) (1 if Trump wins, 0 if Harris wins).
@@
 \[
-y_i|\mu_i \sim \text{Bernoulli}(\mu_i)
-\]
-\[
-\mu_i = \alpha + \beta_i + \gamma_i
-\]
-\[
-\alpha \sim \text{Normal}(0, 2.5)
-\]
-\[
-\beta_i \sim \text{Normal}(0, 2.5)
-\]
-\[
-\gamma_i \sim \text{Normal}(0, 2.5)
+\begin{align}
+y_i | \mu_i &\sim \text{Bernoulli}(\mu_i) \\
+\mu_i &= \alpha + \beta_i + \gamma_i \\
+\alpha &\sim \text{Normal}(0, 2.5) \\
+\beta_i &\sim \text{Normal}(0, 2.5) \\
+\gamma_i &\sim \text{Normal}(0, 2.5)
+\end{align}
 \]

### 2024-10-20 paper/paper.qmd
committing latex changes for model section

@@
 ## Model set-up
-Define \( y_i \) as the binary outcome variable indicating whether Donald Trump is predicted to win state \( i \) (1 if Trump wins, 0 if Harris wins).
-The predictor variables are \( \beta_i \), the average percentage of polling support for Donald Trump in state \( i \), and \( \gamma_i \), the average percentage of polling support for Kamala Harris in state \( i \).
+Define $y_i$ as the binary outcome variable indicating whether Donald Trump is predicted to win state $i$ (1 if Trump wins, 0 if Harris wins).
+The predictor variables are $ \beta_i $, the average percentage of polling support for Donald Trump in state $i$, and $ \gamma_i $, the average percentage of polling support for Kamala Harris in state $i$.
 

### 2024-10-20 paper/paper.qmd
committing text changes
commit changes to render
changes in latex for paper.qmd

@@
 Define $y_i$ as the binary outcome variable indicating whether Donald Trump is predicted to win state $i$ (1 if Trump wins, 0 if Harris wins).
-The predictor variables are $ \beta_i $, the average percentage of polling support for Donald Trump in state $i$, and $ \gamma_i $, the average percentage of polling support for Kamala Harris in state $i$.
+The predictor variables are $\beta_i$, the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of polling support for Kamala Harris in state $i$.
 

### 2024-10-20 paper/paper.qmd
committing changes to rendered file
commit changes

@@
 
-\[
-\begin{align}
+\begin{align*}
 y_i | \mu_i &\sim \text{Bernoulli}(\mu_i) \\
@@
 \gamma_i &\sim \text{Normal}(0, 2.5)
-\end{align}
-\]
+\end{align*}
 
@@
 The model used to predict the 2024 US Presidential Election outcome is a logistic regression model.
-The binary outcome variable, \( y_i \), represents whether Donald Trump (1) or Kamala Harris (0) is predicted to win each state.
-The predictor variables, \( \beta_i \) and \( \gamma_i \), are the average polling percentages for Trump and Harris, respectively, in state \( i \).
+The binary outcome variable, $y_i$, represents whether Donald Trump (1) or Kamala Harris (0) is predicted to win each state.
+The predictor variables, $\beta_i$ and $\gamma_i$, are the average polling percentages for Trump and Harris, respectively, in state $i$.
 We use a logistic regression model because the outcome is binary, making it appropriate for estimating the probability of a Trump victory in each state.
@@
 We expect a positive relationship between polling support for each candidate and the likelihood of winning a state.
-In particular, as the percentage of support for Donald Trump (\(\beta_i\)) increases, the probability of Trump winning state \(i\) also increases.
-Similarly, as Kamala Harris's support (\(\gamma_i\)) increases, the likelihood of her winning the state rises.
+In particular, as the percentage of support for Donald Trump ($\beta_i$) increases, the probability of Trump winning state $i$ also increases.
+Similarly, as Kamala Harris's support ($\gamma_i$) increases, the likelihood of her winning the state rises.
 The logistic regression model is suitable for capturing these dynamics, as it allows us to estimate the probability of a binary outcome (win or lose) using continuous predictor variables.
-The model outputs probabilities that are constrained between 0 and 1, which aligns with the binary nature of our outcome variable (\(y_i\)).
-In essence, we expect a higher \(\beta_i\) to correlate with a higher probability of Trump winning a state, while a higher \(\gamma_i\) suggests a greater likelihood of a Harris victory.
+The model outputs probabilities that are constrained between 0 and 1, which aligns with the binary nature of our outcome variable ($y_i$).
+In essence, we expect a higher $\beta_i$ to correlate with a higher probability of Trump winning a state, while a higher $\gamma_i$ suggests a greater likelihood of a Harris victory.
 

### 2024-10-20 paper/paper.qmd
commit changes to qmd file

@@
 
-### Interpretation of Polling Averages by State
+## Interpretation of Polling Averages by State
 
@@
 # Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-#pp_check(first_model) +
-  #theme_classic() +
-  #theme(legend.position = "bottom")
-
-#posterior_vs_prior(first_model) +
- # theme_minimal() +
-  #scale_color_brewer(palette = "Set1") +
- # theme(legend.position = "bottom") +
-  #coord_flip()
-```
 

### 2024-10-20 paper/paper.qmd
committing changes to the qmd file

@@
 ## Diagnostics
+@tbl-modelsummary presents the results of the logistic regression model predicting the winner of the 2024 US Presidential Election in each state based on polling percentages for Donald Trump and Kamala Harris.
+The coefficients indicate that an increase in Trump’s polling percentage (\( \beta = 8.72 \)) is positively associated with the probability of winning a state, whereas an increase in Harris’s polling percentage (\( \beta = -31.63 \)) is negatively associated with Trump’s probability of winning.
+The standard errors for both predictors are quite large, suggesting high variability in the estimates and possible uncertainty in the model.
 
@@
 
-@tbl-modelsummary presents the results of the logistic regression model predicting the winner of the 2024 US Presidential Election in each state based on polling percentages for Donald Trump and Kamala Harris.
-The coefficients indicate that an increase in Trump’s polling percentage (\( \beta = 8.72 \)) is positively associated with the probability of winning a state, whereas an increase in Harris’s polling percentage (\( \beta = -31.63 \)) is negatively associated with Trump’s probability of winning.
-The standard errors for both predictors are quite large, suggesting high variability in the estimates and possible uncertainty in the model.
-
 The model diagnostics, including the AIC (6.0) and BIC (8.1), suggest that the model fits the data reasonably well given the small sample size of 15 observations.

### 2024-10-20 paper/paper.qmd
commit text changes paper.qmd

@@
 date-format: long
-abstract: "Placeholder: This paper presents a predictive model for the 2024 United States Presidential Election, utilizing polling data from various states.
-The model aggregates high-quality poll data, focusing on support for the two main candidates, Donald Trump and Kamala Harris.
-Using state-level polling averages, we create a linear model to predict the election outcome in each state.
-The model compares the polling percentages of both candidates to identify the likely winner.
-A geographic visualization provides a state-by-state breakdown, showing which candidate is predicted to win based on the polling data.
-The results offer insights into the competitive dynamics of the election and demonstrate the power of polling data in forecasting political outcomes.
-Furthermore, we discuss the methodology used to filter for high-quality pollsters and ensure data accuracy.
-The paper concludes with a reflection on the model’s limitations and the potential impact of polling biases on electoral predictions."
+abstract: "This paper presents a predictive model for the 2024 United States Presidential Election using state-level polling data.
+By aggregating high-quality polls, we predict the likely winner between Donald Trump and Kamala Harris in each state.
+A logistic regression model estimates the probability of Trump winning, with polling percentages as predictor variables.
+The model’s results highlight key battleground states and offer insights into voter dynamics, while also reflecting on the limitations of polling data in forecasting political outcomes."
 format: pdf
@@
 
-This paper develops a predictive model for the 2024 US Presidential Election using high-quality polling data.
-By aggregating state-level poll results for Donald Trump and Kamala Harris, we forecast the likely winner in each state.
-Our analysis leverages a linear regression model, with state-wise polling percentages as predictor variables, to estimate the outcome of the election.
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris.
+We aggregate high-quality poll data to create a logistic regression model that estimates the probability of a Trump or Harris victory in each state.
+By comparing polling percentages for both candidates, we aim to predict election outcomes based on voter support patterns across the states.
 
-The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, based on state-level polling averages.
-The binary outcome variable indicates whether Trump (1) or Harris (0) is predicted to win each state.
+The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, based on aggregated state-level polling averages.
+The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, and the predictor variables are the average polling percentages for each candidate.
 
-Our model predicts the election outcome by comparing the average polling percentages for Trump and Harris in each state.
-The results show a geographic distribution of support, with some states clearly favoring one candidate.
-The map visualization highlights these state-by-state predictions, offering insight into key battleground states and strongholds for both candidates.
+Our model predicts election outcomes by comparing the average polling percentages for Trump and Harris in each state.
+The results highlight the geographic distribution of support, with some states clearly favoring one candidate over the other.
+Swing states emerge as critical battlegrounds, where polling percentages are closely contested and could influence the final election result.
 
-Accurate election predictions are crucial for understanding voter dynamics and campaign strategies.
-By focusing on high-quality polling data, this model provides a more reliable forecast of the election outcome, which can help inform both political analysts and the public about potential election results and voting trends.
+Accurate election predictions provide valuable insights into voter dynamics and help political analysts, campaigns, and the public anticipate electoral outcomes.
+By focusing on high-quality polling data, our model improves the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.
 
 The remainder of this paper is structured as follows.
-In @sec-data, we describe the data and the variables used in the analysis.
-@sec-model outlines the model setup and estimation strategy.
-@sec-results presents the results, including the predicted election outcomes and visualizations.
-Finally, @sec-discussion the implications and limitations of our findings.
+In Section 2, we describe the data and variables used in the analysis.
+Section 3 outlines the model setup and estimation strategy.
+Section 4 presents the results, including the predicted election outcomes and visualizations.
+Finally, Section 5 discusses the implications and limitations of our findings.
 
@@
 
-To predict the election outcome, we will assume that the higher pct in a state between Trump and Harris determines the winner of the election.
-The difference in pct is used as the predictor for which candidate wins each state.
+Placeholder - The measurement of voter support in our dataset is derived from polling data, which serves as a representation of public sentiment during the 2024 U.S.
+Presidential Election.
+Each entry in the dataset corresponds to a poll conducted by various polling organizations, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
+To ensure the quality of the data, only polls from reputable pollsters with a numeric grade of 3 or higher and transparency scores above 6 were included.
+These polls measure voter preferences through structured survey questions, and the results are aggregated to create state-level averages for each candidate.
+The dataset thus transforms individual polling responses into a broader measurement of voter sentiment, reflecting the competitive dynamics between Trump and Harris across different states.
 
@@
 These averages reflect the level of support for each candidate across all polls in each state.
+
 To further illustrate the distribution of support, a side-by-side bar graph is provided, comparing the average polling percentages for both candidates across all states.
-This visualization offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
-
-@fig-barplot illustrates the average polling support for Donald Trump and Kamala Harris in each state.
+@fig-barplot offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
 The side-by-side comparison represents the percentage of support each candidate has received, based on aggregated poll data from pollsters with high-quality scores.
@@
 
-The goal of our modeling strategy is twofold.
-Firstly, we aim to predict the winner of the 2024 US Presidential Election in each state based on aggregated polling data.
-Secondly, we seek to understand the relationship between polling support for each candidate and the likelihood of winning a state.
-Here we briefly describe the Bayesian analysis model used to investigate these relationships.
-This model allows for uncertainty in the predictions by incorporating prior distributions for the parameters, providing a probabilistic framework for estimating the election outcome.
+The goal of our modeling strategy is to predict the winner of the 2024 US Presidential Election in each state based on polling data, while also assessing the likelihood of Donald Trump or Kamala Harris winning.
+We employ a Bayesian logistic regression model, which allows for probabilistic predictions of binary outcomes, making it well-suited for the task of predicting election results.
 Background details and diagnostics are included in [Appendix -@sec-model-details].
@@
 ## Model set-up
-Define $y_i$ as the binary outcome variable indicating whether Donald Trump is predicted to win state $i$ (1 if Trump wins, 0 if Harris wins).
-The predictor variables are $\beta_i$, the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of polling support for Kamala Harris in state $i$.
+The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win.
+The predictor variables are $\beta_i$, which represents the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of support for Kamala Harris in state $i$.
+The relationship between these variables and the outcome is captured through a logistic link function.
 
@@
 \end{align*}
+
+We use weakly informative Normal priors for the intercept ($\alpha_i$) and the coefficients ($\beta_i$) and $\gamma_i$), centered around zero with a standard deviation of 2.5.
+These priors were chosen to reflect the assumption that, before seeing the data, there is no strong reason to favor one candidate over the other across states, but to allow room for the data to adjust predictions as polling percentages differ.
 
@@
 
-The model used to predict the 2024 US Presidential Election outcome is a logistic regression model.
-The binary outcome variable, $y_i$, represents whether Donald Trump (1) or Kamala Harris (0) is predicted to win each state.
-The predictor variables, $\beta_i$ and $\gamma_i$, are the average polling percentages for Trump and Harris, respectively, in state $i$.
-We use a logistic regression model because the outcome is binary, making it appropriate for estimating the probability of a Trump victory in each state.
-The model is implemented in R using the `glm()` function with a binomial family, providing a probabilistic prediction of the election outcome based on polling data.
-
 ## Model Justification
 
-We expect a positive relationship between polling support for each candidate and the likelihood of winning a state.
-In particular, as the percentage of support for Donald Trump ($\beta_i$) increases, the probability of Trump winning state $i$ also increases.
-Similarly, as Kamala Harris's support ($\gamma_i$) increases, the likelihood of her winning the state rises.
-The logistic regression model is suitable for capturing these dynamics, as it allows us to estimate the probability of a binary outcome (win or lose) using continuous predictor variables.
-The model outputs probabilities that are constrained between 0 and 1, which aligns with the binary nature of our outcome variable ($y_i$).
-In essence, we expect a higher $\beta_i$ to correlate with a higher probability of Trump winning a state, while a higher $\gamma_i$ suggests a greater likelihood of a Harris victory.
+We expect a positive relationship between polling support for each candidate and their probability of winning a state.
+As Donald Trump’s polling percentage ($\beta_i$) increases, the likelihood of him winning state $i$ rises.
+Similarly, as Kamala Harris's polling percentage ($\gamma_i$) increases, her chances of winning that state increase.
+
+Bayesian logistic regression was chosen for its ability to model binary outcomes like win/loss predictions while incorporating prior knowledge and uncertainty, helping mitigate overfitting through regularization with priors.
+This model strikes a balance between flexibility and interpretability, given that polling percentages are continuous and the outcome is binary.
+The model assumes polling data accurately reflects voter intentions and that state outcomes are independent of each other, although it does not account for interactions between states or regional voting trends.
+Potential biases in polling data could also affect predictions.
+The model was implemented using the rstanarm package in R, which supports Bayesian inference with default priors for logistic regression.
+Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].
 
@@
 
+## Predicted Election Winner By State
+
 @fig-statemap shows the predicted winner of the 2024 US Presidential Election by state, based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
 States where Trump is predicted to win are shown in red, while states where Harris is predicted to win are displayed in blue.
-Several states appear in gray, indicating that the model could not generate a prediction due to incomplete or missing polling data.
-The results highlight clear regional differences, with Trump expected to perform strongly in the southern and midwestern states, while Harris is favored in parts of the Northeast and Midwest.
-The predicted outcomes underscore the importance of swing states, where polling results are closely contested and could shift the election results.
+States in gray represent missing or incomplete polling data, where predictions were not possible.
+According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.
 
-## Predicted Election Winner By State
+The results highlight the geographic divide between the two candidates.
+Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.
+Harris, on the other hand, is projected to win in the Midwest and along the East Coast, with stronger polling percentages in traditionally Democratic-leaning regions.
+
 ```{r}
@@
 
-It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.
-These uncolored states reflect the absence of reliable or sufficient polling information required for an accurate prediction.
-This underscores the limitations of the data used in the model, where predictions are only made for states with adequate polling coverage.
-As a result, certain regions lack predictions, which could be addressed by including more comprehensive polling data in future analyses.
-
 ### Summary of Predicted Election Results by State
 
-@tbl-modelresults presents a summary of the predicted 2024 US Presidential Election results on a state-by-state basis.
-It includes the average polling percentages for Donald Trump and Kamala Harris, the predicted probability of Trump winning each state based on a logistic regression model, and the predicted winner in each state.
+@tbl-modelresults further breaks down the polling data by state, showing the average polling percentages for each candidate and the predicted win probability for Trump.
+For example, in Florida, Trump’s polling average is 53%, and his predicted win probability is almost certain at 99.93%.
+Conversely, in Michigan, Harris leads with an average of 46.27% of the vote, resulting in a 96.77% predicted probability that she will win the state.
+
+In swing states like North Carolina and Nevada, where the polling percentages are closer, Trump is still predicted to win with probabilities of 94.36% and 51.23%, respectively.
+These battleground states show tighter polling averages and thus greater uncertainty in the predicted outcomes.
 
@@
 
-According to our model results, in **Florida**, Trump’s polling percentage (53.00%) is substantially higher than Harris’s (40.50%), leading to a near-certain predicted win for Trump with a probability of 0.9998.
-Conversely, in **Michigan**, Harris holds a slight polling lead (46.27% vs.
-44.52%), resulting in a predicted win for her with a lower probability of 0.036.
-Several states, such as **Montana** and **Missouri**, have extremely high probabilities of a Trump victory, while others, like **Nebraska CD-2** and **North Carolina**, lean toward Harris.
-The inclusion of predicted probabilities adds nuance by highlighting the confidence levels of the model’s predictions for each state, especially in swing states where polling percentages are closely contested.
+The results of our predictive model suggest that Donald Trump is likely to win key states such as Florida, Missouri, and Montana, while Kamala Harris is expected to secure victories in states like Michigan, Virginia, and Pennsylvania.
+Close contests are predicted in battleground states like Nevada and North Carolina, where Trump holds a slight edge.
 
@@
 
-The results of the logistic regression model provide insights into the predicted winner of the 2024 US Presidential Election based on polling percentages for Donald Trump and Kamala Harris.
-The side-by-side bar graph comparing polling averages across states reveals a clear geographic divide.
-Trump tends to perform better in southern and midwestern states, while Harris shows stronger support in the northeastern and western regions.
-Swing states, where the polling differences between the two candidates are narrower, emerge as crucial battlegrounds that could determine the election outcome.
-The map visualization reinforces these trends, visually highlighting where each candidate is expected to win.
-These results align with historical voting patterns, but the model also emphasizes the close competition in key states that are not traditionally battlegrounds.
+The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast.
+The predictions align with historical voting patterns, with Trump’s strongest support coming from traditionally conservative regions and Harris garnering more support in Democratic-leaning areas.
+Notably, several battleground states, such as Nevada and North Carolina, show a close contest, reflecting the competitive nature of the 2024 election.
+These results suggest that the election may hinge on the outcomes in these key states, where even small shifts in polling percentages could influence the final result.
+The predicted probabilities in states like Michigan, Virginia, and Pennsylvania indicate a strong chance for Harris, while Trump's significant lead in states like Florida and Missouri solidifies his dominance in certain regions.
+Overall, the model provides valuable insights into the competitive landscape, highlighting the states that could be decisive in determining the election outcome.
 
@@
 # Additional data details
+
+It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.
+These uncolored states reflect the absence of reliable or sufficient polling information required for an accurate prediction.
+This underscores the limitations of the data used in the model, where predictions are only made for states with adequate polling coverage.
+As a result, certain regions lack predictions, which could be addressed by including more comprehensive polling data in future analyses.
 

### 2024-10-20 paper/paper.qmd
committing changes to clean_data script

@@
 
-# Remove any states with missing polling data for Trump or Harris
-#state_average <- state_average %>%
-  #filter(!is.na(Trump_pct) & !is.na(Harris_pct))
-
-# Proceed with the prediction for all states
-#state_average <- state_average %>%
-  #mutate(predicted_prob = predict(logistic_model, newdata = state_average, type = "response"))
-
-# Classify the predicted winner: Trump if probability > 0.5, else Harris
-#state_average <- state_average %>%
-  #mutate(predicted_winner = ifelse(predicted_prob > 0.5, "Trump", "Harris"))
-
-#### Map Visualization ####
-#plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
-  #scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
-  #labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
-  #theme(legend.position = "right")
 ```
@@
 # Additional data details
-
 It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.

### 2024-10-20 scripts/03-clean_data.R

@@ -18,6 +18,10 @@ library(dplyr)
 clean_poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
   clean_names()
 
+# Replace "nebraska-cd20" with "nebraska"
+clean_poll_data <- clean_poll_data |>
+  mutate(state = str_replace(state, "Nebraska-CD20", "Nebraska"))
+
 # Filter data to Harris estimates based on high-quality polls after she declared
 clean_poll_data <- clean_poll_data |>
   filter(numeric_grade >= 3, transparency_score > 5, candidate_name %in% c("Donald Trump", "Kamala Harris"))

### 2024-10-20 paper/paper.qmd
committing text changes to data - overview section

@@
 
-Placholder - We use the statistical programming language R [@citeR] for all data manipulation, modeling, and visualization in this study.
-Our data [@shelter] was obtained from the public polling aggregation website FiveThirtyEight, which provides comprehensive polling data for the 2024 US Presidential Election [@fivethirtyeight].
-The dataset includes national polls conducted by various polling organizations, capturing voter support for the two leading candidates, Donald Trump and Kamala Harris.
-The dataset records critical variables such as the pollster, sample size, and percentage of voter support for each candidate.
+We use the statistical programming language R [@citeR] to perform data cleaning and analysis using the datasets we obtained from the Project 538 online database [@fivethirtyeight].
+The database provides a wealth of information on the current cycle of presidential general election polls.
+It consists of the polling results from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
+Information about the transparency of methodology as well as the quality of the data is present as well, denoted as transparency score and numeric grades, which are significant to our data cleaning process to ensure the quality of data.
+Polling results are specific to different states in the US, mostly focusing on the six battleground states that are the likeliest to determine the outcome of the elections.
+Thus, we chose to use states, candidate names, and percentage of support to perform statistical analysis and gather our results.
 

### 2024-10-20 paper/paper.qmd
committing text changes to paper.qmd

@@
 
-We use the statistical programming language R [@citeR] to perform data cleaning and analysis using the datasets we obtained from the Project 538 online database [@fivethirtyeight].
+We use the statistical programming language R [@citeR] to perform data cleaning and analysis using the datasets we obtained from the Project 538 online database [@FiveThirtyEight].
 The database provides a wealth of information on the current cycle of presidential general election polls.
@@
 # Render the simplest possible table
-knitr::kable(average_table, caption = "Summary Table of Polling Averages by State")
+knitr::kable(average_table)
 
@@
   geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
-  labs(title = "Average Poll Percentages for Trump and Harris by State",
-       x = "State", y = "Average Percentage") +
+  labs(x = "State", y = "Average Percentage") +
   scale_fill_manual(name = "Candidate", values = c("Trump_pct" = "red", "Harris_pct" = "blue")) +
@@
 # Model {#sec-model}
-
 The goal of our modeling strategy is to predict the winner of the 2024 US Presidential Election in each state based on polling data, while also assessing the likelihood of Donald Trump or Kamala Harris winning.
@@
 #| label: fig-statemap
-#| fig-cap: Predicted Election Winner by State
+#| fig-cap: Predicted Winner of 2024 US Presidential Election by State
 #| echo: false
@@
   scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
-  labs(title = "Predicted Winner of 2024 US Presidential Election by State") +
   theme(legend.position = "right")
@@
 
-knitr::kable(summary_table, caption = "Summary of Predicted 2024 US Presidential Election Results by State", escape = TRUE)
+knitr::kable(summary_table, escape = TRUE)
 
@@
 
-#### Model Summary Table ####
-# Create and display a model summary table
-#modelsummary::modelsummary(
-  #list("First Model" = logistic_model),
-  #statistic = c("std.error", "p.value"),  # Use standard error and p-value
-  #fmt = 2,  # Format numbers to 2 decimal places
-  #output = "markdown"  # Output format for Quarto or RMarkdown
-#)
-
 ```

### 2024-10-20 paper/paper.qmd
committing changes to readme

@@
 ---
-title: "2024 US Presidential Election Model"
-subtitle: "My subtitle if needed"
+title: "Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest"
 author: 
@@
 
-@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state, based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
+@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on a map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
 States where Trump is predicted to win are shown in red, while states where Harris is predicted to win are displayed in blue.

### 2024-10-20 scripts/03-clean_data.R

@@ -18,9 +18,9 @@ library(dplyr)
 clean_poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
   clean_names()
 
-# Replace "nebraska-cd20" with "nebraska"
+# Replace "Nebraska CD-2" with "Nebraska"
 clean_poll_data <- clean_poll_data |>
-  mutate(state = str_replace(state, "Nebraska-CD20", "Nebraska"))
+  mutate(state = str_replace(state, "Nebraska CD-2", "Nebraska"))
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 clean_poll_data <- clean_poll_data |>

### 2024-10-20 paper/paper.qmd
committing text changes to measurement in data section

@@
 
-Placeholder - The measurement of voter support in our dataset is derived from polling data, which serves as a representation of public sentiment during the 2024 U.S.
+The measurement of voter support in the dataset is derived from raw polling data obtained from the Project 538 online database, which serves as a representation of potential election results for the 2024 U.S.
 Presidential Election.
-Each entry in the dataset corresponds to a poll conducted by various polling organizations, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
-To ensure the quality of the data, only polls from reputable pollsters with a numeric grade of 3 or higher and transparency scores above 6 were included.
+Each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
+Each pollster has their approach of sampling respondents as well as their methodology of approaching respondents, which affects each pollster’s numeric grades and transparency scores.
+Therefore, to ensure the quality of the data, only polls from reputable pollsters with a numeric grade of 3 or higher and transparency scores above 6 were used for our paper.
 These polls measure voter preferences through structured survey questions, and the results are aggregated to create state-level averages for each candidate.
-The dataset thus transforms individual polling responses into a broader measurement of voter sentiment, reflecting the competitive dynamics between Trump and Harris across different states.
+The dataset thus transforms general voter sentiment into individual data points that reflect the competitive dynamics between Trump and Harris across different states.
 

### 2024-10-20 paper/paper.qmd
committ changes to paper.qmd

@@
 
-Our model predicts election outcomes by comparing the average polling percentages for Trump and Harris in each state.
+Our model utilizes the Bayesian logistic regression model to predict election outcomes by comparing the average polling percentages for Trump and Harris in each state.
 The results highlight the geographic distribution of support, with some states clearly favoring one candidate over the other.
@@
 ```
+The comparison shows significant variability in support for each candidate.
+In some states like **Montana** and **Missouri**, Trump’s support is notably higher, with a clear margin over Harris.
+In contrast, **Colorado** stands out as having nearly equal polling support for both candidates, suggesting a closely contested race in that state.
+Swing states such as **Florida**, **North Carolina**, and **Pennsylvania** exhibit relatively close polling percentages, indicating that the election outcomes in these states could be pivotal and difficult to predict.
+The plot also includes states like **Virginia** and **Texas**, where the support levels are more balanced, though Trump seems to hold a slight lead.
+The state labeled as **NA** at the end may indicate missing or incomplete data, emphasizing the need for comprehensive polling coverage in all states for more reliable predictions.
 
 # Model {#sec-model}
+
 The goal of our modeling strategy is to predict the winner of the 2024 US Presidential Election in each state based on polling data, while also assessing the likelihood of Donald Trump or Kamala Harris winning.
@@
 ## Model set-up
+
 The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win.
@@
 The model was implemented using the rstanarm package in R, which supports Bayesian inference with default priors for logistic regression.
-Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].
 
@@
 
-### Summary of Predicted Election Results by State
+## Summary of Predicted Election Results by State
 
@@
 However, states with narrower margins reflect greater uncertainty, which is critical in understanding election dynamics.
+Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].
 
-## Interpretation of Polling Averages by State
-
-The bar graph displays the average polling percentages for Donald Trump (in red) and Kamala Harris (in blue) across various states.
-The comparison shows significant variability in support for each candidate.
-In some states like **Montana** and **Missouri**, Trump’s support is notably higher, with a clear margin over Harris.
-In contrast, **Colorado** stands out as having nearly equal polling support for both candidates, suggesting a closely contested race in that state.
-
-Swing states such as **Florida**, **North Carolina**, and **Pennsylvania** exhibit relatively close polling percentages, indicating that the election outcomes in these states could be pivotal and difficult to predict.
-The plot also includes states like **Virginia** and **Texas**, where the support levels are more balanced, though Trump seems to hold a slight lead.
-
-The state labeled as **NA** at the end may indicate missing or incomplete data, emphasizing the need for comprehensive polling coverage in all states for more reliable predictions.
-
-Overall, the side-by-side comparison highlights the states where each candidate holds a significant lead and where the race remains competitive, offering a detailed view of the election's geographic dynamics.
-
-## Weaknesses and next steps
+## Weaknesses and Next Steps
 

### 2024-10-20 paper/paper.qmd
commit text changes

@@
 ## Predicted Election Winner By State
-
 @fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on a map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
@@
 ## Summary of Predicted Election Results by State
-
 @tbl-modelresults further breaks down the polling data by state, showing the average polling percentages for each candidate and the predicted win probability for Trump.

### 2024-10-20 paper/paper.qmd
commit qmd file

@@
 Harris, on the other hand, is projected to win in the Midwest and along the East Coast, with stronger polling percentages in traditionally Democratic-leaning regions.
-
 ```{r}
@@
 These battleground states show tighter polling averages and thus greater uncertainty in the predicted outcomes.
-
 ```{r}
@@
 ```
-
 The results of our predictive model suggest that Donald Trump is likely to win key states such as Florida, Missouri, and Montana, while Kamala Harris is expected to secure victories in states like Michigan, Virginia, and Pennsylvania.

### 2024-10-20 paper/paper.qmd
commit text changes to paper.qmd

--- 2024-10-20 paper/paper.qmd
@@
 
-The results highlight the geographic divide between the two candidates.
-Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.
-Harris, on the other hand, is projected to win in the Midwest and along the East Coast, with stronger polling percentages in traditionally Democratic-leaning regions.
-
 ```{r}
@@
   theme(legend.position = "right")
-
 ```
+The results highlight the geographic divide between the two candidates.
+Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.
+Harris, on the other hand, is projected to win in the Midwest and along the East Coast, with stronger polling percentages in traditionally Democratic-leaning regions.
 
@@
 knitr::kable(summary_table, escape = TRUE)
-
 ```
-
 The results of our predictive model suggest that Donald Trump is likely to win key states such as Florida, Missouri, and Montana, while Kamala Harris is expected to secure victories in states like Michigan, Virginia, and Pennsylvania.

### 2024-10-20 paper/paper.qmd
commit text changes
committing text changes to paper.qmd

@@
 
+\newpage
+
 ## Summary of Predicted Election Results by State

### 2024-10-20 paper/paper.qmd
committing formatting changes in paper.qmd

--- 2024-10-20 paper/paper.qmd
@@
 # Results {#sec-results}
-
 ## Predicted Election Winner By State
-
 @fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on a map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
@@
 According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.
-
 ```{r}
@@
 
+\newpage
+
 ## Summary of Predicted Election Results by State
-
 @tbl-modelresults further breaks down the polling data by state, showing the average polling percentages for each candidate and the predicted win probability for Trump.

### 2024-10-20 paper/paper.qmd
committing text changes to paper.qmd

@@
 States in gray represent missing or incomplete polling data, where predictions were not possible.
-According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.
 ```{r}
@@
 ```
-The results highlight the geographic divide between the two candidates.
+
+According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.The results highlight the geographic divide between the two candidates.
 Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.

### 2024-10-20 paper/paper.qmd
committing formatting changes to paper.qmd
committing changes to paper.qmd

@@
 
-\newpage
-
 ## Summary of Predicted Election Results by State
@@
 
-The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast.
+Placeholder - The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast.
 The predictions align with historical voting patterns, with Trump’s strongest support coming from traditionally conservative regions and Harris garnering more support in Democratic-leaning areas.
@@
 
-The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state.
+Placeholder - The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state.
 The coefficients for the predictor variables (polling percentages for Trump and Harris) indicate a strong relationship between candidate support and the probability of winning.
@@
 
-Despite the model's predictive power, there are several weaknesses that should be noted.
+Placeholder - Despite the model's predictive power, there are several weaknesses that should be noted.
 First, polling data itself is prone to biases, such as underrepresentation of certain demographic groups or inaccuracies in sampling.
@@
 
-To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data.
+Placeholder - To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data.
 Factors such as voter turnout models, economic conditions, and social media sentiment analysis may provide a more holistic view of voter behavior.

### 2024-10-21 paper/paper.qmd
committing rendered paper.qmd

@@
 ## Diagnostics
-@tbl-modelsummary presents the results of the logistic regression model predicting the winner of the 2024 US Presidential Election in each state based on polling percentages for Donald Trump and Kamala Harris.
-The coefficients indicate that an increase in Trump’s polling percentage (\( \beta = 8.72 \)) is positively associated with the probability of winning a state, whereas an increase in Harris’s polling percentage (\( \beta = -31.63 \)) is negatively associated with Trump’s probability of winning.
-The standard errors for both predictors are quite large, suggesting high variability in the estimates and possible uncertainty in the model.
+
+Out-of-sample testing was conducted to test our model’s reliability in predicting potential presidential election outcomes.
+We train our model on a training set of our data and then test its performance on the test set of our data, we evaluate how well the model generalizes to new data that it was not trained on.
+This process helps reveal whether the model is overfitting to the training data or truly capturing the underlying patterns that apply more broadly.
+The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction.
+Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.
 
 ```{r}
+
 #| echo: false
-#| eval: true
-#| label: tbl-modelsummary
-#| tbl-cap: "Explanatory models of election outcome based on polling percentages for Trump and Harris"
+#| message: false
 #| warning: false
 
-# Ensure the model exists
-if (exists("logistic_model")) {
-  # Create and display a model summary table in LaTeX format for PDF
-  modelsummary::modelsummary(
-    list("First Model" = logistic_model),
-    statistic = c("std.error", "p.value"),  # Use standard error and p-value
-    fmt = 2,  # Format numbers to 2 decimal places
-    output = "latex"  # Use LaTeX for PDF output
-  )
-} else {
-  print("Model not found")
-}
+# Out-of-Sample Testing of Our Model
+# Load necessary libraries
+library(caret)
+library(rstanarm)
+library(ggplot2)
+library(dplyr)
+
+# Split the data into training (70%) and test (30%) sets
+set.seed(123)  # For reproducibility
+trainIndex <- createDataPartition(state_average$winner, p = 0.7, list = FALSE)
+train_data <- state_average[trainIndex, ]
+test_data <- state_average[-trainIndex, ]
+
+# Train the Bayesian logistic regression model on the training data
+logistic_model_train <- stan_glm(winner ~ Trump_pct + Harris_pct, data = train_data, 
+                                 family = binomial, prior = student_t(df = 7), 
+                                 chains = 4, iter = 2000, refresh = 0, verbose = FALSE)
+
+# Make predictions on the test set
+test_predictions <- posterior_predict(logistic_model_train, newdata = test_data, draws = 1)
+
+# Convert predictions to binary outcome (1 if Trump wins, 0 if Harris wins)
+test_pred_binary <- ifelse(colMeans(test_predictions) > 0.5, 1, 0)
+
+# Calculate accuracy
+accuracy <- mean(test_pred_binary == test_data$winner)
+accuracy_message <- paste("Test Accuracy: ", round(accuracy * 100, 2), "%")
+print(accuracy_message)
+
+# Optionally, calculate RMSE
+rmse_value <- sqrt(mean((test_pred_binary - test_data$winner)^2))
+rmse_message <- paste("Test RMSE: ", round(rmse_value, 4))
+print(rmse_message)
+
+# Confusion Matrix for additional diagnostics
+conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)
+print(conf_matrix)
 
@@
 
-The model diagnostics, including the AIC (6.0) and BIC (8.1), suggest that the model fits the data reasonably well given the small sample size of 15 observations.
-However, the low log-likelihood (0.000) and root mean squared error (RMSE = 0.00) indicate that the model may not be well-calibrated, likely due to the small dataset or the structure of the data.
-These results emphasize the importance of using a larger sample size.
+Our accuracy test generated a 75% score for the model, which means that our model is quite reliable for making predictions about state-level election outcomes.
+Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.
+The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes.
+The confusion matrix breaks down the model’s predictions into four categories: 1.
+correctly predicted Trump wins, 2.
+correctly predicted Harris wins, 3.
+Incorrectly predicted Trump wins, 4.
+Incorrectly predicted Harris wins.
+The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
 

### 2024-10-22 paper/paper.qmd
committing qmd file
committing changes to code in model appendix

@@
 
-This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris.
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data [@FiveThirtyEight] to forecast the likely winner between Donald Trump and Kamala Harris.
 We aggregate high-quality poll data to create a logistic regression model that estimates the probability of a Trump or Harris victory in each state.
@@
 
-The measurement of voter support in the dataset is derived from raw polling data obtained from the Project 538 online database, which serves as a representation of potential election results for the 2024 U.S.
+The measurement of voter support in the dataset is derived from raw polling data obtained from the Project 538 online database [@FiveThirtyEight], which serves as a representation of potential election results for the 2024 U.S.
 Presidential Election.
@@
 Potential biases in polling data could also affect predictions.
-The model was implemented using the rstanarm package in R, which supports Bayesian inference with default priors for logistic regression.
+The model was implemented using the [@rstanarm] package in [@citeR], which supports Bayesian inference with default priors for logistic regression.
 
@@
 
-Placeholder - The results of our model reveal clear geographic divides in voter sentiment, with Donald Trump predicted to win in many southern and midwestern states, while Kamala Harris is expected to perform well in the Midwest and on the East Coast.
-The predictions align with historical voting patterns, with Trump’s strongest support coming from traditionally conservative regions and Harris garnering more support in Democratic-leaning areas.
-Notably, several battleground states, such as Nevada and North Carolina, show a close contest, reflecting the competitive nature of the 2024 election.
-These results suggest that the election may hinge on the outcomes in these key states, where even small shifts in polling percentages could influence the final result.
-The predicted probabilities in states like Michigan, Virginia, and Pennsylvania indicate a strong chance for Harris, while Trump's significant lead in states like Florida and Missouri solidifies his dominance in certain regions.
-Overall, the model provides valuable insights into the competitive landscape, highlighting the states that could be decisive in determining the election outcome.
+Since the electoral college system makes state-level forecasts crucial, and the U.S.
+operates under a predominantly two-party system, the focus of election predictions should be on battleground states.
+While many states are strongly aligned with either the Democratic or Republican party, a handful of key battleground states—where neither party has overwhelming dominance—will likely determine the overall outcome of the election.
+These battleground states, often evenly split in voter sentiment, hold immense significance because their electoral votes can swing the election in favor of one candidate or the other.
+The results of our model underscore the critical role that these battleground states will play in the 2024 election.
+Although many states are predictably stable due to their party loyalty, it is the six battleground states that will serve as the main arena for competition.
+In these states, where voter preferences are not as firmly entrenched, even minor shifts in public opinion or turnout could lead to vastly different outcomes.
+As a result, a comprehensive and precise prediction of voting behavior in these battleground states is far more important than in states where the result is a foregone conclusion.
 
@@
 
-Placeholder - The logistic regression model used in this analysis successfully captures the relationship between polling percentages and the likelihood of a candidate winning a state.
-The coefficients for the predictor variables (polling percentages for Trump and Harris) indicate a strong relationship between candidate support and the probability of winning.
-The model outputs probabilities for each state, which are useful for gauging the certainty of predictions, especially in closely contested states.
-The model’s performance is reinforced by the robustness of the results in states where one candidate has a clear polling lead, providing accurate predictions.
-However, states with narrower margins reflect greater uncertainty, which is critical in understanding election dynamics.
-Convergence diagnostics and model checking were conducted to ensure reliable estimates, with further details and alternative models provided in [Appendix -@sec-model-details].
+Out-of-sample testing results [Appendix -@sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
+presidential election.
+Our accuracy test generated a 75% score for the model, indicating our model's high reliability for making predictions about state-level election outcomes.
+Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.
+
+The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes.
+The confusion matrix breaks down the model’s predictions into four categories: 1.
+correctly predicted Trump wins, 2.
+correctly predicted Harris wins, 3.
+Incorrectly predicted Trump wins, 4.
+Incorrectly predicted Harris wins.
+The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
+Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].
 
@@
 
-Placeholder - Despite the model's predictive power, there are several weaknesses that should be noted.
-First, polling data itself is prone to biases, such as underrepresentation of certain demographic groups or inaccuracies in sampling.
-These issues can distort the predictions, particularly in states with fewer or lower-quality polls.
-Additionally, voter turnout and last-minute shifts in public opinion are not accounted for, which can significantly alter the outcome, especially in close races.
-The model also assumes that polling percentages directly translate to votes, overlooking other factors like voter mobilization efforts and campaign dynamics that may influence election outcomes.
+While our model provides valuable insights into the electoral landscape, it faces limitations due to missing data for certain states, which results in some states being left blank in our predictions.
+Although our primary focus is on understanding the dynamics within battleground states, the lack of information from other states can lead to incomplete predictions which may obscure broader trends affecting voter sentiment.
 
-Placeholder - To improve the accuracy and reliability of election forecasts, future iterations of the model could incorporate additional variables beyond polling data.
-Factors such as voter turnout models, economic conditions, and social media sentiment analysis may provide a more holistic view of voter behavior.
-Additionally, real-time polling data integration could allow for dynamic updates to predictions as the election nears, improving the timeliness and accuracy of forecasts.
-Moreover, regional adjustments to account for differences in polling methodology or demographic shifts would help to address some of the biases inherent in polling data.
-Expanding the model to include these aspects would offer a more comprehensive understanding of the election landscape and enhance the precision of predictions.
+To address these weaknesses and improve the prediction, we propose several next steps: actively seeking additional data sources to fill gaps, including polling data and historical records; conducting sensitivity analyses to understand the impact of missing data on predictions; and analyzing existing trends within battleground states for more insights.
+Another weakness of our model is the lack of consideration for the national vote, which may impact our election predictions.
+The national vote refers to the total number of votes cast by citizens across the country during an election, representing the support for each presidential candidate.
+Although the national vote does not directly determine the outcome of the election due to the Electoral College system, it can still provide valuable insights into voter sentiment and trends.
+By not considering the national vote into our analysis, we risk missing potential influences on voting behavior and electoral outcomes.
 
@@
 
-# Additional data details
+# Additional Data Details
 It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.
@@
 
-## Diagnostics
+## Diagnostics {#-sec-model-diagnostics}
 
@@
 ```{r}
-
 #| echo: false
@@
 accuracy_message <- paste("Test Accuracy: ", round(accuracy * 100, 2), "%")
-print(accuracy_message)
 
@@
 rmse_message <- paste("Test RMSE: ", round(rmse_value, 4))
-print(rmse_message)
 
@@
 conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)
-print(conf_matrix)
 
@@
 
-Our accuracy test generated a 75% score for the model, which means that our model is quite reliable for making predictions about state-level election outcomes.
-Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.
-The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes.
-The confusion matrix breaks down the model’s predictions into four categories: 1.
-correctly predicted Trump wins, 2.
-correctly predicted Harris wins, 3.
-Incorrectly predicted Trump wins, 4.
-Incorrectly predicted Harris wins.
-The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
+[@barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
+The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
+The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
+This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
+
+```{r}
+#| label: barplot-diagnostics
+#| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
+#| echo: false
+#| warning: false
+#| message: false
+#| cache: false
+
+# Visualizing the RMSE and accuracy in a plot
+accuracy_rmse_df <- data.frame(
+  Metric = c("Accuracy", "RMSE"),
+  Value = c(accuracy, rmse_value)
+)
+
+# Visualizing the RMSE and accuracy in a plot with skinnier bars
+ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
+  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
+  labs(y = "Value") +
+  theme_light()
+
+```
+[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
+In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
+However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
+This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-confusion-matrix
+#| tbl-cap: "Confusion Matrix of Test Predictions"
+#| warning: false
+
+# Visualizing the confusion matrix as a table
+library(knitr)
+kable(conf_matrix)
+
+```
+
+# Pollster Methodology Overview and Evaluation
+
+The following information about the Siena College/New York Times poll methodology is based on the details provided in [@nyt-siena-methodology].
+
+The Siena College/New York Times poll for the 2024 presidential election uses a robust sampling approach to ensure accuracy and representativeness.
+The poll applies a stratified dual-frame sample, drawing from both landlines and cell phones.
+Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia.
+The sampling population is all registered voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
+Given the nature of the presidential elections, the decision is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are likeliest to decide the outcome of the race.
+The voters’ information is taken from the L-2 voter file, which includes details such as voter registration and history of participation in previous elections.
+
+To refine their sample, Siena College adjusts the data by key demographic and political variables such as region, race/ethnicity, party affiliation, education, and voting patterns from the 2020 election.
+These adjustments help ensure the sample accurately represents the electorate.
+Additionally, they use a “likely voter screen,” which combines self-reported likelihood of voting with historical voter behavior to estimate how likely respondents are to vote in the upcoming election.
+At every step of the survey, Sienna/NYT uses the information in the datasets to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions.
+Once the survey is complete, they compare their respondents to the voter file and use a process known as weighting to ensure that the sample reflects the broader voting population.
+This combination of historical data and weighted adjustments ensures the poll is designed to predict election outcomes as accurately as possible.
+
+The New York Times/Siena College pollsters handle non-response by using weighting and adjusting their sample to correct for any biases that might emerge due to individuals not responding.
+Specifically, they adjust on multiple demographic and political variables such as age, gender, education, race/ethnicity, and party affiliation.
+They also account for variations in voter likelihood and previous voting patterns, ensuring that the sample represents the likely electorate as accurately as possible.
+This helps reduce the potential bias from non-response, especially since certain demographic groups or political affiliations may be less likely to respond to polls.
+By rebalancing the sample, they ensure that even if some groups have lower response rates, their representation in the final poll results aligns with what is expected based on historical trends and current voter enthusiasm.
+
+# Idealized Methodology
+To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via [@US-election-survey].
+
+## Sampling Approach
+
+Using data from voter files, which contain demographic information about registered voters, we aim to ensure that our sample accurately represents the population.
+To achieve this, we will use stratified random sampling.
+Given the importance of state-level forecasts due to the Electoral College system, we will first stratify by state.
+Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographic critical for predicting voter behavior.
+The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).
+Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth.
+Voting history will be classified into those who voted in the previous election and non-voters, while party registration will include Democrats, Republicans, and Independents.
+
+For recruiting participants, we will begin by sending survey invitations to our selected voters via email, offering a $20 incentive to encourage participation.
+To improve response rates and reduce non-response bias, we will send a reminder email three days later.
+This approach helps minimize labor costs to some extent.
+If we do not receive a response within a week(7 days), we will follow up by phone to reach those who did not respond to the email invitation.
+
+## Data Validation
+
+After completing the survey, we will adjust the data through weighting to ensure the sample accurately reflects the broader population for predictive purposes.
+More weight will be assigned proportionally to each stratum based on its size, and additional weight will be given to respondents from strata that are less likely to participate in surveys.
+And to avoid duplicate responses, each voter will be assigned a unique ID with only the first response from each ID being retained.
+To ensure the selected voters are completing the survey correctly, we will  track responses in real-time using a centralized system.
+Additionally, we will include validation questions in the survey to catch careless or fraudulent responses.
 

### 2024-10-22 paper/paper.qmd
committing text changes to idealized methodology appendix section

@@
 ```{r}
+#| echo: false
+#| eval: true
 #| label: barplot-diagnostics
 #| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
-#| echo: false
 #| warning: false
 #| message: false
-#| cache: false
 
@@
 # Idealized Methodology
+
+Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions.
+This approach incorporates best practices and innovative techniques to ensure a comprehensive prediction of voter behavior in the upcoming 2024 U.S.
+presidential election.
 To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via [@US-election-survey].
@@
 This approach helps minimize labor costs to some extent.
-If we do not receive a response within a week(7 days), we will follow up by phone to reach those who did not respond to the email invitation.
+If we do not receive a response within a week (7 days), we will follow up by phone to reach those who did not respond to the email invitation.
 

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd for model testing

@@
 
-Out-of-sample testing results [Appendix -@sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
+Out-of-sample testing results [Appendix @sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
 presidential election.
@@
 The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
-Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].
+Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix @sec-model-diagnostics].
 
@@
 
-## Diagnostics {#-sec-model-diagnostics}
+## Diagnostics {#sec-model-diagnostics}
 
@@
 conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)
+rownames(conf_matrix) <- c("Predicted 0 (Harris)", "Predicted 1 (Trump)")
+colnames(conf_matrix) <- c("Actual 0 (Harris)", "Actual 1 (Trump)")
 
@@
 
-[@barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
+[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
 The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
@@
 #| eval: true
-#| label: barplot-diagnostics
+#| label: fig-barplot-diagnostics
 #| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
@@
 presidential election.
-To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via [@US-election-survey].
+To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the following url: https://docs.google.com/forms/d/1S6aN5Q82orjKYJI4i7001sPyn5jhRkgNIkI9teiQq_o/edit [@US-election-survey].
 

### 2024-10-22 paper/paper.qmd
committing format changes

@@
 ```
+\newpage
 
@@
 ```
+
 [@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).

### 2024-10-22 paper/paper.qmd
committing text changes

@@
 # Additional Data Details
-It is important to note that several states remain uncolored (in gray), indicating missing or incomplete polling data for those states.
-These uncolored states reflect the absence of reliable or sufficient polling information required for an accurate prediction.
-This underscores the limitations of the data used in the model, where predictions are only made for states with adequate polling coverage.
-As a result, certain regions lack predictions, which could be addressed by including more comprehensive polling data in future analyses.
+Placeholder
 
@@
 ```
-\newpage
 

### 2024-10-22 paper/paper.qmd
committing text changes and adding sections under model diagnostics

@@
 ## Diagnostics {#sec-model-diagnostics}
-
 Out-of-sample testing was conducted to test our model’s reliability in predicting potential presidential election outcomes.
@@
 
+### Model Diagnostics Result
 [@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
@@
 ```
-
+### Confusion Matrix
 [@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd

@@
 ## Model set-up
-
 The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win.
@@
 Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.
+The testing was implemented using the following packages in [@citeR]: [@caret], [@rstanarm], [@ggplot2], [@dplyr].
 
@@
   theme_light()
-
 ```

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd

@@
 ```
+\newpage
+
 ### Confusion Matrix
+
 [@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).

### 2024-10-22 paper/paper.qmd
committing text changes to paper.qmd

@@
 ### Model Diagnostics Result
-[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
-The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
-The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
-This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
 
@@
 ```
-\newpage
+[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
+The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
+The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
+This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
 
 ### Confusion Matrix
-
-[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
-In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
-However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
-This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 
@@
 ```
+
+[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
+In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
+However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
+This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 

### 2024-10-22 paper/paper.qmd
commit text changes

@@
 # Additional Data Details
-Placeholder
+Given the nature of the presidential election, the electoral college is more impactful and influential on the possible outcome compared to the nationwide popular votes.
+Therefore, the majority of the pollsters choose to focus on the battleground states to perform their polling, to gain more insight into the public sentiment specific to certain states.
+For instance, Sienna/NYT mentions that they have in-state interviewers who call their respondents in states such as Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
+This could explain the discrepancies in our dataset, where there are many missing values from states other than the battleground ones.
+Furthermore, many pollsters focus on state-wide polls instead of nationwide polls given the nature of the election, resulting in fewer data points or missing data for nationwide poll results.
+Even though the national popular vote and the results in the electoral college don’t line up a lot of the time, the popular votes can still provide information about issues and opinions that are shaping the election as a whole.
+Therefore, even if we are focusing on selective states, results from other states and popular votes still remain relevant to outcome prediction.
 
@@
 
+[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
+The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
+The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
+
 ```{r}
@@
 ```
-[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
-The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
-The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
 This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.

### 2024-10-22 paper/paper.qmd
committing text changes to appendix in paper.qmd and bib

@@
 ```
+
 The results of our predictive model suggest that Donald Trump is likely to win key states such as Florida, Missouri, and Montana, while Kamala Harris is expected to secure victories in states like Michigan, Virginia, and Pennsylvania.
@@
 Therefore, even if we are focusing on selective states, results from other states and popular votes still remain relevant to outcome prediction.
+Additionally, during our data cleaning process, we added constraints of high numeric grades as well as high transparency scores which led to fewer data points for us to build the model on.
+For example, for the state of Colorado, when we set the transparency score to higher than 5, and the numeric grade to larger and equal to 3, there are no data points for Kamala Harris.
+This is represented by the empty row in [@fig-averagetable].
 

### 2024-10-22 paper/paper.qmd
committing formatting changes to paper.qmd

@@
 ### Confusion Matrix
+[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
+In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
+However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
+This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 
@@
 ```
-
-[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
-In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
-However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
-This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 

### 2024-10-22 paper/paper.qmd
commit text changes to paper.qmd

@@
 Presidential Election.
-Each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
+Each entry in the data corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
 Each pollster has their approach of sampling respondents as well as their methodology of approaching respondents, which affects each pollster’s numeric grades and transparency scores.
+
 Therefore, to ensure the quality of the data, only polls from reputable pollsters with a numeric grade of 3 or higher and transparency scores above 6 were used for our paper.
@@
 Potential biases in polling data could also affect predictions.
-The model was implemented using the [@rstanarm] package in [@citeR], which supports Bayesian inference with default priors for logistic regression.
+The model was implemented using the 'rstanarm' [@rstanarm] package in R[@citeR], which supports Bayesian inference with default priors for logistic regression.
 
@@
 ## Predicted Election Winner By State
-@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on a map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
+@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on the US map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
 States where Trump is predicted to win are shown in red, while states where Harris is predicted to win are displayed in blue.
@@
 
-According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.The results highlight the geographic divide between the two candidates.
+According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.
+The results highlight the geographic divide between the two candidates.
 Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.
@@
 These battleground states, often evenly split in voter sentiment, hold immense significance because their electoral votes can swing the election in favor of one candidate or the other.
+
 The results of our model underscore the critical role that these battleground states will play in the 2024 election.
@@
 The confusion matrix breaks down the model’s predictions into four categories: 1.
-correctly predicted Trump wins, 2.
-correctly predicted Harris wins, 3.
+orrectly predicted Trump wins, 2.
+Correctly predicted Harris wins, 3.
 Incorrectly predicted Trump wins, 4.
@@
 To address these weaknesses and improve the prediction, we propose several next steps: actively seeking additional data sources to fill gaps, including polling data and historical records; conducting sensitivity analyses to understand the impact of missing data on predictions; and analyzing existing trends within battleground states for more insights.
+
 Another weakness of our model is the lack of consideration for the national vote, which may impact our election predictions.
@@
 For instance, Sienna/NYT mentions that they have in-state interviewers who call their respondents in states such as Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
-This could explain the discrepancies in our dataset, where there are many missing values from states other than the battleground ones.
+
+This could explain the discrepancies in our data, where there are many missing values from states other than the battleground ones.
 Furthermore, many pollsters focus on state-wide polls instead of nationwide polls given the nature of the election, resulting in fewer data points or missing data for nationwide poll results.
@@
 Therefore, even if we are focusing on selective states, results from other states and popular votes still remain relevant to outcome prediction.
+
 Additionally, during our data cleaning process, we added constraints of high numeric grades as well as high transparency scores which led to fewer data points for us to build the model on.
@@
 We train our model on a training set of our data and then test its performance on the test set of our data, we evaluate how well the model generalizes to new data that it was not trained on.
-This process helps reveal whether the model is overfitting to the training data or truly capturing the underlying patterns that apply more broadly.
+
+This process helps reveal whether the model is over-fitting to the training data or truly capturing the underlying patterns that apply more broadly.
 The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction.
 Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.
-The testing was implemented using the following packages in [@citeR]: [@caret], [@rstanarm], [@ggplot2], [@dplyr].
+The testing was implemented using the following packages in R[@citeR]: caret[@caret], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr].
 
@@
 
-The following information about the Siena College/New York Times poll methodology is based on the details provided in [@nyt-siena-methodology].
+The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times[@nyt-siena-methodology].
 
-The Siena College/New York Times poll for the 2024 presidential election uses a robust sampling approach to ensure accuracy and representativeness.
-The poll applies a stratified dual-frame sample, drawing from both landlines and cell phones.
+The Siena College/New York Times poll for the 2024 presidential election uses a robust sampling approach to ensure accuracy and relevance.
+The poll applies a stratified dual-frame sample, drawing from both land-lines and cell phones.
 Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia.
 The sampling population is all registered voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
+
 Given the nature of the presidential elections, the decision is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are likeliest to decide the outcome of the race.
@@
 Additionally, they use a “likely voter screen,” which combines self-reported likelihood of voting with historical voter behavior to estimate how likely respondents are to vote in the upcoming election.
-At every step of the survey, Sienna/NYT uses the information in the datasets to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions.
+
+At every step of the survey, Sienna/NYT uses the information in the data to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions.
 Once the survey is complete, they compare their respondents to the voter file and use a process known as weighting to ensure that the sample reflects the broader voting population.
@@
 They also account for variations in voter likelihood and previous voting patterns, ensuring that the sample represents the likely electorate as accurately as possible.
+
 This helps reduce the potential bias from non-response, especially since certain demographic groups or political affiliations may be less likely to respond to polls.
@@
 presidential election.
-To represent our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the following url: https://docs.google.com/forms/d/1S6aN5Q82orjKYJI4i7001sPyn5jhRkgNIkI9teiQq_o/edit [@US-election-survey].
+
+To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@sec-references] [@US-election-survey].
 
@@
 Given the importance of state-level forecasts due to the Electoral College system, we will first stratify by state.
+
 Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographic critical for predicting voter behavior.
 The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).
+
 Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth.
@@
 
-For recruiting participants, we will begin by sending survey invitations to our selected voters via email, offering a $20 incentive to encourage participation.
+For the recruitment of participants, we will begin by sending survey invitations to our selected voters via email, offering a $20 incentive to encourage participation.
 To improve response rates and reduce non-response bias, we will send a reminder email three days later.
-This approach helps minimize labor costs to some extent.
+This approach will help minimize expenses to some extent.
 If we do not receive a response within a week (7 days), we will follow up by phone to reach those who did not respond to the email invitation.
@@
 
-# References
+# References {sec-references}

### 2024-10-22 paper/paper.qmd
committing text changes to the paper.qmd

@@
 These priors were chosen to reflect the assumption that, before seeing the data, there is no strong reason to favor one candidate over the other across states, but to allow room for the data to adjust predictions as polling percentages differ.
-
 We run the model in R [@citeR] using the `rstanarm` package [@rstanarm].
@@
 Potential biases in polling data could also affect predictions.
-The model was implemented using the 'rstanarm' [@rstanarm] package in R[@citeR], which supports Bayesian inference with default priors for logistic regression.
+The model was implemented using the 'rstanarm' [@rstanarm] package in R [@citeR], which supports Bayesian inference with default priors for logistic regression.
 
@@
 Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.
-The testing was implemented using the following packages in R[@citeR]: caret[@caret], rstanarm[@rstanarm], ggplot2[@ggplot2], dplyr[@dplyr].
+The testing was implemented using the following packages in R [@citeR]: caret [@caret], rstanarm [@rstanarm], ggplot2 [@ggplot2], dplyr [@dplyr].
 
@@
 
-The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times[@nyt-siena-methodology].
+The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times [@nyt-siena-methodology].
 
@@
 presidential election.
-
 To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@sec-references] [@US-election-survey].
@@
 The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).
-
 Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth.

### 2024-10-22 paper/paper.qmd
committing formatting and text changes to paper.qmd
committing text changes and cross references to paper.qmd

@@
 
-# Introduction
+# Introduction {#sec-introduction}
 
@@
 The remainder of this paper is structured as follows.
-In Section 2, we describe the data and variables used in the analysis.
-Section 3 outlines the model setup and estimation strategy.
-Section 4 presents the results, including the predicted election outcomes and visualizations.
-Finally, Section 5 discusses the implications and limitations of our findings.
+In [@sec-data], we describe the data and variables used in the analysis.
+[@sec-model] outlines the model setup and estimation strategy.
+[@sec-results] presents the results, including the predicted election outcomes and visualizations.
+Finally, [@sec-discussion] discusses the implications and limitations of our findings.
 
@@
 Conversely, in Michigan, Harris leads with an average of 46.27% of the vote, resulting in a 96.77% predicted probability that she will win the state.
-
 In swing states like North Carolina and Nevada, where the polling percentages are closer, Trump is still predicted to win with probabilities of 94.36% and 51.23%, respectively.
@@
 [@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
-In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
-However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
-This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 
@@
 ```
+In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
+However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
+This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
 
@@
 presidential election.
-To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@sec-references] [@US-election-survey].
+To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [@sec-references].
 

### 2024-10-31 paper/paper.qmd
committing text changes
committing changes

@@
 
-@fig-averagetable provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model.
+[@fig-averagetable] provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model.
 
@@
 To further illustrate the distribution of support, a side-by-side bar graph is provided, comparing the average polling percentages for both candidates across all states.
-@fig-barplot offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
+[@fig-barplot] offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
 The side-by-side comparison represents the percentage of support each candidate has received, based on aggregated poll data from pollsters with high-quality scores.

### 2024-10-31 paper/paper.qmd
committing changes to paper

@@
 ---
-title: "Forecasting the 2024 Election: Trump Dominates the South, Harris Leads in the Midwest"
+title: "Predicting a Trump Victory: A State-Level Analysis of the 2024 US Presidential Election Using Bayesian Modeling"
+subtitle: "State-by-State Insights on Voter Dynamics Favoring Trump Over Harris in Key Battlgrounds"
 author: 
@@
 A logistic regression model estimates the probability of Trump winning, with polling percentages as predictor variables.
-The model’s results highlight key battleground states and offer insights into voter dynamics, while also reflecting on the limitations of polling data in forecasting political outcomes."
+The model’s results highlight key battleground states and offer insights into voter dynamics, while also reflecting on the limitations of polling data in forecasting political outcomes.
+Our findings suggest that Trump holds an adantage in critical swing states, positinoning him as the likely winner of the election."
 format: pdf
@@
 library(rstanarm)
-library(readr)
-library(dplyr)
-library(ggplot2)
 library(usmap)
+library(arrow)
 ```
@@
 By focusing on high-quality polling data, our model improves the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.
+
+The results of our model predict a victory for Donald Trump in the 2024 U.S.
+Presidential Election.
+Key battleground states, such as Florida, Ohio, and North Carolina, lean toward Trump, emphasizing his advantage in critical regions.
+Our analysis points to Trump as the probable winner, with swing states playing a decisive role in determining the final outcome.
 
@@
 # Load cleaned data
-clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
+clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
 
@@
 average_table <- state_average |>
-  select(state, Trump_pct, Harris_pct, winner) |>
-  mutate(predicted_winner = ifelse(winner == 1, "Trump", "Harris"))
+  select(
+    State = state, 
+    Trump = Trump_pct, 
+    Harris = Harris_pct, 
+    Winner = winner
+    ) |>
+  mutate(Winner = ifelse(Winner == 1, "Trump", "Harris"))
 
@@
 
-ggplot(stateaverage_long, aes(x=reorder(state, -support_pct), y= support_pct, fill = candidate)) + 
+ggplot(stateaverage_long, aes(x = reorder(state, -support_pct), y = support_pct, fill = candidate)) + 
   geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
   labs(x = "State", y = "Average Percentage") +
-  scale_fill_manual(name = "Candidate", values = c("Trump_pct" = "red", "Harris_pct" = "blue")) +
+  scale_fill_manual(
+    name = "Candidate", 
+    values = c("Trump_pct" = "red", "Harris_pct" = "blue"),
+    labels = c("Trump_pct" = "Trump", "Harris_pct" = "Harris")
+  ) +
   theme_minimal() +
@@
 
-The results of our predictive model suggest that Donald Trump is likely to win key states such as Florida, Missouri, and Montana, while Kamala Harris is expected to secure victories in states like Michigan, Virginia, and Pennsylvania.
-Close contests are predicted in battleground states like Nevada and North Carolina, where Trump holds a slight edge.
+The predictive model forecasts Donald Trump as the likely winner of the 2024 U.S.
+Presidential Election.
+By analyzing state-level polling data through a Bayesian logistic regression model, the study estimates the probability of Trump or Kamala Harris winning in each state.
+The model identifies a significant geographic split in voter support: Trump is expected to dominate in southern and midwestern states, where his polling percentages consistently surpass those of Harris.
+Key states projected to favor Trump include Florida, Georgia, and Ohio, which are pivotal in securing an Electoral College majority.
+Harris, on the other hand, holds stronger polling percentages in East Coast states and parts of the Midwest, yet these leads appear insufficient to overcome Trump’s advantage across the broader landscape of competitive states.
+
+In particular, battleground states show varied support but lean overall toward Trump, reflecting his edge in areas like Arizona, Nevada, and North Carolina.
+By aggregating high-quality polls filtered for transparency and reliability, the model underscores Trump’s stable base in regions that carry substantial electoral weight.
+The predictions suggest that while Harris may perform well in several Democratic strongholds, Trump’s consistent support across critical swing states positions him as the probable victor in the upcoming election.
 
@@
 
-Additionally, during our data cleaning process, we added constraints of high numeric grades as well as high transparency scores which led to fewer data points for us to build the model on.
-For example, for the state of Colorado, when we set the transparency score to higher than 5, and the numeric grade to larger and equal to 3, there are no data points for Kamala Harris.
-This is represented by the empty row in [@fig-averagetable].
+During our data cleaning process, we added constraints of high numeric grades and high transparency scores, leading to fewer data points for us to build the model on.
+A numeric grade is a numeric rating given to the pollster indicating their quality or reliability, with a highest rating of 3.0.
+The transparency score is a score reflecting the pollster’s transparency about their methodology where the highest score is 10.
+Therefore, to ensure high-quality and reliable data to build our model on, we filtered pollsters with numeric grades of 3.0 and transparency scores of 5 or higher.
+This is one of the reasons why there are no data points available for Kamala Harris for the state of Colorado, represented by the empty row @fig-averagetable.
+
+Furthermore, in the full raw data set from 538 [@FiveThirtyEight], for the state of Colorado, even without any constraints for numeric grade or transparency scores, there are only 3 data points for Kamala Harris.
+After reviewing the datasets, we found that except for the 3 available data points, all the rest of the polls ended in June 2024, which is a month earlier than when Biden dropped out of the race before Harris entered the presidential election race.
+It would not be possible that there are data points available for Harris if she was not participating in the race yet.
+The state of Colorado is a blue state, with the Democrats winning the state in every presidential election since 2008.
+Since our paper focuses on mainly the battleground states to predict election outcomes, limited data from a blue state like Colorado would not be too influential on our results.
 

### 2024-10-31 scripts/03-clean_data.R

@@ -11,8 +11,8 @@
 library(tidyverse)
 library(janitor)
 library(lubridate)
-library(readr)
-library(dplyr)
+library(arrow)
+
 
 #### Prepare Dataset ####
 clean_poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
@@ -24,10 +24,10 @@ clean_poll_data <- clean_poll_data |>
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 clean_poll_data <- clean_poll_data |>
-  filter(numeric_grade >= 3, transparency_score > 5, candidate_name %in% c("Donald Trump", "Kamala Harris"))
+  filter(numeric_grade >= 2.7, transparency_score > 5, candidate_name %in% c("Donald Trump", "Kamala Harris"))
   
 head(clean_poll_data)
 
 #### Save data ####
-write_csv(clean_poll_data, "data/02-analysis_data/analysis_data.csv")
+write_parquet(clean_poll_data, "data/02-analysis_data/analysis_data.parquet")
 

### 2024-11-03 paper/paper.qmd
committing changes to model and results

@@
 library(arrow)
+library(knitr)
+library(Matrix)
+library(splines)
+library(kableExtra)
+library(broom.mixed)
 ```
@@
 #| label: fig-averagetable
-#| fig-cap: Summary Table of Polling Averages by State
+#| fig-cap: Summary of 2024 U.S.
+Presidential Election Polling Averages by State for Trump and Harris, with Indicated Leading Candidate.
+States with missing data, such as Colorado, are labeled as "NA."
 #| echo: false
@@
 #| message: false
-
 # Load cleaned data
@@
   summarise(
-    Trump_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE),
-    Harris_pct = mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE)
+    Trump_pct = mean(Trump_pct, na.rm = TRUE),
+    Harris_pct = mean(Harris_pct, na.rm = TRUE)
   )
@@
 
+# Prepare table for display with rounding to 1 decimal place
 average_table <- state_average |>
+  mutate(
+    Trump_pct = round(Trump_pct, 1),
+    Harris_pct = round(Harris_pct, 1)
+  ) |>
   select(
     State = state, 
-    Trump = Trump_pct, 
-    Harris = Harris_pct, 
+    "Trump %" = Trump_pct, 
+    "Harris %" = Harris_pct, 
     Winner = winner
@@
 
-# Render the simplest possible table
+# Render the table
 knitr::kable(average_table)
@@
 ```
+
+Based on the table provided, Trump is predicted to win in the states of Arizona, Florida, Georgia, Missouri, Montana, Nevada, Ohio, and Texas, while Harris is predicted to win in Michigan, Nebraska, North Carolina, Pennsylvania, Virginia, and Wisconsin.
+However, this does not necessarily determine the overall winner of the election, as U.S.
+presidential elections are decided by the Electoral College rather than by a simple majority in each state.
+Each state carries a different number of electoral votes, and a candidate must secure at least 270 electoral votes to win the presidency.
+To predict the overall winner, it would be necessary to assign each state’s electoral votes to the respective predicted winner and calculate the total for each candidate.
+With the electoral vote counts for each state, we could determine which candidate is projected to win the election.
 
@@
 #| label: fig-barplot
-#| fig-cap: Average Poll Percentage by State
+#| fig-cap: "Comparison of 2024 U.S.
+Presidential Election Polling Averages for Trump and Harris by State.
+Bars represent the average polling percentages for each candidate, with states missing data, such as Colorado, labeled accordingly."
 #| echo: false
@@
     values = c("Trump_pct" = "red", "Harris_pct" = "blue"),
-    labels = c("Trump_pct" = "Trump", "Harris_pct" = "Harris")
+    labels = c("Trump_pct" = "Trump %", "Harris_pct" = "Harris %")
   ) +
@@
 ```
-The comparison shows significant variability in support for each candidate.
-In some states like **Montana** and **Missouri**, Trump’s support is notably higher, with a clear margin over Harris.
-In contrast, **Colorado** stands out as having nearly equal polling support for both candidates, suggesting a closely contested race in that state.
-Swing states such as **Florida**, **North Carolina**, and **Pennsylvania** exhibit relatively close polling percentages, indicating that the election outcomes in these states could be pivotal and difficult to predict.
-The plot also includes states like **Virginia** and **Texas**, where the support levels are more balanced, though Trump seems to hold a slight lead.
-The state labeled as **NA** at the end may indicate missing or incomplete data, emphasizing the need for comprehensive polling coverage in all states for more reliable predictions.
+
+@fig-bar-plot illustrates the state-by-state polling data by comparing the average support percentages for Trump and Harris.
+In each state, the red bar represents Trump's average polling percentage, while the blue bar represents Harris's.
+Trump shows stronger support in states like **Montana**, **Missouri**, **Florida**, and **Nevada**, where his average polling percentage surpasses that of Harris.
+Conversely, Harris has a slight edge in states such as **Wisconsin**, **North Carolina**, **Nebraska**, **Pennsylvania**, and **Virginia**.
+Some states, like Ohio and **Georgia**, exhibit closely matched polling averages, indicating a tighter race.
+This visualization underscores the regional variation in candidate support across states, reflecting diverse voter preferences that could influence the final outcome based on the Electoral College distribution.
 
@@
 The goal of our modeling strategy is to predict the winner of the 2024 US Presidential Election in each state based on polling data, while also assessing the likelihood of Donald Trump or Kamala Harris winning.
-We employ a Bayesian logistic regression model, which allows for probabilistic predictions of binary outcomes, making it well-suited for the task of predicting election results.
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+We use a Bayesian hierarchical modeling approach with natural splines to capture time-dependent trends in support and random effects to account for state-level variability.
+This strategy enables us to model changes in candidate support over time across states, allowing for more nuanced, probabilistic predictions of each candidate’s support levels and overall election outcomes.
+
+Separate models are fit for each candidate’s polling support, with state-specific random intercepts and a natural spline for time, allowing the model to capture state-level deviations from the national average as well as time-based changes in support.
+This hierarchical Bayesian approach enables partial pooling across states, stabilizing predictions for states with limited polling data.
+Further background details, model specifications, and diagnostics are included in [Appendix -@sec-model-details].
 
 ## Model set-up
-The binary outcome variable $y_i$ is defined as 1 if Donald Trump is predicted to win state $i$, and 0 if Kamala Harris is predicted to win.
-The predictor variables are $\beta_i$, which represents the average percentage of polling support for Donald Trump in state $i$, and $\gamma_i$, the average percentage of support for Kamala Harris in state $i$.
-The relationship between these variables and the outcome is captured through a logistic link function.
+The binary outcome variable \( y_i \) represents whether Donald Trump is predicted to lead in state \( i \) based on polling averages for both candidates over time.
+Here, \( y_i = 1 \) indicates a Trump lead in state \( i \), and \( y_i = 0 \) indicates a lead for Kamala Harris.
+The model estimates the support levels for each candidate based on polling data using a Gaussian likelihood function, with time trends captured by a natural spline and state-level variation captured through random intercepts.
 
+The predictor variables are \( \text{Trump\_pct}_i \), the predicted percentage of support for Donald Trump in state \( i \), and \( \text{Harris\_pct}_i \), the predicted percentage of support for Kamala Harris in state \( i \).
+A separate model is fit for each candidate’s support over time, capturing the time-varying nature of support with a natural spline function and allowing for state-specific random effects.
+
+The model is specified as follows:
+
+\[
 \begin{align*}
-y_i | \mu_i &\sim \text{Bernoulli}(\mu_i) \\
-\mu_i &= \alpha + \beta_i + \gamma_i \\
-\alpha &\sim \text{Normal}(0, 2.5) \\
-\beta_i &\sim \text{Normal}(0, 2.5) \\
-\gamma_i &\sim \text{Normal}(0, 2.5)
+\text{Trump\_pct}_i &\sim \text{Normal}(\mu_{\text{Trump}, i}, \sigma_{\text{Trump}}) \\
+\mu_{\text{Trump}, i} &= \alpha_{\text{Trump}} + f(\text{end\_date\_num}_i) + u_{\text{state}}^{\text{Trump}} \\
+u_{\text{state}}^{\text{Trump}} &\sim \text{Normal}(0, \tau_{\text{Trump}}) \\
+\text{Harris\_pct}_i &\sim \text{Normal}(\mu_{\text{Harris}, i}, \sigma_{\text{Harris}}) \\
+\mu_{\text{Harris}, i} &= \alpha_{\text{Harris}} + f(\text{end\_date\_num}_i) + u_{\text{state}}^{\text{Harris}} \\
+u_{\text{state}}^{\text{Harris}} &\sim \text{Normal}(0, \tau_{\text{Harris}})
 \end{align*}
+\]
 
-We use weakly informative Normal priors for the intercept ($\alpha_i$) and the coefficients ($\beta_i$) and $\gamma_i$), centered around zero with a standard deviation of 2.5.
-These priors were chosen to reflect the assumption that, before seeing the data, there is no strong reason to favor one candidate over the other across states, but to allow room for the data to adjust predictions as polling percentages differ.
-We run the model in R [@citeR] using the `rstanarm` package [@rstanarm].
-The model uses default priors from `rstanarm`, and estimates the likelihood of Donald Trump winning each state based on polling averages for both candidates.
+where:
+- \( \text{Trump\_pct}_i \) and \( \text{Harris\_pct}_i \) are the polling support levels for Donald Trump and Kamala Harris in state \( i \).
+- \( f(\text{end\_date\_num}_i) \) is a natural spline function of time (measured as `end_date_num`), with 4 degrees of freedom, capturing the non-linear time trends in support.
+- \( \alpha_{\text{Trump}} \) and \( \alpha_{\text{Harris}} \) are intercepts for the models, representing the baseline support levels for each candidate.
+- \( u_{\text{state}}^{\text{Trump}} \) and \( u_{\text{state}}^{\text{Harris}} \) are state-level random effects, accounting for differences in baseline support across states.
+
+For both models, we use weakly informative Normal priors.
+The intercepts \( \alpha_{\text{Trump}} \) and \( \alpha_{\text{Harris}} \) are given Normal priors with a mean of 0 and a standard deviation of 10, reflecting prior uncertainty about the baseline support levels.
+The time trend coefficients are also given Normal priors with mean 0 and standard deviation 5, allowing for flexibility in capturing the temporal changes in support.
+State-level random effects for each candidate are modeled with a Normal distribution centered at 0, with the variance estimated from the data.
+
+The model was implemented in R using the `brms` package, which allows for Bayesian inference using `Stan` as the backend.
+This approach captures the dynamic nature of polling support over time, while accounting for state-specific variability and allowing us to estimate the probability of Trump or Harris leading in each state.
 
@@
 
-logistic_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+trump_model <-
+  readRDS(file = here::here("models/trump_model.rds"))
+
+harris_model <-
+  readRDS(file = here::here("models/harris_model.rds"))
 ```
@@
 
-We expect a positive relationship between polling support for each candidate and their probability of winning a state.
-As Donald Trump’s polling percentage ($\beta_i$) increases, the likelihood of him winning state $i$ rises.
-Similarly, as Kamala Harris's polling percentage ($\gamma_i$) increases, her chances of winning that state increase.
+We expect a positive relationship between each candidate's polling support over time and their likelihood of winning a given state.
+As Donald Trump’s polling percentage in a state increases, his probability of leading in that state rises, while higher support for Kamala Harris increases her probability of winning.
+By modeling each candidate's support trends over time, we capture shifts in voter sentiment that reflect campaign dynamics, national events, and other time-dependent factors influencing support levels.
 
-Bayesian logistic regression was chosen for its ability to model binary outcomes like win/loss predictions while incorporating prior knowledge and uncertainty, helping mitigate overfitting through regularization with priors.
-This model strikes a balance between flexibility and interpretability, given that polling percentages are continuous and the outcome is binary.
-The model assumes polling data accurately reflects voter intentions and that state outcomes are independent of each other, although it does not account for interactions between states or regional voting trends.
-Potential biases in polling data could also affect predictions.
-The model was implemented using the 'rstanarm' [@rstanarm] package in R [@citeR], which supports Bayesian inference with default priors for logistic regression.
+A Bayesian hierarchical model was chosen for its ability to handle the complexity of election forecasting, where both temporal trends and state-specific variations play significant roles.
+Bayesian inference allows us to incorporate prior beliefs and uncertainties about candidate support, while the hierarchical structure enables partial pooling across states.
+This is particularly useful for states with sparse data, as the model can "borrow strength" from the national trend, stabilizing predictions.
 
-# Results {#sec-results}
-## Predicted Election Winner By State
-@fig-statemap shows the predicted winner of the 2024 US Presidential Election by state on the US map [@usmap], based on a logistic regression model using aggregated polling data for Donald Trump and Kamala Harris.
-States where Trump is predicted to win are shown in red, while states where Harris is predicted to win are displayed in blue.
-States in gray represent missing or incomplete polling data, where predictions were not possible.
-```{r}
-#| label: fig-statemap
-#| fig-cap: Predicted Winner of 2024 US Presidential Election by State
-#| echo: false
-#| warning: false
-#| message: false
-#| cache: false
+The inclusion of a natural spline for time allows us to model non-linear changes in support, capturing important shifts in polling trends as the election approaches.
+Random effects for each state account for baseline differences in support levels, reflecting the unique political climate of each state and reducing the risk of overfitting by allowing state-level deviations from the national average.
 
-#### Workspace setup ####
-library(tidyverse)
-library(usmap)
-library(ggplot2)
-library(rstanarm)
+This model assumes that polling data reasonably reflects voter intentions and that each state’s outcome can be modeled independently of others, though regional trends are partially addressed by state-level random effects.
+Potential biases inherent in polling data, such as underrepresented demographics or sampling error, could affect our predictions.
+The model was implemented in R using the brms package, which provides Bayesian inference via Stan, allowing for flexible, probabilistic predictions of each candidate's support across states.
 
-# Check for missing values in state_average and ensure data is clean
-state_average <- state_average %>%
-  filter(!is.na(Trump_pct) & !is.na(Harris_pct)) %>%
-  ungroup()
-
-# Proceed with prediction after ensuring the data is clean
-predicted_prob <- predict(logistic_model, newdata = state_average, type = "response")
-
-# Ensure predicted_prob matches the number of rows in state_average
-if (length(predicted_prob) != nrow(state_average)) {
-  stop("Mismatch between predicted_prob and state_average rows")
-}
-
-# Add predicted_prob to state_average
-state_average <- state_average %>%
-  mutate(predicted_prob = predicted_prob)
-
-# Classify the predicted winner: Trump if probability > 0.5, else Harris
-state_average <- state_average %>%
-  mutate(predicted_winner = ifelse(predicted_prob > 0.5, "Trump", "Harris"))
-
-# Map visualization showing predicted winner by state
-plot_usmap(data = state_average, values = "predicted_winner", regions = "states") +
-  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue"), name = "Predicted Winner") +
-  theme(legend.position = "right")
-```
-
-According to the model, Trump is expected to win key states, including Florida, Missouri, and Montana, while Harris is forecasted to lead in Michigan, Virginia, and Pennsylvania.
-The results highlight the geographic divide between the two candidates.
-Trump is predicted to perform well in southern and midwestern states, where his polling averages are significantly higher.
-Harris, on the other hand, is projected to win in the Midwest and along the East Coast, with stronger polling percentages in traditionally Democratic-leaning regions.
-
-## Summary of Predicted Election Results by State
-@tbl-modelresults further breaks down the polling data by state, showing the average polling percentages for each candidate and the predicted win probability for Trump.
-For example, in Florida, Trump’s polling average is 53%, and his predicted win probability is almost certain at 99.93%.
-Conversely, in Michigan, Harris leads with an average of 46.27% of the vote, resulting in a 96.77% predicted probability that she will win the state.
-In swing states like North Carolina and Nevada, where the polling percentages are closer, Trump is still predicted to win with probabilities of 94.36% and 51.23%, respectively.
-These battleground states show tighter polling averages and thus greater uncertainty in the predicted outcomes.
+## Model Summary Results
 ```{r}
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Predicted 2024 US Presidential Election outcomes by state, based on polling percentages for Trump and Harris"
+#| label: tbl-fixed-effects-summary
+#| tbl-cap: "Fixed Effects Summary for Trump and Harris Models."
 #| warning: false
 
-#### Create and Display the Summary Table ####
-summary_table <- state_average %>%
-  select(state, Trump_pct, Harris_pct, predicted_prob, predicted_winner) %>%
+# Load necessary libraries
+library(dplyr)
+library(broom)        # For the tidy function
+library(knitr)
+library(kableExtra)
+
+#| echo: false
+#| eval: true
+#| label: tbl-fixed-effects-summary
+#| tbl-cap: "Fixed Effects Summary for Trump and Harris Models."
+#| warning: false
+
+# Load necessary libraries
+library(dplyr)
+library(broom)        # For the tidy function
+library(knitr)
+library(kableExtra)
+
+# Summarize fixed effects for Trump model
+trump_summary <- tidy(trump_model, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
   rename(
-    "State" = state,
-    "Trump Polling (%)" = Trump_pct,
-    "Harris Polling (%)" = Harris_pct,
-    "Predicted Trump Win Probability" = predicted_prob,
-    "Predicted Winner" = predicted_winner
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `Lower 95% CI` = conf.low,
+    `Upper 95% CI` = conf.high
+  ) %>%
+  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Remove non-alphanumeric characters
+
+# Summarize fixed effects for Harris model
+harris_summary <- tidy(harris_model, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `Lower 95% CI` = conf.low,
+    `Upper 95% CI` = conf.high
+  ) %>%
+  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Remove non-alphanumeric characters
+
+# Combine summaries into one table
+model_summary_table <- full_join(trump_summary, harris_summary, by = "term", suffix = c(" Trump", " Harris"))
+
+# Display the table with kable
+kable(model_summary_table, caption = "Fixed Effects Summary for Trump and Harris Models") %>%
+  kable_styling(full_width = FALSE, position = "center", font_size = 8)
+
+```
+
+# Results {#sec-results}
+
+## National Support Trends Over Time
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-national-trend
+#| fig-cap: "National Support Trends Over Time for the 2024 U.S.
+Presidential Election.
+The plot displays the predicted national average support for Donald Trump (red) and Kamala Harris (blue) over time, based on a Bayesian hierarchical model with a natural spline fit.
+The trends show changes in polling support for both candidates throughout 2024, highlighting key shifts in public opinion as the election approaches."
+#| warning: false
+
+# Generate a sequence of dates for national predictions
+prediction_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
+
+# Predict national support for each date
+national_trend <- prediction_dates %>%
+  map_dfr(~ {
+    data <- data.frame(state = unique(clean_poll_data$state), end_date = .x)
+    data$end_date_num <- as.numeric(data$end_date - min(clean_poll_data$end_date))
+
+    trump_pred <- mean(posterior_predict(trump_model, newdata = data))
+    harris_pred <- mean(posterior_predict(harris_model, newdata = data))
+
+    tibble(date = .x, Trump = trump_pred, Harris = harris_pred)
+  })
+
+# Plot national support trends
+ggplot(national_trend, aes(x = date)) +
+  geom_line(aes(y = Trump, color = "Trump")) +
+  geom_line(aes(y = Harris, color = "Harris")) +
+  labs(title = "National Support Trends Over Time", y = "Predicted Support", x = "Date") +
+  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
+  theme_minimal()
+
+```
+
+## State Level Support Trends Over Time
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-state-trend
+#| fig-cap: "State-Level Support Trends for Trump and Harris Over Time."
+#| warning: false
+
+# Generate predictions for each state and date
+state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
+state_prediction_grid <- expand.grid(state = unique(clean_poll_data$state), end_date = state_dates) %>%
+  mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))
+
+# Predict support for each state and date
+state_predictions <- state_prediction_grid %>%
+  rowwise() %>%
+  mutate(
+    Trump_Support = mean(posterior_predict(trump_model, newdata = cur_data())),
+    Harris_Support = mean(posterior_predict(harris_model, newdata = cur_data()))
   )
 
-knitr::kable(summary_table, escape = TRUE)
+# Plot state-level support trends over time with adjusted x-axis labels
+ggplot(state_predictions, aes(x = end_date)) +
+  geom_line(aes(y = Trump_Support, color = "Trump")) +
+  geom_line(aes(y = Harris_Support, color = "Harris")) +
+  labs(title = "State-Level Support Trends for Trump and Harris Over Time", y = "Predicted Support", x = "Date") +
+  facet_wrap(~ state) +
+  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
+  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  # Set date format and break frequency
+  theme_minimal() +
+  theme(
+    legend.position = "none",
+    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
+    strip.text = element_text(size = 10),  # Control facet label size if needed
+    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
+    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
+  )
+
+```
+## Election Day Predictions
+
+## Probability of Winning by State on Election Day
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: fig-state-probabilities
+#| fig-cap: "Probability of Trump Winning by State on Election Day."
+#| warning: false
+
+# Set election day or target prediction date
+election_day <- as.Date("2024-11-05")  # Example election day
+
+# Predict probability of Trump winning by comparing posterior predictions for each state
+state_predictions <- clean_poll_data %>%
+  distinct(state) %>%
+  mutate(end_date = election_day, end_date_num = as.numeric(election_day - min(clean_poll_data$end_date)))
+
+# Calculate probability of Trump winning in each state
+state_predictions <- state_predictions %>%
+  rowwise() %>%
+  mutate(
+    prob_trump_wins = mean(posterior_predict(trump_model, newdata = cur_data()) >
+                           posterior_predict(harris_model, newdata = cur_data()))
+  )
+
+# Plot probability of winning by state
+ggplot(state_predictions, aes(x = reorder(state, prob_trump_wins), y = prob_trump_wins, fill = prob_trump_wins > 0.5)) +
+  geom_col() +
+  coord_flip() +
+  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"), labels = c("Harris", "Trump")) +
+  labs(title = "Probability of Trump Winning by State on Election Day", x = "State", y = "Probability Trump Wins") +
+  theme_minimal() +
+  theme(legend.position = "none")
+
 ```
 
-The predictive model forecasts Donald Trump as the likely winner of the 2024 U.S.
-Presidential Election.
-By analyzing state-level polling data through a Bayesian logistic regression model, the study estimates the probability of Trump or Kamala Harris winning in each state.
-The model identifies a significant geographic split in voter support: Trump is expected to dominate in southern and midwestern states, where his polling percentages consistently surpass those of Harris.
-Key states projected to favor Trump include Florida, Georgia, and Ohio, which are pivotal in securing an Electoral College majority.
-Harris, on the other hand, holds stronger polling percentages in East Coast states and parts of the Midwest, yet these leads appear insufficient to overcome Trump’s advantage across the broader landscape of competitive states.
+## US Map Showing Predicted Winner by State on Election Day
 
-In particular, battleground states show varied support but lean overall toward Trump, reflecting his edge in areas like Arizona, Nevada, and North Carolina.
-By aggregating high-quality polls filtered for transparency and reliability, the model underscores Trump’s stable base in regions that carry substantial electoral weight.
-The predictions suggest that while Harris may perform well in several Democratic strongholds, Trump’s consistent support across critical swing states positions him as the probable victor in the upcoming election.
+```{r}
+
+#| echo: false
+#| eval: true
+#| label: fig-us-map
+#| fig-cap: "Predicted Winner by State on Election Day."
+#| warning: false
+
+# Add predicted winner column based on `prob_trump_wins`
+state_predictions <- state_predictions %>%
+  mutate(predicted_winner = ifelse(prob_trump_wins > 0.5, "Trump", "Harris"))
+
+# Plot the US map with predicted winner for each state
+plot_usmap(data = state_predictions, values = "predicted_winner", regions = "states") +
+  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
+  labs(title = "Predicted Winner by State on Election Day", fill = "Predicted Winner") +
+  theme(legend.position = "right")
+
+```
 
@@
 
-```{r}
-#| echo: false
-#| message: false
-#| warning: false
-
-# Out-of-Sample Testing of Our Model
-# Load necessary libraries
-library(caret)
-library(rstanarm)
-library(ggplot2)
-library(dplyr)
-
-# Split the data into training (70%) and test (30%) sets
-set.seed(123)  # For reproducibility
-trainIndex <- createDataPartition(state_average$winner, p = 0.7, list = FALSE)
-train_data <- state_average[trainIndex, ]
-test_data <- state_average[-trainIndex, ]
-
-# Train the Bayesian logistic regression model on the training data
-logistic_model_train <- stan_glm(winner ~ Trump_pct + Harris_pct, data = train_data, 
-                                 family = binomial, prior = student_t(df = 7), 
-                                 chains = 4, iter = 2000, refresh = 0, verbose = FALSE)
-
-# Make predictions on the test set
-test_predictions <- posterior_predict(logistic_model_train, newdata = test_data, draws = 1)
-
-# Convert predictions to binary outcome (1 if Trump wins, 0 if Harris wins)
-test_pred_binary <- ifelse(colMeans(test_predictions) > 0.5, 1, 0)
-
-# Calculate accuracy
-accuracy <- mean(test_pred_binary == test_data$winner)
-accuracy_message <- paste("Test Accuracy: ", round(accuracy * 100, 2), "%")
-
-# Optionally, calculate RMSE
-rmse_value <- sqrt(mean((test_pred_binary - test_data$winner)^2))
-rmse_message <- paste("Test RMSE: ", round(rmse_value, 4))
-
-# Confusion Matrix for additional diagnostics
-conf_matrix <- table(Predicted = test_pred_binary, Actual = test_data$winner)
-rownames(conf_matrix) <- c("Predicted 0 (Harris)", "Predicted 1 (Trump)")
-colnames(conf_matrix) <- c("Actual 0 (Harris)", "Actual 1 (Trump)")
-
-```
-
 ### Model Diagnostics Result
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: fig-barplot-diagnostics
-#| fig-cap: "Model: Diagnostics: Accuracy and RMSE"
-#| warning: false
-#| message: false
-
-# Visualizing the RMSE and accuracy in a plot
-accuracy_rmse_df <- data.frame(
-  Metric = c("Accuracy", "RMSE"),
-  Value = c(accuracy, rmse_value)
-)
-
-# Visualizing the RMSE and accuracy in a plot with skinnier bars
-ggplot(accuracy_rmse_df, aes(x = Metric, y = Value)) +
-  geom_bar(stat = "identity", fill = "steelblue", width = 0.2) +  # Adjusted width for skinnier bars
-  labs(y = "Value") +
-  theme_light()
-```
 This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-confusion-matrix
-#| tbl-cap: "Confusion Matrix of Test Predictions"
-#| warning: false
-
-# Visualizing the confusion matrix as a table
-library(knitr)
-kable(conf_matrix)
-
-```
 In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).

### 2024-11-03 scripts/03-clean_data.R

@@ -7,27 +7,47 @@
 # Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 # Any other information needed? N/A
 
-#### Workspace setup ####
+
+#### Data Cleaning Script ####
+# Load necessary libraries
 library(tidyverse)
 library(janitor)
 library(lubridate)
-library(arrow)
 
+# Load the dataset
+poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
+  clean_names()  
 
-#### Prepare Dataset ####
-clean_poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
-  clean_names()
+# Filter and select relevant columns
+clean_poll_data <- poll_data |>
+  select(state, end_date, pollster, numeric_grade, transparency_score, candidate_name, pct) |>
+  filter(
+    !is.na(state),
+    !is.na(numeric_grade),
+    !is.na(transparency_score),
+    numeric_grade >= 3.0,          # numeric_grade threshold
+    transparency_score > 5,        # transparency_score threshold
+    candidate_name %in% c("Donald Trump", "Kamala Harris")
+  ) |>
+  mutate(
+    end_date = as.Date(end_date, format = "%m/%d/%y"),
+    end_date_num = as.numeric(end_date - min(end_date, na.rm = TRUE))
+  )
 
 # Replace "Nebraska CD-2" with "Nebraska"
 clean_poll_data <- clean_poll_data |>
   mutate(state = str_replace(state, "Nebraska CD-2", "Nebraska"))
 
-# Filter data to Harris estimates based on high-quality polls after she declared
+# Aggregate pct values by averaging for unique combinations before pivoting
 clean_poll_data <- clean_poll_data |>
-  filter(numeric_grade >= 2.7, transparency_score > 5, candidate_name %in% c("Donald Trump", "Kamala Harris"))
-  
-head(clean_poll_data)
+  group_by(state, end_date, end_date_num, candidate_name) |>
+  summarize(pct = mean(pct, na.rm = TRUE), .groups = "drop") |>
+  pivot_wider(names_from = candidate_name, values_from = pct, values_fill = 0) |>
+  rename(Trump_pct = `Donald Trump`, Harris_pct = `Kamala Harris`) |>
+  filter(Trump_pct > 0 & Harris_pct > 0)  # Retain rows with both candidates’ support data
 
-#### Save data ####
[…]

### 2024-11-03 scripts/06-model_data.R

@@ -6,37 +6,52 @@
 # License: MIT
 # Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 # Any other information needed? N/A
-#### Workspace setup ####
+
+
+#### Modeling Script ####
+# Load necessary libraries
 library(tidyverse)
 library(rstanarm)
-
-#### Read data ####
-# Load the cleaned polling data
-clean_poll_data <- read_csv("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.csv")
+library(splines)
 
 
-### Prepare Model Data ####
-# Aggregating polling percentages for Trump and Harris by state
-state_average <- clean_poll_data %>%
-  group_by(state) %>%
-  summarise(
-    Trump_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE),
-    Harris_pct = mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE)
-  )
+# Load cleaned data
+clean_poll_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
-# Create a binary outcome for prediction: 1 if Trump is predicted to win, 0 if Harris
-state_average <- state_average %>%
-  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
+# Fit Trump model with time spline and state random effects
+trump_model <- brm(
+  Trump_pct ~ ns(end_date_num, df = 4) + (1 | state),  # Natural spline with 4 degrees of freedom
+  data = clean_poll_data,
+  family = gaussian(),
+  prior = c(
+    prior(normal(0, 5), class = "b"),
+    prior(normal(0, 10), class = "Intercept")
+  ),
+  seed = 1234,
+  iter = 3000,
+  chains = 4,
+  control = list(adapt_delta = 0.95)
+)
 
-#### Bayesian Logistic Regression Model with Regularization ####
-# Using a regularized logistic regression model from rstanarm to avoid convergence issues
-logistic_model <- stan_glm(winner ~ Trump_pct + Harris_pct, data = state_average, family = binomial, 
-                           prior = student_t(df = 7), chains = 4, iter = 2000)
+# Fit Harris model with time spline and state random effects
+harris_model <- brm(
+  Harris_pct ~ ns(end_date_num, df = 4) + (1 | state),
+  data = clean_poll_data,
+  family = gaussian(),
+  prior = c(
[…]

### 2024-11-03 paper/paper.qmd
committing changes to git

@@
 ---
-title: "Predicting a Trump Victory: A State-Level Analysis of the 2024 US Presidential Election Using Bayesian Modeling"
-subtitle: "State-by-State Insights on Voter Dynamics Favoring Trump Over Harris in Key Battlgrounds"
+title: "Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis"
+subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
 author: 
@@
 
-## Model Summary Results
+## Model Summary - Harris Model
 ```{r}
@@
 #| eval: true
-#| label: tbl-fixed-effects-summary
-#| tbl-cap: "Fixed Effects Summary for Trump and Harris Models."
+#| label: tbl-harris-fixed-effects
+#| tbl-cap: "Fixed Effects Summary for Harris Model."
 #| warning: false
 
-# Load necessary libraries
-library(dplyr)
-library(broom)        # For the tidy function
-library(knitr)
-library(kableExtra)
+# Summarize fixed effects for Harris model
+harris_summary <- tidy(harris_model, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `Lower 95% CI` = conf.low,
+    `Upper 95% CI` = conf.high
+  ) %>%
+  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Clean up term names
 
+# Display the Harris model summary table
+knitr::kable(harris_summary) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+EDIT - explain model summary table
+
+## Model summary - Trump Model
+```{r}
 #| echo: false
 #| eval: true
-#| label: tbl-fixed-effects-summary
-#| tbl-cap: "Fixed Effects Summary for Trump and Harris Models."
+#| label: tbl-trump-fixed-effects
+#| tbl-cap: "Fixed Effects Summary for Trump Model."
 #| warning: false
-
-# Load necessary libraries
-library(dplyr)
-library(broom)        # For the tidy function
-library(knitr)
-library(kableExtra)
 
@@
   ) %>%
-  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Remove non-alphanumeric characters
+  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Clean up term names
 
-# Summarize fixed effects for Harris model
-harris_summary <- tidy(harris_model, effects = "fixed", conf.int = TRUE) %>%
-  select(term, estimate, std.error, conf.low, conf.high) %>%
-  rename(
-    Estimate = estimate,
-    `Std. Error` = std.error,
-    `Lower 95% CI` = conf.low,
-    `Upper 95% CI` = conf.high
-  ) %>%
-  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Remove non-alphanumeric characters
+# Display the Trump model summary table
+knitr::kable(trump_summary) %>%
+  kable_styling(full_width = FALSE, position = "center")
+```
 
-# Combine summaries into one table
-model_summary_table <- full_join(trump_summary, harris_summary, by = "term", suffix = c(" Trump", " Harris"))
-
-# Display the table with kable
-kable(model_summary_table, caption = "Fixed Effects Summary for Trump and Harris Models") %>%
-  kable_styling(full_width = FALSE, position = "center", font_size = 8)
-
-```
+EDIT - explain model summary table
 
@@
 ```
+EDIT - explain results on graph
 
@@
 ```
-## Election Day Predictions
+EDIT - explain results on graph
 
 ## Probability of Winning by State on Election Day
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-state-probabilities
+#| fig-cap: "Probability of Trump Winning by State on Election Day."
+#| warning: false
+
+# Set election day or target prediction date
+election_day <- as.Date("2024-11-05")  # Example election day
+
+# Predict probability of Trump and Harris winning by comparing posterior predictions for each state
+state_predictions <- clean_poll_data %>%
+  distinct(state) %>%
+  mutate(end_date = election_day, end_date_num = as.numeric(election_day - min(clean_poll_data$end_date)))
+
+# Calculate probabilities of Trump and Harris winning in each state
+state_predictions <- state_predictions %>%
+  rowwise() %>%
+  mutate(
+    prob_trump_wins = mean(posterior_predict(trump_model, newdata = cur_data()) > 
+                           posterior_predict(harris_model, newdata = cur_data())),
+    prob_harris_wins = mean(posterior_predict(harris_model, newdata = cur_data()) > 
+                            posterior_predict(trump_model, newdata = cur_data()))
+  )
+
+# Prepare the state predictions for the table
+state_summary_table <- state_predictions %>%
+  select(state, prob_trump_wins, prob_harris_wins) %>%
+  mutate(
+    prob_trump_wins = round(prob_trump_wins * 100, 2),  # Convert to percentage and round
+    prob_harris_wins = round(prob_harris_wins * 100, 2)  # Convert to percentage and round
+  )
+
+# Display state predictions i n table
+knitr::kable(state_summary_table) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+EDIT - explain table results
 
@@
 
-# Set election day or target prediction date
-election_day <- as.Date("2024-11-05")  # Example election day
+# Reshape the data for plotting
+state_predictions_long <- state_predictions %>%
+  select(state, prob_trump_wins, prob_harris_wins) %>%
+  pivot_longer(cols = starts_with("prob"), 
+               names_to = "Candidate", 
+               values_to = "Probability") %>%
+  mutate(Candidate = recode(Candidate, 
+                            prob_trump_wins = "Trump", 
+                            prob_harris_wins = "Harris"))
 
-# Predict probability of Trump winning by comparing posterior predictions for each state
-state_predictions <- clean_poll_data %>%
-  distinct(state) %>%
-  mutate(end_date = election_day, end_date_num = as.numeric(election_day - min(clean_poll_data$end_date)))
-
-# Calculate probability of Trump winning in each state
-state_predictions <- state_predictions %>%
-  rowwise() %>%
-  mutate(
-    prob_trump_wins = mean(posterior_predict(trump_model, newdata = cur_data()) >
-                           posterior_predict(harris_model, newdata = cur_data()))
-  )
-
-# Plot probability of winning by state
-ggplot(state_predictions, aes(x = reorder(state, prob_trump_wins), y = prob_trump_wins, fill = prob_trump_wins > 0.5)) +
-  geom_col() +
-  coord_flip() +
-  scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue"), labels = c("Harris", "Trump")) +
-  labs(title = "Probability of Trump Winning by State on Election Day", x = "State", y = "Probability Trump Wins") +
+# Plot double bar graph of predicted winning percentages by state
+ggplot(state_predictions_long, aes(x = reorder(state, Probability), y = Probability * 100, fill = Candidate)) +
+  geom_col(position = "dodge") +  # Use dodge for side-by-side bars
+  coord_flip() +  # Flip coordinates for better readability
+  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
+  labs(title = "Predicted Winning Percentages by State on Election Day", 
+       x = "State", 
+       y = "Winning Probability (%)") +
   theme_minimal() +
-  theme(legend.position = "none")
-
+  theme(legend.position = "right")
 ```
+EDIT - explain graph results
 
@@
 ```
+
+EDIT - explain map results
 

### 2024-11-03 paper/paper.qmd
committing edited version of paper

@@
 date-format: long
-abstract: "This paper presents a predictive model for the 2024 United States Presidential Election using state-level polling data.
-By aggregating high-quality polls, we predict the likely winner between Donald Trump and Kamala Harris in each state.
-A logistic regression model estimates the probability of Trump winning, with polling percentages as predictor variables.
-The model’s results highlight key battleground states and offer insights into voter dynamics, while also reflecting on the limitations of polling data in forecasting political outcomes.
-Our findings suggest that Trump holds an adantage in critical swing states, positinoning him as the likely winner of the election."
+abstract: "This paper presents a predictive model for the 2024 United States Presidential Election using state-level polling data, with a Bayesian approach to estimate the outcomes between Donald Trump and Kamala Harris.
+By focusing on high-quality polls from reputable sources, we aggregated data to predict the likely winner in each state.
+Our Bayesian hierarchical model, incorporating time-dependent splines and state-specific random effects, captures shifts in candidate support over time.
+The results underscore Trump’s advantage in critical swing states such as Florida, Ohio, and North Carolina, positioning him as the predicted overall winner of the 2024 election.
+We discuss the model’s performance, accuracy, and limitations, reflecting on the challenges inherent in electoral forecasting."
 format: pdf
@@
 
-# Introduction {#sec-introduction}
-
+# Introduction
 This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data [@FiveThirtyEight] to forecast the likely winner between Donald Trump and Kamala Harris.
-We aggregate high-quality poll data to create a logistic regression model that estimates the probability of a Trump or Harris victory in each state.
+We aggregate high-quality poll data to create a Bayesian logistic regression model that estimates the probability of a Trump or Harris victory in each state.
 By comparing polling percentages for both candidates, we aim to predict election outcomes based on voter support patterns across the states.
@@
 
-Our model utilizes the Bayesian logistic regression model to predict election outcomes by comparing the average polling percentages for Trump and Harris in each state.
-The results highlight the geographic distribution of support, with some states clearly favoring one candidate over the other.
+Our model utilizes a Bayesian approach with a hierarchical structure to predict election outcomes by comparing the average polling percentages for Trump and Harris across states.
+This methodology incorporates time-dependent splines and state-specific random effects, which allow for capturing variations in support trends over time.
 Swing states emerge as critical battlegrounds, where polling percentages are closely contested and could influence the final election result.
@@
 Accurate election predictions provide valuable insights into voter dynamics and help political analysts, campaigns, and the public anticipate electoral outcomes.
-By focusing on high-quality polling data, our model improves the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.
+By focusing on high-quality polling data, our model enhances the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.
 
@@
 Key battleground states, such as Florida, Ohio, and North Carolina, lean toward Trump, emphasizing his advantage in critical regions.
-Our analysis points to Trump as the probable winner, with swing states playing a decisive role in determining the final outcome.
+Our analysis points to Trump as the probable overall winner, with swing states playing a decisive role in determining the final outcome.
 
@@
 
-We use the statistical programming language R [@citeR] to perform data cleaning and analysis using the datasets we obtained from the Project 538 online database [@FiveThirtyEight].
-The database provides a wealth of information on the current cycle of presidential general election polls.
-It consists of the polling results from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
-Information about the transparency of methodology as well as the quality of the data is present as well, denoted as transparency score and numeric grades, which are significant to our data cleaning process to ensure the quality of data.
-Polling results are specific to different states in the US, mostly focusing on the six battleground states that are the likeliest to determine the outcome of the elections.
-Thus, we chose to use states, candidate names, and percentage of support to perform statistical analysis and gather our results.
+REWRITE
 
@@
 
-The measurement of voter support in the dataset is derived from raw polling data obtained from the Project 538 online database [@FiveThirtyEight], which serves as a representation of potential election results for the 2024 U.S.
-Presidential Election.
-Each entry in the data corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for Donald Trump or Kamala Harris.
-Each pollster has their approach of sampling respondents as well as their methodology of approaching respondents, which affects each pollster’s numeric grades and transparency scores.
-
-Therefore, to ensure the quality of the data, only polls from reputable pollsters with a numeric grade of 3 or higher and transparency scores above 6 were used for our paper.
-These polls measure voter preferences through structured survey questions, and the results are aggregated to create state-level averages for each candidate.
-The dataset thus transforms general voter sentiment into individual data points that reflect the competitive dynamics between Trump and Harris across different states.
+REWRITE
 
@@
 
-The outcome variable in our analysis is the binary variable winner, which represents the predicted winner of the 2024 US Presidential Election in each state.
-The value of winner is set to 1 if Donald Trump is predicted to win the state, and 0 if Kamala Harris is predicted to win.
-This binary outcome is determined by comparing the polling support for both candidates within each state.
-By setting up this binary variable, we aim to forecast which candidate will secure more votes in each state based on the aggregated polling data.
-
-[@fig-averagetable] provides an overview of the average support for Trump and Harris across all states, along with the predicted winner for each state based on our model.
+REWRITE
 
@@
 
-Based on the table provided, Trump is predicted to win in the states of Arizona, Florida, Georgia, Missouri, Montana, Nevada, Ohio, and Texas, while Harris is predicted to win in Michigan, Nebraska, North Carolina, Pennsylvania, Virginia, and Wisconsin.
-However, this does not necessarily determine the overall winner of the election, as U.S.
-presidential elections are decided by the Electoral College rather than by a simple majority in each state.
-Each state carries a different number of electoral votes, and a candidate must secure at least 270 electoral votes to win the presidency.
-To predict the overall winner, it would be necessary to assign each state’s electoral votes to the respective predicted winner and calculate the total for each candidate.
-With the electoral vote counts for each state, we could determine which candidate is projected to win the election.
+REWRITE
 
@@
 
-The predictor variables used in the model are the aggregated polling percentages for Donald Trump (Trump_pct) and Kamala Harris (Harris_pct).
-These variables are calculated as the average support for each candidate, using polling data filtered to include only high-quality pollsters with a numeric grade of 3 or higher and transparency scores above 6.
-These averages reflect the level of support for each candidate across all polls in each state.
-
-To further illustrate the distribution of support, a side-by-side bar graph is provided, comparing the average polling percentages for both candidates across all states.
-[@fig-barplot] offers a clear comparison of the support levels, helping to visualize the competitive dynamics within each state.
-The side-by-side comparison represents the percentage of support each candidate has received, based on aggregated poll data from pollsters with high-quality scores.
-Trump's support is shown in red, while Harris's support is depicted in red.
+REWRITE
 
@@
 
-@fig-bar-plot illustrates the state-by-state polling data by comparing the average support percentages for Trump and Harris.
-In each state, the red bar represents Trump's average polling percentage, while the blue bar represents Harris's.
-Trump shows stronger support in states like **Montana**, **Missouri**, **Florida**, and **Nevada**, where his average polling percentage surpasses that of Harris.
-Conversely, Harris has a slight edge in states such as **Wisconsin**, **North Carolina**, **Nebraska**, **Pennsylvania**, and **Virginia**.
-Some states, like Ohio and **Georgia**, exhibit closely matched polling averages, indicating a tighter race.
-This visualization underscores the regional variation in candidate support across states, reflecting diverse voter preferences that could influence the final outcome based on the Electoral College distribution.
+REWRITE
 
@@
 ## Model set-up
-The binary outcome variable \( y_i \) represents whether Donald Trump is predicted to lead in state \( i \) based on polling averages for both candidates over time.
-Here, \( y_i = 1 \) indicates a Trump lead in state \( i \), and \( y_i = 0 \) indicates a lead for Kamala Harris.
+The binary outcome variable $y_i$ represents whether Donald Trump is predicted to lead in state $i$ based on polling averages for both candidates over time.
+Here, $y_i = 1$ indicates a Trump lead in state $i$, and $y_i = 0$ indicates a lead for Kamala Harris.
 The model estimates the support levels for each candidate based on polling data using a Gaussian likelihood function, with time trends captured by a natural spline and state-level variation captured through random intercepts.
 
-The predictor variables are \( \text{Trump\_pct}_i \), the predicted percentage of support for Donald Trump in state \( i \), and \( \text{Harris\_pct}_i \), the predicted percentage of support for Kamala Harris in state \( i \).
+The predictor variables are $Trump\_pct_i$ and $Harris\_pct_i$, which represent the predicted percentages of support for Donald Trump and Kamala Harris in state $i$.
 A separate model is fit for each candidate’s support over time, capturing the time-varying nature of support with a natural spline function and allowing for state-specific random effects.
@@
 
-where:
-- \( \text{Trump\_pct}_i \) and \( \text{Harris\_pct}_i \) are the polling support levels for Donald Trump and Kamala Harris in state \( i \).
-- \( f(\text{end\_date\_num}_i) \) is a natural spline function of time (measured as `end_date_num`), with 4 degrees of freedom, capturing the non-linear time trends in support.
-- \( \alpha_{\text{Trump}} \) and \( \alpha_{\text{Harris}} \) are intercepts for the models, representing the baseline support levels for each candidate.
-- \( u_{\text{state}}^{\text{Trump}} \) and \( u_{\text{state}}^{\text{Harris}} \) are state-level random effects, accounting for differences in baseline support across states.
+Where:
+
+- $ \text{Trump\_pct}_i $ and $ \text{Harris\_pct}_i $ are the polling support levels for Donald Trump and Kamala Harris in state $ i $.
+- $ f(\text{end\_date\_num}_i) $ is a natural spline function of time (measured as `end_date_num`), with 4 degrees of freedom, capturing the non-linear time trends in support.
+- $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are intercepts for the models, representing the baseline support levels for each candidate.
+- $ u_{\text{state}}^{\text{Trump}} $ and $ u_{\text{state}}^{\text{Harris}} $ are state-level random effects, accounting for differences in baseline support across states.
 
 For both models, we use weakly informative Normal priors.
-The intercepts \( \alpha_{\text{Trump}} \) and \( \alpha_{\text{Harris}} \) are given Normal priors with a mean of 0 and a standard deviation of 10, reflecting prior uncertainty about the baseline support levels.
+The intercepts $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are given Normal priors with a mean of 0 and a standard deviation of 10, reflecting prior uncertainty about the baseline support levels.
 The time trend coefficients are also given Normal priors with mean 0 and standard deviation 5, allowing for flexibility in capturing the temporal changes in support.

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

@@
 
-- $ \text{Trump\_pct}_i $ and $ \text{Harris\_pct}_i $ are the polling support levels for Donald Trump and Kamala Harris in state $ i $.
-- $ f(\text{end\_date\_num}_i) $ is a natural spline function of time (measured as `end_date_num`), with 4 degrees of freedom, capturing the non-linear time trends in support.
+- $ \text{Trump\_pct}\_i $ and $ \text{Harris\_pct}\_i $ are the polling support levels for Donald Trump and Kamala Harris in state $ i $.
+- $ f(\text{end\_date\_num}\_i) $ is a natural spline function of time (measured as $ \text{end\_date\_num} $), with 4 degrees of freedom, capturing the non-linear time trends in support.
 - $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are intercepts for the models, representing the baseline support levels for each candidate.

### 2024-11-03 paper/paper.qmd
committing text changes

--- 2024-11-03 paper/paper.qmd
@@
 - $ \text{Trump\_pct}_i $ and $ \text{Harris\_pct}_i $ are the polling support levels for Donald Trump and Kamala Harris in state $ i $.
-- $ f(\text{end\_date\_num}_i) $ is a natural spline function of time (measured as `end_date_num`), with 4 degrees of freedom, capturing the non-linear time trends in support.
+- $ f(\text{end\_date\_num}_i) $ is a natural spline function of time (measured as $ \text{end\_date\_num} $), with 4 degrees of freedom, capturing the non-linear time trends in support.
 - $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are intercepts for the models, representing the baseline support levels for each candidate.
@@
 # Display state predictions i n table
-knitr::kable(state_summary_table) %>%
-  kable_styling(full_width = FALSE, position = "center")
+knitr::kable(state_summary_table)
 

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

@@
 
+## Other Variables
+
+REWRITE
+
 # Model {#sec-model}
@@
 
-- $ \text{Trump\_pct}_i $ and $ \text{Harris\_pct}_i $ are the polling support levels for Donald Trump and Kamala Harris in state $ i $.
-- $ f(\text{end\_date\_num}_i) $ is a natural spline function of time (measured as $ \text{end\_date\_num} $), with 4 degrees of freedom, capturing the non-linear time trends in support.
-- $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are intercepts for the models, representing the baseline support levels for each candidate.
-- $ u_{\text{state}}^{\text{Trump}} $ and $ u_{\text{state}}^{\text{Harris}} $ are state-level random effects, accounting for differences in baseline support across states.
+- \begin{equation}
+  \text{Trump\_pct}_i \text{ and } \text{Harris\_pct}_i \text{ are the polling support levels for Donald Trump and Kamala Harris in state } i.
+  \end{equation}
+
+- \begin{equation}
+  f(\text{end\_date\_num}_i) \text{ is a natural spline function of time (measured as } \text{end\_date\_num}, \text{ with 4 degrees of freedom, capturing the non-linear time trends in support.}
+  \end{equation}
+
+- \begin{equation}
+  \alpha_{\text{Trump}} \text{ and } \alpha_{\text{Harris}} \text{ are intercepts for the models, representing the baseline support levels for each candidate.}
+  \end{equation}
+
+- \begin{equation}
+  u_{\text{state}}^{\text{Trump}} \text{ and } u_{\text{state}}^{\text{Harris}} \text{ are state-level random effects, accounting for differences in baseline support across states.}
+  \end{equation}
 

### 2024-11-03 paper/paper.qmd
committing changes to data section in paper.qmd and  03-clean_data.R

@@
 title: "Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis"
-subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
+subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key Battleground States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
 author: 
@@
 date-format: long
-abstract: "This paper presents a predictive model for the 2024 United States Presidential Election using state-level polling data, with a Bayesian approach to estimate the outcomes between Donald Trump and Kamala Harris.
-By focusing on high-quality polls from reputable sources, we aggregated data to predict the likely winner in each state.
-Our Bayesian hierarchical model, incorporating time-dependent splines and state-specific random effects, captures shifts in candidate support over time.
-The results underscore Trump’s advantage in critical swing states such as Florida, Ohio, and North Carolina, positioning him as the predicted overall winner of the 2024 election.
-We discuss the model’s performance, accuracy, and limitations, reflecting on the challenges inherent in electoral forecasting."
+abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, utilizing state-level polling data to focus specifically on key battleground states.
+By employing a Bayesian approach, we estimate the outcomes between Donald Trump and Kamala Harris, incorporating high-quality polls from reputable sources.
+Our Bayesian hierarchical model, featuring time-dependent splines and state-specific random effects, effectively captures shifts in candidate support over time.
+The analysis reveals that Trump is projected to win in critical battleground states such as Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, highlighting his advantage in these pivotal regions.
+These findings underscore the significance of battleground states in determining electoral outcomes, particularly in a highly polarized political landscape.
+We also discuss the model’s performance, accuracy, and limitations, reflecting on the unique challenges inherent in forecasting electoral outcomes in these crucial areas."
 format: pdf
@@
 # Introduction
-This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data [@FiveThirtyEight] to forecast the likely winner between Donald Trump and Kamala Harris.
-We aggregate high-quality poll data to create a Bayesian logistic regression model that estimates the probability of a Trump or Harris victory in each state.
-By comparing polling percentages for both candidates, we aim to predict election outcomes based on voter support patterns across the states.
 
-The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, based on aggregated state-level polling averages.
-The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, and the predictor variables are the average polling percentages for each candidate.
+EDIT & ELABORATE - gpt generated
+The 2024 United States Presidential Election presents a unique opportunity to examine the dynamics of voter preferences in key battleground states.
+With a highly polarized electorate and the stakes at an all-time high, accurately predicting the outcome of this election is crucial for political analysts, campaigns, and the public.
+This paper develops a predictive model that leverages state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, using a Bayesian hierarchical approach to accommodate variations across states.
 
-Our model utilizes a Bayesian approach with a hierarchical structure to predict election outcomes by comparing the average polling percentages for Trump and Harris across states.
-This methodology incorporates time-dependent splines and state-specific random effects, which allow for capturing variations in support trends over time.
-Swing states emerge as critical battlegrounds, where polling percentages are closely contested and could influence the final election result.
+The central estimand of this study is the probability of victory for each candidate in battleground states, defined by the binary outcome variable indicating a predicted win for Trump (1) or Harris (0).
+To achieve this, we utilize aggregated polling percentages as predictor variables, allowing us to capture shifts in voter sentiment over time.
+By employing logistic regression and accounting for state-specific random effects, our model not only identifies trends in polling data but also highlights critical regions that could determine the overall election outcome.
 
-Accurate election predictions provide valuable insights into voter dynamics and help political analysts, campaigns, and the public anticipate electoral outcomes.
-By focusing on high-quality polling data, our model enhances the reliability of predictions and highlights key regions where voter sentiment may shift, ultimately affecting the election.
+In our analysis, we focused on key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—where the margins are often razor-thin and can swing the election in favor of one candidate or another.
+Our findings indicate that Trump is predicted to win in these crucial states, positioning him as the likely overall victor in the 2024 election.
+This is significant as it underscores the importance of battleground states in shaping the electoral landscape and highlights the intricate dynamics of voter preferences that can influence the final result.
 
-The results of our model predict a victory for Donald Trump in the 2024 U.S.
+The structure of this paper is as follows: we begin by detailing the data sources and methodology used for the analysis.
+Next, we present the results, focusing on the predicted probabilities of each candidate winning in the targeted battleground states.
+Finally, we discuss the implications of our findings and the limitations of our model, contributing to a better understanding of voter dynamics in the 2024 U.S.
 Presidential Election.
-Key battleground states, such as Florida, Ohio, and North Carolina, lean toward Trump, emphasizing his advantage in critical regions.
-Our analysis points to Trump as the probable overall winner, with swing states playing a decisive role in determining the final outcome.
-
-The remainder of this paper is structured as follows.
-In [@sec-data], we describe the data and variables used in the analysis.
-[@sec-model] outlines the model setup and estimation strategy.
-[@sec-results] presents the results, including the predicted election outcomes and visualizations.
-Finally, [@sec-discussion] discusses the implications and limitations of our findings.
 
 # Data {#sec-data}
+## Overview
+This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight].
+The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
+It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
+For every pollster, relevant variables such as state, polling date, transparency score, candidate name, and percentage of support for the appropriate candidate are included.
+The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
 
-## Overview
-
-REWRITE
+We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.
 
 ## Measurement
+In the current presidential election cycle,  Americans’ opinions on voting for their preferred candidate are transformed into data points through a series of steps that involve surveying, data processing, and structuring responses.
+Pollsters then store these responses in structured databases for analysis.
+In the dataset obtained from Project 538, each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate.
+Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.
 
-REWRITE
+Pollsters begin by selecting a sample from the registered voter population, often using voter files with demographic and contact information.
+They then apply methods of stratified random sampling to ensure the sample reflects the electorate’s diversity in terms of age, gender, race, education, voter behavior, etc.
 
-## Outcome variables
+Then, pollsters reach out to voters through various channels, such as phone calls, emails, text messages, or online panels.
+These are called recruiting methods and usually, each of them has its trade-offs.
+If the respondents are connected through any of these channels, they are asked a series of questions either by a live interviewer or a preset of questions.
 
-REWRITE
+The types of questions often include:
 
+- “If the election were held today, would you vote for [Candidate A] or [Candidate B]?” 
+- “How strongly do you support [Candidate]?” (e.g., on a scale of 1 to 10, or strongly, somewhat, undecided)
+- “Do you support [Policy A] proposed by [Candidate A] or [Policy B] proposed by [Candidate B]?”
+
+Responses are recorded digitally, either by survey software or directly into a database by interviewers, capturing both the answers and relevant demographic data.
+Each response becomes a data point in the database.
+For example, if a respondent prefers Candidate A, their answer may be recorded as a binary variable (e.g., “1” for Candidate A, “0” for Candidate B).
+Along with candidate preference, pollsters capture demographic and behavioral details, such as age, gender, race, education, income, voter history, and party registration.
+
+Each respondent’s answers and demographic details are grouped and saved in a structured database.
+Pollsters then clean the data by checking for inconsistencies, removing incomplete responses, and addressing non-responses.
+Pollsters also perform weighting to ensure the representativeness of data collected, details of weighting are further explained in {appendix 1}.
+
+Once cleaned and weighted, data points are aggregated to determine overall candidate support, often producing metrics such as the percentage of respondents supporting each candidate by state or other specific demographics.
+Each pollster would have a database with all the weighted and aggregated data, and Project 538 presents a compilation of the databases obtained from various pollsters.
+
+## Variables
+To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
 ```{r}
-#| label: fig-averagetable
-#| fig-cap: Summary of 2024 U.S.
-Presidential Election Polling Averages by State for Trump and Harris, with Indicated Leading Candidate.
-States with missing data, such as Colorado, are labeled as "NA."
+#| label: fig-summarystatistics
+#| fig-cap: "Vertical Summary Statistics for Trump and Harris Polling Data.
+This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
+All values are rounded to three decimal places, providing a concise overview of the polling landscape as of the selected election date."
 #| echo: false
@@
 #| message: false
+
 # Load cleaned data
 clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
+
+# Summary statistics for the relevant columns in clean_poll_data
+summary_statistics <- clean_poll_data %>%
+  summarize(
+    `Total Polls` = n(),
+    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
+    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
+    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),
+
+    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
+    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
+    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
+  )
+
+# Reshape the summary statistics to a vertical format
+summary_statistics_long <- pivot_longer(
+  summary_statistics,
+  cols = everything(),
+  names_to = "Statistic",
+  values_to = "Value"
+)
+
+# Display vertical summary statistics in a kable table
+kable(summary_statistics_long) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+@tbl-summarystatistics reveals a total of 34 polls conducted, with an average support level of 46.321% for Trump and 46.661% for Harris.
+Trump's support peaked at 51.000%, while the lowest recorded support was 42.500%.
+For Harris, the maximum support also reached 51.000%, with a minimum of 42.000%.
+These statistics illustrate the competitive landscape in these critical states, emphasizing the narrow margins that could ultimately influence the election outcome.
+The data reflects the volatility of voter sentiment leading up to the 2024 Presidential Election, underscoring the importance of these battleground states in determining the final result.
+
+### Outcome Variable
+The outcome variable in this study is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S.
+Presidential Election based on polling data from key battleground states.
+This variable categorizes the election outcome into two distinct categories: a value of "Trump" indicates that Donald Trump is predicted to win, while a value of "Harris" denotes a predicted victory for Kamala Harris.
+The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris.
+This binary outcome variable serves a crucial role in facilitating the analysis of voter support dynamics and the effectiveness of the predictive modeling employed in this paper.
+
+By focusing on this dichotomy, we can glean insights into how polling data might translate into electoral outcomes, highlighting the competitive nature of the election in these pivotal states.
+The construction of the outcome variable not only enhances our understanding of voter preferences but also underscores the significance of binary classification in electoral forecasting, especially in states that could ultimately sway the overall election results.
+
+### Predictive Variable
+The predictive variable in this study is the polling support percentage for each candidate, specifically for Donald Trump and Kamala Harris, measured across various battleground states.
+This variable represents the predicted likelihood of each candidate winning in the specified states based on the polling data collected during the election cycle.
+```{r}
+#| label: tbl-outcomevar
+#| fig-cap: "Average polling support percentages for Donald Trump and Kamala Harris across key battleground states.
+The table presents the average support levels, indicating Trump’s and Harris’s polling figures in states such as Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada.
+The comparative analysis highlights the varying levels of candidate support in these critical regions, providing insight into the electoral landscape ahead of the 2024 U.S.
+Presidential Election."
+#| echo: false
+#| warning: false
+#| message: false
 
@@
   )
-
-# Create winner column: 1 if Trump_pct is greater than Harris_pct, 0 otherwise
-state_average <- state_average |>
-  mutate(winner = ifelse(Trump_pct > Harris_pct, 1, 0))
 
@@
     "Harris %" = Harris_pct, 
-    Winner = winner
-  ) |>
-  mutate(Winner = ifelse(Winner == 1, "Trump", "Harris"))
+  )
 
@@
 ```
+@tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada.
+The values reflect the polling percentages for each candidate, providing a straightforward comparison of their standings within these critical regions.
 
-REWRITE
-
-## Predictor variables
-
-REWRITE
-
-```{r}
-#| label: fig-barplot
-#| fig-cap: "Comparison of 2024 U.S.
-Presidential Election Polling Averages for Trump and Harris by State.
-Bars represent the average polling percentages for each candidate, with states missing data, such as Colorado, labeled accordingly."
-#| echo: false
-#| warning: false
-
-# Reshape data to a long format for side-by-side bars
-stateaverage_long <- state_average %>%
-  pivot_longer(cols = c(Trump_pct, Harris_pct),
-               names_to = "candidate",
-               values_to = "support_pct")
-
-ggplot(stateaverage_long, aes(x = reorder(state, -support_pct), y = support_pct, fill = candidate)) + 
-  geom_bar(stat = "identity", position = position_dodge(width = 0.8), color = "black") +
-  labs(x = "State", y = "Average Percentage") +
-  scale_fill_manual(
-    name = "Candidate", 
-    values = c("Trump_pct" = "red", "Harris_pct" = "blue"),
-    labels = c("Trump_pct" = "Trump %", "Harris_pct" = "Harris %")
-  ) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1))
-```
-
-REWRITE
+Understanding these polling percentages is essential as they offer insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election.
+The battleground states depicted in the table are particularly significant due to their historical tendency to swing between parties, making them crucial targets for campaign efforts.
+Analyzing the outcome variable through these polling percentages informs our predictive modeling and enhances our comprehension of the dynamics at play during the electoral process.
+This comparison not only contextualizes the current political climate but also underlines the importance of these battleground states in shaping the overall election results.
+As the campaign progresses, these insights will be invaluable for strategizing and understanding voter sentiment in these pivotal areas.
 
@@
 
-REWRITE
+WRITE SECTION
 
@@
   \end{equation}
-
 - \begin{equation}
@@
 
+# Define battleground states
+battleground_states <- c("Arizona", "Georgia", "Michigan", "Nevada", 
+                          "North Carolina", "Pennsylvania", "Wisconsin")
+
 # Generate predictions for each state and date
 state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
-state_prediction_grid <- expand.grid(state = unique(clean_poll_data$state), end_date = state_dates) %>%
+state_prediction_grid <- expand.grid(state = battleground_states, end_date = state_dates) %>%
   mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))
@@
   geom_line(aes(y = Harris_Support, color = "Harris")) +
-  labs(title = "State-Level Support Trends for Trump and Harris Over Time", y = "Predicted Support", x = "Date") +
+  labs(title = "State-Level Support Trends for Trump and Harris Over Time (Battleground States)", 
+       y = "Predicted Support", 
+       x = "Date") +
   facet_wrap(~ state) +
@@
   distinct(state) %>%
-  mutate(end_date = election_day, end_date_num = as.numeric(election_day - min(clean_poll_data$end_date)))
+  filter(state %in% battleground_states) %>%  # Only include battleground states
+  mutate(end_date = election_day, 
+         end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))
 
@@
 
-# Display state predictions i n table
-knitr::kable(state_summary_table)
+# Display state predictions in table
+knitr::kable(state_summary_table) %>%
+  kable_styling(full_width = FALSE, position = "center")
 
@@
 
-# Reshape the data for plotting
+# Reshape the data for plotting only for battleground states
 state_predictions_long <- state_predictions %>%
   select(state, prob_trump_wins, prob_harris_wins) %>%
+  filter(state %in% battleground_states) %>%  # Ensure we're only working with battleground states
   pivot_longer(cols = starts_with("prob"), 
@@
   scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  labs(title = "Predicted Winning Percentages by State on Election Day", 
+  labs(title = "Predicted Winning Percentages by Battleground State on Election Day", 
        x = "State", 
@@
   theme(legend.position = "right")
+
 ```
@@
 #| label: fig-us-map
-#| fig-cap: "Predicted Winner by State on Election Day."
+#| fig-cap: "Predicted Winner by Battleground State on Election Day."
 #| warning: false
@@
 state_predictions <- state_predictions %>%
+  filter(state %in% battleground_states) %>%  # Ensure we only include battleground states
   mutate(predicted_winner = ifelse(prob_trump_wins > 0.5, "Trump", "Harris"))
@@
   scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  labs(title = "Predicted Winner by State on Election Day", fill = "Predicted Winner") +
+  labs(title = "Predicted Winner by Battleground State on Election Day", fill = "Predicted Winner") +
   theme(legend.position = "right")

### 2024-11-03 scripts/03-clean_data.R

@@ -8,17 +8,20 @@
 # Any other information needed? N/A
 
 
-#### Data Cleaning Script ####
-# Load necessary libraries
+# Load the necessary libraries
 library(tidyverse)
 library(janitor)
 library(lubridate)
+
+# Define battleground states
+battleground_states <- c("Arizona", "Georgia", "Michigan", "Nevada", 
+                         "North Carolina", "Pennsylvania", "Wisconsin")
 
 # Load the dataset
 poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
   clean_names()  
 
-# Filter and select relevant columns
+# Filter and select relevant columns, including only battleground states
 clean_poll_data <- poll_data |>
   select(state, end_date, pollster, numeric_grade, transparency_score, candidate_name, pct) |>
   filter(
@@ -27,14 +30,15 @@ clean_poll_data <- poll_data |>
     !is.na(transparency_score),
     numeric_grade >= 3.0,          # numeric_grade threshold
     transparency_score > 5,        # transparency_score threshold
-    candidate_name %in% c("Donald Trump", "Kamala Harris")
+    candidate_name %in% c("Donald Trump", "Kamala Harris"),
+    state %in% battleground_states   # Filter to include only battleground states
   ) |>
   mutate(
     end_date = as.Date(end_date, format = "%m/%d/%y"),
     end_date_num = as.numeric(end_date - min(end_date, na.rm = TRUE))
   )
 
-# Replace "Nebraska CD-2" with "Nebraska"
+# Replace "Nebraska CD-2" with "Nebraska" if it appears in the data
 clean_poll_data <- clean_poll_data |>
   mutate(state = str_replace(state, "Nebraska CD-2", "Nebraska"))
 
@@ -48,6 +52,4 @@ clean_poll_data <- clean_poll_data |>
 
 #### Save Data ####
 write_parquet(clean_poll_data, "data/02-analysis_data/analysis_data.parquet")
-
-
-
+print(clean_poll_data)

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd in results section

@@
 
-## National Support Trends Over Time
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: fig-national-trend
-#| fig-cap: "National Support Trends Over Time for the 2024 U.S.
-Presidential Election.
-The plot displays the predicted national average support for Donald Trump (red) and Kamala Harris (blue) over time, based on a Bayesian hierarchical model with a natural spline fit.
-The trends show changes in polling support for both candidates throughout 2024, highlighting key shifts in public opinion as the election approaches."
-#| warning: false
-
-# Generate a sequence of dates for national predictions
-prediction_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
-
-# Predict national support for each date
-national_trend <- prediction_dates %>%
-  map_dfr(~ {
-    data <- data.frame(state = unique(clean_poll_data$state), end_date = .x)
-    data$end_date_num <- as.numeric(data$end_date - min(clean_poll_data$end_date))
-
-    trump_pred <- mean(posterior_predict(trump_model, newdata = data))
-    harris_pred <- mean(posterior_predict(harris_model, newdata = data))
-
-    tibble(date = .x, Trump = trump_pred, Harris = harris_pred)
-  })
-
-# Plot national support trends
-ggplot(national_trend, aes(x = date)) +
-  geom_line(aes(y = Trump, color = "Trump")) +
-  geom_line(aes(y = Harris, color = "Harris")) +
-  labs(title = "National Support Trends Over Time", y = "Predicted Support", x = "Date") +
-  scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  theme_minimal()
-
-```
-EDIT - explain results on graph
-
 ## State Level Support Trends Over Time
@@
 #| eval: true
-#| label: fig-state-trend
-#| fig-cap: "State-Level Support Trends for Trump and Harris Over Time."
+#| label: fig-state-polling-average
+#| fig-cap: "Polling Averages for Trump and Harris Over Time in Battleground States."
 #| warning: false
@@
 
-# Generate predictions for each state and date
-state_dates <- seq.Date(min(clean_poll_data$end_date), max(clean_poll_data$end_date), by = "1 week")
-state_prediction_grid <- expand.grid(state = battleground_states, end_date = state_dates) %>%
-  mutate(end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))
-
-# Predict support for each state and date
-state_predictions <- state_prediction_grid %>%
-  rowwise() %>%
-  mutate(
-    Trump_Support = mean(posterior_predict(trump_model, newdata = cur_data())),
-    Harris_Support = mean(posterior_predict(harris_model, newdata = cur_data()))
+# Calculate average polling percentages over time for battleground states
+polling_average <- clean_poll_data %>%
+  filter(state %in% battleground_states) %>%
+  group_by(end_date, state) %>%
+  summarise(
+    Average_Trump = mean(Trump_pct, na.rm = TRUE),
+    Average_Harris = mean(Harris_pct, na.rm = TRUE),
+    .groups = 'drop'
   )
 
-# Plot state-level support trends over time with adjusted x-axis labels
-ggplot(state_predictions, aes(x = end_date)) +
-  geom_line(aes(y = Trump_Support, color = "Trump")) +
-  geom_line(aes(y = Harris_Support, color = "Harris")) +
-  labs(title = "State-Level Support Trends for Trump and Harris Over Time (Battleground States)", 
-       y = "Predicted Support", 
+# Plot polling averages over time for each battleground state
+ggplot(polling_average, aes(x = end_date)) +
+  geom_line(aes(y = Average_Trump, color = "Trump")) +
+  geom_line(aes(y = Average_Harris, color = "Harris")) +
+  labs(title = "Polling Averages for Trump and Harris Over Time in Battleground States", 
+       y = "Average Support (%)", 
        x = "Date") +
@@
   scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months") +  # Set date format and break frequency
+  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  # Set date format and break frequency
   theme_minimal() +
@@
 ```
+
 EDIT - explain results on graph

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

@@
 
-We use the statistical programming language R [cite R]... to perform data simulation, testing, cleaning, and analysis.
+The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
+For all the following variables mentioned, we collected observations for both of the candidates.
+We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris.
+We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’ and ‘state’.
+Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data quality are further explained in {additional data details}.
+The electorate support for each candidate is denoted as ‘pct’.
+
+All of the variables mentioned above are used during the steps of data cleaning, simulation, testing, and analysis.
+We use the statistical programming language R [cite R]...
+to perform data simulation, testing, cleaning, and analysis.
 
@@
 ### Outcome Variable
-The outcome variable in this study is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S.
+The outcome variable in our model is ‘predicted_winner’, which is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S.
 Presidential Election based on polling data from key battleground states.
@@
 The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris.
-This binary outcome variable serves a crucial role in facilitating the analysis of voter support dynamics and the effectiveness of the predictive modeling employed in this paper.
 
-By focusing on this dichotomy, we can glean insights into how polling data might translate into electoral outcomes, highlighting the competitive nature of the election in these pivotal states.
-The construction of the outcome variable not only enhances our understanding of voter preferences but also underscores the significance of binary classification in electoral forecasting, especially in states that could ultimately sway the overall election results.
+The ‘predicted_winner’ variable in our model is a constructed binary indicator derived from the pct variable based on our model.
+The “pct” variable represents the percentage of support for each candidate, Harris and Trump, within each state.
+To create the “winner” variable, we compare the “pct” values: if Trump’s percentage of support in a given state is higher than Harris’s, the “winner” variable is assigned a value indicating Trump as the predicted winner for the given state; otherwise, it is assigned to Harris.
+This construction allows us to translate individual support levels from the raw data into a binary prediction of each state’s likely outcome.
 
 ### Predictive Variable
-The predictive variable in this study is the polling support percentage for each candidate, specifically for Donald Trump and Kamala Harris, measured across various battleground states.
-This variable represents the predicted likelihood of each candidate winning in the specified states based on the polling data collected during the election cycle.
+The predictive variables for this paper are each candidate’s support rate (‘pct’), which directly influences the “winner” variable.
+The ‘pct’ variable represents the percentage of the vote or support that each candidate received in the poll.
+Understanding these polling percentages is crucial, as they provide insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election.
+The battleground states presented in the table are particularly significant due to their historical tendency to swing between parties, making them vital focal points for campaign efforts.
+By examining the outcome variable through these polling percentages, we enhance our predictive modeling and gain a deeper understanding of the factors influencing the election.
+This comparison not only contextualizes the current political climate but also highlights the critical role these battleground states play in determining the overall election results.
+As the campaign unfolds, these insights will be invaluable for strategizing and gauging voter sentiment in these pivotal areas.
+
 ```{r}
@@
 @tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada.
-The values reflect the polling percentages for each candidate, providing a straightforward comparison of their standings within these critical regions.
-
-Understanding these polling percentages is essential as they offer insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election.
-The battleground states depicted in the table are particularly significant due to their historical tendency to swing between parties, making them crucial targets for campaign efforts.
-Analyzing the outcome variable through these polling percentages informs our predictive modeling and enhances our comprehension of the dynamics at play during the electoral process.
-This comparison not only contextualizes the current political climate but also underlines the importance of these battleground states in shaping the overall election results.
-As the campaign progresses, these insights will be invaluable for strategizing and understanding voter sentiment in these pivotal areas.
+The values reflect the polling percentages for each candidate, based on 'pct', providing a straightforward comparison of the predictive variables in each battleground state.
+It is important to note that since ‘pct’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
+In reality, as seen in Arizona, the combined ‘pct’ for Trump and Harris is 92.9%, with the remaining 7.1% likely reflecting support for other candidates in the race.
 

### 2024-11-03 paper/paper.qmd
committing changes to model and predictions in paper.qmd

@@
 library(broom.mixed)
+library(brms)
 ```
@@
 #| label: fig-summarystatistics
-#| fig-cap: "Vertical Summary Statistics for Trump and Harris Polling Data.
+#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
 This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
-All values are rounded to three decimal places, providing a concise overview of the polling landscape as of the selected election date."
+All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
 #| echo: false
@@
 #| label: tbl-outcomevar
-#| fig-cap: "Average polling support percentages for Donald Trump and Kamala Harris across key battleground states.
-The table presents the average support levels, indicating Trump’s and Harris’s polling figures in states such as Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada.
-The comparative analysis highlights the varying levels of candidate support in these critical regions, providing insight into the electoral landscape ahead of the 2024 U.S.
-Presidential Election."
+#| fig-cap: "Summary of fixed effects estimates from the Bayesian hierarchical models for both Donald Trump and Kamala Harris.
+The table includes the estimated coefficients, standard errors, and 95% confidence intervals for each predictor, illustrating how various factors influence polling percentages.
+The significant intercept and time-related variables are crucial for understanding candidate support dynamics over the election period."
+
 #| echo: false
@@
 
-## State Level Support Trends Over Time
+## Polling Averages for Trump and Harris Over Time in Battleground States
 ```{r}
@@
 ## Probability of Winning by State on Election Day 
-
 ```{r}
@@
 
-# Set election day or target prediction date
-election_day <- as.Date("2024-11-05")  # Example election day
+# Step 1: Prepare new data frame for November 5, 2024
+election_day <- as.Date("2024-11-05")
+min_date <- min(clean_poll_data$end_date, na.rm = TRUE)
+end_date_num_election <- as.numeric(election_day - min_date)
 
-# Predict probability of Trump and Harris winning by comparing posterior predictions for each state
-state_predictions <- clean_poll_data %>%
-  distinct(state) %>%
-  filter(state %in% battleground_states) %>%  # Only include battleground states
-  mutate(end_date = election_day, 
-         end_date_num = as.numeric(end_date - min(clean_poll_data$end_date)))
-
-# Calculate probabilities of Trump and Harris winning in each state
-state_predictions <- state_predictions %>%
-  rowwise() %>%
-  mutate(
-    prob_trump_wins = mean(posterior_predict(trump_model, newdata = cur_data()) > 
-                           posterior_predict(harris_model, newdata = cur_data())),
-    prob_harris_wins = mean(posterior_predict(harris_model, newdata = cur_data()) > 
-                            posterior_predict(trump_model, newdata = cur_data()))
+# Create a data frame for predictions, one row per state
+new_data <- data.frame(
+  state = unique(clean_poll_data$state),
+  end_date_num = rep(end_date_num_election, length(unique(clean_poll_data$state)))
 )
 
-# Prepare the state predictions for the table
-state_summary_table <- state_predictions %>%
-  select(state, prob_trump_wins, prob_harris_wins) %>%
-  mutate(
-    prob_trump_wins = round(prob_trump_wins * 100, 2),  # Convert to percentage and round
-    prob_harris_wins = round(prob_harris_wins * 100, 2)  # Convert to percentage and round
+# Step 2: Generate predictions for both models
+trump_predictions <- posterior_predict(trump_model, newdata = new_data)
+harris_predictions <- posterior_predict(harris_model, newdata = new_data)
+
+# Step 3: Calculate winning percentages
+# Calculate mean predictions and convert to percentages
+trump_mean_pct <- apply(trump_predictions, 2, mean) * 100
+harris_mean_pct <- apply(harris_predictions, 2, mean) * 100
+
+# Combine results into a data frame
+state_predictions <- data.frame(
+  state = new_data$state,
+  trump_pct = trump_mean_pct,
+  harris_pct = harris_mean_pct
 )
 
-# Display state predictions in table
-knitr::kable(state_summary_table) %>%
-  kable_styling(full_width = FALSE, position = "center")
+# Round the percentages to two decimal points
+state_predictions <- state_predictions %>%
+  mutate(
+    trump_pct = round(trump_pct, 2),
+    harris_pct = round(harris_pct, 2),
+    winner = ifelse(trump_pct > harris_pct, "Donald Trump", "Kamala Harris")
+  )
+
+# Print results in table
+kable(state_predictions, 
+      col.names = c("State", 
+                    "Predicted Harris Win (%)", 
+                    "Predicted Trump Win (%)", 
+                    "Predicted Winner")) %>%
+  kable_styling("striped", full_width = F)
 
@@
 
-# Reshape the data for plotting only for battleground states
-state_predictions_long <- state_predictions %>%
-  select(state, prob_trump_wins, prob_harris_wins) %>%
-  filter(state %in% battleground_states) %>%  # Ensure we're only working with battleground states
-  pivot_longer(cols = starts_with("prob"), 
-               names_to = "Candidate", 
-               values_to = "Probability") %>%
-  mutate(Candidate = recode(Candidate, 
-                            prob_trump_wins = "Trump", 
-                            prob_harris_wins = "Harris"))
+# Reshape the data for ggplot
+state_predictions_long <- predicted_results %>%
+  pivot_longer(cols = c(harris_pct, trump_pct), 
+               names_to = "candidate", 
+               values_to = "winning_probability") %>%
+  mutate(candidate = recode(candidate, 
+                            harris_pct = "Kamala Harris", 
+                            trump_pct = "Donald Trump"))
 
-# Plot double bar graph of predicted winning percentages by state
-ggplot(state_predictions_long, aes(x = reorder(state, Probability), y = Probability * 100, fill = Candidate)) +
-  geom_col(position = "dodge") +  # Use dodge for side-by-side bars
-  coord_flip() +  # Flip coordinates for better readability
-  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
+# Create the bar chart
+ggplot(state_predictions_long, aes(x = state, y = winning_probability, fill = candidate)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  coord_flip() +  # Flip coordinates for horizontal bars
+  scale_fill_manual(values = c("blue", "red")) +  # Specify colors
   labs(title = "Predicted Winning Percentages by Battleground State on Election Day",
-       x = "State", 
-       y = "Winning Probability (%)") +
+       x = "Winning Probability (%)",
+       y = "State") +
   theme_minimal() +
-  theme(legend.position = "right")
+  theme(legend.title = element_blank())  # Remove legend title
 
@@
 
-# Add predicted winner column based on `prob_trump_wins`
-state_predictions <- state_predictions %>%
-  filter(state %in% battleground_states) %>%  # Ensure we only include battleground states
-  mutate(predicted_winner = ifelse(prob_trump_wins > 0.5, "Trump", "Harris"))
+# Step 4: Plot the US map with predicted winner for each state
+# Convert state names to lowercase to match the map data
+state_predictions$state <- tolower(state_predictions$state)
 
-# Plot the US map with predicted winner for each state
-plot_usmap(data = state_predictions, values = "predicted_winner", regions = "states") +
-  scale_fill_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  labs(title = "Predicted Winner by Battleground State on Election Day", fill = "Predicted Winner") +
+# Create the map
+us_map <- plot_usmap(data = state_predictions, values = "winner", regions = "states") +
+  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue", "NA" = "gray"), 
+                    na.value = "gray") +
+  labs(title = "Predicted Winner by Battleground State on Election Day", 
+       fill = "Predicted Winner") +
   theme(legend.position = "right")
 
+# Print the map
+print(us_map)
 ```

### 2024-11-03 scripts/03-clean_data.R

@@ -7,11 +7,11 @@
 # Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 # Any other information needed? N/A
 
-
 # Load the necessary libraries
 library(tidyverse)
 library(janitor)
 library(lubridate)
+library(arrow)
 
 # Define battleground states
 battleground_states <- c("Arizona", "Georgia", "Michigan", "Nevada", 
@@ -23,20 +23,22 @@ poll_data <- read_csv("data/01-raw_data/raw_data.csv") |>
 
 # Filter and select relevant columns, including only battleground states
 clean_poll_data <- poll_data |>
-  select(state, end_date, pollster, numeric_grade, transparency_score, candidate_name, pct) |>
+  select(state, start_date, end_date, pollster, numeric_grade, transparency_score, candidate_name, pct) |>
   filter(
     !is.na(state),
     !is.na(numeric_grade),
     !is.na(transparency_score),
-    numeric_grade >= 3.0,          # numeric_grade threshold
+    numeric_grade >= 2.5,          # numeric_grade threshold
     transparency_score > 5,        # transparency_score threshold
     candidate_name %in% c("Donald Trump", "Kamala Harris"),
     state %in% battleground_states   # Filter to include only battleground states
   ) |>
   mutate(
+    start_date = as.Date(start_date, format = "%m/%d/%y"),
     end_date = as.Date(end_date, format = "%m/%d/%y"),
     end_date_num = as.numeric(end_date - min(end_date, na.rm = TRUE))
-  )
+  ) |>
+  filter(start_date > as.Date("2024-07-21"))  # Filter for start_date after July 21st
 
 # Replace "Nebraska CD-2" with "Nebraska" if it appears in the data
 clean_poll_data <- clean_poll_data |>
@@ -52,4 +54,4 @@ clean_poll_data <- clean_poll_data |>
 
 #### Save Data ####
 write_parquet(clean_poll_data, "data/02-analysis_data/analysis_data.parquet")
-print(clean_poll_data)
+

### 2024-11-03 scripts/06-model_data.R

@@ -7,22 +7,27 @@
 # Pre-requisites: Downloaded data from url "https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 # Any other information needed? N/A
 
-
-#### Modeling Script ####
 # Load necessary libraries
 library(tidyverse)
 library(rstanarm)
 library(splines)
+library(arrow)  # For reading/writing parquet files
+library(brms)
 
+# Create a dataset for Donald Trump
+trump_data <- clean_poll_data %>%
+  select(state, end_date, Trump_pct) %>%
+  rename(pct = Trump_pct)
 
-# Load cleaned data
-clean_poll_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
+# Convert end_date to numeric (days since the minimum date)
+trump_data <- trump_data %>%
+  mutate(end_date_num = as.numeric(end_date - min(end_date, na.rm = TRUE)))
 
-# Fit Trump model with time spline and state random effects
+# Fit a Bayesian model for Donald Trump using natural splines for end_date_num
 trump_model <- brm(
-  Trump_pct ~ ns(end_date_num, df = 4) + (1 | state),  # Natural spline with 4 degrees of freedom
-  data = clean_poll_data,
-  family = gaussian(),
+  bf(pct / 100 ~ ns(end_date_num, df = 4) + (1 | state)),  # Natural splines for end_date_num
+  data = trump_data,
+  family = Beta(link = "logit"),  # Use beta regression for proportions
   prior = c(
     prior(normal(0, 5), class = "b"),
     prior(normal(0, 10), class = "Intercept")
@@ -33,25 +38,34 @@ trump_model <- brm(
   control = list(adapt_delta = 0.95)
 )
 
-# Fit Harris model with time spline and state random effects
-harris_model <- brm(
-  Harris_pct ~ ns(end_date_num, df = 4) + (1 | state),
-  data = clean_poll_data,
-  family = gaussian(),
-  prior = c(
-    prior(normal(0, 5), class = "b"),
-    prior(normal(0, 10), class = "Intercept")
-  ),
-  seed = 1234,
-  iter = 3000,
-  chains = 4,
-  control = list(adapt_delta = 0.95)
-)
-
-
-
-# Save the models
+# Save the Trump model if needed
 saveRDS(trump_model, "models/trump_model.rds")
-saveRDS(harris_model, "models/harris_model.rds")
 
[…]

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

@@
 # Reshape the data for ggplot
-state_predictions_long <- predicted_results %>%
+state_predictions_long <- state_predictions %>%
   pivot_longer(cols = c(harris_pct, trump_pct), 
@@
   coord_flip() +  # Flip coordinates for horizontal bars
-  scale_fill_manual(values = c("blue", "red")) +  # Specify colors
+  scale_fill_manual(values = c("red", "blue")) +  # Specify colors
   labs(title = "Predicted Winning Percentages by Battleground State on Election Day",

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

@@
 ---
-title: "Predicting the 2024 U.S. Presidential Election: Trump’s Likely Victory Based on Polling Analysis"
-subtitle: "A Bayesian Approach Reveals Strong Support for Donald Trump Across Key Battleground States, Highlighting Critical Trends in Voter Preferences Leading Up to Election Day"
+title: "Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory"
+subtitle: A Bayesian Analysis of Polling Data from Key Battleground States"
 author: 
@@
 date-format: long
-abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, utilizing state-level polling data to focus specifically on key battleground states.
-By employing a Bayesian approach, we estimate the outcomes between Donald Trump and Kamala Harris, incorporating high-quality polls from reputable sources.
-Our Bayesian hierarchical model, featuring time-dependent splines and state-specific random effects, effectively captures shifts in candidate support over time.
-The analysis reveals that Trump is projected to win in critical battleground states such as Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, highlighting his advantage in these pivotal regions.
-These findings underscore the significance of battleground states in determining electoral outcomes, particularly in a highly polarized political landscape.
-We also discuss the model’s performance, accuracy, and limitations, reflecting on the unique challenges inherent in forecasting electoral outcomes in these crucial areas."
+abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, focusing on critical battleground states.
+Utilizing a Bayesian approach with state-level polling data, we estimate the winning probabilities for candidates Donald Trump and Kamala Harris.
+Our findings reveal that Harris is projected to win in six out of seven key battleground states, including Michigan, Nevada, and North Carolina, with predicted winning percentages surpassing 50%.
+Given this significant advantage, we predict Kamala Harris to be the likely winner of the election.
+This analysis underscores the importance of battleground states in shaping electoral outcomes and provides valuable insights into voter sentiment as Election Day approaches."
 format: pdf
@@
 # Introduction
+The 2024 United States Presidential Election is shaping up to be a critical event, particularly as battleground states emerge as pivotal areas influencing the electoral outcome.
+As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions becomes essential for predicting the election results.
+This paper utilizes state-level polling data to build a predictive model that estimates the likelihood of victory for each candidate in several key battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
 
-EDIT & ELABORATE - gpt generated
-The 2024 United States Presidential Election presents a unique opportunity to examine the dynamics of voter preferences in key battleground states.
-With a highly polarized electorate and the stakes at an all-time high, accurately predicting the outcome of this election is crucial for political analysts, campaigns, and the public.
-This paper develops a predictive model that leverages state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, using a Bayesian hierarchical approach to accommodate variations across states.
+The primary estimand in this study is the probability that either Donald Trump or Kamala Harris will win each battleground state, based on aggregated polling averages.
+Our model employs a Bayesian hierarchical approach with natural splines to capture the nuances of voter sentiment over time, allowing for the estimation of winning probabilities across these crucial states.
+By comparing the predicted winning probabilities for both candidates, we can ascertain the likely winner based on the prevailing voter support patterns.
 
-The central estimand of this study is the probability of victory for each candidate in battleground states, defined by the binary outcome variable indicating a predicted win for Trump (1) or Harris (0).
-To achieve this, we utilize aggregated polling percentages as predictor variables, allowing us to capture shifts in voter sentiment over time.
-By employing logistic regression and accounting for state-specific random effects, our model not only identifies trends in polling data but also highlights critical regions that could determine the overall election outcome.
+Our analysis reveals that Kamala Harris is projected to win in six out of the seven battleground states examined, indicating a favorable voter sentiment in states such as Michigan and Nevada.
+Conversely, Donald Trump shows a competitive stance in Arizona, where he is also predicted to secure a slight edge.
+The calculation of the predicted winner is determined by comparing the estimated probabilities for each candidate, which underscores Harris's overall advantage in the election.
 
-In our analysis, we focused on key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—where the margins are often razor-thin and can swing the election in favor of one candidate or another.
-Our findings indicate that Trump is predicted to win in these crucial states, positioning him as the likely overall victor in the 2024 election.
-This is significant as it underscores the importance of battleground states in shaping the electoral landscape and highlights the intricate dynamics of voter preferences that can influence the final result.
+Understanding these predictions is vital as it offers insights into the dynamics of voter behavior and the potential implications for election outcomes.
+This information can help political analysts, campaign strategists, and the public anticipate how these battleground states may influence the overall results.
+Accurate predictions not only inform campaign strategies but also highlight regions where voter sentiment may shift in the lead-up to Election Day.
 
-The structure of this paper is as follows: we begin by detailing the data sources and methodology used for the analysis.
-Next, we present the results, focusing on the predicted probabilities of each candidate winning in the targeted battleground states.
-Finally, we discuss the implications of our findings and the limitations of our model, contributing to a better understanding of voter dynamics in the 2024 U.S.
-Presidential Election.
+The remainder of this paper is structured as follows.
+In Section 2, we describe the data and variables used in the analysis.
+Section 3 outlines the model setup and estimation strategy.
+Section 4 presents the results, including the predicted election outcomes and visualizations.
+Finally, Section 5 discusses the implications and limitations of our findings
 
@@
 The significant intercept and time-related variables are crucial for understanding candidate support dynamics over the election period."
-
 #| echo: false
@@
       col.names = c("State", 
+                    "Predicted Trump Win (%)", 
                     "Predicted Harris Win (%)", 
-                    "Predicted Trump Win (%)", 
                     "Predicted Winner")) %>%

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

@@
 
-The goal of our modeling strategy is to predict the winner of the 2024 US Presidential Election in each state based on polling data, while also assessing the likelihood of Donald Trump or Kamala Harris winning.
-We use a Bayesian hierarchical modeling approach with natural splines to capture time-dependent trends in support and random effects to account for state-level variability.
-This strategy enables us to model changes in candidate support over time across states, allowing for more nuanced, probabilistic predictions of each candidate’s support levels and overall election outcomes.
+In this analysis, we develop separate Bayesian models for candidates Donald Trump and Kamala Harris to predict their respective probabilities of winning in the 2024 U.S.
+Presidential Election.
+The use of distinct models allows us to capture the unique polling dynamics and voter preferences for each candidate across key battleground states, which include Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+By analyzing polling data collected from various sources, we aim to provide a comprehensive assessment of each candidate’s electoral prospects.
 
-Separate models are fit for each candidate’s polling support, with state-specific random intercepts and a natural spline for time, allowing the model to capture state-level deviations from the national average as well as time-based changes in support.
-This hierarchical Bayesian approach enables partial pooling across states, stabilizing predictions for states with limited polling data.
+The models are built using state-level polling data, which is aggregated to ensure robustness in the predictions.
+Each candidate's model incorporates temporal factors through natural splines applied to the numeric representation of the end date, allowing us to observe changes in voter support over time.
+This approach provides a nuanced understanding of how public sentiment evolves as the election date approaches.
 Further background details, model specifications, and diagnostics are included in [Appendix -@sec-model-details].
 
-## Model set-up
-The binary outcome variable $y_i$ represents whether Donald Trump is predicted to lead in state $i$ based on polling averages for both candidates over time.
-Here, $y_i = 1$ indicates a Trump lead in state $i$, and $y_i = 0$ indicates a lead for Kamala Harris.
-The model estimates the support levels for each candidate based on polling data using a Gaussian likelihood function, with time trends captured by a natural spline and state-level variation captured through random intercepts.
+## Model Setup
+The primary estimand in our models is the probability of victory for each candidate in the selected battleground states, expressed through a logistic regression framework.
+The model is formulated as follows:
 
-The predictor variables are $Trump\_pct_i$ and $Harris\_pct_i$, which represent the predicted percentages of support for Donald Trump and Kamala Harris in state $i$.
-A separate model is fit for each candidate’s support over time, capturing the time-varying nature of support with a natural spline function and allowing for state-specific random effects.
+$$
+P(\text{Victory}_i) = \frac{e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num}_i + u_{\text{state}}}}{1 + e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num}_i + u_{\text{state}}}}
+$$
 
-The model is specified as follows:
+Where:
+- \(P(\text{Victory}_i)\) is the predicted probability of winning for candidate \(i\).
+- \(\beta_0\) represents the intercept, reflecting the baseline log-odds of winning.
+- \(\beta_1\) is the coefficient for the natural spline transformation of the numeric end date (\( \text{end\_date\_num}_i \)), capturing nonlinear trends in voter support over time.
+- \(u_{\text{state}}\) is the random effect associated with each state, accounting for variations in voter preferences.
+
+Data preparation involved creating two distinct datasets from the cleaned polling data: one for Donald Trump and another for Kamala Harris.
+Each dataset included the state, end date, and corresponding polling percentage, with the end date converted into a numeric format representing the number of days since the earliest date in the dataset.
+This allows the model to effectively capture the evolving nature of voter sentiment as the election approaches.
+
+We implemented the models using the `brms` package in R, selecting a Beta distribution with a logit link function to accurately model the proportions of support for each candidate:
 
 \[
-\begin{align*}
-\text{Trump\_pct}_i &\sim \text{Normal}(\mu_{\text{Trump}, i}, \sigma_{\text{Trump}}) \\
-\mu_{\text{Trump}, i} &= \alpha_{\text{Trump}} + f(\text{end\_date\_num}_i) + u_{\text{state}}^{\text{Trump}} \\
-u_{\text{state}}^{\text{Trump}} &\sim \text{Normal}(0, \tau_{\text{Trump}}) \\
-\text{Harris\_pct}_i &\sim \text{Normal}(\mu_{\text{Harris}, i}, \sigma_{\text{Harris}}) \\
-\mu_{\text{Harris}, i} &= \alpha_{\text{Harris}} + f(\text{end\_date\_num}_i) + u_{\text{state}}^{\text{Harris}} \\
-u_{\text{state}}^{\text{Harris}} &\sim \text{Normal}(0, \tau_{\text{Harris}})
-\end{align*}
+y_i \sim \text{Beta}(\mu_i, \phi)
 \]
 
-Where:
-
-- \begin{equation}
-  \text{Trump\_pct}_i \text{ and } \text{Harris\_pct}_i \text{ are the polling support levels for Donald Trump and Kamala Harris in state } i.
-  \end{equation}
-- \begin{equation}
-  f(\text{end\_date\_num}_i) \text{ is a natural spline function of time (measured as } \text{end\_date\_num}, \text{ with 4 degrees of freedom, capturing the non-linear time trends in support.}
-  \end{equation}
-- \begin{equation}
-  \alpha_{\text{Trump}} \text{ and } \alpha_{\text{Harris}} \text{ are intercepts for the models, representing the baseline support levels for each candidate.}
-  \end{equation}
-- \begin{equation}
-  u_{\text{state}}^{\text{Trump}} \text{ and } u_{\text{state}}^{\text{Harris}} \text{ are state-level random effects, accounting for differences in baseline support across states.}
-  \end{equation}
-
-For both models, we use weakly informative Normal priors.
-The intercepts $ \alpha_{\text{Trump}} $ and $ \alpha_{\text{Harris}} $ are given Normal priors with a mean of 0 and a standard deviation of 10, reflecting prior uncertainty about the baseline support levels.
-The time trend coefficients are also given Normal priors with mean 0 and standard deviation 5, allowing for flexibility in capturing the temporal changes in support.
-State-level random effects for each candidate are modeled with a Normal distribution centered at 0, with the variance estimated from the data.
-
-The model was implemented in R using the `brms` package, which allows for Bayesian inference using `Stan` as the backend.
-This approach captures the dynamic nature of polling support over time, while accounting for state-specific variability and allowing us to estimate the probability of Trump or Harris leading in each state.
+Here, \(y_i\) represents the proportion of support for candidate \(i\) in a given state.
 
@@
 
-We expect a positive relationship between each candidate's polling support over time and their likelihood of winning a given state.
-As Donald Trump’s polling percentage in a state increases, his probability of leading in that state rises, while higher support for Kamala Harris increases her probability of winning.
-By modeling each candidate's support trends over time, we capture shifts in voter sentiment that reflect campaign dynamics, national events, and other time-dependent factors influencing support levels.
+The decision to utilize Bayesian logistic regression models stems from their flexibility and the ability to incorporate prior beliefs into the estimation process.
+This approach is particularly valuable in the context of electoral forecasting, where uncertainty is a significant factor.
+We defined the following priors for the model parameters:
 
-A Bayesian hierarchical model was chosen for its ability to handle the complexity of election forecasting, where both temporal trends and state-specific variations play significant roles.
-Bayesian inference allows us to incorporate prior beliefs and uncertainties about candidate support, while the hierarchical structure enables partial pooling across states.
-This is particularly useful for states with sparse data, as the model can "borrow strength" from the national trend, stabilizing predictions.
+- For the intercept \(\beta_0\), we specified \( \beta_0 \sim \text{Normal}(0, 10)\), providing a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds.
+- For the slope \(\beta_1\), we used \( \beta_1 \sim \text{Normal}(0, 5)\), reflecting reasonable expectations regarding the influence of polling trends over time.
 
-The inclusion of a natural spline for time allows us to model non-linear changes in support, capturing important shifts in polling trends as the election approaches.
-Random effects for each state account for baseline differences in support levels, reflecting the unique political climate of each state and reducing the risk of overfitting by allowing state-level deviations from the national average.
-
-This model assumes that polling data reasonably reflects voter intentions and that each state’s outcome can be modeled independently of others, though regional trends are partially addressed by state-level random effects.
-Potential biases inherent in polling data, such as underrepresented demographics or sampling error, could affect our predictions.
-The model was implemented in R using the brms package, which provides Bayesian inference via Stan, allowing for flexible, probabilistic predictions of each candidate's support across states.
+Natural splines were chosen to model the effect of the end date due to the observed nonlinearities in voter support, allowing the model to adapt to changing public opinion.
+Key assumptions include the independence of observations within states and the appropriateness of the Beta distribution for modeling proportions.
+However, potential limitations include the reliance on the accuracy of polling data and the assumption that historical voting patterns are indicative of future behavior.
 
@@
 
-EDIT - explain model summary table
+The model summary for Kamala Harris indicates that the intercept estimate is \(-0.1508\) with a standard error of \(0.0279\).
+This intercept value suggests a negative baseline log-odds for her victory, reflecting the starting point of her support in the absence of any other predictors.
+The model shows that the coefficients for the natural spline terms (\(nsenddatenum\)) are particularly informative, with the first spline term (\(nsendatenum\_df41\)) yielding a positive estimate of \(0.1225\), suggesting that as time progresses, Harris’s probability of winning increases, particularly during key moments captured by the polling data.
+This is also reflected in the confidence intervals, with the upper limit of the confidence interval for this term reaching \(0.1903\).
+In contrast, the estimates for the second spline term (\(nsenddatenum\_df42\)) indicate a smaller effect, suggesting fluctuations in her support that may w
 
@@
 
-EDIT - explain model summary table
+In comparison, the model summary for Donald Trump presents a different picture.
+The intercept for Trump's model is \(-0.0923\) with a standard error of \(0.0306\), indicating a similarly negative baseline log-odds of victory as Harris.
+However, Trump's model demonstrates more pronounced increases in support as reflected by the estimates for the natural spline terms.
+For example, the second spline term (\(nsenddatenum\_df42\)) has a positive coefficient of \(0.0803\), highlighting a significant upward trend in support over time.
+This trend is further supported by the confidence interval, which ranges from \(0.0179\) to \(0.1426\), indicating a substantial increase in his predicted probabilities as the election approaches.
+The fourth spline term (\(nsenddatenum\_df44\)) also shows a positive coefficient (\(0.0789\)), suggesting that the support for Trump remains resilient over the polling period.
+Overall, both models reveal complex interactions between time and voter sentiment, with implications for the electoral dynamics leading up to the election.
 
@@
 ## Polling Averages for Trump and Harris Over Time in Battleground States
+@fit-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024.
+Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support.
+Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October.
+This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 4th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
 ```{r}
@@
 ```
-
-EDIT - explain results on graph
+Out of the seven battleground states, Arizona and Georgia display a clear leading win by Trump while Nevada and North Carolina display a clear leading win by Harris.
+Michigan, Pennsylvania, and Wisconsin are the three states that don’t have a clear leading winner.
+Therefore, these three states are the key states that could determine the election outcome as it approaches election day.
+Election outcomes could be easily flipped if either of the candidates can sweep more than one out of the three states.
+Candidates should focus on strategizing their campaigns to increase their voters' support at critical moments.
+The frequent fluctuations of greater margins also indicate the uncertainty and instability of voters’ attitudes towards both candidates.
+Thus, it is difficult to conclude a definitive predictive outcome by simply observing the current data.
 

### 2024-11-03 paper/paper.qmd
committing new version of paper.qmd

@@
 ## Variables
+The collected dataset contains several key variables relevant to the analysis:
+
+- **State**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
+- **Polling Date**: The date when the poll was conducted, which helps in analyzing trends over time.
+- **Pollster**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
+- **Numeric Grade**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
+- **Transparency Score**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
+- **Candidate Name**: Identifies the candidate being polled, specifically focusing on Donald Trump and Kamala Harris.
+- **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.
+
+To ensure the quality of the analysis, only those polls that meet specific thresholds were included - refer to [@sec-appendix].
+
+### Outcome Variable
+
 To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
 ```{r}
-#| label: fig-summarystatistics
+#| label: tbl-outcomevar
 #| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
@@
 
-@tbl-summarystatistics reveals a total of 34 polls conducted, with an average support level of 46.321% for Trump and 46.661% for Harris.
-Trump's support peaked at 51.000%, while the lowest recorded support was 42.500%.
-For Harris, the maximum support also reached 51.000%, with a minimum of 42.000%.
-These statistics illustrate the competitive landscape in these critical states, emphasizing the narrow margins that could ultimately influence the election outcome.
-The data reflects the volatility of voter sentiment leading up to the 2024 Presidential Election, underscoring the importance of these battleground states in determining the final result.
+MISSING - explain table
 
-### Outcome Variable
-The outcome variable in our model is ‘predicted_winner’, which is a binary indicator, specifically constructed to represent the predicted winner of the 2024 U.S.
-Presidential Election based on polling data from key battleground states.
-This variable categorizes the election outcome into two distinct categories: a value of "Trump" indicates that Donald Trump is predicted to win, while a value of "Harris" denotes a predicted victory for Kamala Harris.
-The classification is derived from a conditioning statement that assigns the predicted winner based on the probability of Trump winning; if this probability exceeds 0.5, the predicted winner is labeled as Trump, otherwise, it is labeled as Harris.
-
-The ‘predicted_winner’ variable in our model is a constructed binary indicator derived from the pct variable based on our model.
-The “pct” variable represents the percentage of support for each candidate, Harris and Trump, within each state.
-To create the “winner” variable, we compare the “pct” values: if Trump’s percentage of support in a given state is higher than Harris’s, the “winner” variable is assigned a value indicating Trump as the predicted winner for the given state; otherwise, it is assigned to Harris.
-This construction allows us to translate individual support levels from the raw data into a binary prediction of each state’s likely outcome.
-
-### Predictive Variable
-The predictive variables for this paper are each candidate’s support rate (‘pct’), which directly influences the “winner” variable.
-The ‘pct’ variable represents the percentage of the vote or support that each candidate received in the poll.
-Understanding these polling percentages is crucial, as they provide insight into the electoral landscape and help evaluate which candidate may hold an advantage in the upcoming election.
-The battleground states presented in the table are particularly significant due to their historical tendency to swing between parties, making them vital focal points for campaign efforts.
-By examining the outcome variable through these polling percentages, we enhance our predictive modeling and gain a deeper understanding of the factors influencing the election.
-This comparison not only contextualizes the current political climate but also highlights the critical role these battleground states play in determining the overall election results.
-As the campaign unfolds, these insights will be invaluable for strategizing and gauging voter sentiment in these pivotal areas.
-
-```{r}
-#| label: tbl-outcomevar
-#| fig-cap: "Summary of fixed effects estimates from the Bayesian hierarchical models for both Donald Trump and Kamala Harris.
-The table includes the estimated coefficients, standard errors, and 95% confidence intervals for each predictor, illustrating how various factors influence polling percentages.
-The significant intercept and time-related variables are crucial for understanding candidate support dynamics over the election period."
-#| echo: false
-#| warning: false
-#| message: false
-
-# Aggregating polling percentages for Trump and Harris by state
-state_average <- clean_poll_data %>%
-  group_by(state) %>%
-  summarise(
-    Trump_pct = mean(Trump_pct, na.rm = TRUE),
-    Harris_pct = mean(Harris_pct, na.rm = TRUE)
-  )
-
-# Prepare table for display with rounding to 1 decimal place
-average_table <- state_average |>
-  mutate(
-    Trump_pct = round(Trump_pct, 1),
-    Harris_pct = round(Harris_pct, 1)
-  ) |>
-  select(
-    State = state, 
-    "Trump %" = Trump_pct, 
-    "Harris %" = Harris_pct, 
-  )
-
-# Render the table
-knitr::kable(average_table)
-
-```
-@tbl-predictivevar highlights the average support levels for both Donald Trump and Kamala Harris across key battleground states, including Wisconsin, North Carolina, Georgia, Arizona, Pennsylvania, Michigan, and Nevada.
-The values reflect the polling percentages for each candidate, based on 'pct', providing a straightforward comparison of the predictive variables in each battleground state.
-It is important to note that since ‘pct’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
-In reality, as seen in Arizona, the combined ‘pct’ for Trump and Harris is 92.9%, with the remaining 7.1% likely reflecting support for other candidates in the race.
+### Predictive Var
+The predictive variable in our model are 'time' and 'state'
 
@@
 
-WRITE SECTION
+'numeric grade', 'transparency score', 'pollster'
 
@@
 The frequent fluctuations of greater margins also indicate the uncertainty and instability of voters’ attitudes towards both candidates.
-Thus, it is difficult to conclude a definitive predictive outcome by simply observing the current data.
 
-## Probability of Winning by State on Election Day 
+Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in [@sec-model] to predict the possible outcomes by incorporating the given data.
+By using the current data, the model can identify the state-specific trends thus generating more accurate predictions of the outcome.
+Relevant prediction results are provided below.
+
+## State-Level Probability of Trump Winning on Election Day
+By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024).
+@fig-state-probabilities provides a visual summary of these predicted probabilities.
 ```{r}
@@
 #| eval: true
-#| label: tbl-state-probabilities
+#| label: fig-state-probabilities
 #| fig-cap: "Probability of Trump Winning by State on Election Day."
@@
 
-# Round the percentages to two decimal points
-state_predictions <- state_predictions %>%
-  mutate(
-    trump_pct = round(trump_pct, 2),
-    harris_pct = round(harris_pct, 2),
-    winner = ifelse(trump_pct > harris_pct, "Donald Trump", "Kamala Harris")
-  )
-
-# Print results in table
-kable(state_predictions, 
-      col.names = c("State", 
-                    "Predicted Trump Win (%)", 
-                    "Predicted Harris Win (%)", 
-                    "Predicted Winner")) %>%
-  kable_styling("striped", full_width = F)
-
-```
-
-EDIT - explain table results
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: fig-state-probabilities
-#| fig-cap: "Probability of Trump Winning by State on Election Day."
-#| warning: false
-
 # Reshape the data for ggplot
@@
 ```
-EDIT - explain graph results
+These predictions are based on polling data collected up to October 19th, 2024 to forecast each candidate's support within these critical states.
+Each state’s political dynamics are unique, and this is reflected in the model’s predictions.
+States like Arizona and Georgia show relatively close probabilities for both candidates, marking them as key states to focus on as they are the most influential on the election outcomes.
+The five other states all show relatively distinct leading support for Harris, suggesting that Harris’s support base in those states has been consistently strong in the polls leading up to Election Day.
+
+## Probability of Winning by State on Election Day 
+@tbl-fig-probabilities displays the numeric percentage of the probabilities of winning for Trump and Harris.
+By comparing the predicted probabilities, the predicted winner for each state is assigned accordingly.
+For instance, the table displays that Trump has a 49.48% predicted probability of winning Arizona compared to 49.24% of Harris’s, making Trump the predicted winner in Arizona.
+For all six other states, Harris has a higher predicted probability of winning than Trump.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-state-probabilities
+#| fig-cap: "Probability of Trump Winning by State on Election Day."
+#| warning: false
+
+# Round the percentages to two decimal points
+state_predictions <- state_predictions %>%
+  mutate(
+    trump_pct = round(trump_pct, 2),
+    harris_pct = round(harris_pct, 2),
+    winner = ifelse(trump_pct > harris_pct, "Donald Trump", "Kamala Harris")
+  )
+
+# Print results in table
+kable(state_predictions, 
+      col.names = c("State", 
+                    "Predicted Trump Win (%)", 
+                    "Predicted Harris Win (%)", 
+                    "Predicted Winner")) %>%
+  kable_styling("striped", full_width = F)
+
+```
 
 ## US Map Showing Predicted Winner by State on Election Day
+To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.
+
+- 'Predicted Winner' column is Donald Trump: State filled red
+- 'Predicted Winner' column is Kamala Harris: State filled blue
 
@@
 ```
-
-EDIT - explain map results
+According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
+Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.
 
 # Discussion {#sec-discussion}
-
 ## Interpretation of Results {#sec-results-interpretation}
-
 Since the electoral college system makes state-level forecasts crucial, and the U.S.

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

@@
 ## Overview
-This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight].
+NEEDS EDIT & ELABORATE - This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight].
 The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
@@
 
-The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
-For all the following variables mentioned, we collected observations for both of the candidates.
-We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris.
-We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’ and ‘state’.
-Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data quality are further explained in {additional data details}.
-The electorate support for each candidate is denoted as ‘pct’.
+To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
+```{r}
+#| label: tbl-outcomevar
+#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
+This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
+All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
+#| echo: false
+#| warning: false
+#| message: false
 
-All of the variables mentioned above are used during the steps of data cleaning, simulation, testing, and analysis.
-We use the statistical programming language R [cite R]...
-to perform data simulation, testing, cleaning, and analysis.
+# Load cleaned data
+clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
+
+# Summary statistics for the relevant columns in clean_poll_data
+summary_statistics <- clean_poll_data %>%
+  summarize(
+    `Total Polls` = n(),
+    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
+    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
+    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),
+
+    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
+    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
+    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
+  )
+
+# Reshape the summary statistics to a vertical format
+summary_statistics_long <- pivot_longer(
+  summary_statistics,
+  cols = everything(),
+  names_to = "Statistic",
+  values_to = "Value"
+)
+
+# Display vertical summary statistics in a kable table
+kable(summary_statistics_long) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
 
@@
 
-To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
 ```{r}
-#| label: tbl-outcomevar
-#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
-This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
-All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
 #| echo: false
@@
 #| message: false
+#| label: tbl-stateaverage
+#| fig-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
+The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day.
+Notably, Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in states like Michigan and Nevada.
+This data provides insight into the current polling dynamics and potential electoral outcomes in these critical states."
 
-# Load cleaned data
-clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
-
-# Summary statistics for the relevant columns in clean_poll_data
-summary_statistics <- clean_poll_data %>%
-  summarize(
-    `Total Polls` = n(),
-    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
-    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
-    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),
-
-    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
-    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
-    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
+# Aggregating polling percentages for Trump and Harris by state
+state_average <- clean_poll_data %>%
+   group_by(state) %>%
+   summarise(
+     Trump_pct = mean(Trump_pct, na.rm = TRUE),
+     Harris_pct = mean(Harris_pct, na.rm = TRUE)
    )
 
-# Reshape the summary statistics to a vertical format
-summary_statistics_long <- pivot_longer(
-  summary_statistics,
-  cols = everything(),
-  names_to = "Statistic",
-  values_to = "Value"
+ # Prepare table for display with rounding to 1 decimal place
+ average_table <- state_average |>
+   mutate(
+     Trump_pct = round(Trump_pct, 1),
+     Harris_pct = round(Harris_pct, 1)
+   ) |>
+   select(
+     State = state, 
+     "Trump %" = Trump_pct, 
+     "Harris %" = Harris_pct, 
    )
 
-# Display vertical summary statistics in a kable table
-kable(summary_statistics_long) %>%
-  kable_styling(full_width = FALSE, position = "center")
-
+ # Render the table
+ knitr::kable(average_table)
 ```
-
-MISSING - explain table
 

### 2024-11-03 paper/paper.qmd
committing new file onto git for paper.qmd

@@
 library(brms)
+
+# Load cleaned data
+clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
 ```
@@
 The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
-
-To give an overview of the data, we provide summary statistics of the key variables, highlighting the characteristics of the polling support for each candidate across the battleground states.
-```{r}
-#| label: tbl-outcomevar
-#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
-This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
-All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
-#| echo: false
-#| warning: false
-#| message: false
-
-# Load cleaned data
-clean_poll_data <- read_parquet("/Users/jamielee/2024_US_Elections/data/02-analysis_data/analysis_data.parquet")
-
-# Summary statistics for the relevant columns in clean_poll_data
-summary_statistics <- clean_poll_data %>%
-  summarize(
-    `Total Polls` = n(),
-    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
-    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
-    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),
-
-    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
-    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
-    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
-  )
-
-# Reshape the summary statistics to a vertical format
-summary_statistics_long <- pivot_longer(
-  summary_statistics,
-  cols = everything(),
-  names_to = "Statistic",
-  values_to = "Value"
-)
-
-# Display vertical summary statistics in a kable table
-kable(summary_statistics_long) %>%
-  kable_styling(full_width = FALSE, position = "center")
-
-```
 
@@
 ### Outcome Variable
+Outcome variable is 'pct'
 
+To give an overview of the outcome variable, divided per candidate, we provide summary statistics of this key variable, highlighting the characteristics of the polling support for each candidate across the battleground states.
+```{r}
+#| label: tbl-outcomevar
+#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
+This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
+All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
+#| echo: false
+#| warning: false
+#| message: false
+
+# Summary statistics for the relevant columns in clean_poll_data
+summary_statistics <- clean_poll_data %>%
+  summarize(
+    `Total Polls` = n(),
+    `Average Trump Support` = round(mean(Trump_pct, na.rm = TRUE), 3),
+    `Max Trump Support` = round(max(Trump_pct, na.rm = TRUE), 3),
+    `Min Trump Support` = round(min(Trump_pct, na.rm = TRUE), 3),
+
+    `Average Harris Support` = round(mean(Harris_pct, na.rm = TRUE), 3),
+    `Max Harris Support` = round(max(Harris_pct, na.rm = TRUE), 3),
+    `Min Harris Support` = round(min(Harris_pct, na.rm = TRUE), 3)
+  )
+
+# Reshape the summary statistics to a vertical format
+summary_statistics_long <- pivot_longer(
+  summary_statistics,
+  cols = everything(),
+  names_to = "Statistic",
+  values_to = "Value"
+)
+
+# Display vertical summary statistics in a kable table
+kable(summary_statistics_long) %>%
+  kable_styling(full_width = FALSE, position = "center")
+
+```
+
+### Predictive Variables
+The predictive variables is 'state' 
 ```{r}
@@
 #| message: false
-#| label: tbl-stateaverage
-#| fig-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
-The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day.
-Notably, Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in states like Michigan and Nevada.
-This data provides insight into the current polling dynamics and potential electoral outcomes in these critical states."
+#| label: fig-state-distribution
+#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data.
+The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity.
+This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."
 
-# Aggregating polling percentages for Trump and Harris by state
-state_average <- clean_poll_data %>%
-   group_by(state) %>%
-   summarise(
-     Trump_pct = mean(Trump_pct, na.rm = TRUE),
-     Harris_pct = mean(Harris_pct, na.rm = TRUE)
+# Visualize the distribution of the state variable
+ggplot(clean_poll_data, aes(x = state)) +
+  geom_bar(fill = "steelblue") +  # Bar plot with specified fill color
+  labs(
+    title = "Distribution of State Variable in Polling Data",
+    x = "State",
+    y = "Count of Observations"
+  ) +
+  theme_minimal() +  # Use a minimal theme
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
+    plot.title = element_text(hjust = 0.5)  # Center the plot title
+  )
+```
+
+Another predictive variable is 'end-date'
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-end-date-distribution
+#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
+The histogram shows the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches.
+This visualization is crucial for understanding how polling data correlates with the election timeline."
+
+# Visualize the distribution of end_date
+ggplot(clean_poll_data, aes(x = end_date)) +
+  geom_histogram(binwidth = 5, fill = "steelblue", color = "white") +  # Histogram with specified binwidth
+  labs(
+    title = "Distribution of Polling End Dates",
+    x = "End Date",
+    y = "Count of Polls"
+  ) +
+  scale_x_date(date_labels = "%b %d, %Y", date_breaks = "1 week") +  # Format date labels and breaks
+  theme_minimal() +  # Use a minimal theme
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),  # Rotate x-axis labels for better readability
+    plot.title = element_text(hjust = 0.5)  # Center the plot title
   )
 
- # Prepare table for display with rounding to 1 decimal place
- average_table <- state_average |>
-   mutate(
-     Trump_pct = round(Trump_pct, 1),
-     Harris_pct = round(Harris_pct, 1)
-   ) |>
-   select(
-     State = state, 
-     "Trump %" = Trump_pct, 
-     "Harris %" = Harris_pct, 
-   )
-
- # Render the table
- knitr::kable(average_table)
 ```
 
-### Predictive Var
-The predictive variable in our model are 'time' and 'state'
+###Other Variables
 
-## Other Variables
+'numeric grade'
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-numgrade-distribution
+#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
+Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches.
+This visualization is crucial for understanding how polling data correlates with the election timeline."
 
-'numeric grade', 'transparency score', 'pollster'
+# Density plot for numeric grade
+ggplot(clean_poll_data, aes(x = numeric_grade)) +
+  geom_density(fill = "lightgreen", alpha = 0.5) +
+  labs(
+       x = "Numeric Grade",
+       y = "Density") +
+  theme_minimal()
+
+```
+
+'transparency score'
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-transparency-score-distribution
+#| fig-cap: "Histogram illustrating the distribution of transparency scores for polls in key battleground states.
+Reveals the frequency of different transparency ratings, providing insight into the reliability of polling data.
+Understanding the distribution of transparency scores is crucial for evaluating the credibility of the polling results as Election Day approaches."
+
+# Distribution of transparency score
+ggplot(clean_poll_data, aes(x = transparency_score)) +
+  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
+  labs(
+       x = "Transparency Score",
+       y = "Count") +
+  theme_minimal()
+
+```
+
+'pollster'
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-pollster-distribution
+#| fig-cap: "Horizontal bar graph illustrating the distribution of polls by various pollsters in key battleground states.
+Displays the count of polls conducted by each pollster, highlighting the most frequently utilized polling organizations.
+Understanding the diversity of pollsters is essential for assessing the reliability and variance in polling data as Election Day approaches."
+
+# Horizontal bar graph for the distribution of pollsters
+ggplot(clean_poll_data, aes(x = factor(pollster))) +  # Convert to factor to ensure proper categorization
+  geom_bar(fill = "purple") +
+  labs(
+       x = "Count of Polls",
+       y = "Pollster") +  # Notice the labels are switched for horizontal orientation
+  theme_minimal() +
+  coord_flip() +  # Flip the coordinates for a horizontal bar graph
+  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Adjust y-axis text if necessary
+```
+
+'start_date'
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+#| label: fig-start-date-distribution
+#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
+This visualization displays the frequency of polls conducted over time, providing insights into when polling activities were initiated and potential trends in voter sentiment as Election Day approaches.
+Understanding the timing of polls is crucial for evaluating the relevance of the polling data."
+
+# Distribution of start_date with x-axis labels every 15 days
+ggplot(clean_poll_data, aes(x = start_date)) +
+  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
+  labs(
+       x = "Start Date",
+       y = "Count") +
+  scale_x_date(date_breaks = "15 days", date_labels = "%b %d") +  # Adjust x-axis labels
+  theme_minimal()
+```
 
@@
 # Results {#sec-results}
+
+## State-Level Polling Averages for Trump and Harris
+WRITE - explain why we are showing this average table first and what the table reveals
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: tbl-stateaverage
+#| fig-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
+The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day.
+Notably, Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in states like Michigan and Nevada.
+This data provides insight into the current polling dynamics and potential electoral outcomes in these critical states."
+
+# Aggregating polling percentages for Trump and Harris by state
+state_average <- clean_poll_data %>%
+   group_by(state) %>%
+   summarise(
+     Trump_pct = mean(Trump_pct, na.rm = TRUE),
+     Harris_pct = mean(Harris_pct, na.rm = TRUE)
+   )
+
+ # Prepare table for display with rounding to 1 decimal place
+ average_table <- state_average |>
+   mutate(
+     Trump_pct = round(Trump_pct, 1),
+     Harris_pct = round(Harris_pct, 1)
+   ) |>
+   select(
+     State = state, 
+     "Trump %" = Trump_pct, 
+     "Harris %" = Harris_pct, 
+   )
+
+ # Render the table
+ knitr::kable(average_table)
+```
 
@@
 # Additional Data Details
+
 Given the nature of the presidential election, the electoral college is more impactful and influential on the possible outcome compared to the nationwide popular votes.

### 2024-11-03 scripts/03-clean_data.R

@@ -44,12 +44,18 @@ clean_poll_data <- poll_data |>
 clean_poll_data <- clean_poll_data |>
   mutate(state = str_replace(state, "Nebraska CD-2", "Nebraska"))
 
-# Aggregate pct values by averaging for unique combinations before pivoting
+# Aggregate pct values by averaging for unique combinations, while keeping additional columns
 clean_poll_data <- clean_poll_data |>
-  group_by(state, end_date, end_date_num, candidate_name) |>
-  summarize(pct = mean(pct, na.rm = TRUE), .groups = "drop") |>
-  pivot_wider(names_from = candidate_name, values_from = pct, values_fill = 0) |>
-  rename(Trump_pct = `Donald Trump`, Harris_pct = `Kamala Harris`) |>
+  group_by(state, end_date, end_date_num) |>
+  summarize(
+    Trump_pct = mean(pct[candidate_name == "Donald Trump"], na.rm = TRUE),
+    Harris_pct = mean(pct[candidate_name == "Kamala Harris"], na.rm = TRUE),
+    start_date = first(start_date),
+    pollster = first(pollster),
+    numeric_grade = first(numeric_grade),
+    transparency_score = first(transparency_score),
+    .groups = "drop"
+  ) |>
   filter(Trump_pct > 0 & Harris_pct > 0)  # Retain rows with both candidates’ support data
 
 #### Save Data ####

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

@@
 ## Overview
-NEEDS EDIT & ELABORATE - This paper utilizes the dataset of the 2024 Presidential Election cycle, obtained from [@FiveThirtyEight].
+The dataset of the 2024 Presidential Election cycle is obtained from Project 538 (cite).
 The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
 It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
-For every pollster, relevant variables such as state, polling date, transparency score, candidate name, and percentage of support for the appropriate candidate are included.
+For every poll conducted, relevant variables such as state, start_date, end_date, transparency_score, candidate_name, and percentage of support (pct) for the appropriate candidates are included.
 The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
+
+The paper focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
+For all the following variables mentioned, we collected observations for both of the candidates.
+We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris.
+We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’, and ‘state’, as well as the electorate support for each candidate is denoted as ‘pct’.
+Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in {additional data details}.
 
@@
 #| label: tbl-outcomevar
-#| fig-cap: "Summary Statistics for Trump and Harris Polling Data.
-This table presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
+#| fig-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
 All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
@@
 Where:
-- \(P(\text{Victory}_i)\) is the predicted probability of winning for candidate \(i\).
-- \(\beta_0\) represents the intercept, reflecting the baseline log-odds of winning.
-- \(\beta_1\) is the coefficient for the natural spline transformation of the numeric end date (\( \text{end\_date\_num}_i \)), capturing nonlinear trends in voter support over time.
-- \(u_{\text{state}}\) is the random effect associated with each state, accounting for variations in voter preferences.
+
+- \( P(\text{Victory}_i) \) is the predicted probability of winning for candidate \( i \).
+- \( \beta_0 \) represents the intercept, reflecting the baseline log-odds of winning.
+- \( \beta_1 \) is the coefficient for the natural spline transformation of the numeric end date \( (\text{end\_date\_num}_i) \), capturing nonlinear trends in voter support over time.
+- \( u_{\text{state}} \) is the random effect associated with each state, accounting for variations in voter preferences.
 
@@
 
-We implemented the models using the `brms` package in R, selecting a Beta distribution with a logit link function to accurately model the proportions of support for each candidate:
+We implemented the models using the **`brms`** package in R, selecting a **Beta distribution** with a **logit link function** to accurately model the proportions of support for each candidate:
 
@@
 
-Here, \(y_i\) represents the proportion of support for candidate \(i\) in a given state.
+In this model:
+
+- \(y_i\) represents the proportion of support for candidate \(i\) in a given state.
+- \(\mu_i\) is the mean of the Beta distribution, reflecting the expected support proportion.
+- \(\phi\) is the precision parameter, indicating the variability around the mean.
+
+This approach allows us to effectively capture the dynamics of voter support in various states as we approach the election.
 

### 2024-11-03 paper/paper.qmd
committing changes in model section to paper.qmd

@@
 
-The paper focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
+The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
 For all the following variables mentioned, we collected observations for both of the candidates.
@@
 We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’, and ‘state’, as well as the electorate support for each candidate is denoted as ‘pct’.
-Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in {additional data details}.
+Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [@sec-data-details]
 
@@
 
-- \( P(\text{Victory}_i) \) is the predicted probability of winning for candidate \( i \).
-- \( \beta_0 \) represents the intercept, reflecting the baseline log-odds of winning.
-- \( \beta_1 \) is the coefficient for the natural spline transformation of the numeric end date \( (\text{end\_date\_num}_i) \), capturing nonlinear trends in voter support over time.
-- \( u_{\text{state}} \) is the random effect associated with each state, accounting for variations in voter preferences.
+- $P(\text{Victory}_i)$ is the predicted probability of winning for candidate $i$.
+- $\beta_0$ represents the intercept, reflecting the baseline log-odds of winning.
+- $\beta_1$ is the coefficient for the natural spline transformation of the numeric end date $(\text{end\_date\_num}_i)$, capturing nonlinear trends in voter support over time.
+- $u_{\text{state}}$ is the random effect associated with each state, accounting for variations in voter preferences.
 
@@
 
-- \(y_i\) represents the proportion of support for candidate \(i\) in a given state.
-- \(\mu_i\) is the mean of the Beta distribution, reflecting the expected support proportion.
-- \(\phi\) is the precision parameter, indicating the variability around the mean.
+- $y_i$ represents the proportion of support for candidate $i$ in a given state.
+- $\mu_i$ is the mean of the Beta distribution, reflecting the expected support proportion.
+- $\phi$ is the precision parameter, indicating the variability around the mean.
 

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd

@@
 ### Outcome Variable
-Outcome variable is 'pct'
+As our focus is to analyze and predict voter support for Kamala Harris and Donald Trump during the current presidential election cycle, the primary outcome variables of interest are the percentages of support for each candidate, denoted as ‘pct’.The pct is categorized by candidate and scaled to a proportion by diving by 100.
+This variable represents the proportion of voter support for each candidate based on polling data.
+Through understanding the percentage of support (pct) for each candidate, we can observe and infer the candidate that has a leading advantage in winning the presidential election race.
 
@@
 ```
+For instance, @fig-summarystatistics reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
+‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris.
+The 107 polls conducted display an average support level of 47.801% for Trump and 47.999% for Harris.
+This suggests that based on the dataset, Trump has a leading advantage in winning the presidential election race by having a slighty higher average level of support than Harris.
+The highest percentage of support for Trump in the 107 polls was 51.700% while the lowest recorded support was 43%.
+For Harris, the highest percentage of support also reached 51.200%, with a minimum of 43.600%.
+These statistics illustrate the competitive landscape of the presidential election race, without a clear leading candidate in the current race.
+
+It is important to note that since ‘pct’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
+For instance, in the summary statistics presented in Table 1, the average support for Donald Trump is 47.801%, while the average support for Kamala Harris is 47.999%.
+This results in a combined average support of 95.800%, indicating that there is a significant portion of respondents who are either undecided or supporting other candidates.
+This highlights the competitive nature of the election and emphasizes the necessity of examining not only the percentages attributed to the main candidates but also the context of voter preferences that could influence the final outcome.
 
@@
 
-\[
+$$
 y_i \sim \text{Beta}(\mu_i, \phi)
-\]
+$$
 
@@
 
-- For the intercept \(\beta_0\), we specified \( \beta_0 \sim \text{Normal}(0, 10)\), providing a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds.
-- For the slope \(\beta_1\), we used \( \beta_1 \sim \text{Normal}(0, 5)\), reflecting reasonable expectations regarding the influence of polling trends over time.
+- For the intercept $\beta_0$, we specified $\beta_0 \sim \text{Normal}(0, 10)$, providing a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds.
+- For the slope $\beta_1$, we used $\beta_1 \sim \text{Normal}(0, 5)$, reflecting reasonable expectations regarding the influence of polling trends over time.
 
@@
 
-The model summary for Kamala Harris indicates that the intercept estimate is \(-0.1508\) with a standard error of \(0.0279\).
+The model summary for Kamala Harris indicates that the intercept estimate is $-0.1508$ with a standard error of $0.0279$.
 This intercept value suggests a negative baseline log-odds for her victory, reflecting the starting point of her support in the absence of any other predictors.
-The model shows that the coefficients for the natural spline terms (\(nsenddatenum\)) are particularly informative, with the first spline term (\(nsendatenum\_df41\)) yielding a positive estimate of \(0.1225\), suggesting that as time progresses, Harris’s probability of winning increases, particularly during key moments captured by the polling data.
-This is also reflected in the confidence intervals, with the upper limit of the confidence interval for this term reaching \(0.1903\).
-In contrast, the estimates for the second spline term (\(nsenddatenum\_df42\)) indicate a smaller effect, suggesting fluctuations in her support that may w
+The model shows that the coefficients for the natural spline terms ($nsendate_{num}$) are particularly informative, with the first spline term ($nsendate_{num\_df41}$) yielding a positive estimate of $0.1225$, suggesting that as time progresses, Harris’s probability of winning increases, particularly during key moments captured by the polling data.
+This is also reflected in the confidence intervals, with the upper limit of the confidence interval for this term reaching $0.1903$.
+In contrast, the estimates for the second spline term ($nsendate_{num\_df42}$) indicate a smaller effect, suggesting fluctuations in her support that may warrant further investigation.
 
@@
 In comparison, the model summary for Donald Trump presents a different picture.
-The intercept for Trump's model is \(-0.0923\) with a standard error of \(0.0306\), indicating a similarly negative baseline log-odds of victory as Harris.
+The intercept for Trump's model is $-0.0923$ with a standard error of $0.0306$, indicating a similarly negative baseline log-odds of victory as Harris.
 However, Trump's model demonstrates more pronounced increases in support as reflected by the estimates for the natural spline terms.
-For example, the second spline term (\(nsenddatenum\_df42\)) has a positive coefficient of \(0.0803\), highlighting a significant upward trend in support over time.
-This trend is further supported by the confidence interval, which ranges from \(0.0179\) to \(0.1426\), indicating a substantial increase in his predicted probabilities as the election approaches.
-The fourth spline term (\(nsenddatenum\_df44\)) also shows a positive coefficient (\(0.0789\)), suggesting that the support for Trump remains resilient over the polling period.
+For example, the second spline term ($nsendate_{num\_df42}$) has a positive coefficient of $0.0803$, highlighting a significant upward trend in support over time.
+This trend is further supported by the confidence interval, which ranges from $0.0179$ to $0.1426$, indicating a substantial increase in his predicted probabilities as the election approaches.
+The fourth spline term ($nsendate_{num\_df44}$) also shows a positive coefficient of $0.0789$, suggesting that the support for Trump remains resilient over the polling period.
 Overall, both models reveal complex interactions between time and voter sentiment, with implications for the electoral dynamics leading up to the election.
@@
 # Results {#sec-results}
-
 ## State-Level Polling Averages for Trump and Harris

### 2024-11-03 paper/paper.qmd
committing text changes to paper.qmd

@@
 
-# Introduction
+# Introduction {#sec-intro}
 The 2024 United States Presidential Election is shaping up to be a critical event, particularly as battleground states emerge as pivotal areas influencing the electoral outcome.
 As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions becomes essential for predicting the election results.
-This paper utilizes state-level polling data to build a predictive model that estimates the likelihood of victory for each candidate in several key battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+This paper utilizes state-level polling data from Project 538 [@FiveThirtyEight] and the statistical programming language R [@citeR] to build a predictive model that estimates the likelihood of victory for each candidate in several key battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
 
@@
 The remainder of this paper is structured as follows.
-In Section 2, we describe the data and variables used in the analysis.
-Section 3 outlines the model setup and estimation strategy.
-Section 4 presents the results, including the predicted election outcomes and visualizations.
-Finally, Section 5 discusses the implications and limitations of our findings
+In @sec-data, we describe the data and variables used in the analysis.
+@sec-model outlines the model setup and estimation strategy.
+@sec-results presents the results, including the predicted election outcomes and visualizations.
+Finally, @sec-discussion discusses the implications and limitations of our findings.
 
@@
 ## Overview
-The dataset of the 2024 Presidential Election cycle is obtained from Project 538 (cite).
+The dataset of the 2024 Presidential Election cycle is obtained from Project 538 [@FiveThirtyEight].
 The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
@@
 We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’, and ‘state’, as well as the electorate support for each candidate is denoted as ‘pct’.
-Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [@sec-data-details]
+Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [Appendix -@sec-data-details].
 
@@
 
-To ensure the quality of the analysis, only those polls that meet specific thresholds were included - refer to [@sec-appendix].
+To ensure the quality of the analysis, only those polls that meet specific thresholds were included - for more details, refer to [Appendix -@sec-data-details].
 
@@
 ```
-For instance, @fig-summarystatistics reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
+For instance, [@fig-summarystatistics] reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
 ‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris.
@@
 
-###Other Variables
+### Other Variables
 
@@
 
-## Model Summary - Harris Model
+## Model Summary
+### Harris Model Summary
 ```{r}
@@
 
-## Model summary - Trump Model
+### Trump Model Summary
 ```{r}
@@
 ## Polling Averages for Trump and Harris Over Time in Battleground States
-@fit-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024.
+[@fit-state-polling-average] provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024.
 Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support.
@@
 By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024).
-@fig-state-probabilities provides a visual summary of these predicted probabilities.
+[@fig-state-probabilities] provides a visual summary of these predicted probabilities.
 ```{r}
@@
 ## Probability of Winning by State on Election Day 
-@tbl-fig-probabilities displays the numeric percentage of the probabilities of winning for Trump and Harris.
+[@tbl-fig-probabilities] displays the numeric percentage of the probabilities of winning for Trump and Harris.
 By comparing the predicted probabilities, the predicted winner for each state is assigned accordingly.
@@
 ## US Map Showing Predicted Winner by State on Election Day
-To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.
+To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from [@tbl-state-probabilities].
 
@@
 ```
-According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
+According to [@fig-us-map], Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
 Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.
@@
 ## Model Performance {#sec-model-performance}
-
-Out-of-sample testing results [Appendix @sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
+Out-of-sample testing results [Appendix -@sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
 presidential election.
@@
 The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
-Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix @sec-model-diagnostics].
+Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].
 
 ## Weaknesses and Next Steps
-
 While our model provides valuable insights into the electoral landscape, it faces limitations due to missing data for certain states, which results in some states being left blank in our predictions.
@@
 
-# Additional Data Details
+# Additional Data Details {#sec-data-details}
 

### 2024-11-03 paper/paper.qmd
committing code changes to paper.qmd

@@
 ```
-For instance, [@fig-summarystatistics] reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
+For instance, @tbl-outcomevar reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
 ‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris.
@@
 ## Polling Averages for Trump and Harris Over Time in Battleground States
-[@fit-state-polling-average] provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024.
+@fig-state-polling-average provides a visual comparison of the current polling support for Donald Trump and Kamala Harris across key battleground states, tracked over time from August 2024 to October 2024.
 Each line plot represents the average polling percentages for both candidates in a specific state, with red lines denoting Trump’s support and blue lines representing Harris’s support.
@@
 
-Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in [@sec-model] to predict the possible outcomes by incorporating the given data.
+Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in @sec-model to predict the possible outcomes by incorporating the given data.
 By using the current data, the model can identify the state-specific trends thus generating more accurate predictions of the outcome.
@@
 By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024).
-[@fig-state-probabilities] provides a visual summary of these predicted probabilities.
+@fig-state-probabilities provides a visual summary of these predicted probabilities.
 ```{r}
@@
 ## Probability of Winning by State on Election Day 
-[@tbl-fig-probabilities] displays the numeric percentage of the probabilities of winning for Trump and Harris.
+@tbl-state-probabilities displays the numeric percentage of the probabilities of winning for Trump and Harris.
 By comparing the predicted probabilities, the predicted winner for each state is assigned accordingly.
@@
 ## US Map Showing Predicted Winner by State on Election Day
-To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from [@tbl-state-probabilities].
+To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.
 
@@
 ```
-According to [@fig-us-map], Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
+According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
 Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd

@@
 ### Predictive Variables
-The predictive variables is 'state' 
+#### State
+The state variable indicates the U.S.
+state where the poll was conducted or targeted.
+In the context of the U.S.
+presidential election, the outcome is frequently determined by several key states.
+In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
+To enhance our model and refine our predictions, we focused exclusively on these seven battleground states.
+The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
 ```{r}
@@
 ```
+The distribution of the state variable is visualized in @fig-state-distribution and indicates that some states, such as Pennsylvania, North Carolina and Wisconsin have more polling observations compared to others, like Nevada.
+This variability reflects differences in polling frequency and interest among these battleground states.
+States with more polling observations provide a larger sample size, potentially leading to more stable and reliable predictions for those states in the model.
 
+#### End Date
 Another predictive variable is 'end-date'

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd

@@
 ```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 #| label: tbl-outcomevar
@@
 All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
-#| echo: false
-#| warning: false
-#| message: false
 
@@
 #| echo: false
+#| eval: true
 #| warning: false
@@
   labs(
-    title = "Distribution of State Variable in Polling Data",
     x = "State",
@@
   labs(
-    title = "Distribution of Polling End Dates",
     x = "End Date",
@@
 
-'numeric grade'
+#### Numeric Grade
 ```{r}
@@
 
-'transparency score'
+#### Transparency Score
 ```{r}
@@
 
-'pollster'
+#### Pollster
 ```{r}
@@
        x = "Count of Polls",
-       y = "Pollster") +  # Notice the labels are switched for horizontal orientation
+       y = "Pollster") +  
   theme_minimal() +
-  coord_flip() +  # Flip the coordinates for a horizontal bar graph
-  theme(axis.text.y = element_text(angle = 0, hjust = 1))  # Adjust y-axis text if necessary
+  coord_flip() +  
+  theme(axis.text.y = element_text(angle = 0, hjust = 1))  
 ```
 
-'start_date'
+#### Start Date
 ```{r}

### 2024-11-03 paper/paper.qmd
adding table of contents to paper

@@
 format: pdf
+toc: true
 number-sections: true

### 2024-11-03 paper/paper.qmd
commit changes to paper.qmd

@@
 
+# Workplace SetUp
 library(tidyverse)
@@
 #| label: tbl-outcomevar
-#| fig-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
+#| tbl-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
 All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
@@
 #| label: tbl-stateaverage
-#| fig-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
+#| tbl-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
 The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day.
@@
 #| label: tbl-state-probabilities
-#| fig-cap: "Probability of Trump Winning by State on Election Day."
+#| tbl-cap: "Probability of Trump Winning by State on Election Day."
 #| warning: false

### 2024-11-03 paper/paper.qmd
committing changes to paper.qmd in variables

@@
 #### State
-The state variable indicates the U.S.
+One of the paper's predictive variables is the 'state' variable indicates the U.S.
 state where the poll was conducted or targeted.
@@
 #### End Date
-Another predictive variable is 'end-date'
+The variable 'end_date' represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
+In the context of our paper, the end date of each poll serves as one of the predictive variables and is used to assess changes in voter support over time.
+Including end date as a variable in our model can help us identify patterns in voter support as Election Day approaches, potentially reflecting the impact of campaign events, debates, or other political developments.
+@fig-end-date-distribution displays the distribution of polling end dates from July 2024 to October 2024.
 ```{r}
@@
 ```
+We filtered for the earliest end date as July 2024 since Harris officially announced her presidential campaign on July 21, 2024.
+Therefore, any polls that has an end date before July would not contain polling results for Harris.
+The data from [@FiveThirtyEight] is last updated on October 15, 2024, thus the latest end date being October 21, 2024.
+Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time.
+There are clear peaks where the number of polls conducted is higher, notably in early September, late Semptember, as well as early October.
+This concentration of polls likely corresponds to period of heightened political interest or significant campaign events that prompt increased polling.
 
 ### Other Variables
-
 #### Numeric Grade
@@
 
-# References {sec-references}
+# References {#sec-references}

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd in model set-up

@@
 ### Predictive Variables
+In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over ti
 #### State
-One of the paper's predictive variables is the 'state' variable indicates the U.S.
+The state ('state') variable indicates the U.S.
 state where the poll was conducted or targeted.
 In the context of the U.S.
-presidential election, the outcome is frequently determined by several key states.
+presidential election, the outcome is frequently determined by several key battlegroud states.
 In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
@@
 
-#### End Date
-The variable 'end_date' represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
+#### End Date {#sec-end-date}
+End date ('end_date') represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
 In the context of our paper, the end date of each poll serves as one of the predictive variables and is used to assess changes in voter support over time.
@@
 Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time.
-There are clear peaks where the number of polls conducted is higher, notably in early September, late Semptember, as well as early October.
+There are clear peaks where the number of polls conducted is higher, notably in early September, late September, as well as early October.
 This concentration of polls likely corresponds to period of heightened political interest or significant campaign events that prompt increased polling.
@@
 ### Other Variables
+This section details additional variables that provide essential context for our analysis, including numeric grade, which evaluates the quality of the polls; transparency score, indicating the openness of poll methodologies; pollster, identifying the organizations conducting the polls; and start date, which denotes the initiation of the polling period for each survey.
 #### Numeric Grade
+Numeric Grade ('numeric_grade') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
+In the context of our paper, numeric grade allows us to weigh the reliability of different polls.
+By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support.
+@fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0.
+Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions.
+The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
 ```{r}
@@
 ```
+This variability in numeric grade highlights the diversity in polling quality within our dataset.
+The presence of peaks suggests that some pollsters consistently receive higher ratings, likely due to their use of more rigorous methods or transparent reporting standards.
+Conversely, the valleys in the distribution may indicate fewer polls from sources with lower numeric grades, reflecting a potential tendency for lower-quality polls to be less prevalent or less frequently conducted.
+By accounting for numeric grade, we can improve the accuracy of predictions by emphasizing data from more reliable sources.
+This consideration is crucial, as polling quality can impact the stability of the model’s forecasts, especially in closely contested states where polling accuracy is essential for reliable predictions.
 
 #### Transparency Score
+Transparency Score ('transparency_score') represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
+This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error.
+In the context of our paper, transparency score is an essential variable for evaluating the reliability of polling data, as higher transparency scores generally indicate a greater level of methodological rigor and openness.
+@fig-transparency-score-distribution displays the distribution of transparency scores across the polling data for key battleground states.
+The plot shows that transparency scores cluster heavily around ratings of  7 and 9, suggesting that a majority of polls in our dataset come from sources that maintain a relatively high level of transparency.
+There are smaller peaks at lower scores (5 and 10), indicating that while some pollsters offer limited transparency, a few polls also achieve perfect transparency scores.
 ```{r}
@@
 ```
+This distribution is important for understanding the credibility of our data sources.
+Polls with higher transparency scores are more likely to provide consistent and reliable results, as they openly disclose their methodologies, allowing us to assess potential biases or limitations in their data.
+In contrast, polls with lower transparency scores may lack sufficient methodological detail, making it challenging to gauge their accuracy.
+Incorporating transparency scores into our model helps differentiate between high- and low-transparency polls.
+This adjustment is crucial as transparency can influence the stability of the model’s predictions, especially in closely contested states where the reliability of polling data may impact the accuracy of our forecasted election outcomes.
 
 #### Pollster
+Pollster ('pollster') represents the various organizations responsible for conducting polls across key battleground states in the lead-up to the 2024 United States Presidential Election.
+In the context of our paper, the diversity and frequency of polling by different pollsters are essential for understanding the reliability and variance in the data.
+Each pollster may use different methodologies, sample populations, and data collection techniques, which can affect the results and introduce variability in polling outcomes.
 ```{r}
@@
 ```
+@fig-pollster-distribution displays the count of polls conducted by each pollster, illustrating the distribution of polling efforts across multiple organizations.
+Notably, Emerson is the most frequent pollster, conducting the highest number of polls, followed by Quinnipiac and Siena/NYT.
+Other pollsters, such as YouGov, CNN/SSRS, and Marist, also contribute a significant number of polls, whereas some organizations like The Washington Post and SurveyUSA have comparatively fewer entries in the dataset.This distribution is crucial for evaluating the consistency and breadth of the polling data.
+Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others.
+On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.
 
 #### Start Date
+Start date ('start_date') represents the date when each poll began collecting responses.
+In the context of our paper, start date is a key variable as it allows us to focus on polling data collected after July 21, 2024—the date Kamala Harris officially announced her presidential campaign.
+By filtering the dataset to include only polls with a start date after this announcement, we ensure that each poll reflects voter sentiment with Harris as an active candidate, enabling a more accurate comparison between her and Donald Trump.
+@fig-start-date-distribution displays the distribution of polling start dates from late July to mid-October 2024, showing when polling efforts began across key battleground states.
+The histogram reveals peaks in polling activity, particularly in mid-August, early September, and late September, indicating periods of heightened interest or campaign events that likely prompted increased polling.
 ```{r}
@@
 ```
+This distribution provides insights into the timing and intensity of polling efforts throughout the campaign season.
+By including only polls initiated after Harris’s campaign launch, we capture voter preferences in a context where both major candidates are actively campaigning, thus making the polling data more relevant for direct comparisons.
+This filtered timeline ensures our model focuses on data that reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as we approach Election Day.
 
 # Model {#sec-model}
+The Bayesian Beta regression models constructed in this paper aim to predict the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S.
+Presidential Election by analyzing polling data of the 7 battleground states.
+One Bayesian Beta regression model is constructed for each candidate, and each of them estimates the proportion of support for each candidate based on polling data, accounting for both polling trends over time, and state-level differences in voter sentiment.
+These models allow us to evaluate how support for each candidate fluctuates over time in key states, and predict potential electoral outcomes.
 
-In this analysis, we develop separate Bayesian models for candidates Donald Trump and Kamala Harris to predict their respective probabilities of winning in the 2024 U.S.
-Presidential Election.
-The use of distinct models allows us to capture the unique polling dynamics and voter preferences for each candidate across key battleground states, which include Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-By analyzing polling data collected from various sources, we aim to provide a comprehensive assessment of each candidate’s electoral prospects.
+One of the goals of our model is to analyze changes in voter support for each candidate over time, as Election Day approaches.
+By using end_date_num (days the earliest poll date) as a predictor with natural splines, the model captures non-linear trends in support.
+The time-based analysis provides us with information about how voter preference may shift due to campaign events, news cycles, and other influential factors in the months leading up to the election.
 
-The models are built using state-level polling data, which is aggregated to ensure robustness in the predictions.
-Each candidate's model incorporates temporal factors through natural splines applied to the numeric representation of the end date, allowing us to observe changes in voter support over time.
-This approach provides a nuanced understanding of how public sentiment evolves as the election date approaches.
+Another goal of our model is to predict the potential winner of the Presidential Election.
+We recognize that battleground states have unique political landscapes, so we include state as a random effect in each model.
+This allows each of the 7 battleground states to have a unique baseline level of support for each candidate, capturing regional differences in voter preferences.
+By incorporating state-level effects, the model can show variations in support across different states, revealing the potential winning candidate for the presidential election.
 Further background details, model specifications, and diagnostics are included in [Appendix -@sec-model-details].
 
+## Model Assumptions
+In order to ensure the validity and reliability of our Bayesian models for predicting voter support percentages for Donald Trump and Kamala Harris, we establish several key assumptions that guide the modeling process:
+
+- **Linearity in the Log-Odds**: The relationship between predictor variables (end_date and state) and the log-odds of the outcome variable (percentage of support) is linear.
+To account for non-linear trends over time, nature splines are used.
+- **Independence**: The outcome for each polling observation is assumed to be independent of others. Each poll represents a distinct sample of voter sentiment at a specific time and location. 
+- **Appropriate Distribution for Proportions**: Since the model predicts support percentages (bounded between 0 and 100), a Beta distribution with a logit link is applied, which is ideal for data that are restricted to this range.
+- **Random Effects**: To capture variation across states, state-level random effects are included, allowing each state to have its unique baseline level of support, acknowledging regional political differences
+- **Prior Distributions**: The model uses weakly informative priors for the intercept and coefficients, which helps in generating stable estimates and reducing the influence of outliers or unusual polling data.
+
 ## Model Setup
-The primary estimand in our models is the probability of victory for each candidate in the selected battleground states, expressed through a logistic regression framework.
-The model is formulated as follows:
+In this section, we outline the statistical models employed to predict the probabilities of victory for Donald Trump and Kamala Harris in the 2024 presidential election within key battleground states.
+Our analysis utilizes a Bayesian framework, which allows us to incorporate uncertainty and prior information into the modeling process, yielding more robust predictions.
+The primary estimand for both models is the probability of each candidate's victory, expressed through a logistic regression approach.
+By examining polling data and its temporal dynamics, we aim to capture the evolving nature of voter support leading up to the election.
+We present separate subsections for the models of Donald Trump and Kamala Harris, detailing the structure, variables, and assumptions underlying each model.
+This division allows for a focused discussion on how each model has been tailored to reflect the unique electoral contexts and polling dynamics for each candidate.
+
+### Trump Model
+The primary estimand for Donald Trump's model is the probability of victory in the selected battleground states, derived from a logistic regression framework. The model is structured as follows:
 
 $$
-P(\text{Victory}_i) = \frac{e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num}_i + u_{\text{state}}}}{1 + e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num}_i + u_{\text{state}}}}
+P(\text{Victory}_{Trump} = 1) = \frac{e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num} + u_{\text{state}}}}{1 + e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num} + u_{\text{state}}}}
 $$
@@
 
-- $P(\text{Victory}_i)$ is the predicted probability of winning for candidate $i$.
-- $\beta_0$ represents the intercept, reflecting the baseline log-odds of winning.
-- $\beta_1$ is the coefficient for the natural spline transformation of the numeric end date $(\text{end\_date\_num}_i)$, capturing nonlinear trends in voter support over time.
-- $u_{\text{state}}$ is the random effect associated with each state, accounting for variations in voter preferences.
+- $P(\text{Victory}_{Trump})$ is the predicted probability of Donald Trump winning.
+- $\beta_0$ represents the intercept, indicating the baseline log-odds of winning.
+- $\beta_1$ is the coefficient for the natural spline transformation of the numeric end date ($\text{end\_date\_num}$), capturing nonlinear trends in voter support over time.
+- $u_{\text{state}}$ accounts for random effects specific to each state, reflecting variations in voter preferences.
 
-Data preparation involved creating two distinct datasets from the cleaned polling data: one for Donald Trump and another for Kamala Harris.
-Each dataset included the state, end date, and corresponding polling percentage, with the end date converted into a numeric format representing the number of days since the earliest date in the dataset.
-This allows the model to effectively capture the evolving nature of voter sentiment as the election approaches.
+The dataset for the Trump model was prepared by extracting the relevant polling data, including the state, end date, and corresponding percentage of support for Trump ($Trump\_pct$).
+The end date was converted into a numeric format, denoting the number of days since the earliest polling date in the dataset.
+This transformation allows the model to effectively capture the evolving nature of voter sentiment as the election approaches.
 
-We implemented the models using the **`brms`** package in R, selecting a **Beta distribution** with a **logit link function** to accurately model the proportions of support for each candidate:
+The model was implemented using the **`brms`** package in R, applying a **Beta distribution** with a **logit link function** to accurately model the proportions of support:
 
@@
 
-In this model:
+Here, $y_i$ is the proportion of support for Trump in a given state, $\mu_i$ is the mean of the Beta distribution, and $\phi$ is the precision parameter reflecting variability.
 
-- $y_i$ represents the proportion of support for candidate $i$ in a given state.
-- $\mu_i$ is the mean of the Beta distribution, reflecting the expected support proportion.
-- $\phi$ is the precision parameter, indicating the variability around the mean.
+We specified priors for the model parameters as follows:
 
-This approach allows us to effectively capture the dynamics of voter support in various states as we approach the election.
+- For the intercept $\beta_0$, we set a prior of $\beta_0 \sim \text{Normal}(0, 10)$, allowing the data to inform the estimation of baseline log-odds without imposing strong assumptions.
+- For the slope $\beta_1$, we applied a prior of $\beta_1 \sim \text{Normal}(0, 5)$, reflecting reasonable expectations regarding the influence of polling trends.
+
+Natural splines were used to model the effect of the end date, accommodating observed non-linearity in voter support.
+Key assumptions include the independence of observations within states and the suitability of the Beta distribution for modeling proportions.
+However, limitations may arise from the accuracy of polling data and the assumption that historical voting patterns will predict future behavior.
+
+### Harris Model
+Similar to the Trump model, the estimand for Kamala Harris's model is the probability of her victory in the selected battleground states, modeled within a logistic regression framework.
+The formulation is analogous:
+
+$$
+P(\text{Victory}_{Harris} = 1) = \frac{e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num} + u_{\text{state}}}}{1 + e^{\beta_0 + \beta_1 \cdot \text{end\_date\_num} + u_{\text{state}}}}
+$$
+
+Where the components mirror those described for the Trump model, tailored to reflect the probability of victory for Harris.
+
+For this model, the dataset was constructed similarly, extracting polling data specifically for Harris ($Harris\_pct$).
+The same transformations were applied, converting end dates into a numeric format to facilitate the modeling of trends over time.
+
+The Harris model was also implemented using the **`brms`** package, employing a **Beta distribution** with a **logit link function** to represent the proportions of support accurately.
+The same mathematical representation and priors were applied as in the Trump model:
+
+$$
+y_i \sim \text{Beta}(\mu_i, \phi)
+$$
+
+With the same definitions for $y_i$, $\mu_i$, and $\phi$.
+
+Natural splines were employed to adapt to the non-linear trends in voter support over time. The assumptions and limitations discussed for the Trump model apply equally to the Harris model.
+
+By utilizing Bayesian beta regression models for both candidates, we leverage their flexibility and capacity to incorporate prior beliefs, which is particularly advantageous in the context of electoral forecasting where uncertainty plays a critical role.
 
@@
 ```
-
 In comparison, the model summary for Donald Trump presents a different picture.
@@
 # Results {#sec-results}
+In this section, we present the results of our analysis, focusing on the electoral support for Donald Trump and Kamala Harris across key battleground states during the 2024 Presidential Election cycle.
+We begin by examining the state-level polling averages for both candidates, which provide a snapshot of voter preferences as captured by various pollsters.
+Next, we explore how polling averages have evolved over time, highlighting trends that may influence election outcomes.
+We then assess the predicted probabilities of Trump winning on Election Day, followed by a detailed look at the likelihood of winning by state.
+Finally, we visualize the predicted winner by state on a US map, summarizing our findings and their implications for the electoral landscape.
+
 ## State-Level Polling Averages for Trump and Harris
-WRITE - explain why we are showing this average table first and what the table reveals
+@tbl-outcomevar displays overview of the percentage of support for Trump and Harris by each battleground state.
+It provides us with a high level understanding of the current voter dynamic.
+The pct results are aggregated averages of all pct data points for Trump and Harris in each state.
+By comparing the percentage of support for T.rump and Harris, we can determine the candidate that has a leading advantage in winning each state.
+For example,Trump has 49.2% support in Arizona while Harris has 46.8% support, which means Trump has won the state of Arizona.
 ```{r}
@@
 ```
+This distribution is crucial for evaluating the consistency and breadth of the polling data.
+Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others.
+On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.
 
@@
 
+### Model Performance - Trump Model
 The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes.
@@
 Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].
+
+### Model Performance - Harris Model
 
@@
 
-## Diagnostics {#sec-model-diagnostics}
-Out-of-sample testing was conducted to test our model’s reliability in predicting potential presidential election outcomes.
-We train our model on a training set of our data and then test its performance on the test set of our data, we evaluate how well the model generalizes to new data that it was not trained on.
+## Model Diagnostics {#sec-model-diagnostics}
+In order to assess the validity and reliability of our Bayesian models for predicting voter support for Kamala Harris and Donald Trump, we conducted posterior predictive checks for both models.
+Posterior predictive checks are a crucial diagnostic tool that allows us to evaluate how well our models can replicate observed data based on the parameters estimated during the modeling process.
+By comparing the distribution of the observed support percentages with the predicted support values, we can identify potential discrepancies and gauge the overall fit of our models.
 
-This process helps reveal whether the model is over-fitting to the training data or truly capturing the underlying patterns that apply more broadly.
-The test produces an accuracy score, RMSE, and a confusion matrix, each measuring different aspects of the model’s prediction.
-Accuracy is the proportion of correctly predicted outcomes compared to the total number of predictions made in the test set.
-The testing was implemented using the following packages in R [@citeR]: caret [@caret], rstanarm [@rstanarm], ggplot2 [@ggplot2], dplyr [@dplyr].
+These checks not only help us verify whether the model assumptions hold but also provide insights into how well the models capture the underlying data-generating processes.
+A strong alignment between observed and predicted values would suggest that our models are robust and capable of accurately reflecting the dynamics of voter support in the current electoral context.
+Conversely, significant deviations may indicate areas for model refinement or highlight limitations in our current approach.
+Through these diagnostics, we aim to ensure the credibility of our findings and enhance our understanding of the electoral landscape as we approach the 2024 Presidential Election.
 
-### Model Diagnostics Result
+## Posterior Predictive Check for the Trump Model
+The posterior predictive check for the Trump model is illustrated @fig-trump-check, showcasing the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+The plot displays a density estimate for both the actual observed data (in dark blue) and the predicted values generated by the model (in light blue).
+The primary aim of this diagnostic check is to assess how well the model captures the distribution of observed support for Donald Trump.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-trump-check
+#| fig-cap: "Posterior predictive check for Donald Trump's model, illustrating the distribution of observed support percentages $y$ alongside the predicted support values $y_{\text{rep}}$.
+The graph shows a strong overlap between the observed and predicted distributions, indicating that the model accurately captures the dynamics of voter support for Trump."
 
-[@fig-barplot-diagnostics] displays the performance of the Bayesian logistic regression model used to predict the outcome of the 2024 presidential election.
-The model's accuracy is approximately 75%, indicating that 75% of the test set's state-level outcomes were correctly predicted.
-The RMSE (Root Mean Squared Error) is about 0.5, showing the average error in predicting the winning candidate across states.
+# Generate posterior predictive checks for Trump model
+posterior_predict_trump <- posterior_predict(trump_model)
 
-This suggests the model performs reasonably well, but there is some error in predicting the exact outcome in certain states, which may be due to variations in polling data or other unaccounted factors.
+# Plot the posterior predictive check
+pp_check_trump <- bayesplot::pp_check(trump_model, nsamples = 100) +
+  xlab("Support (%)") +
+  ylab("Density") +
+  theme_minimal()
 
-### Confusion Matrix
-[@tbl-confusion-matrix] displays the model's predictive performance in distinguishing between state-level wins for Trump (coded as 1) and Harris (coded as 0).
+print(pp_check_trump)
+```
+From the visualization, we can observe that the predicted values closely align with the actual observed support percentages.
+The dark blue curve representing the observed data indicates a concentrated support range around 48% to 52%, suggesting that the model effectively captures the central tendency of voter support for Trump.
+The light blue curves depict a variety of predicted densities, indicating the model's ability to simulate a range of possible outcomes around the observed values.
+The alignment of the $y$ and $y_{\text{rep}}$ distributions indicates that the model is successfully reflecting the dynamics of voter support, thereby validating the choice of model and its structure.
+However, the presence of some variability in the predicted values suggests that while the model fits the observed data well, there may still be unexplained fluctuations in support levels that could warrant further investigation.
+These limitations are further discussed in @sec-model-diagnostics.
+This examination enhances our confidence in the model’s utility for forecasting electoral outcomes, ensuring it is robust enough to inform strategic decisions as the election approaches.
 
-In this case, the model correctly predicted 12 Harris wins and 15 Trump wins (true positives and true negatives).
-However, there were 5 instances where the model incorrectly predicted a Trump win when Harris won (false positives), and 3 instances where it predicted a Harris win when Trump won (false negatives).
-This breakdown helps assess not only the overall accuracy but also how well the model distinguishes between close contests in different states, which is critical for predicting the outcome of the 2024 presidential election.
+### Model Diagnostics - Harris Model
+The posterior predictive check for the Harris model is presented in Figure X, illustrating the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+@fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
+The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-harris-check
+#| fig-cap: "Posterior predictive check for Kamala Harris's model, illustrating the distribution of observed support percentages $y$ against the predicted support values $y_{\text{rep}}$.
+The graph shows a strong correspondence between the observed and predicted distributions, indicating that the model accurately captures the dynamics of voter support for Harris in the given dataset."
+
+# Generate posterior predictive checks for Harris model
+posterior_predict_harris <- posterior_predict(harris_model)
+
+# Plot the posterior predictive check
+pp_check_harris <- bayesplot::pp_check(harris_model, nsamples = 100) +
+  xlab("Support (%)") +
+  ylab("Density") +
+  theme_minimal()
+
+print(pp_check_harris)
+```
+
+## Posterior Predictive Check for the Harris Model
+The posterior predictive check for the Harris model is presented in Figure X, illustrating the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+This visualization features a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
+The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
+
+In the plot, we observe a strong overlap between the $y$ and $y_{\text{rep}}$ distributions, with the dark blue curve representing the observed support percentages primarily concentrated around 48% to 52%.
+This indicates that the model successfully approximates the actual voting support dynamics for Harris.
+The light blue curves signify the distribution of predicted values, demonstrating the model's ability to replicate the observed trends effectively.
+The close alignment of the observed and predicted densities suggests that the model adequately reflects the underlying support for Harris among voters, thereby affirming the appropriateness of the model structure.
+However, similar to the Trump model, the presence of some variability in the predicted values indicates that while the model performs well, there may still be unexplained variations in support levels that could benefit from further analysis.
+Such limitations are also discussed in @sec-model-diagnostics.
 
 # Pollster Methodology Overview and Evaluation
-
 The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times [@nyt-siena-methodology].

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 ### Predictive Variables
-In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over ti
+In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.
+
 #### State
@@
 This section details additional variables that provide essential context for our analysis, including numeric grade, which evaluates the quality of the polls; transparency score, indicating the openness of poll methodologies; pollster, identifying the organizations conducting the polls; and start date, which denotes the initiation of the polling period for each survey.
+
 #### Numeric Grade
@@
 ## Model Summary
+### Trump Model Summary
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-trump-fixed-effects
+#| tbl-cap: "Fixed Effects Summary for Trump Model."
+#| warning: false
+
+# Summarize fixed effects for Trump model
+trump_summary <- tidy(trump_model, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `Lower 95% CI` = conf.low,
+    `Upper 95% CI` = conf.high
+  ) %>%
+  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Clean up term names
+
+# Display the Trump model summary table
+knitr::kable(trump_summary) %>%
+  kable_styling(full_width = FALSE, position = "center")
+```
+
 ### Harris Model Summary
@@
 ```
-
-The model summary for Kamala Harris indicates that the intercept estimate is $-0.1508$ with a standard error of $0.0279$.
-This intercept value suggests a negative baseline log-odds for her victory, reflecting the starting point of her support in the absence of any other predictors.
-The model shows that the coefficients for the natural spline terms ($nsendate_{num}$) are particularly informative, with the first spline term ($nsendate_{num\_df41}$) yielding a positive estimate of $0.1225$, suggesting that as time progresses, Harris’s probability of winning increases, particularly during key moments captured by the polling data.
-This is also reflected in the confidence intervals, with the upper limit of the confidence interval for this term reaching $0.1903$.
-In contrast, the estimates for the second spline term ($nsendate_{num\_df42}$) indicate a smaller effect, suggesting fluctuations in her support that may warrant further investigation.
-
-### Trump Model Summary
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-trump-fixed-effects
-#| tbl-cap: "Fixed Effects Summary for Trump Model."
-#| warning: false
-
-# Summarize fixed effects for Trump model
-trump_summary <- tidy(trump_model, effects = "fixed", conf.int = TRUE) %>%
-  select(term, estimate, std.error, conf.low, conf.high) %>%
-  rename(
-    Estimate = estimate,
-    `Std. Error` = std.error,
-    `Lower 95% CI` = conf.low,
-    `Upper 95% CI` = conf.high
-  ) %>%
-  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Clean up term names
-
-# Display the Trump model summary table
-knitr::kable(trump_summary) %>%
-  kable_styling(full_width = FALSE, position = "center")
-```
-In comparison, the model summary for Donald Trump presents a different picture.
-The intercept for Trump's model is $-0.0923$ with a standard error of $0.0306$, indicating a similarly negative baseline log-odds of victory as Harris.
-However, Trump's model demonstrates more pronounced increases in support as reflected by the estimates for the natural spline terms.
-For example, the second spline term ($nsendate_{num\_df42}$) has a positive coefficient of $0.0803$, highlighting a significant upward trend in support over time.
-This trend is further supported by the confidence interval, which ranges from $0.0179$ to $0.1426$, indicating a substantial increase in his predicted probabilities as the election approaches.
-The fourth spline term ($nsendate_{num\_df44}$) also shows a positive coefficient of $0.0789$, suggesting that the support for Trump remains resilient over the polling period.
-Overall, both models reveal complex interactions between time and voter sentiment, with implications for the electoral dynamics leading up to the election.
 

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 #| label: fig-state-distribution
-#| fig-cap: "Bar plot showing the distribution of observations across key battleground states in the 2024 United States Presidential Election polling data.
-The plot illustrates the count of polling observations for each state, highlighting states with more frequent polling activity.
-This visualization aids in understanding the level of engagement and data collection in each battleground state leading up to the election."
+#| fig-cap: "Distribution of Observations for Each State in Polling Data.
+Illustrates the number of polling observations across key battleground states in the lead-up to the 2024 United States Presidential Election.
+Notably, Pennsylvania has the highest count of observations, suggesting a more robust polling presence, while Nevada shows fewer observations.
+"
 
@@
 #| label: fig-end-date-distribution
-#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
-The histogram shows the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches.
-This visualization is crucial for understanding how polling data correlates with the election timeline."
+#| fig-cap: "Distribution of Polling End Dates.
+This histogram displays the frequency of polling end dates for key battleground states leading up to the 2024 United States Presidential Election, highlighting peaks in polling activity around late September."
 
@@
 The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
+
 ```{r}
@@
 #| label: fig-numgrade-distribution
-#| fig-cap: "Histogram illustrating the distribution of polling end dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
-Displays the frequency of polls conducted over time, providing insight into the timing of polling activities and potential trends in voter sentiment as Election Day approaches.
-This visualization is crucial for understanding how polling data correlates with the election timeline."
+#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data as the 2024 United States Presidential Election approaches.
+The graph highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality.
+"
 
@@
 #| label: fig-transparency-score-distribution
-#| fig-cap: "Histogram illustrating the distribution of transparency scores for polls in key battleground states.
-Reveals the frequency of different transparency ratings, providing insight into the reliability of polling data.
-Understanding the distribution of transparency scores is crucial for evaluating the credibility of the polling results as Election Day approaches."
+#| fig-cap: "Histogram displaying the distribution of transparency scores among pollsters in the polling data leading up to the 2024 United States Presidential Election.
+The plot indicates the frequency of various transparency scores, highlighting a concentration of scores around 7 and 9, which suggests a predominance of high transparency among the assessed pollsters.
+"
 
@@
 #| label: fig-pollster-distribution
-#| fig-cap: "Horizontal bar graph illustrating the distribution of polls by various pollsters in key battleground states.
-Displays the count of polls conducted by each pollster, highlighting the most frequently utilized polling organizations.
-Understanding the diversity of pollsters is essential for assessing the reliability and variance in polling data as Election Day approaches."
+#| fig-cap: "Horizontal bar chart displaying the distribution of polls conducted by various pollsters in key battleground states for the 2024 United States Presidential Election.
+The chart highlights the frequency of polling activities across different organizations, with Marist and Emerson standing out as the most prolific pollsters."
 
@@
 #| label: fig-start-date-distribution
-#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states in the lead-up to the 2024 United States Presidential Election.
-This visualization displays the frequency of polls conducted over time, providing insights into when polling activities were initiated and potential trends in voter sentiment as Election Day approaches.
-Understanding the timing of polls is crucial for evaluating the relevance of the polling data."
+#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states leading up to the 2024 United States Presidential Election.
+The chart highlights the frequency of polling activities initiated over time, with a notable peak on September 14, indicating a surge in polling efforts as the election approached."
 
@@
 
-The decision to utilize Bayesian logistic regression models stems from their flexibility and the ability to incorporate prior beliefs into the estimation process.
-This approach is particularly valuable in the context of electoral forecasting, where uncertainty is a significant factor.
-We defined the following priors for the model parameters:
-
-- For the intercept $\beta_0$, we specified $\beta_0 \sim \text{Normal}(0, 10)$, providing a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds.
-- For the slope $\beta_1$, we used $\beta_1 \sim \text{Normal}(0, 5)$, reflecting reasonable expectations regarding the influence of polling trends over time.
-
-Natural splines were chosen to model the effect of the end date due to the observed nonlinearities in voter support, allowing the model to adapt to changing public opinion.
-Key assumptions include the independence of observations within states and the appropriateness of the Beta distribution for modeling proportions.
-However, potential limitations include the reliance on the accuracy of polling data and the assumption that historical voting patterns are indicative of future behavior.
+The Bayesian Beta Regression model with a logit link function is the most suitable model for our prediction of the 2024 U.S.
+Presidential Election because it aligns closely with both the nature of the data and the goals of our paper, which include understanding voter support trends over time and accounting for regional differences across battleground states.
+
+Our primary outcome variable is the percentage of support for each candidate in 7 key battleground states, scaled to a proportion between 0 and 1.
+Support percentages are naturally bounded between 0% and 100%, and this bounded nature poses challenges for traditional linear regression, which assumes an unbounded continuous outcome.
+The Beta distribution is ideal for modeling proportion data because it is inherently defined on the interval (0, 1).
+This allows us to model the proportions of support without the risk of generating unrealistic values outside this range.
+Unlike a linear model, which could predict support percentages exceeding 100% or falling below 0%, the Beta distribution respects the boundaries of our data.
+The logit link function further ensures that the model predicts in terms of log odds, which is suitable for interpreting support as probabilities—an essential consideration given our goal of forecasting potential victory in each state.
+
+One of our key goals is to examine how support for each candidate evolves over time as Election Day approaches.
+Polling data often show non-linear trends, with fluctuations in support that may correspond to campaign events, debates, or shifts in the political landscape.
+Using natural splines with end_date_num (the number of days since the earliest poll date) allows the model to capture these non-linear trends in a flexible way.
+By including natural splines with 4 degrees of freedom, we allow the model to capture gradual changes in voter sentiment while avoiding overfitting to noise or abrupt fluctuations in the data.
+This is particularly important in the context of election polling, where voter opinions can change either slowly or rapidly depending on unpredictable external events.
+Alternative models, such as standard logistic regression without splines, would fail to capture these changes, potentially leading to a less accurate depiction of how support varies over time.
+
+A unique aspect of our analysis is the focus on battleground states, which have distinct political dynamics that can heavily influence election outcomes.
+Incorporating state as a random effect provides flexibility in estimating state-specific baseline support levels for each candidate, while still allowing us to identify overall trends across all states.
+This approach is more effective than treating state as a fixed effect because it avoids overparameterizing the model with individual state coefficients, which could lead to overfitting in cases with limited data for some states.
+The random effect structure also helps generalize the model’s predictions, making it more robust when predicting support in battleground states with varying polling frequencies.
+
+The Bayesian approach offers advantages for our model by allowing us to incorporate prior information, enhancing the stability and reliability of our estimates.
+Given that polling data can be sparse for certain states or dates, Bayesian inference helps stabilize parameter estimates, especially when there are limited data points for certain states or specific dates.
+By using weakly informative priors, we let the data primarily determine the model’s outcomes, while still providing some structure to prevent overfitting or extreme parameter values.
+The Bayesian framework also produces credible intervals, which provide a clear interpretation of uncertainty around our predictions.
+In the context of election forecasting, this is valuable as it enables us to quantify the range of plausible outcomes for each candidate’s support, adding a layer of transparency to our predictions.
+This model provides interpretability that aligns well with the goals of our paper.
+By structuring the model with a logit link function, we can interpret the results in terms of odds or probability of victory, which is highly relevant in an election context.
+This interpretability allows us to understand the likelihood of each candidate winning in a specific state at a given time.
+
+Additionally, the Bayesian Beta regression model offers flexibility in how we examine each candidate’s trends individually.
+By constructing separate models for Trump and Harris, we can directly compare their temporal trends and state-level variations without confounding effects.
+
+Overall, the Bayesian Beta regression model with a logit link function is the most suitable model for our study.
+It allows for flexible non-linear trends over time, accounts for regional variation across battleground states, and provides interpretable results in terms of victory probabilities.
+The Bayesian framework stabilizes estimates in the presence of sparse data and provides meaningful uncertainty estimates.
+The model’s structure aligns well with our goal of forecasting each candidate’s probability of victory, justifying the choice for achieving a comprehensive analysis and prediction of the 2024 U.S.
+Presidential Election.
 
 ## Model Summary
+
 ### Trump Model Summary

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 Pollsters then store these responses in structured databases for analysis.
-In the dataset obtained from Project 538, each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate.
+In the dataset obtained from Project 538 [@FiveThirtyEight], each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate.
 Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.
@@
 States with more polling observations provide a larger sample size, potentially leading to more stable and reliable predictions for those states in the model.
+
+\newpage
 
@@
 
+\newpage
+
 ### Other Variables
@@
 
+\newpage
+
 #### Transparency Score
@@
 This adjustment is crucial as transparency can influence the stability of the model’s predictions, especially in closely contested states where the reliability of polling data may impact the accuracy of our forecasted election outcomes.
+
+\newpage
 
@@
 This filtered timeline ensures our model focuses on data that reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as we approach Election Day.
+
+\newpage
 
@@
 
-# Pollster Methodology Overview and Evaluation
-The following information about the Siena College/New York Times poll methodology is based on the details provided in the article "How The Times/Siena Poll Is Conducted" by the New York Times [@nyt-siena-methodology].
-
-The Siena College/New York Times poll for the 2024 presidential election uses a robust sampling approach to ensure accuracy and relevance.
-The poll applies a stratified dual-frame sample, drawing from both land-lines and cell phones.
-Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia.
-The sampling population is all registered voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
-
-Given the nature of the presidential elections, the decision is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are likeliest to decide the outcome of the race.
-The voters’ information is taken from the L-2 voter file, which includes details such as voter registration and history of participation in previous elections.
-
-To refine their sample, Siena College adjusts the data by key demographic and political variables such as region, race/ethnicity, party affiliation, education, and voting patterns from the 2020 election.
+# Pollster methodology overview and evaluation: NYT/Siena College
+
+## Background of NYT/Siena College Polling
+The NYT/Siena College Polling is a U.S.
+polling organization where the New York Times and Siena College collaborate to deliver precise and timely poll results for the presidential elections [@nyt-siena-methodology].
+For the 2024 election cycle, The NYT/Siena College poll uses a robust sampling approach to ensure accuracy and representativeness.
+The poll applies a stratified dual-frame sample, drawing from both landlines and cell phones.
+
+## Target population, Sampling frame, and Sample
+- Target population: Target population refers to “the collection of all items about which we would like to speak” [@telling-stories-data].
+In this case, NYT/Siena College’s polling target population is all registered American voters who live in the six battleground states: Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
+Given the nature of the presidential elections, the election result is based on the electoral college instead of the popular vote, thus the pollster focuses on polling on the states that are the likeliest to decide the outcome of the race, which are called battleground states.
+The battleground states are states that have similar voter support for both the Democratic or the Republican party, a slight difference in support could determine which party wins a state.
+- Frame: Sampling frame refers to “a list of all items from the target population that we could get data about” [@telling-stories-data].
+NYT/Siena College’s sampling frame is a comprehensive list of registered American voters obtained from the L-2 voter file, it includes details such as demographic characteristics, contact information, and voting history.
+By using the L-2 voter file as their sampling frame, the pollster can ensure representativeness by allowing them to draw samples that accurately reflect the demographics of the registered voter population.
+- Samples: Sample refers to “the items from the sampling frame that we get data about” [@telling-stories-data].
+In this case, the samples of NYT/Sienna College are individuals who are registered American voters obtained from the L-2 voter file from each stratum from each state who have also answered and completed the questionnaires.
+
+## Sampling Methodology
+NYT/Siena College uses stratified sampling as their methodology where the population is divided into distinct strata based on shared characteristics.
+These strata include age, gender, race, geographic region, and other various relevant variables.
+Once the population is divided, random samples are taken from each stratum, in proportion to their size in the population.
+For each poll, NYT/Siena College takes a sample size of 1000 registered American voters.
+To refine their sample, NYT/Siena College adjusts the data by key demographic and political variables such as region, race/ethnicity, party affiliation, education, and voting patterns from the 2020 election.
 These adjustments help ensure the sample accurately represents the electorate.
 Additionally, they use a “likely voter screen,” which combines self-reported likelihood of voting with historical voter behavior to estimate how likely respondents are to vote in the upcoming election.
-
-At every step of the survey, Sienna/NYT uses the information in the data to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions.
+At every step of the survey, Siena/NYT uses the information in the datasets to try to ensure that they have the right number of Democrats and Republicans, young people and old people, the right ratio of people with different income levels, and a diverse mix of different races and regions.
 Once the survey is complete, they compare their respondents to the voter file and use a process known as weighting to ensure that the sample reflects the broader voting population.
@@
 
-The New York Times/Siena College pollsters handle non-response by using weighting and adjusting their sample to correct for any biases that might emerge due to individuals not responding.
+## Trade-off of Sampling Methodology
+Although stratified sampling has many strengths such as appropriate representation of the target population, reduction of sampling bias, and sampling efficiency, it also has some weaknesses.
+First, its implementation can be complex, stratified sampling requires detailed knowledge of the population to define the strata appropriately.
+This can be challenging to implement if accurate, up-to-date information on demographic or geographic characteristics is unavailable or difficult to access.
+For example, if there are significant changes in the L-2 voter file since the last election, it is difficult to take account into the unknown changes.
+Second, there could be potential for mis-stratification.
+If the chosen strata do not capture meaningful differences within the population, or if important subgroups are overlooked, the results might be biased or inaccurate.
+For example, if new voting trends emerge that were not captured in the strata definitions, the poll might miss important shifts in voter opinion.
+Third, stratified sampling can be time-consuming and resource-intensive.
+Diving the population into strata and ensuring proportionate sampling within each group can be more time-consuming and resource-intensive compared to simple random sampling.
+Pollsters may need to gather detailed data on the target population, which adds to the cost and complexity of the survey.
+Lastly, there is the risk of over-emphasis on stratification.
+If the population is stratified too finely, the sample size in each stratum may become too small to yield meaningful results, increasing the margin of error for each subgroup.
+
+## Sample Recruitment Methods
+Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia.
+Most people that the live interviewers call do not answer the call, and some don’t finish the survey, leaving only about two percent both answer their phone and complete the poll.
+To mitigate this:
+First, the pollster keeps calling numbers until someone picks up, even though it can take hours–a reason why a high-quality poll is time-consuming and expensive to conduct. 
+Second, voter files are used to help make sure the pollster is reaching a representative sample; for example, they may reach out to a higher percentage of people who are less likely to respond. 
+Third, pollster mathematically adjusts the responses they do receive so they are representative of the wider population, it includes adjusting by age, race, gender, voting history and etc.
+
+## Non-response Handling
+Non-responses are critical factors in polling because they affect the representativeness of the sample and, consequently, the accuracy of the poll results.
+When a significant portion of selected individuals doesn’t respond, the poll risks becoming unrepresentative of the broader population, especially if certain groups are systematically more or less likely to respond.
+It is also important to know how pollsters handle non-response since it provides transparency about the polling methodology.
+It also gives confidence that the results are not skewed by an unrepresentative sample when pollsters clearly explain their non-response handling.
+Therefore, polls with thorough non-response strategies are generally more reliable than those without.
+The NYT/Siena College pollsters handle non-response by using weighting and adjusting their sample to correct for any biases that might emerge due to individuals not responding.
 Specifically, they adjust on multiple demographic and political variables such as age, gender, education, race/ethnicity, and party affiliation.
 They also account for variations in voter likelihood and previous voting patterns, ensuring that the sample represents the likely electorate as accurately as possible.
-
 This helps reduce the potential bias from non-response, especially since certain demographic groups or political affiliations may be less likely to respond to polls.
@@
 
+## Questionnaire Pros and Cons
+Based on the NYT/Siena College’s past practices on presidential election polls, they often ask questions about candidate preference, approval ratings, voter opinions on key issues, economic perceptions, as well as social and cultural issues.
+They also ask about the respondents’ demographic information and voting behavior to categorize respondents for analysis.
+
+### Strengths of the NYT/Siena College questionnaires
+First, they have a comprehensive question design.
+NYT/Sienna College polls ask a wide range of questions that provides a detailed view of the respondent’s preferences and opinions.
+It helps produce multi-dimensional data that allows for in-depth analysis and segmentation of voter groups.
+Second, the questionnaires are tailored to likely voters, as identified through the L-2 voter file.
+This means that the voters are more likely to be up-to-date with relevant policies proposed by different parties, and have formed their own opinions about them, this improves the predictive reliability of the poll.
+Third, the questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.
+
+### Weaknesses of the NYT/Siena College questionnaires
+First, the length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls.
+Respondents may rush through questions or drop out before completion.
+Respondents may also be prone to moderacy response bias where they have the tendency to choose middle options regardless of question content, as well as response order bias where respondents choose answers based on their order.
+Second, questionnaires have a dependence on self-reported likelihood to vote.
+Self-reported likelihood can be unreliable, especially for individuals whose intentions to vote may change close to the actual election date.
+This can reduce the accuracy of predictions based on likely voter models.
+Third, there are limited open-ended questions.
+Many voters may have mixed opinions on specific issues or policies that cannot be explained in a single-choice answer.
+However, answers to open-ended questions are often difficult to convert to actionable data, so questionnaires tend to rely on closed-ended questions for scalability.
+This might restrict the ability to capture complex reasonings behind their choices, which could add extra strain on the time and resources used to interpret the answers.
+
 # Idealized Methodology
-
 Building on our discussion of the New York Times/Siena College Poll, we now present an idealized methodology that aims to enhance the accuracy and reliability of our predictions.

### 2024-11-04 paper/paper.qmd
commmiting changes to paper.qmd

@@
 
-\newpage
-
 #### Pollster
@@
 ## Model Setup
-In this section, we outline the statistical models employed to predict the probabilities of victory for Donald Trump and Kamala Harris in the 2024 presidential election within key battleground states.
-Our analysis utilizes a Bayesian framework, which allows us to incorporate uncertainty and prior information into the modeling process, yielding more robust predictions.
-The primary estimand for both models is the probability of each candidate's victory, expressed through a logistic regression approach.
-By examining polling data and its temporal dynamics, we aim to capture the evolving nature of voter support leading up to the election.
-We present separate subsections for the models of Donald Trump and Kamala Harris, detailing the structure, variables, and assumptions underlying each model.
-This division allows for a focused discussion on how each model has been tailored to reflect the unique electoral contexts and polling dynamics for each candidate.
+Our models were ran in R [@citeR] using the 'rstanarm' package of @rstanarm.
 
@@
 # Additional Data Details {#sec-data-details}
-
 Given the nature of the presidential election, the electoral college is more impactful and influential on the possible outcome compared to the nationwide popular votes.

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 
-\newpage
-
 # Model {#sec-model}
@@
 
-The model was implemented using the **`brms`** package in R, applying a **Beta distribution** with a **logit link function** to accurately model the proportions of support:
+The model was implemented using the 'brms' package [@brms] in R [@citeR], applying a Beta distribution with a logit link function to accurately model the proportions of support:
 

### 2024-11-04 paper/paper.qmd
committing testing scripts and changes to paper.qmd

@@
 
-- “If the election were held today, would you vote for [Candidate A] or [Candidate B]?” 
-- “How strongly do you support [Candidate]?” (e.g., on a scale of 1 to 10, or strongly, somewhat, undecided)
-- “Do you support [Policy A] proposed by [Candidate A] or [Policy B] proposed by [Candidate B]?”
+- “If the election were held today, would you vote for Candidate A or Candidate B?” 
+- “How strongly do you support Candidate?” (e.g., on a scale of 1 to 10, or strongly, somewhat, undecided)
+- “Do you support Policy A proposed by Candidate A or Policy B proposed by Candidate B?”
 
@@
 Pollsters then clean the data by checking for inconsistencies, removing incomplete responses, and addressing non-responses.
-Pollsters also perform weighting to ensure the representativeness of data collected, details of weighting are further explained in {appendix 1}.
+Pollsters also perform weighting to ensure the representativeness of data collected, details of weighting are further explained in @sec-pollster-methodology.
 
@@
 This consideration is crucial, as polling quality can impact the stability of the model’s forecasts, especially in closely contested states where polling accuracy is essential for reliable predictions.
-
 \newpage
@@
 
-# Pollster methodology overview and evaluation: NYT/Siena College
+# Pollster methodology overview and evaluation: NYT/Siena College {#sec-pollster-methodology}
 

### 2024-11-04 scripts/00-simulate_data.R

@@ -50,3 +50,4 @@ head(simulated_data)
 
 #### Save data ####
 write_csv(simulated_data, "data/00-simulated_data/simulated_data.csv")
+

### 2024-11-04 scripts/01-test_simulated_data.R

@@ -15,75 +15,43 @@
 library(tidyverse)
 
 analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
+# Test 1: Check Support Percentage Consistency
+test_support_percentage <- function(data) {
+  support_summary <- data %>%
+    group_by(candidate) %>%
+    summarise(avg_support = mean(support_percent), .groups = 'drop')
+  
+  # Check that the support for "Donald Trump" is between 45 and 55
+  trump_support <- support_summary %>% filter(candidate == "Donald Trump") %>% pull(avg_support)
+  kamala_support <- support_summary %>% filter(candidate == "Kamala Harris") %>% pull(avg_support)
+  
+  if (trump_support >= 45 && trump_support <= 55) {
+    message("Test 1 Passed: Donald Trump's support percentage is within the expected range.")
+  } else {
+    stop("Test 1 Failed: Donald Trump's support percentage is outside the expected range.")
+  }
+  
+  # Check that Kamala Harris' support is correctly calculated as well
+  if (kamala_support >= 45 && kamala_support <= 55) {
+    message("Test 1 Passed: Kamala Harris' support percentage is within the expected range.")
+  } else {
+    stop("Test 1 Failed: Kamala Harris' support percentage is outside the expected range.")
+  }
 }
 
+# Run the test
+test_support_percentage(simulated_data)
 
-#### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+# Test 2: Check Sample Size Range
+test_sample_size <- function(data) {
+  if (all(data$sample_size >= 500 & data$sample_size <= 3000)) {
+    message("Test 2 Passed: All sample sizes are within the expected range.")
+  } else {
+    stop("Test 2 Failed: There are sample sizes outside the expected range of 500 to 3000.")
+  }
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
+# Run the test
[…]

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -11,59 +11,90 @@
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+clean_poll_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 
+# Test 1: Check for Required Columns
+test_required_columns <- function(data) {
+  required_columns <- c("state", "start_date", "end_date", "pollster", 
+                        "numeric_grade", "transparency_score", 
+                        "Trump_pct", "Harris_pct")
+  missing_columns <- setdiff(required_columns, colnames(data))
+  if (length(missing_columns) > 0) {
+    stop(paste("Test Failed: Missing columns:", paste(missing_columns, collapse = ", ")))
+  } else {
+    message("Test Passed: All required columns are present.")
+  }
+}
 
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
+# Run the test
+test_required_columns(clean_poll_data)
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
+# Test 2: Check Data Types
+test_data_types <- function(data) {
+  expected_types <- list(
+    state = "character",
+    start_date = "Date",
+    end_date = "Date",
+    pollster = "character",
+    numeric_grade = "numeric",
+    transparency_score = "numeric",
+    Trump_pct = "numeric",
+    Harris_pct = "numeric"
+  )
+  
+  for (col in names(expected_types)) {
+    if (!inherits(data[[col]], expected_types[[col]])) {
+      stop(paste("Test Failed: Column", col, "is not of type", expected_types[[col]]))
+    }
+  }
+  message("Test Passed: All columns have expected data types.")
+}
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
+# Run the test
+test_data_types(clean_poll_data)
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
+# Test 3: Check for NA Values in Critical Columns
+test_na_values <- function(data) {
+  critical_columns <- c("state", "start_date", "end_date", "Trump_pct", "Harris_pct")
[…]

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 date-format: long
-abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, focusing on critical battleground states.
-Utilizing a Bayesian approach with state-level polling data, we estimate the winning probabilities for candidates Donald Trump and Kamala Harris.
-Our findings reveal that Harris is projected to win in six out of seven key battleground states, including Michigan, Nevada, and North Carolina, with predicted winning percentages surpassing 50%.
-Given this significant advantage, we predict Kamala Harris to be the likely winner of the election.
-This analysis underscores the importance of battleground states in shaping electoral outcomes and provides valuable insights into voter sentiment as Election Day approaches."
+abstract: "This paper presents a predictive model for the 2024 United States Presidential Election, focusing on essential battleground states.
+Utilizing a Bayesian framework and state-level polling data, we estimate the probabilities of victory for candidates Donald Trump and Kamala Harris.
+Our findings indicate that Harris is projected to secure a win in six out of seven key battleground states, such as Michigan, Nevada, and North Carolina, with predicted support exceeding 50%.
+This analysis not only forecasts the likely winner of the U.S.
+election but also equips society and the economy to better undertstand the current U.S.
+political climate."
 format: pdf
@@
 # Introduction {#sec-intro}
-The 2024 United States Presidential Election is shaping up to be a critical event, particularly as battleground states emerge as pivotal areas influencing the electoral outcome.
-As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions becomes essential for predicting the election results.
-This paper utilizes state-level polling data from Project 538 [@FiveThirtyEight] and the statistical programming language R [@citeR] to build a predictive model that estimates the likelihood of victory for each candidate in several key battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-
-The primary estimand in this study is the probability that either Donald Trump or Kamala Harris will win each battleground state, based on aggregated polling averages.
-Our model employs a Bayesian hierarchical approach with natural splines to capture the nuances of voter sentiment over time, allowing for the estimation of winning probabilities across these crucial states.
-By comparing the predicted winning probabilities for both candidates, we can ascertain the likely winner based on the prevailing voter support patterns.
-
-Our analysis reveals that Kamala Harris is projected to win in six out of the seven battleground states examined, indicating a favorable voter sentiment in states such as Michigan and Nevada.
-Conversely, Donald Trump shows a competitive stance in Arizona, where he is also predicted to secure a slight edge.
-The calculation of the predicted winner is determined by comparing the estimated probabilities for each candidate, which underscores Harris's overall advantage in the election.
-
-Understanding these predictions is vital as it offers insights into the dynamics of voter behavior and the potential implications for election outcomes.
-This information can help political analysts, campaign strategists, and the public anticipate how these battleground states may influence the overall results.
-Accurate predictions not only inform campaign strategies but also highlight regions where voter sentiment may shift in the lead-up to Election Day.
-
-The remainder of this paper is structured as follows.
-In @sec-data, we describe the data and variables used in the analysis.
-@sec-model outlines the model setup and estimation strategy.
-@sec-results presents the results, including the predicted election outcomes and visualizations.
-Finally, @sec-discussion discusses the implications and limitations of our findings.
+The 2024 United States Presidential Election is anticipated to be a pivotal moment in American political history, particularly as battleground states emerge as key determinants of the electoral outcome.
+With candidates Donald Trump and Kamala Harris intensifying their campaigns, a thorough understanding of voter preferences in these critical regions becomes essential for forecasting election results.
+This paper leverages state-level polling data from Project 538 [@FiveThirtyEight] and employs the statistical programming language R [@citeR] to construct a predictive model that estimates the likelihood of victory for each candidate across several crucial battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+
+The primary estimand of this study is the probability of winning for either Donald Trump or Kamala Harris in these battleground states, derived from aggregated polling averages.
+Our analysis utilizes a Bayesian hierarchical approach combined with natural splines to accurately capture the temporal dynamics of voter sentiment, enabling us to estimate winning probabilities in these decisive areas.
+By comparing the predicted winning probabilities for both candidates, we can ascertain the likely winner based on prevailing voter support trends.
+Our findings indicate a competitive landscape, with Kamala Harris projected to secure victories in six of the seven battleground states analyzed, including key states like Michigan and Nevada, where voter sentiment appears to favor her candidacy.
+However, Donald Trump is also predicted to maintain a competitive edge in Arizona, showcasing the variability in voter preferences across these critical regions.
+The predicted winner is determined by comparing the estimated probabilities for each candidate, highlighting Harris’s overall advantage leading into the election.
+
+Understanding these predictions is crucial as they not only forecast the likely outcome of the U.S.
+election but also help society and the economy to better understand the current state of the US political climate.
+The implications of these findings extend beyond mere predictions, providing a framework for analyzing how shifts in voter sentiment may occur as the election draws near.
+The remainder of this paper is organized as follows: Section @sec-data details the data and variables utilized in the analysis; Section @sec-model outlines the model setup and estimation strategies employed; Section @sec-results presents the results, including visualizations of the predicted election outcomes; and Section @sec-discussion discusses the implications and limitations of our findings.
 
@@
 ## Overview
-The dataset of the 2024 Presidential Election cycle is obtained from Project 538 [@FiveThirtyEight].
-The dataset is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
+The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538's polling project [@FiveThirtyEight] on October 19th, 2024.
+The data is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
 It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
@@
 The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
-For all the following variables mentioned, we collected observations for both of the candidates.
+For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates.
 We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris.
@@
 ## Measurement
-In the current presidential election cycle,  Americans’ opinions on voting for their preferred candidate are transformed into data points through a series of steps that involve surveying, data processing, and structuring responses.
-Pollsters then store these responses in structured databases for analysis.
-In the dataset obtained from Project 538 [@FiveThirtyEight], each entry in the dataset corresponds to a poll conducted by each pollster, capturing the percentage of respondents who express support for each candidate.
-Each pollster has different approaches to sampling respondents, recruiting respondents, as well as questionnaire types, which affects each pollster’s reliability and data quality.
-
-Pollsters begin by selecting a sample from the registered voter population, often using voter files with demographic and contact information.
-They then apply methods of stratified random sampling to ensure the sample reflects the electorate’s diversity in terms of age, gender, race, education, voter behavior, etc.
-
-Then, pollsters reach out to voters through various channels, such as phone calls, emails, text messages, or online panels.
-These are called recruiting methods and usually, each of them has its trade-offs.
-If the respondents are connected through any of these channels, they are asked a series of questions either by a live interviewer or a preset of questions.
-
-The types of questions often include:
-
-- “If the election were held today, would you vote for Candidate A or Candidate B?” 
-- “How strongly do you support Candidate?” (e.g., on a scale of 1 to 10, or strongly, somewhat, undecided)
-- “Do you support Policy A proposed by Candidate A or Policy B proposed by Candidate B?”
-
-Responses are recorded digitally, either by survey software or directly into a database by interviewers, capturing both the answers and relevant demographic data.
-Each response becomes a data point in the database.
-For example, if a respondent prefers Candidate A, their answer may be recorded as a binary variable (e.g., “1” for Candidate A, “0” for Candidate B).
-Along with candidate preference, pollsters capture demographic and behavioral details, such as age, gender, race, education, income, voter history, and party registration.
-
-Each respondent’s answers and demographic details are grouped and saved in a structured database.
-Pollsters then clean the data by checking for inconsistencies, removing incomplete responses, and addressing non-responses.
-Pollsters also perform weighting to ensure the representativeness of data collected, details of weighting are further explained in @sec-pollster-methodology.
-
-Once cleaned and weighted, data points are aggregated to determine overall candidate support, often producing metrics such as the percentage of respondents supporting each candidate by state or other specific demographics.
-Each pollster would have a database with all the weighted and aggregated data, and Project 538 presents a compilation of the databases obtained from various pollsters.
-
-## Variables
-The collected dataset contains several key variables relevant to the analysis:
+Voters’ opinions are multifaceted and can be influenced by numerous factors, including media coverage, recent political events, personal beliefs, and social pressures.
+Each voter’s decision is shaped by a unique mix of these factors, which can change frequently, especially close to an election.
+Pollsters start by converting the complex, multi-layered opinions in voters' heads into a simple polling response.
+This is the first challenge: each voter’s thoughts might not fit neatly into options like “Trump,” “Harris,” or “Undecided.” People may lean toward one candidate but have reservations, or may change their views over time.
+By asking for a single answer, the poll essentially “forces” an opinion that may not fully capture the voter’s internal conflict.
+
+Once collected, these individual responses are combined into percentages of support.
+This aggregation abstracts further, treating all “Trump” responses equally, regardless of how confident or reluctant the voter might have been.
+In effect, the complexity of each opinion is smoothed over into a uniform data point, with many variations of opinions lost in the process.
+
+Finally, these support percentages are fed into a model to make a predictive statement about the election outcome.
+The model’s goal is to determine who is likely to win, based on the assumption that the aggregated opinions in our data reflect real-world voter behavior on Election Day.
+However, this step relies heavily on simplification, assuming that the snapshot of opinions captured in the polls will hold steady and that all voters will follow through on their expressed preferences.
+
+The raw polling data gives us percentages of support for each candidate in each state.
+To measure our outcome variable (the likelihood of winning), we convert these percentages into a binary outcome in our model (predicting Trump or Harris as the winner based on who has higher support in each state).
+Our model uses these support percentages to generate a probability of each candidate winning.
+This transformation helps translate fluctuating polling data into a clear prediction of the likely winner, but it also involves a level of simplification, as it assumes these support percentages accurately reflect voting intentions.
+
+Using state as a predictor variable, which influences our outcome variable, assumes that all responses within a state are comparable and that regional differences are consistent across the state.
+However, each state has internal diversity—urban vs.
+rural areas, differences in income levels, or cultural values—that the variable “state” may not fully capture.
+This simplification is necessary for the model, but it means that nuanced, within-state variations in opinion are lost in the abstraction.
+Another predictor variable, end_date is also measured, it is designed to capture how voter sentiment changes over time, especially as the election approaches.
+By using the end date of each poll, the model can recognize patterns or shifts in support due to campaign events, news coverage, or debates.
+While “end_date” provides temporal context, it simplifies complex, often continuous changes in opinion into a discrete snapshot.
+This abstraction assumes that voter sentiment is relatively stable between poll dates, which may not reflect reality.
+For example, an event occurring immediately after a poll’s end date could dramatically shift opinions that the model wouldn’t capture until the next poll.
+Additionally, the importance of time in influencing voter sentiment is assumed to be similar across states, but certain states might react differently to events depending on their local context or demographic makeup.
+
+## Variables {#sec-variables}
+The collected data from Project538 [@FiveThirtyEight] contains several key variables relevant to the analysis:
 
@@
 
-To ensure the quality of the analysis, only those polls that meet specific thresholds were included - for more details, refer to [Appendix -@sec-data-details].
+To ensure the quality of the analysis, only those polls that meet specific thresholds were included - for more details on the data cleaning process, refer to @sec-data-cleaning.
 
@@
 
-# Display vertical summary statistics in a kable table
-kable(summary_statistics_long) %>%
-  kable_styling(full_width = FALSE, position = "center")
-
-```
-For instance, @tbl-outcomevar reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
+# Put table into kable
+knitr::kable(summary_statistics_long)
+
+```
+@tbl-outcomevar reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
 ‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris.
@@
 #| fig-cap: "Distribution of Observations for Each State in Polling Data.
-Illustrates the number of polling observations across key battleground states in the lead-up to the 2024 United States Presidential Election.
+Illustrates the number of polling observations across key battleground statesn.
 Notably, Pennsylvania has the highest count of observations, suggesting a more robust polling presence, while Nevada shows fewer observations.
@@
 
-\newpage
-
 #### End Date {#sec-end-date}
@@
 #| label: fig-end-date-distribution
-#| fig-cap: "Distribution of Polling End Dates.
-This histogram displays the frequency of polling end dates for key battleground states leading up to the 2024 United States Presidential Election, highlighting peaks in polling activity around late September."
+#| fig-cap: "Distribution of Polling End Dates. This histogram displays the frequency of polling end dates for key battleground states, highlighting peaks in polling activity around late September."
 
@@
 #| label: fig-numgrade-distribution
-#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data as the 2024 United States Presidential Election approaches.
-The graph highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality.
+#| fig-cap: "Density plot illustrating the distribution of numeric grades assigned to pollsters within the polling data.
+Highlights the frequency of numeric grades, indicating key concentrations that suggest variability in pollster quality.
 "
@@
 #| label: fig-transparency-score-distribution
-#| fig-cap: "Histogram displaying the distribution of transparency scores among pollsters in the polling data leading up to the 2024 United States Presidential Election.
-The plot indicates the frequency of various transparency scores, highlighting a concentration of scores around 7 and 9, which suggests a predominance of high transparency among the assessed pollsters.
+#| fig-cap: "Histogram displaying the distribution of transparency scores among pollsters in the polling data.
+Indicates the frequency of various transparency scores, highlighting a concentration of scores around 7 and 9, which suggests a predominance of high transparency among the assessed pollsters.
 "
@@
 #| label: fig-pollster-distribution
-#| fig-cap: "Horizontal bar chart displaying the distribution of polls conducted by various pollsters in key battleground states for the 2024 United States Presidential Election.
-The chart highlights the frequency of polling activities across different organizations, with Marist and Emerson standing out as the most prolific pollsters."
+#| fig-cap: "Horizontal bar chart displaying the distribution of polls conducted by various pollsters in key battleground states.
+Highlights the frequency of polling activities across different organizations, with Marist and Emerson standing out as the most prolific pollsters."
 
@@
 #| label: fig-start-date-distribution
-#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states leading up to the 2024 United States Presidential Election.
-The chart highlights the frequency of polling activities initiated over time, with a notable peak on September 14, indicating a surge in polling efforts as the election approached."
+#| fig-cap: "Histogram illustrating the distribution of polling start dates for key battleground states.
+Displays the frequency of polling activities initiated over time, with a notable peak on September 14, indicating a surge in polling efforts as the election approached."
 
@@
 ## Model Justification
-
-The Bayesian Beta Regression model with a logit link function is the most suitable model for our prediction of the 2024 U.S.
-Presidential Election because it aligns closely with both the nature of the data and the goals of our paper, which include understanding voter support trends over time and accounting for regional differences across battleground states.
-
-Our primary outcome variable is the percentage of support for each candidate in 7 key battleground states, scaled to a proportion between 0 and 1.
-Support percentages are naturally bounded between 0% and 100%, and this bounded nature poses challenges for traditional linear regression, which assumes an unbounded continuous outcome.
-The Beta distribution is ideal for modeling proportion data because it is inherently defined on the interval (0, 1).
-This allows us to model the proportions of support without the risk of generating unrealistic values outside this range.
-Unlike a linear model, which could predict support percentages exceeding 100% or falling below 0%, the Beta distribution respects the boundaries of our data.
-The logit link function further ensures that the model predicts in terms of log odds, which is suitable for interpreting support as probabilities—an essential consideration given our goal of forecasting potential victory in each state.
-
-One of our key goals is to examine how support for each candidate evolves over time as Election Day approaches.
-Polling data often show non-linear trends, with fluctuations in support that may correspond to campaign events, debates, or shifts in the political landscape.
-Using natural splines with end_date_num (the number of days since the earliest poll date) allows the model to capture these non-linear trends in a flexible way.
-By including natural splines with 4 degrees of freedom, we allow the model to capture gradual changes in voter sentiment while avoiding overfitting to noise or abrupt fluctuations in the data.
-This is particularly important in the context of election polling, where voter opinions can change either slowly or rapidly depending on unpredictable external events.
-Alternative models, such as standard logistic regression without splines, would fail to capture these changes, potentially leading to a less accurate depiction of how support varies over time.
-
-A unique aspect of our analysis is the focus on battleground states, which have distinct political dynamics that can heavily influence election outcomes.
-Incorporating state as a random effect provides flexibility in estimating state-specific baseline support levels for each candidate, while still allowing us to identify overall trends across all states.
-This approach is more effective than treating state as a fixed effect because it avoids overparameterizing the model with individual state coefficients, which could lead to overfitting in cases with limited data for some states.
-The random effect structure also helps generalize the model’s predictions, making it more robust when predicting support in battleground states with varying polling frequencies.
-
-The Bayesian approach offers advantages for our model by allowing us to incorporate prior information, enhancing the stability and reliability of our estimates.
-Given that polling data can be sparse for certain states or dates, Bayesian inference helps stabilize parameter estimates, especially when there are limited data points for certain states or specific dates.
-By using weakly informative priors, we let the data primarily determine the model’s outcomes, while still providing some structure to prevent overfitting or extreme parameter values.
-The Bayesian framework also produces credible intervals, which provide a clear interpretation of uncertainty around our predictions.
-In the context of election forecasting, this is valuable as it enables us to quantify the range of plausible outcomes for each candidate’s support, adding a layer of transparency to our predictions.
-This model provides interpretability that aligns well with the goals of our paper.
-By structuring the model with a logit link function, we can interpret the results in terms of odds or probability of victory, which is highly relevant in an election context.
-This interpretability allows us to understand the likelihood of each candidate winning in a specific state at a given time.
-
-Additionally, the Bayesian Beta regression model offers flexibility in how we examine each candidate’s trends individually.
-By constructing separate models for Trump and Harris, we can directly compare their temporal trends and state-level variations without confounding effects.
-
-Overall, the Bayesian Beta regression model with a logit link function is the most suitable model for our study.
-It allows for flexible non-linear trends over time, accounts for regional variation across battleground states, and provides interpretable results in terms of victory probabilities.
-The Bayesian framework stabilizes estimates in the presence of sparse data and provides meaningful uncertainty estimates.
-The model’s structure aligns well with our goal of forecasting each candidate’s probability of victory, justifying the choice for achieving a comprehensive analysis and prediction of the 2024 U.S.
+The Bayesian Beta Regression model with a logit link function is the most suitable model for predicting voter support for Kamala Harris and Donald Trump in the 2024 U.S.
 Presidential Election.
+Our primary outcome variable is the percentage of support for each candidate in seven key battleground states, which is naturally bounded between 0% and 100%.
+The Beta distribution is ideal for modeling this proportion data, as it allows us to avoid unrealistic predictions outside the (0, 1) range.
+To capture non-linear trends in voter support as Election Day approaches, we employe natural splines with **end_date_num**, which represents the number of days since the earliest poll date.
+Additionally, incorporating state-specific random effects acknowledges the variations in voter preferences across different regions, allowing for more accurate modeling of baseline support levels.
+The Bayesian framework enhances our model by enabling the incorporation of prior information, leading to more stable estimates, especially in the context of sparse polling data.
+Furthermore, the model's structure allows for clear interpretation in terms of victory probabilities, aligning well with our goals of forecasting electoral outcomes.
+Further details are discussed in [Appendix -@sec-model-details].
 
 ## Model Summary
-
 ### Trump Model Summary
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-trump-fixed-effects
+@tbl-trump-model-summary fixed effects estimates for several key terms that capture the influence of time on his support in the battleground states.
+The intercept, with an estimate of -0.0923, represents the baseline log-odds of Trump’s support when other predictors are held constant.
+This negative intercept suggests a slightly lower baseline level of support.
+The term nsenddatenumdfEQ42, with an estimate of 0.0803 and a statistically significant positive confidence interval (0.0179 to 0.1426), indicates an increase in support around this time period.
+Similarly, nsenddatenumdfEQ44 has an estimate of 0.0789 and a confidence interval above zero (0.0014 to 0.1589), also showing a positive effect on support.
+Other terms such as nsenddatenumdfEQ41 and nsenddatenumdfEQ43 have less significant impacts, suggesting that certain time periods were more influential than others in shaping Trump’s support.
+The standard errors for the fixed effects terms range from 0.0306 to 0.0676 for nsenddatenumdfEQ43.
+The relatively low standard error of 0.0306 on the intercept suggests that the baseline estimate is quite stable, meaning we can be more confident about this baseline support level.
+However, the standard error for nsenddatenumdfEQ43 (0.0676) is higher than for other terms, indicating that this estimate is less precise.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-trump-model-summary
 #| tbl-cap: "Fixed Effects Summary for Trump Model."
@@
 ```
+These results imply that support for Trump varied over time in ways that may correspond to specific campaign events or other influences affecting voter sentiment.
+The significance of terms like nsenddatenumdfEQ42 and nsenddatenumdfEQ44 suggests that changes in voter preference could be captured effectively by this model.
+And by checking standard errors, helping us determine how much we can rely on those data.
 
 ### Harris Model Summary
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-harris-fixed-effects
+@tbl-harris-model-summary similarly shows the impact of time on her support in the battleground states.
+The intercept for the Harris model is estimated at -0.1508, suggesting a slightly lower baseline support level compared to Trump(-0.0923).
+The term nsenddatenumdfEQ41 has an estimate of 0.1225 and a confidence interval 0.0553 to 0.1903, indicating a significant increase in support during this period.
+Another positive and significant effect is seen in nsenddatenumdfEQ43, with an estimate of 0.1191 and confidence interval from -0.0071 to 0.2417, which suggests a trend toward increased support.
+While other terms stay positive, they have less impacts on support.
+Additionally, the standard errors for fixed effects are generally smaller, ranging from 0.0279 for the intercept to 0.0631 for nsenddatenumdfEQ43.
+So the larger standard error for nsenddatenumdfEQ43 (0.0631) suggests that there would be some uncertainty in how this time point influences support, though the positive estimate still points toward an upward trend.
+In comparison to the Trump model, the standard errors in the Harris model are slightly smaller for most terms, indicating potentially more consistent support trends over time in the data analyzed.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-harris-model-summary
 #| tbl-cap: "Fixed Effects Summary for Harris Model."
@@
 ```
+These results highlight that Harris's support also fluctuated over time, with nsenddatenumdfEQ41 and nsenddatenumdfEQ43 marking periods of notable increases.
+Helping us understand the pattern, this model predicts Harris's probability of victory and provides insights for future changes in support as the election nears.
 
@@
 # Discussion {#sec-discussion}
-## Interpretation of Results {#sec-results-interpretation}
-Since the electoral college system makes state-level forecasts crucial, and the U.S.
-operates under a predominantly two-party system, the focus of election predictions should be on battleground states.
-While many states are strongly aligned with either the Democratic or Republican party, a handful of key battleground states—where neither party has overwhelming dominance—will likely determine the overall outcome of the election.
-These battleground states, often evenly split in voter sentiment, hold immense significance because their electoral votes can swing the election in favor of one candidate or the other.
-
-The results of our model underscore the critical role that these battleground states will play in the 2024 election.
-Although many states are predictably stable due to their party loyalty, it is the six battleground states that will serve as the main arena for competition.
-In these states, where voter preferences are not as firmly entrenched, even minor shifts in public opinion or turnout could lead to vastly different outcomes.
-As a result, a comprehensive and precise prediction of voting behavior in these battleground states is far more important than in states where the result is a foregone conclusion.
-
-## Model Performance {#sec-model-performance}
-Out-of-sample testing results [Appendix -@sec-model-diagnostics] demonstrate the effectiveness and reliability of our Bayesian logistic regression model in predicting state-level outcomes for the 2024 U.S.
-presidential election.
-Our accuracy test generated a 75% score for the model, indicating our model's high reliability for making predictions about state-level election outcomes.
-Since presidential elections hinge on state-level wins, a reliable state-by-state prediction would imply a reliable national outcome prediction.
-
-### Model Performance - Trump Model
-The RMSE value generated for our model is 0.5, which indicates that the model-predicted probabilities are close to the true outcomes.
-The confusion matrix breaks down the model’s predictions into four categories: 1.
-orrectly predicted Trump wins, 2.
-Correctly predicted Harris wins, 3.
-Incorrectly predicted Trump wins, 4.
-Incorrectly predicted Harris wins.
-The generated confusion matrix for our model all had values of 0 to 2 for all four categories, indicating a balanced prediction between the two candidates, and suggesting the reliability of our model.
-Out-of-sample testing was conducted to test the reliability of our model, with further details provided in [Apppendix -@sec-model-diagnostics].
-
-### Model Performance - Harris Model
-
-## Weaknesses and Next Steps
-While our model provides valuable insights into the electoral landscape, it faces limitations due to missing data for certain states, which results in some states being left blank in our predictions.
-Although our primary focus is on understanding the dynamics within battleground states, the lack of information from other states can lead to incomplete predictions which may obscure broader trends affecting voter sentiment.
-
-To address these weaknesses and improve the prediction, we propose several next steps: actively seeking additional data sources to fill gaps, including polling data and historical records; conducting sensitivity analyses to understand the impact of missing data on predictions; and analyzing existing trends within battleground states for more insights.
-
-Another weakness of our model is the lack of consideration for the national vote, which may impact our election predictions.
-The national vote refers to the total number of votes cast by citizens across the country during an election, representing the support for each presidential candidate.
-Although the national vote does not directly determine the outcome of the election due to the Electoral College system, it can still provide valuable insights into voter sentiment and trends.
-By not considering the national vote into our analysis, we risk missing potential influences on voting behavior and electoral outcomes.
+
+## Key Findings {#key-findings}
+This paper presents a forecast of the 2024 U.S, Presidential Election, leveraging Bayesian Beta regression models trained on polling data.
+We were able to model both the electoral support trend over time and the different political dynamics across the battleground states.
+The model’s predictions reveal that, as Election Day approaches, Kamala Harris holds a lead in six of the seven battleground states–Pennsylvania, Michigan, Wisconsin, Nevada, North Carolina, and Georgia.
+However, her margins are consistently narrow, with predicted probabilities only slightly above 50% in each of these states.
+Harris shows minimal leads over Trump, with differences between the probability of winning of no more than 3 percent across all six states.
+Meanwhile, Trump also holds a narrow advantage in Arizona, suggesting a highly competitive electoral dynamic in this election.
+
+The small margins in the predicted outcomes emphasize the potential volatility of the race; a slight shift in voter sentiment or late-breaking events could easily flip the outcome in these battleground states.
+This narrow margin of victory suggests a level of electoral fragility for Harris, as her support base may not be solidly locked in.
+The close race in each state implies that both candidates must maintain their voter engagement and strategic focus, as any shift in these states could have significant implications for the election’s overall outcome.
+Thus, while Harris is favored in the majority of battleground states, the slight leads indicate a need for cautious interpretation, with both campaigns facing high stakes in these key states.
+
+## Real World Implications
+If Kamala Harris wins the election by a narrow margin, the results could have significant real-world implications, reflecting the deeply polarized state of the nation.
+A close race indicates not only a tight political divide but also underscores broader cultural, economic, and social differences among Americans, often manifested through political affiliations.
+One major challenge for Harris would be governing with such a divided mandate; she could face resistance from nearly half the population, complicating her efforts to implement her agenda.
+This suggests that any significant policy shifts may encounter strong opposition, potentially deepening divisions rather than bridging them (Yourish and Gamio, 2024).
+
+Moreover, a narrow victory for Harris would reinforce national polarization, revealing the extent to which the country is split.
+The nearly equal support for both candidates highlights entrenched ideological divides, with voters committed to opposing perspectives on critical issues such as healthcare, immigration, climate policy, and the role of government (Heltzel & Laurin, 2020).
+Economic concerns, including job security, wage equality, and tax policy, alongside cultural values related to family, education, and identity, play significant roles in voter alignment.
+Thus, a close election result suggests that these differences are sharply felt across the electorate, with each side viewing the outcome as an opportunity to uphold their values and address their unique challenges.
+
+## Uncertainty Beyond Data and Modeling
+There are external factors other than data and modeling that could influence the actual outcome of the election. These factors are often hard to control and predict.
+
+Firstly, recency bias influences public perception and even polling responses.
+Many people expect Trump to win because of his unexpected victory in 2016 and his close race in 2020, despite lagging in the polls.
+This bias conveys the tendency to focus on recent, high-profile events while overlooking historical trends, such as Obama’s strong polling performance in 2012, where he outperformed polls.
+This bias can affect how people respond to polls or even how pollsters interpret results, as certain events may be weighted more heavily than warranted.
+As our model depends on polling data, any systematic bias introduced by recency expectations can skew our predictions, making it challenging to account for potential polling errors.
+
+Social desirability bias and non-response bias are additional sources of uncertainty, particularly with Trump’s supporters.
+Some voters may be hesitant to disclose their support for a conservative candidate due to social stigma, leading to underreporting of support in polls.
+This reluctance can be especially strong in areas where conservative views are less socially accepted.
+Trump’s support base also includes individuals with lower social trust and civic engagement, which means they are less likely to participate in polls altogether.
+If certain groups systematically avoid participating, polling data may underrepresent these voters, resulting in an underestimation of Trump’s support in our model.
+This non-response bias is difficult to quantify, adding a layer of uncertainty to our predictions.
+On the other hand, Harris faces potential biases such as the Bradley effect and gender-related bias.
+As a Black female candidate, Harris might encounter the Bradley effect, where some voters express indecision rather than explicitly stating they won’t support a minority candidate.
+Historically, this effect has impacted Black candidates, as undecided voters tend to lean against them in the actual vote count, despite polling responses suggesting neutrality.
+Additionally, Clinton’s 2016 campaign demonstrated that undecided voters can heavily swing against female candidates.
+This trend could similarly impact Harris’s polling and election results.
+If these voters decide against Harris in the end, our model might overestimate her support based on the polling data, leading to inaccuracies in predictions.
+These biases are inherently difficult to measure or adjust for, as they reflect deeper societal attitudes and individual decision-making processes.
+Pollsters may attempt to correct for non-response or social desirability bias, but such adjustments are often based on assumptions that may not fully capture the complex social dynamics at play.
+Consequently, our model’s predictions should be interpreted with caution, acknowledging that these biases can lead to unexpected shifts in voter behavior that fall outside the range of our data-driven forecasts.
+
+## Limitations and Weaknesses
+Although the model is suitable and performs well on achieving the goal of predicting the outcome, it has several limitations that could be addressed in future studies. Here are some key areas:
+
+- **Pollster Ratings and Credibility**: Polls from different organizations often vary in reliability, as some pollsters have more accurate or less biased methodologies.
+By integrating pollster ratings, the model could assign weights to different polls based on the historical accuracy and reputation of each organization.
+This adjustment would allow the model to downplay the influence of potentially biased or less credible polls, particularly in competitive battleground states, where small differences in predicted support can significantly affect the outcome.
+This approach would lead to a more calibrated model that accounts for poll variability and thus provides a more reliable prediction.
+- **Timing and Recency of Polls**: Voter sentiment is not static, and the timing of polls can significantly impact their relevance.
+As the election date nears, polls generally become more reflective of the final outcome.
+Incorporating a temporal weighting scheme that prioritizes recent polls could enhance the model’s responsiveness to last-minute shifts in voter preferences, which often occur due to campaign events, debates, or emergent political issues.
+This weighting approach could also help mitigate the risk of using outdated data that no longer accurately reflects current voter sentiment.
+- **Economic and Political Context**: The political landscape is shaped by various external factors, such as economic indicators (e.g., unemployment rates, inflation) and prominent political events.
+By incorporating economic data or major political issues, such as health care or immigration policies, the model could capture contextual variables that influence voter sentiment.
+Such contextual factors could be especially useful in a Bayesian framework, allowing us to adjust prior distributions based on known historical influences on voter behavior.
+- **Inclusive of National/Popular votes**: our model has exclusive focus on the seven battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+By narrowing the dataset to only these critical swing states, our model omits data from the rest of the nation, including states with predictable voting patterns (solidly Democratic or Republican) and states outside the battleground category.
+This omission is significant because it restricts the model’s capacity to capture the broader national voting landscape and limits the ability to provide a holistic prediction of the 2024 U.S.
+Presidential Election outcome.
+In particular, the exclusion of national popular vote data limits our model’s understanding of overall voter sentiment across the country.
+While battleground states are pivotal in determining the electoral outcome due to the Electoral College system, popular vote trends from non-battleground states can still reveal important national shifts in sentiment that may indirectly influence battleground dynamics.
+For example, a nationwide trend towards or away from a candidate could suggest emerging political momentum or reflect the impact of major events that also resonate within battleground states.
+Without incorporating this popular vote data, our model may overlook broader trends that could subtly shape battleground results, especially if voters in battleground states are influenced by national sentiment.
+
+Moreover, data from non-battleground states could add valuable context and enhance model robustness by reflecting a range of voter behaviors and regional influences.
+While these states may not directly impact the electoral outcome due to their predictability, understanding regional dynamics across all states could help the model detect shifts in voter sentiment that cut across state lines.
+For instance, trends in nearby states or those with similar demographic or economic characteristics might mirror those in battleground states, providing additional insights into potential voting patterns.
+Including these states in the model could also reveal hidden correlations and strengthen the model’s predictive power by capturing a wider range of political and demographic factors.
+The focus on battleground states also means that the model may amplify the idiosyncrasies of these seven states without balancing them against broader national trends.
+Each battleground state has unique political, economic, and demographic characteristics, and relying solely on these states could cause the model to disproportionately weigh regional dynamics that may not be representative of the country as a whole.
+As a result, predictions based only on these states might overestimate or underestimate certain voting behaviors, particularly if the election outcome depends on trends that are not captured within these seven states alone.
+
+## Implications for Future Modeling
+For future modeling, incorporating data from a broader range of states or using a weighted approach that balances battleground states with national polling data could address these limitations.
+For example, incorporating popular vote trends and data from states with similar demographics or political contexts could help the model detect and account for national sentiment shifts.
+Additionally, adding more factors, such as ratings for pollsters, demographic details, and economic conditions, could help the model capture different influences on how people vote.
+Giving more weight to recent polls would also make the model more responsive to last-minute changes in public opinion as Election Day gets closer.
+By expanding the data and including a wider range of details, future models could better balance the unique characteristics of battleground states with trends across the nation, resulting in a stronger and more reliable prediction tool for elections.
 
@@
 # Additional Data Details {#sec-data-details}
-Given the nature of the presidential election, the electoral college is more impactful and influential on the possible outcome compared to the nationwide popular votes.
-Therefore, the majority of the pollsters choose to focus on the battleground states to perform their polling, to gain more insight into the public sentiment specific to certain states.
-For instance, Sienna/NYT mentions that they have in-state interviewers who call their respondents in states such as Arizona, Georgia, Michigan, Nevada, Pennsylvania, and Wisconsin.
-
-This could explain the discrepancies in our data, where there are many missing values from states other than the battleground ones.
-Furthermore, many pollsters focus on state-wide polls instead of nationwide polls given the nature of the election, resulting in fewer data points or missing data for nationwide poll results.
-Even though the national popular vote and the results in the electoral college don’t line up a lot of the time, the popular votes can still provide information about issues and opinions that are shaping the election as a whole.
-Therefore, even if we are focusing on selective states, results from other states and popular votes still remain relevant to outcome prediction.
-
-During our data cleaning process, we added constraints of high numeric grades and high transparency scores, leading to fewer data points for us to build the model on.
-A numeric grade is a numeric rating given to the pollster indicating their quality or reliability, with a highest rating of 3.0.
-The transparency score is a score reflecting the pollster’s transparency about their methodology where the highest score is 10.
-Therefore, to ensure high-quality and reliable data to build our model on, we filtered pollsters with numeric grades of 3.0 and transparency scores of 5 or higher.
-This is one of the reasons why there are no data points available for Kamala Harris for the state of Colorado, represented by the empty row @fig-averagetable.
-
-Furthermore, in the full raw data set from 538 [@FiveThirtyEight], for the state of Colorado, even without any constraints for numeric grade or transparency scores, there are only 3 data points for Kamala Harris.
-After reviewing the datasets, we found that except for the 3 available data points, all the rest of the polls ended in June 2024, which is a month earlier than when Biden dropped out of the race before Harris entered the presidential election race.
-It would not be possible that there are data points available for Harris if she was not participating in the race yet.
-The state of Colorado is a blue state, with the Democrats winning the state in every presidential election since 2008.
-Since our paper focuses on mainly the battleground states to predict election outcomes, limited data from a blue state like Colorado would not be too influential on our results.
-
-# Model details {#sec-model-details}
+## Data Cleaning {#sec-data-cleaning}
+The data cleaning process for our analysis focused on refining the raw polling data obtained from Project 538's polling project [@FiveThirtyEight].
+The raw data contains polling information for various U.S.
+states, but for our specific analysis, we concentrated on seven designated battleground states: Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+This is due to the fact that these battleground states have been recognized as pivotal in determining this year's election outcome.
+
+To prepare the data for modeling, we performed several key steps.
+First, we filtered the data to retain only relevant columns, including state, polling dates, pollster information, numeric grades, transparency scores, candidate names, and their respective support percentages.
+This filtering process ensured that we excluded any unnecessary data that could complicate our analysis.
+Next, we implemented thresholds for the numeric grade (minimum of 2.5) and transparency score (greater than 5) to ensure that only high-quality polling data was considered.
+We also converted polling dates into a standardized date format and created a new numeric variable representing the number of days since the earliest polling date.
+Additionally, we handled specific data inconsistencies, such as replacing "Nebraska CD-2" with "Nebraska." The polling percentages were then aggregated by averaging the values for each candidate while preserving other important attributes such as pollster and state.
+This aggregation allowed us to focus on polling data that represented a balanced view of voter support across the selected battleground states.
+The cleaned dataset was then saved in a parquet format for efficient processing in subsequent analyses.
+
+To ensure the integrity of our cleaned polling dataset, we implemented a comprehensive data testing process.
+This included verifying the presence of all required columns, checking data types for consistency, and ensuring there were no missing values in critical fields.
+We validated that candidate support percentages fell within the logical range of 0 to 100 and confirmed that numeric grades and transparency scores met predefined thresholds.
+By rigorously applying these tests, we ensured that the dataset is reliable and suitable for subsequent modeling and analysis, providing a solid foundation for our exploration of voter behavior in the battleground states.
+
+## Limitations and Future Directions
+Xiyan
+
+# Additional Model details {#sec-model-details}
+## Model Justification - Further Details
+The Bayesian Beta Regression model with a logit link function is particularly suited for our prediction of the 2024 U.S.
+Presidential Election due to its alignment with the nature of our outcome data and the goals of our analysis.
+Our primary outcome variable is the percentage of support for each candidate in the seven battleground states, which we scaled to a proportion between 0 and 1.
+Since support percentages are naturally bounded between 0% and 100%, modeling these data with traditional linear regression presents challenges, as linear regression assumes an unbounded continuous outcome.
+The Beta distribution is ideal for this type of data because it is defined on the interval (0, 1), allowing us to model proportions without the risk of generating unrealistic values outside this range.
+Unlike a linear model, which could predict support percentages exceeding 100% or falling below 0%, the Beta distribution respects the boundaries of our data.
+
+To effectively capture the dynamics of voter support, we utilized natural splines with the variable end_date_num, which represents the number of days since the earliest poll date.
+This choice allows the model to adapt to non-linear trends in voter support that can fluctuate due to various campaign events, debates, or shifts in the political landscape.
+By specifying four degrees of freedom for the natural splines, we enable the model to flexibly capture gradual changes in voter sentiment while mitigating the risk of overfitting to noise or abrupt fluctuations in the data.
+This flexibility is particularly critical in electoral polling, where opinions may shift slowly or rapidly in response to unpredictable external events.
+Alternative models, such as standard logistic regression without splines, would fail to capture these complex changes, potentially leading to a less accurate depiction of how support varies over time.
+
+A unique aspect of our analysis is the focus on battleground states, which exhibit distinct political dynamics that can significantly influence election outcomes.
+Incorporating state as a random effect provides the flexibility necessary to estimate state-specific baseline support levels for each candidate while still allowing us to identify overall trends across all states.
+This approach is more effective than treating state as a fixed effect, as it avoids overparameterizing the model with individual state coefficients, which could lead to overfitting, particularly in cases with limited polling data for certain states.
+The random effect structure enhances the model’s generalizability, making it more robust when predicting support in battleground states with varying polling frequencies.
+
+The Bayesian approach offers substantial advantages for our model by allowing us to incorporate prior information, which enhances the stability and reliability of our estimates.
+Given that polling data can be sparse for certain states or dates, Bayesian inference helps stabilize parameter estimates, especially when limited data points are available.
+We defined weakly informative priors for the model parameters: for the intercept $\beta_0$, we specified $\beta_0 \sim \text{Normal}(0, 10)$ to provide a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds; for the slope $\beta_1$, we used $\beta_1 \sim \text{Normal}(0, 5)$, reflecting reasonable expectations regarding the influence of polling trends over time.
+
+This model structure not only allows for capturing the dynamics of voter support but also provides interpretability that aligns well with the goals of our paper.
+The logit link function enables us to interpret results in terms of odds or probabilities of victory, which is crucial in the election context.
+By constructing separate models for Trump and Harris, we can directly compare their temporal trends and state-level variations without confounding effects.
+Thus, the Bayesian Beta regression model with a logit link function is justified as the most suitable model for our study.
+It accommodates flexible non-linear trends over time, accounts for regional variation across battleground states, and delivers interpretable results concerning victory probabilities.
+The Bayesian framework not only stabilizes estimates in the presence of sparse data but also provides meaningful uncertainty estimates, ultimately aligning with our objective of forecasting each candidate’s probability of victory in the 2024 U.S.
+Presidential Election.
 
 ## Model Diagnostics {#sec-model-diagnostics}
-In order to assess the validity and reliability of our Bayesian models for predicting voter support for Kamala Harris and Donald Trump, we conducted posterior predictive checks for both models.
-Posterior predictive checks are a crucial diagnostic tool that allows us to evaluate how well our models can replicate observed data based on the parameters estimated during the modeling process.
-By comparing the distribution of the observed support percentages with the predicted support values, we can identify potential discrepancies and gauge the overall fit of our models.
-
-These checks not only help us verify whether the model assumptions hold but also provide insights into how well the models capture the underlying data-generating processes.
-A strong alignment between observed and predicted values would suggest that our models are robust and capable of accurately reflecting the dynamics of voter support in the current electoral context.
-Conversely, significant deviations may indicate areas for model refinement or highlight limitations in our current approach.
-Through these diagnostics, we aim to ensure the credibility of our findings and enhance our understanding of the electoral landscape as we approach the 2024 Presidential Election.
+To assess the validity and reliability of our Bayesian models for predicting voter support for Kamala Harris and Donald Trump, we conducted posterior predictive checks for both models.
+These checks are essential for evaluating the models' ability to replicate observed polling data based on the estimated parameters.
+By comparing the distributions of observed support percentages with the predicted values, we can identify discrepancies and assess the overall fit of our models.
+
+Posterior predictive checks help verify model assumptions and provide insights into how well the models capture the underlying voter dynamics.
+A close alignment between observed and predicted values suggests that our models effectively reflect voter support trends in the electoral context.
+In contrast, significant deviations may reveal areas for model improvement or highlight limitations in our approach.
+This diagnostic process aims to enhance the credibility of our findings and deepen our understanding of the electoral landscape as the 2024 Presidential Election approaches
 
@@
 ### Model Diagnostics - Harris Model
-The posterior predictive check for the Harris model is presented in Figure X, illustrating the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
 @fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
@@
 ```
-
-## Posterior Predictive Check for the Harris Model
-The posterior predictive check for the Harris model is presented in Figure X, illustrating the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
-This visualization features a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
-The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
-
 In the plot, we observe a strong overlap between the $y$ and $y_{\text{rep}}$ distributions, with the dark blue curve representing the observed support percentages primarily concentrated around 48% to 52%.
@@
 Such limitations are also discussed in @sec-model-diagnostics.
+
+### Model Performance
+Based on the posterior predictive checks performed for both Trump and Harris models, we are able to make the following conclusions on the performance and reliability of our Bayesian regression models.
+The strong overlap of the observed distribution of support percentages and the predicted distribution from posterior samples suggests that the models are well-calibrated, meaning they accurately reflect the variability and central tendency of voter support for Trump and Harris, respectively.
+This alignment is particularly important given the fluctuating nature of polling data and the importance of accurately capturing voter sentiment as we approach Election Day.
+
+In the context of our paper, these predictive checks provide evidence of the model’s robustness and reliability in forecasting.
+Given our goal of predicting the likelihood of each candidate’s victory in key battleground states, it is important that our models can not only predict the general trend in support but also approximate the changes in voter sentiment captured by the polling data.
+The successful predictive checks validate our choice of a Bayesian Beta regression model with a logit link function, demonstrating that this approach is indeed suitable for analyzing proportion data constrained within the 0-1 range.
+
+Furthermore, these results support the use of state-level random effects and time-based splines, as the models are shown to accommodate both state-specific variations and temporal shifts in support.
+By accurately modeling these elements, our predictions provide more reliable insights into which candidate holds an advantage in each state.
+This model performance enhances confidence in our forecasts, suggesting that our predictions can serve as a meaningful tool for understanding electoral dynamics in these critical regions.
+These findings suggest the model’s capacity to provide robust, data-driven projections of the election outcome, while also accounting for the inherent uncertainty and volatility within polling data.
 

### 2024-11-04 paper/paper.qmd
commmitting newly rendered paper.qmd

@@
 # Introduction {#sec-intro}
-The 2024 United States Presidential Election is anticipated to be a pivotal moment in American political history, particularly as battleground states emerge as key determinants of the electoral outcome.
-With candidates Donald Trump and Kamala Harris intensifying their campaigns, a thorough understanding of voter preferences in these critical regions becomes essential for forecasting election results.
-This paper leverages state-level polling data from Project 538 [@FiveThirtyEight] and employs the statistical programming language R [@citeR] to construct a predictive model that estimates the likelihood of victory for each candidate across several crucial battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-
-The primary estimand of this study is the probability of winning for either Donald Trump or Kamala Harris in these battleground states, derived from aggregated polling averages.
-Our analysis utilizes a Bayesian hierarchical approach combined with natural splines to accurately capture the temporal dynamics of voter sentiment, enabling us to estimate winning probabilities in these decisive areas.
-By comparing the predicted winning probabilities for both candidates, we can ascertain the likely winner based on prevailing voter support trends.
-Our findings indicate a competitive landscape, with Kamala Harris projected to secure victories in six of the seven battleground states analyzed, including key states like Michigan and Nevada, where voter sentiment appears to favor her candidacy.
-However, Donald Trump is also predicted to maintain a competitive edge in Arizona, showcasing the variability in voter preferences across these critical regions.
-The predicted winner is determined by comparing the estimated probabilities for each candidate, highlighting Harris’s overall advantage leading into the election.
-
-Understanding these predictions is crucial as they not only forecast the likely outcome of the U.S.
-election but also help society and the economy to better understand the current state of the US political climate.
-The implications of these findings extend beyond mere predictions, providing a framework for analyzing how shifts in voter sentiment may occur as the election draws near.
+The 2024 United States Presidential Election is set to be a crucial moment in American political history, with battleground states serving as pivotal determinants of the electoral outcome.
+As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions is essential for accurate election forecasting.
+This paper utilizes state-level polling data from FiveThirtyEight [@FiveThirtyEight] and the statistical programming language R [@citeR] to develop a predictive model estimating each candidate's likelihood of victory in critical battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+
+The primary focus of this study is the probability of victory for Trump or Harris in these battleground states, derived from aggregated polling averages.
+Employing a Bayesian hierarchical model combined with natural splines, our analysis effectively captures the temporal dynamics of voter sentiment, allowing us to estimate winning probabilities in these decisive areas.
+Our findings reveal a competitive landscape, with Harris projected to win in six of the seven battleground states analyzed, notably in Michigan and Nevada, where support leans in her favor.
+However, Trump remains competitive in Arizona, illustrating the variability of voter preferences across these regions.
+
+These predictions are critical not only for forecasting the likely outcome of the election but also for enhancing societal and economic understanding of the current political climate in the U.S.
+The implications of our findings extend beyond mere predictions, offering a framework for analyzing potential shifts in voter sentiment as Election Day approaches.
 The remainder of this paper is organized as follows: Section @sec-data details the data and variables utilized in the analysis; Section @sec-model outlines the model setup and estimation strategies employed; Section @sec-results presents the results, including visualizations of the predicted election outcomes; and Section @sec-discussion discusses the implications and limitations of our findings.
@@
 It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
-For every poll conducted, relevant variables such as state, start_date, end_date, transparency_score, candidate_name, and percentage of support (pct) for the appropriate candidates are included.
-The dataset compiles data from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
+For every poll conducted, relevant variables such as state, start_date, end_date, transparency_score, candidate_name, and percentage of support (**pct**) for the appropriate candidates are included.
+The data compiles information from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
 
@@
 For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates.
-We collect observations from the dataset where the ‘candidate_name’ is Trump and Harris.
+We collect observations from the data where the ‘candidate_name’ is Trump and Harris.
 We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’, and ‘state’, as well as the electorate support for each candidate is denoted as ‘pct’.
-Furthermore, since there are many pollsters present in the dataset, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [Appendix -@sec-data-details].
+Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [Appendix -@sec-data-details].
 
@@
 In effect, the complexity of each opinion is smoothed over into a uniform data point, with many variations of opinions lost in the process.
-
 Finally, these support percentages are fed into a model to make a predictive statement about the election outcome.
@@
 However, this step relies heavily on simplification, assuming that the snapshot of opinions captured in the polls will hold steady and that all voters will follow through on their expressed preferences.
-
 The raw polling data gives us percentages of support for each candidate in each state.
@@
 
-It is important to note that since ‘pct’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
+It is important to note that since ‘**pct**’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
 For instance, in the summary statistics presented in Table 1, the average support for Donald Trump is 47.801%, while the average support for Kamala Harris is 47.999%.
@@
 ### Predictive Variables
-In this section, we outline the two key predictor variables used in our analysis: 'state', which captures the geographical context of the polling data, and 'end_date', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.
+In this section, we outline the two key predictor variables used in our analysis: '**state**', which captures the geographical context of the polling data, and '**end_date**', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.
 
 #### State
-The state ('state') variable indicates the U.S.
+The state (**state**) variable indicates the U.S.
 state where the poll was conducted or targeted.
@@
 #### End Date {#sec-end-date}
-End date ('end_date') represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
+End date ('**end_date**') represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
 In the context of our paper, the end date of each poll serves as one of the predictive variables and is used to assess changes in voter support over time.
@@
 
-\newpage
-
 ### Other Variables
-This section details additional variables that provide essential context for our analysis, including numeric grade, which evaluates the quality of the polls; transparency score, indicating the openness of poll methodologies; pollster, identifying the organizations conducting the polls; and start date, which denotes the initiation of the polling period for each survey.
+This section details additional variables that provide essential context for our analysis, including **numeric grade**, which evaluates the quality of the polls; **transparency score**, indicating the openness of poll methodologies; **pollster**, identifying the organizations conducting the polls; and **start date**, which denotes the initiation of the polling period for each survey.
 
 #### Numeric Grade
-Numeric Grade ('numeric_grade') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
+Numeric Grade ('**numeric_grade**') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
 In the context of our paper, numeric grade allows us to weigh the reliability of different polls.
@@
 This consideration is crucial, as polling quality can impact the stability of the model’s forecasts, especially in closely contested states where polling accuracy is essential for reliable predictions.
-\newpage
 
 #### Transparency Score
-Transparency Score ('transparency_score') represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
+Transparency Score ('**transparency_score**') represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
 This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error.
@@
 #### Pollster
-Pollster ('pollster') represents the various organizations responsible for conducting polls across key battleground states in the lead-up to the 2024 United States Presidential Election.
+Pollster ('**pollster**') represents the various organizations responsible for conducting polls across key battleground states in the lead-up to the 2024 United States Presidential Election.
 In the context of our paper, the diversity and frequency of polling by different pollsters are essential for understanding the reliability and variance in the data.
@@
 #### Start Date
-Start date ('start_date') represents the date when each poll began collecting responses.
+Start date ('**start_date**') represents the date when each poll began collecting responses.
 In the context of our paper, start date is a key variable as it allows us to focus on polling data collected after July 21, 2024—the date Kamala Harris officially announced her presidential campaign.
@@
 # Model {#sec-model}
+
+## Model Overview
 The Bayesian Beta regression models constructed in this paper aim to predict the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S.
@@
 ### Trump Model Summary
-@tbl-trump-model-summary fixed effects estimates for several key terms that capture the influence of time on his support in the battleground states.
+@tbl-trump-model-summary displays fixed effects estimates for several key terms that capture the influence of time on his support in the battleground states.
 The intercept, with an estimate of -0.0923, represents the baseline log-odds of Trump’s support when other predictors are held constant.
@@
 #| label: tbl-trump-model-summary
-#| tbl-cap: "Fixed Effects Summary for Trump Model."
+#| tbl-cap: "Coefficient estimates from the Bayesian Beta regression model for Donald Trump, showing the impact of the natural spline transformation of end date on voter support.
+The intercept indicates baseline log-odds, while spline terms reflect nonlinear trends in voter sentiment, with positive coefficients suggesting increased support as Election Day approaches."
 #| warning: false
@@
 Additionally, the standard errors for fixed effects are generally smaller, ranging from 0.0279 for the intercept to 0.0631 for nsenddatenumdfEQ43.
-So the larger standard error for nsenddatenumdfEQ43 (0.0631) suggests that there would be some uncertainty in how this time point influences support, though the positive estimate still points toward an upward trend.
+The larger standard error for nsenddatenumdfEQ43 (0.0631) suggests that there would be some uncertainty in how this time point influences support, though the positive estimate still points toward an upward trend.
 In comparison to the Trump model, the standard errors in the Harris model are slightly smaller for most terms, indicating potentially more consistent support trends over time in the data analyzed.
@@
 #| label: tbl-harris-model-summary
-#| tbl-cap: "Fixed Effects Summary for Harris Model."
+#| tbl-cap: "Coefficient estimates from the Bayesian Beta regression model for Kamala Harris, illustrating the effects of the natural spline transformation of end date on voter support.
+The intercept represents the baseline log-odds, while the spline terms indicate nonlinear trends in voter sentiment, with positive coefficients suggesting an increase in support as Election Day nears."
 #| warning: false
@@
   ) %>%
-  mutate(term = gsub("[^[:alnum:] ]", "", term))  # Clean up term names
+  mutate(term = gsub("[^[:alnum:] ]", "", term))
 
@@
 #| label: tbl-stateaverage
-#| tbl-cap: "Table displaying the percentage of support for Donald Trump and Kamala Harris across key battleground states in the lead-up to the 2024 United States Presidential Election.
-The table lists states with their respective support percentages, highlighting the competitive landscape as voters prepare to make their decisions on Election Day.
-Notably, Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in states like Michigan and Nevada.
-This data provides insight into the current polling dynamics and potential electoral outcomes in these critical states."
+#| tbl-cap: "Percentage of support for Donald Trump and Kamala Harris across key battleground states.
+Trump maintains a slight lead in states such as Arizona and Georgia, while Harris shows stronger support in 5 of 7 key battleground states."
 
@@
 #| label: fig-state-polling-average
-#| fig-cap: "Polling Averages for Trump and Harris Over Time in Battleground States."
+#| fig-cap: "Polling averages for Donald Trump (red) and Kamala Harris (blue) in battleground states from August to October 2024.
+Captures the varying fluctuations in average support percentages across Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, indicating the competitive dynamics of the election."
 #| warning: false
@@
   geom_line(aes(y = Average_Harris, color = "Harris")) +
-  labs(title = "Polling Averages for Trump and Harris Over Time in Battleground States", 
+  labs(
        y = "Average Support (%)", 
@@
   scale_color_manual(values = c("Trump" = "red", "Harris" = "blue")) +
-  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  # Set date format and break frequency
+  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +  
   theme_minimal() +
@@
     legend.position = "none",
-    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels
-    strip.text = element_text(size = 10),  # Control facet label size if needed
-    panel.spacing = unit(0.5, "lines"),    # Add spacing between facets
-    panel.grid.major = element_line(linewidth = 0.3)  # Adjust gridline thickness
+    axis.text.x = element_text(angle = 45, hjust = 1), 
+    strip.text = element_text(size = 10), 
+    panel.spacing = unit(0.5, "lines"),
+    panel.grid.major = element_line(linewidth = 0.3) 
   )
@@
 
-Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in @sec-model to predict the possible outcomes by incorporating the given data.
+Since it is difficult to conclude a definitive predictive outcome by observing the current data, we utilize our models, described in @sec-model, to predict the possible outcomes by incorporating the given data.
 By using the current data, the model can identify the state-specific trends thus generating more accurate predictions of the outcome.
-Relevant prediction results are provided below.
+Relevant prediction results are provided below in @fig-state-probabilities.
 
 ## State-Level Probability of Trump Winning on Election Day
-By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (November 5th, 2024).
+By incorporating our Bayesian models, we generate the predicted probabilities of winner for both Donald Trump and Kamala Harris in key battleground states, on Election Day (**November 5th, 2024**).
 @fig-state-probabilities provides a visual summary of these predicted probabilities.
@@
 #| label: fig-state-probabilities
-#| fig-cap: "Probability of Trump Winning by State on Election Day."
+#| fig-cap: "Predicted winning percentages for Donald Trump (red) and Kamala Harris (blue) in key battleground states on Election Day.
+Illustrates the estimated probability of victory for each candidate across Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, highlighting Harris's projected advantages in the majority of these states."
 #| warning: false
@@
   scale_fill_manual(values = c("red", "blue")) +  # Specify colors
-  labs(title = "Predicted Winning Percentages by Battleground State on Election Day",
+  labs(
        x = "Winning Probability (%)",
@@
 #| label: tbl-state-probabilities
-#| tbl-cap: "Probability of Trump Winning by State on Election Day."
+#| tbl-cap: "Predicted winning probabilities for Donald Trump and Kamala Harris in key battleground states, with percentages representing the likelihood of each candidate's victory.
+Summarized predicted support for each candidate in Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin, indicating a competitive race, particularly with Harris expected to win in most states by narrow margins."
 #| warning: false
@@
 
+According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
+Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.
 ```{r}
@@
 #| label: fig-us-map
-#| fig-cap: "Predicted Winner by Battleground State on Election Day."
+#| fig-cap: "Predicted winners by battleground state on Election Day, illustrating the anticipated electoral outcomes for Donald Trump and Kamala Harris across key states.
+Highlights states where Trump is expected to win in red and states favoring Harris in blue, with gray indicating states where data is not applicable (NA).
+Underscores the competitive nature of the election and the distribution of predicted support across the country."
 #| warning: false
@@
                     na.value = "gray") +
-  labs(title = "Predicted Winner by Battleground State on Election Day", 
+  labs(
        fill = "Predicted Winner") +
@@
 ```
-According to @fig-us-map, Arizona is the only state where Trump has a leading win while all the other six states are led by Harris.
-Since Harris has a six-to-one lead in the predicted percentage of support for all battleground states, we conclude that Harris is the predicted winner for the 2024 cycle of Presidential Election.
 
@@
 One major challenge for Harris would be governing with such a divided mandate; she could face resistance from nearly half the population, complicating her efforts to implement her agenda.
-This suggests that any significant policy shifts may encounter strong opposition, potentially deepening divisions rather than bridging them (Yourish and Gamio, 2024).
+This suggests that any significant policy shifts may encounter strong opposition, potentially deepening divisions rather than bridging them [@.
 
@@
 By narrowing the dataset to only these critical swing states, our model omits data from the rest of the nation, including states with predictable voting patterns (solidly Democratic or Republican) and states outside the battleground category.
-This omission is significant because it restricts the model’s capacity to capture the broader national voting landscape and limits the ability to provide a holistic prediction of the 2024 U.S.
+This omission is significant as it restricts the model’s capacity to capture the broader national voting landscape and limits the ability to provide a holistic prediction of the 2024 U.S.
 Presidential Election outcome.
@@
 Without incorporating this popular vote data, our model may overlook broader trends that could subtly shape battleground results, especially if voters in battleground states are influenced by national sentiment.
-
-Moreover, data from non-battleground states could add valuable context and enhance model robustness by reflecting a range of voter behaviors and regional influences.
-While these states may not directly impact the electoral outcome due to their predictability, understanding regional dynamics across all states could help the model detect shifts in voter sentiment that cut across state lines.
-For instance, trends in nearby states or those with similar demographic or economic characteristics might mirror those in battleground states, providing additional insights into potential voting patterns.
-Including these states in the model could also reveal hidden correlations and strengthen the model’s predictive power by capturing a wider range of political and demographic factors.
-The focus on battleground states also means that the model may amplify the idiosyncrasies of these seven states without balancing them against broader national trends.
-Each battleground state has unique political, economic, and demographic characteristics, and relying solely on these states could cause the model to disproportionately weigh regional dynamics that may not be representative of the country as a whole.
-As a result, predictions based only on these states might overestimate or underestimate certain voting behaviors, particularly if the election outcome depends on trends that are not captured within these seven states alone.
 
@@
 ## Limitations and Future Directions
-Xiyan
+To improve data selection for future studies, data from non-battleground states could add valuable context and enhance model robustness by reflecting a range of voter behaviors and regional influences.
+While these states may not directly impact the electoral outcome due to their predictability, understanding regional dynamics across all states could help the model detect shifts in voter sentiment that cut across state lines.
+For instance, trends in nearby states or those with similar demographic or economic characteristics might mirror those in battleground states, providing additional insights into potential voting patterns.
+Including these states in the model could also reveal hidden correlations and strengthen the model’s predictive power by capturing a wider range of political and demographic factors.
+The focus on battleground states also means that the model may amplify the idiosyncrasies of these seven states without balancing them against broader national trends.
+Each battleground state has unique political, economic, and demographic characteristics, and relying solely on these states could cause the model to disproportionately weigh regional dynamics that may not be representative of the country as a whole.
+As a result, predictions based only on these states might overestimate or underestimate certain voting behaviors, particularly if the election outcome depends on trends that are not captured within these seven states alone.
 

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 
+The analysis, modeling, and visualizations of this paper were conducted using the programming language `R` [@citeR] and the following packages: `tidyverse` @Tidyverse , `janitor` @Janitor for data cleaning, `kableExtra`@kableExtra for dynamic report generation and table formatting, `modelsummary` @Modelsummary for summarizing models, `usmap` @usmap for mapping U.S.
+states, `arrow` @arrow for data integration with Apache Arrow, `lubridate` @Lubridate for date handling, `rstanarm` @rstanarm and `brms` @brms for Bayesian regression modeling, `Matrix` @Matrix for handling sparse and dense matrices, `splines` @splines for regression splines, `caret` @caret for machine learning support, and `broom.mixed`  @broommixed for tidying mixed models.
+
 # Data {#sec-data}
@@
 It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
-For every poll conducted, relevant variables such as state, start_date, end_date, transparency_score, candidate_name, and percentage of support (**pct**) for the appropriate candidates are included.
+For every poll conducted, relevant variables such as **'state'**, **'start_date'**, **'end_date'**, **'transparency_score'**, **'candidate_name'**, and percentage of support (**'pct'**) for the appropriate candidates are included.
 The data compiles information from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
@@
 For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates.
-We collect observations from the data where the ‘candidate_name’ is Trump and Harris.
-We are interested in the support for both candidates over time, for each state, therefore, we collect observations of ‘end_date’, and ‘state’, as well as the electorate support for each candidate is denoted as ‘pct’.
-Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high ‘numeric_grade’ and ‘transparency_score’, details of data cleaning are further explained in [Appendix -@sec-data-details].
+We collect observations from the data where the **‘candidate_name’** is Trump and Harris.
+We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’**, and **‘state’**, as well as the electorate support for each candidate is denoted as **‘pct’**.
+Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high **‘numeric_grade’** and **‘transparency_score’**, details of data cleaning are further explained in [Appendix -@sec-data-details].
 
@@
 
-- **State**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
-- **Polling Date**: The date when the poll was conducted, which helps in analyzing trends over time.
-- **Pollster**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
-- **Numeric Grade**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
-- **Transparency Score**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
-- **Candidate Name**: Identifies the candidate being polled, specifically focusing on Donald Trump and Kamala Harris.
+- **'State'**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
+- **'Polling Date'**: The date when the poll was conducted, which helps in analyzing trends over time.
+- **'Pollster'**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
+- **'N'umeric Grade'**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
+- **'Transparency Score'**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
+- **'Candidate Name'**: Identifies the candidate being polled, specifically focusing on Donald Trump and Kamala Harris.
 - **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.
@@
 
-It is important to note that since ‘**pct**’ stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
+It is important to note that since **'pct'** stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
 For instance, in the summary statistics presented in Table 1, the average support for Donald Trump is 47.801%, while the average support for Kamala Harris is 47.999%.
@@
 ### Predictive Variables
-In this section, we outline the two key predictor variables used in our analysis: '**state**', which captures the geographical context of the polling data, and '**end_date**', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.
+In this section, we outline the two key predictor variables used in our analysis: **'state'**, which captures the geographical context of the polling data, and '**end_date**', which reflects the timing of the polls relative to the election, allowing us to assess trends in voter support over time.
 
 #### State
-The state (**state**) variable indicates the U.S.
-state where the poll was conducted or targeted.
-In the context of the U.S.
-presidential election, the outcome is frequently determined by several key battlegroud states.
-In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
-To enhance our model and refine our predictions, we focused exclusively on these seven battleground states.
-The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
+The state (**'state'**) variable indicates the U.S. state where the poll was conducted or targeted. 
 ```{r}
@@
 Illustrates the number of polling observations across key battleground statesn.
-Notably, Pennsylvania has the highest count of observations, suggesting a more robust polling presence, while Nevada shows fewer observations.
-"
+Pennsylvania has the highest count of observations, suggesting a more robust polling presence, while Nevada shows fewer observations."
 
@@
 ```
+In the context of the U.S.
+presidential election, the outcome is frequently determined by several key battlegroud states.
+In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
+To enhance our model and refine our predictions, we focused exclusively on these seven battleground states.
+The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
 The distribution of the state variable is visualized in @fig-state-distribution and indicates that some states, such as Pennsylvania, North Carolina and Wisconsin have more polling observations compared to others, like Nevada.
@@
 #### End Date {#sec-end-date}
-End date ('**end_date**') represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
+End date (**'end_date'**) represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
 In the context of our paper, the end date of each poll serves as one of the predictive variables and is used to assess changes in voter support over time.
@@
 ### Other Variables
-This section details additional variables that provide essential context for our analysis, including **numeric grade**, which evaluates the quality of the polls; **transparency score**, indicating the openness of poll methodologies; **pollster**, identifying the organizations conducting the polls; and **start date**, which denotes the initiation of the polling period for each survey.
+This section details additional variables that provide essential context for our analysis, including **'numeric_grade'**, which evaluates the quality of the polls; **'transparency_score'**, indicating the openness of poll methodologies; **'pollster'**, identifying the organizations conducting the polls; and **'start_date'**, which denotes the initiation of the polling period for each survey.
 
 #### Numeric Grade
-Numeric Grade ('**numeric_grade**') represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
+Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
 In the context of our paper, numeric grade allows us to weigh the reliability of different polls.
 By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support.
-@fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0.
-Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions.
-The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
-
 ```{r}
@@
 ```
+@fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0.
+Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions.
+The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
 This variability in numeric grade highlights the diversity in polling quality within our dataset.
@@
 #### Transparency Score
-Transparency Score ('**transparency_score**') represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
+Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
 This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error.
@@
 #### Pollster
-Pollster ('**pollster**') represents the various organizations responsible for conducting polls across key battleground states in the lead-up to the 2024 United States Presidential Election.
+Pollster (**'pollster'**) represents the various organizations responsible for conducting polls across key battleground states in the lead-up to the 2024 United States Presidential Election.
 In the context of our paper, the diversity and frequency of polling by different pollsters are essential for understanding the reliability and variance in the data.
@@
 #### Start Date
-Start date ('**start_date**') represents the date when each poll began collecting responses.
+Start date (**'start_date'**) represents the date when each poll began collecting responses.
 In the context of our paper, start date is a key variable as it allows us to focus on polling data collected after July 21, 2024—the date Kamala Harris officially announced her presidential campaign.

### 2024-11-04 paper/paper.qmd
committing text changes to intro in paper.qmd

@@
 
-The analysis, modeling, and visualizations of this paper were conducted using the programming language `R` [@citeR] and the following packages: `tidyverse` @Tidyverse , `janitor` @Janitor for data cleaning, `kableExtra`@kableExtra for dynamic report generation and table formatting, `modelsummary` @Modelsummary for summarizing models, `usmap` @usmap for mapping U.S.
-states, `arrow` @arrow for data integration with Apache Arrow, `lubridate` @Lubridate for date handling, `rstanarm` @rstanarm and `brms` @brms for Bayesian regression modeling, `Matrix` @Matrix for handling sparse and dense matrices, `splines` @splines for regression splines, `caret` @caret for machine learning support, and `broom.mixed`  @broommixed for tidying mixed models.
+The analysis, modeling, and visualizations of this paper were conducted using the programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
+states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.
 

### 2024-11-04 paper/paper.qmd
committing formatting changes to paper.qmd

@@
 #### State
-The state (**'state'**) variable indicates the U.S. state where the poll was conducted or targeted. 
 ```{r}
@@
 ```
+The state (**'state'**) variable indicates the U.S.
+state where the poll was conducted or targeted.
 In the context of the U.S.

### 2024-11-04 paper/paper.qmd
commit changes to paper.qmd and adding pdf of survey

@@
 ## State-Level Polling Averages for Trump and Harris
-@tbl-outcomevar displays overview of the percentage of support for Trump and Harris by each battleground state.
+@tbl-stateaverage displays overview of the percentage of support for Trump and Harris by each battleground state.
 It provides us with a high level understanding of the current voter dynamic.
@@
 One major challenge for Harris would be governing with such a divided mandate; she could face resistance from nearly half the population, complicating her efforts to implement her agenda.
-This suggests that any significant policy shifts may encounter strong opposition, potentially deepening divisions rather than bridging them [@.
+This suggests that any significant policy shifts may encounter strong opposition, potentially deepening divisions rather than bridging them [@discussion-article-1].
 
 Moreover, a narrow victory for Harris would reinforce national polarization, revealing the extent to which the country is split.
-The nearly equal support for both candidates highlights entrenched ideological divides, with voters committed to opposing perspectives on critical issues such as healthcare, immigration, climate policy, and the role of government (Heltzel & Laurin, 2020).
+The nearly equal support for both candidates highlights entrenched ideological divides, with voters committed to opposing perspectives on critical issues such as healthcare, immigration, climate policy, and the role of government [@discussion-article-2].
 Economic concerns, including job security, wage equality, and tax policy, alongside cultural values related to family, education, and identity, play significant roles in voter alignment.
@@
 presidential election.
-To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [@sec-references].
+To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [References -@sec-references].
 

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd and reference.bib

@@
 ## Uncertainty Beyond Data and Modeling
-There are external factors other than data and modeling that could influence the actual outcome of the election. These factors are often hard to control and predict.
-
-Firstly, recency bias influences public perception and even polling responses.
-Many people expect Trump to win because of his unexpected victory in 2016 and his close race in 2020, despite lagging in the polls.
-This bias conveys the tendency to focus on recent, high-profile events while overlooking historical trends, such as Obama’s strong polling performance in 2012, where he outperformed polls.
-This bias can affect how people respond to polls or even how pollsters interpret results, as certain events may be weighted more heavily than warranted.
-As our model depends on polling data, any systematic bias introduced by recency expectations can skew our predictions, making it challenging to account for potential polling errors.
-
-Social desirability bias and non-response bias are additional sources of uncertainty, particularly with Trump’s supporters.
-Some voters may be hesitant to disclose their support for a conservative candidate due to social stigma, leading to underreporting of support in polls.
-This reluctance can be especially strong in areas where conservative views are less socially accepted.
-Trump’s support base also includes individuals with lower social trust and civic engagement, which means they are less likely to participate in polls altogether.
-If certain groups systematically avoid participating, polling data may underrepresent these voters, resulting in an underestimation of Trump’s support in our model.
-This non-response bias is difficult to quantify, adding a layer of uncertainty to our predictions.
-On the other hand, Harris faces potential biases such as the Bradley effect and gender-related bias.
-As a Black female candidate, Harris might encounter the Bradley effect, where some voters express indecision rather than explicitly stating they won’t support a minority candidate.
-Historically, this effect has impacted Black candidates, as undecided voters tend to lean against them in the actual vote count, despite polling responses suggesting neutrality.
-Additionally, Clinton’s 2016 campaign demonstrated that undecided voters can heavily swing against female candidates.
-This trend could similarly impact Harris’s polling and election results.
-If these voters decide against Harris in the end, our model might overestimate her support based on the polling data, leading to inaccuracies in predictions.
-These biases are inherently difficult to measure or adjust for, as they reflect deeper societal attitudes and individual decision-making processes.
-Pollsters may attempt to correct for non-response or social desirability bias, but such adjustments are often based on assumptions that may not fully capture the complex social dynamics at play.
-Consequently, our model’s predictions should be interpreted with caution, acknowledging that these biases can lead to unexpected shifts in voter behavior that fall outside the range of our data-driven forecasts.
+Beyond data and modeling, several external factors can influence the actual outcome of the election.
+Recency bias is one of the external factors, and it can significantly shape public perception and polling responses.
+Many people expect Trump to win due to his unexpected victory in 2016 and his narrow loss in 2020, despite lagging in the polls.
+This bias reflects a tendency to focus on recent, high-profile events while overlooking broader historical trends.
+For instance, in 2012, Barack Obama not only won but outperformed polling predictions, a fact that may be forgotten amidst recent election narratives [@discussion-article3].
+Such a bias can affect both how people respond to polls and how pollsters interpret results, potentially leading to an overemphasis on recent events.
+Given that our model depends on polling data, any systematic bias from these recency expectations can skew predictions, making it challenging to account for potential polling errors.
+
+Social desirability bias and non-response bias add further layers of uncertainty, particularly regarding Trump’s support base.
+Some voters may be hesitant to openly express support for a conservative candidate due to social stigma, leading to an under-reporting of Trump support in polls.
+This effect may be stronger in regions where conservative views are less socially accepted.
+Additionally, Trump’s supporters often include individuals with lower social trust and civic engagement, making them less likely to participate in polls altogether [@discussion-article3].
+If specific groups systematically avoid participation, polling data may under-represent these voters, potentially underestimating Trump’s support in our model.
+Non-response bias is difficult to quantify, introducing additional uncertainty to our predictions.
+
+Harris, on the other hand, may face challenges like the Bradley effect and gender-related bias.
+As a Black female candidate, Harris could encounter the Bradley effect, where some voters indicate indecision rather than openly stating that they will not support a minority candidate.
+Historically, undecided voters have leaned against Black candidates in the actual vote count, despite neutral polling responses.
+Additionally, the 2016 election demonstrated that undecided voters could heavily swing against female candidates, as seen with Hillary Clinton [@discussion-article3].
+This trend could similarly impact Harris’s results if undecided voters ultimately decide against her.
+Consequently, our model may overestimate Harris’s support based on polling data alone, leading to potential inaccuracies in prediction.
+
+These biases are inherently challenging to measure or adjust for, as they reflect deeper societal attitudes and individual decision-making patterns.
+While pollsters may attempt to correct for non-response or social desirability bias, such adjustments are often based on assumptions that cannot fully capture the complex social dynamics at play.
+Thus, our model’s predictions should be interpreted with caution, recognizing that these biases may cause unexpected shifts in voter behavior that go beyond what data-driven forecasts can capture.
 
@@
 
-## Sampling Approach
-
+## Survey Demo
+
+**Title: "2024 U.S. Presidential Election Survey"**
+
+### Survey Introduction
+We are conducting an important survey to understand voter opinions and predict the outcome of the upcoming 2024 U.S.
+Presidential Election.
+Your participation will help ensure that our findings accurately reflect the views of voters like you.
+
+This survey will take about 5–10 minutes to complete.
+Your responses will remain completely anonymous and confidential.
+We are committed to protecting your privacy, and no personal information will be shared or used for any purpose beyond this research.
+
+Thank you for your time and contribution.
+
+### Survey Questions
+1. **Thinking ahead to the presidential general election, are you going to vote?**  
+   *On a scale from 1 (Not at all likely) to 10 (Almost certain).*  
+   - 1 (Not at all likely)
+   - 2
+   - 3
+   - 4
+   - 5
+   - 6
+   - 7
+   - 8
+   - 9
+   - 10 (Almost certain)
+
+2. **Who do you want to vote for?**  
+   - Kamala Harris (Democratic Party)
+   - Donald Trump (Republican Party)
+   - Prefer not to answer
+
+3. **How likely are you to change your mind before the election?**  
+   *On a scale from 1 (Very unlikely) to 10 (Almost certain).*  
+   - 1 (Very unlikely)
+   - 2
+   - 3
+   - 4
+   - 5
+   - 6
+   - 7
+   - 8
+   - 9
+   - 10 (Almost certain)
+
+4. **Which of the following issues is most important to you in this election?**  
+   - Economy
+   - Healthcare
+   - Immigration
+   - Climate Change
+   - Safety
+   - Education
+   - Other (please specify)
+
+5. **If you have to decide between Kamala Harris and Donald Trump, who will you vote for?**  
+   - Kamala Harris
+   - Donald Trump
+   - Prefer not to answer
+
+6. **Who do you expect to win the upcoming elections?**  
+   - Kamala Harris
+   - Donald Trump
+   - Prefer not to answer
+
+7. **How often do you follow political news?**  
+   *On a scale from 1 (Rarely) to 10 (Daily).*  
+   - 1 (Rarely)
+   - 2
+   - 3
+   - 4
+   - 5
+   - 6
+   - 7
+   - 8
+   - 9
+   - 10 (Daily)
+
+8. **Would you be open to commenting on the issues in this survey and be interested in being contacted by a reporter?**  
+   - Yes
+   - No
+
+9. **For validation purposes, please select "Agree" from the options below.**  
+   - Agree
+   - Disagree
+
+### Sampling Approach
 Using data from voter files, which contain demographic information about registered voters, we aim to ensure that our sample accurately represents the population.

### 2024-11-04 paper/paper.qmd
commit changes to paper.qmd

@@
 ## Measurement
-Voters’ opinions are multifaceted and can be influenced by numerous factors, including media coverage, recent political events, personal beliefs, and social pressures.
-Each voter’s decision is shaped by a unique mix of these factors, which can change frequently, especially close to an election.
-Pollsters start by converting the complex, multi-layered opinions in voters' heads into a simple polling response.
-This is the first challenge: each voter’s thoughts might not fit neatly into options like “Trump,” “Harris,” or “Undecided.” People may lean toward one candidate but have reservations, or may change their views over time.
-By asking for a single answer, the poll essentially “forces” an opinion that may not fully capture the voter’s internal conflict.
-
-Once collected, these individual responses are combined into percentages of support.
-This aggregation abstracts further, treating all “Trump” responses equally, regardless of how confident or reluctant the voter might have been.
-In effect, the complexity of each opinion is smoothed over into a uniform data point, with many variations of opinions lost in the process.
-Finally, these support percentages are fed into a model to make a predictive statement about the election outcome.
-The model’s goal is to determine who is likely to win, based on the assumption that the aggregated opinions in our data reflect real-world voter behavior on Election Day.
-However, this step relies heavily on simplification, assuming that the snapshot of opinions captured in the polls will hold steady and that all voters will follow through on their expressed preferences.
-The raw polling data gives us percentages of support for each candidate in each state.
-To measure our outcome variable (the likelihood of winning), we convert these percentages into a binary outcome in our model (predicting Trump or Harris as the winner based on who has higher support in each state).
-Our model uses these support percentages to generate a probability of each candidate winning.
-This transformation helps translate fluctuating polling data into a clear prediction of the likely winner, but it also involves a level of simplification, as it assumes these support percentages accurately reflect voting intentions.
-
-Using state as a predictor variable, which influences our outcome variable, assumes that all responses within a state are comparable and that regional differences are consistent across the state.
-However, each state has internal diversity—urban vs.
-rural areas, differences in income levels, or cultural values—that the variable “state” may not fully capture.
-This simplification is necessary for the model, but it means that nuanced, within-state variations in opinion are lost in the abstraction.
-Another predictor variable, end_date is also measured, it is designed to capture how voter sentiment changes over time, especially as the election approaches.
-By using the end date of each poll, the model can recognize patterns or shifts in support due to campaign events, news coverage, or debates.
-While “end_date” provides temporal context, it simplifies complex, often continuous changes in opinion into a discrete snapshot.
-This abstraction assumes that voter sentiment is relatively stable between poll dates, which may not reflect reality.
-For example, an event occurring immediately after a poll’s end date could dramatically shift opinions that the model wouldn’t capture until the next poll.
-Additionally, the importance of time in influencing voter sentiment is assumed to be similar across states, but certain states might react differently to events depending on their local context or demographic makeup.
+Voter opinions are influenced by various factors, such as media, political events, personal beliefs, and social pressures, which can change frequently, especially near an election.
+Pollsters simplify these complex opinions into basic polling responses like “Trump,” “Harris,” or “Undecided,” which can obscure nuanced voter sentiment.
+This initial conversion may not capture voters’ internal conflicts or reservations, as respondents must choose a definitive option.
+
+These individual responses are aggregated into percentages, treating each “Trump” or “Harris” response equally, regardless of confidence level, further smoothing out complexities.
+The model uses these aggregated support percentages to predict election outcomes by estimating the likelihood of each candidate winning.
+This step assumes that poll data accurately reflects voter behavior on Election Day, which may not always hold true, as it simplifies support into binary outcomes (predicting Trump or Harris as the winner based on higher support).
+
+The use of “state” as a predictor assumes that all responses within a state are comparable, despite internal diversity, such as urban-rural divides or cultural differences, which may not be fully captured.
+Similarly, the “end_date” variable tracks changes in voter sentiment over time to capture patterns influenced by campaign events or news.
+However, this variable simplifies continuous shifts into discrete snapshots, assuming stability between poll dates.
+This may overlook rapid opinion changes following significant events.
+Additionally, the model presumes time’s influence on voter sentiment is consistent across states, despite varying local contexts and demographics that may respond differently to events.
 
@@
 - **'Pollster'**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
-- **'N'umeric Grade'**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
+- **'Numeric Grade'**: A score assigned to each poll based on its methodology and historical accuracy, which assists in filtering out less reliable polls.
 - **'Transparency Score'**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
@@
 ## Model Overview
-The Bayesian Beta regression models constructed in this paper aim to predict the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S.
-Presidential Election by analyzing polling data of the 7 battleground states.
-One Bayesian Beta regression model is constructed for each candidate, and each of them estimates the proportion of support for each candidate based on polling data, accounting for both polling trends over time, and state-level differences in voter sentiment.
-These models allow us to evaluate how support for each candidate fluctuates over time in key states, and predict potential electoral outcomes.
-
-One of the goals of our model is to analyze changes in voter support for each candidate over time, as Election Day approaches.
-By using end_date_num (days the earliest poll date) as a predictor with natural splines, the model captures non-linear trends in support.
-The time-based analysis provides us with information about how voter preference may shift due to campaign events, news cycles, and other influential factors in the months leading up to the election.
-
-Another goal of our model is to predict the potential winner of the Presidential Election.
-We recognize that battleground states have unique political landscapes, so we include state as a random effect in each model.
-This allows each of the 7 battleground states to have a unique baseline level of support for each candidate, capturing regional differences in voter preferences.
-By incorporating state-level effects, the model can show variations in support across different states, revealing the potential winning candidate for the presidential election.
-Further background details, model specifications, and diagnostics are included in [Appendix -@sec-model-details].
+The Bayesian Beta regression models in this paper estimate the probability of victory for Donald Trump and Kamala Harris in the 2024 U.S.
+Presidential Election by analyzing polling data from seven battleground states.
+Separate models are constructed for each candidate to estimate their support, considering polling trends over time and state-level differences in voter sentiment.
+This approach captures how support fluctuates leading up to Election Day and predicts potential electoral outcomes.
+To analyze changes in voter support over time, the models use natural splines with the **'end_date_num'** (days from the earliest poll date) as a predictor, allowing for the identification of non-linear trends influenced by campaign events and news cycles.
+Additionally, state-level effects are incorporated as random factors, enabling the models to account for unique political landscapes across the battleground states.
+This regional adjustment helps highlight variations in candidate support and enhances the accuracy of predictions.
+Further detailed model specifications and diagnostics are provided in [Appendix -@sec-model-details].
 
@@
 ## Model Setup
-Our models were ran in R [@citeR] using the 'rstanarm' package of @rstanarm.
 
@@
 ## Uncertainty Beyond Data and Modeling
-Beyond data and modeling, several external factors can influence the actual outcome of the election.
-Recency bias is one of the external factors, and it can significantly shape public perception and polling responses.
-Many people expect Trump to win due to his unexpected victory in 2016 and his narrow loss in 2020, despite lagging in the polls.
-This bias reflects a tendency to focus on recent, high-profile events while overlooking broader historical trends.
-For instance, in 2012, Barack Obama not only won but outperformed polling predictions, a fact that may be forgotten amidst recent election narratives [@discussion-article3].
-Such a bias can affect both how people respond to polls and how pollsters interpret results, potentially leading to an overemphasis on recent events.
-Given that our model depends on polling data, any systematic bias from these recency expectations can skew predictions, making it challenging to account for potential polling errors.
-
-Social desirability bias and non-response bias add further layers of uncertainty, particularly regarding Trump’s support base.
-Some voters may be hesitant to openly express support for a conservative candidate due to social stigma, leading to an under-reporting of Trump support in polls.
-This effect may be stronger in regions where conservative views are less socially accepted.
-Additionally, Trump’s supporters often include individuals with lower social trust and civic engagement, making them less likely to participate in polls altogether [@discussion-article3].
-If specific groups systematically avoid participation, polling data may under-represent these voters, potentially underestimating Trump’s support in our model.
-Non-response bias is difficult to quantify, introducing additional uncertainty to our predictions.
-
-Harris, on the other hand, may face challenges like the Bradley effect and gender-related bias.
-As a Black female candidate, Harris could encounter the Bradley effect, where some voters indicate indecision rather than openly stating that they will not support a minority candidate.
-Historically, undecided voters have leaned against Black candidates in the actual vote count, despite neutral polling responses.
-Additionally, the 2016 election demonstrated that undecided voters could heavily swing against female candidates, as seen with Hillary Clinton [@discussion-article3].
-This trend could similarly impact Harris’s results if undecided voters ultimately decide against her.
-Consequently, our model may overestimate Harris’s support based on polling data alone, leading to potential inaccuracies in prediction.
-
-These biases are inherently challenging to measure or adjust for, as they reflect deeper societal attitudes and individual decision-making patterns.
-While pollsters may attempt to correct for non-response or social desirability bias, such adjustments are often based on assumptions that cannot fully capture the complex social dynamics at play.
-Thus, our model’s predictions should be interpreted with caution, recognizing that these biases may cause unexpected shifts in voter behavior that go beyond what data-driven forecasts can capture.
+Beyond data and modeling, external factors such as recency bias, social desirability bias, and non-response bias can influence the election outcome.
+Recency bias significantly shapes public perception, with many expecting Trump to win due to his 2016 victory and narrow 2020 loss, despite lagging in polls.
+This bias highlights a tendency to focus on recent events while overlooking broader trends.
+For example, in 2012, Barack Obama outperformed polls, a fact often overshadowed by recent election narratives[@discussion-article3].
+This can skew how people respond to polls and how pollsters interpret results, introducing potential polling errors that our model cannot fully account for.
+
+Social desirability bias and non-response bias add further uncertainty, especially regarding Trump’s support base.
+Social stigma may deter some voters from expressing conservative support openly, leading to under-reported Trump support in polls.
+Additionally, Trump’s supporters often have lower social trust, making them less likely to participate in polls altogether[@discussion-article3].
+If certain groups systematically avoid participation, polling data may under-represent these voters, resulting in an underestimation of Trump’s support in our model.
+
+Harris faces potential challenges such as the Bradley effect and gender-related bias.
+As a Black female candidate, Harris may encounter undecided voters leaning against her due to underlying biases, similar to what Hillary Clinton faced in 2016 [@discussion-article3].
+If undecided voters ultimately vote against Harris, our model may overestimate her support based on polling data alone, affecting prediction accuracy.
+
+These biases are difficult to measure or correct, reflecting deeper societal attitudes and complexities in voter behavior.
+Although pollsters may attempt adjustments, these are often based on assumptions that cannot fully capture the dynamics at play.
+Therefore, our model’s predictions should be interpreted cautiously, acknowledging that biases may lead to unexpected shifts beyond the range of data-driven forecasts.
 
@@
 Such contextual factors could be especially useful in a Bayesian framework, allowing us to adjust prior distributions based on known historical influences on voter behavior.
-- **Inclusive of National/Popular votes**: our model has exclusive focus on the seven battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-By narrowing the dataset to only these critical swing states, our model omits data from the rest of the nation, including states with predictable voting patterns (solidly Democratic or Republican) and states outside the battleground category.
-This omission is significant as it restricts the model’s capacity to capture the broader national voting landscape and limits the ability to provide a holistic prediction of the 2024 U.S.
-Presidential Election outcome.
-In particular, the exclusion of national popular vote data limits our model’s understanding of overall voter sentiment across the country.
-While battleground states are pivotal in determining the electoral outcome due to the Electoral College system, popular vote trends from non-battleground states can still reveal important national shifts in sentiment that may indirectly influence battleground dynamics.
-For example, a nationwide trend towards or away from a candidate could suggest emerging political momentum or reflect the impact of major events that also resonate within battleground states.
-Without incorporating this popular vote data, our model may overlook broader trends that could subtly shape battleground results, especially if voters in battleground states are influenced by national sentiment.
+- **Inclusive of National/Popular votes**: Our model focuses exclusively on the seven battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—omitting data from solidly Democratic or Republican states and the national popular vote.
+This exclusion limits the model’s ability to capture broader voter sentiment across the country, restricting its capacity for a holistic prediction of the 2024 Presidential Election.
+While battleground states are crucial due to the Electoral College system, popular vote trends from other states can indicate national shifts that may indirectly influence battleground dynamics.
+For example, a nationwide trend favoring or opposing a candidate could signal political momentum or reflect impactful events resonating with voters in battleground states.
+Without national data, our model may overlook these broader trends, potentially affecting the accuracy of its predictions in key states.
 
@@
 ## Model Justification - Further Details
-The Bayesian Beta Regression model with a logit link function is particularly suited for our prediction of the 2024 U.S.
-Presidential Election due to its alignment with the nature of our outcome data and the goals of our analysis.
-Our primary outcome variable is the percentage of support for each candidate in the seven battleground states, which we scaled to a proportion between 0 and 1.
-Since support percentages are naturally bounded between 0% and 100%, modeling these data with traditional linear regression presents challenges, as linear regression assumes an unbounded continuous outcome.
-The Beta distribution is ideal for this type of data because it is defined on the interval (0, 1), allowing us to model proportions without the risk of generating unrealistic values outside this range.
-Unlike a linear model, which could predict support percentages exceeding 100% or falling below 0%, the Beta distribution respects the boundaries of our data.
-
-To effectively capture the dynamics of voter support, we utilized natural splines with the variable end_date_num, which represents the number of days since the earliest poll date.
-This choice allows the model to adapt to non-linear trends in voter support that can fluctuate due to various campaign events, debates, or shifts in the political landscape.
-By specifying four degrees of freedom for the natural splines, we enable the model to flexibly capture gradual changes in voter sentiment while mitigating the risk of overfitting to noise or abrupt fluctuations in the data.
-This flexibility is particularly critical in electoral polling, where opinions may shift slowly or rapidly in response to unpredictable external events.
-Alternative models, such as standard logistic regression without splines, would fail to capture these complex changes, potentially leading to a less accurate depiction of how support varies over time.
-
-A unique aspect of our analysis is the focus on battleground states, which exhibit distinct political dynamics that can significantly influence election outcomes.
-Incorporating state as a random effect provides the flexibility necessary to estimate state-specific baseline support levels for each candidate while still allowing us to identify overall trends across all states.
-This approach is more effective than treating state as a fixed effect, as it avoids overparameterizing the model with individual state coefficients, which could lead to overfitting, particularly in cases with limited polling data for certain states.
-The random effect structure enhances the model’s generalizability, making it more robust when predicting support in battleground states with varying polling frequencies.
-
-The Bayesian approach offers substantial advantages for our model by allowing us to incorporate prior information, which enhances the stability and reliability of our estimates.
-Given that polling data can be sparse for certain states or dates, Bayesian inference helps stabilize parameter estimates, especially when limited data points are available.
-We defined weakly informative priors for the model parameters: for the intercept $\beta_0$, we specified $\beta_0 \sim \text{Normal}(0, 10)$ to provide a relatively uninformative prior that allows the data to inform the estimation of baseline log-odds; for the slope $\beta_1$, we used $\beta_1 \sim \text{Normal}(0, 5)$, reflecting reasonable expectations regarding the influence of polling trends over time.
-
-This model structure not only allows for capturing the dynamics of voter support but also provides interpretability that aligns well with the goals of our paper.
-The logit link function enables us to interpret results in terms of odds or probabilities of victory, which is crucial in the election context.
-By constructing separate models for Trump and Harris, we can directly compare their temporal trends and state-level variations without confounding effects.
-Thus, the Bayesian Beta regression model with a logit link function is justified as the most suitable model for our study.
-It accommodates flexible non-linear trends over time, accounts for regional variation across battleground states, and delivers interpretable results concerning victory probabilities.
-The Bayesian framework not only stabilizes estimates in the presence of sparse data but also provides meaningful uncertainty estimates, ultimately aligning with our objective of forecasting each candidate’s probability of victory in the 2024 U.S.
-Presidential Election.
+The Bayesian Beta regression model with a logit link function is well-suited for predicting the 2024 U.S.
+Presidential Election due to the nature of our outcome data and analysis goals.
+Our primary outcome variable is the proportion of support for each candidate in seven battleground states, scaled between 0 and 1.
+The Beta distribution is ideal for modeling such bounded data, unlike linear regression, which assumes an unbounded outcome and risks unrealistic predictions outside the 0-100% range.
+
+To capture non-linear trends in voter support over time, we use natural splines with the variable **'end_date_num'** (days since the earliest poll date), allowing the model to adapt to shifts due to campaign events or debates.
+Specifying four degrees of freedom for the splines balances flexibility in modeling gradual changes with reducing noise or overfitting.
+Standard logistic regression lacks this adaptability and would be less effective in capturing the evolving nature of voter sentiment.
+Additionally, incorporating state as a random effect enables the model to estimate state-specific baseline support levels, reflecting the distinct political dynamics of each battleground state while preserving overall trend analysis.
+This approach prevents overfitting by avoiding an overparameterized model, particularly useful when polling data for some states is sparse.
+The random effect structure thus enhances model robustness and generalizability across states with varying polling frequencies.
+
+The Bayesian approach adds value by incorporating prior information, stabilizing parameter estimates where data are limited.
+We applied weakly informative priors to ensure data-driven estimation: the intercept ($\beta_0 \sim \text{Normal}(0, 10)$) and slope ($\beta_1 \sim \text{Normal}(0, 5)$).
+This setup supports stable inference, crucial in states or periods with sparse polling data.
+This model structure provides clear interpretability, essential for comparing the probability of victory for Trump and Harris.
+The logit link function facilitates results interpretation in terms of odds or victory probabilities, aligning with our election analysis goals.
+Constructing separate models for each candidate allows for direct comparisons of trends and state-level variations without confounding.
+Thus, the Bayesian Beta regression model with a logit link function is justified as the optimal choice.
+It captures non-linear temporal trends, accounts for regional variations, and delivers interpretable victory probabilities, supported by Bayesian inference that enhances reliability and reflects uncertainty in forecasts.
 
@@
 By comparing the distributions of observed support percentages with the predicted values, we can identify discrepancies and assess the overall fit of our models.
-
 Posterior predictive checks help verify model assumptions and provide insights into how well the models capture the underlying voter dynamics.

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 By filtering the dataset to include only polls with a start date after this announcement, we ensure that each poll reflects voter sentiment with Harris as an active candidate, enabling a more accurate comparison between her and Donald Trump.
-@fig-start-date-distribution displays the distribution of polling start dates from late July to mid-October 2024, showing when polling efforts began across key battleground states.
-The histogram reveals peaks in polling activity, particularly in mid-August, early September, and late September, indicating periods of heightened interest or campaign events that likely prompted increased polling.
 ```{r}
@@
 ```
+@fig-start-date-distribution displays the distribution of polling start dates from late July to mid-October 2024, showing when polling efforts began across key battleground states.
+The histogram reveals peaks in polling activity, particularly in mid-August, early September, and late September, indicating periods of heightened interest or campaign events that likely prompted increased polling.
 This distribution provides insights into the timing and intensity of polling efforts throughout the campaign season.

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd

@@
 
-The analysis, modeling, and visualizations of this paper were conducted using the programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
+All programming for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
 states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.
@@
 
-To ensure the quality of the analysis, only those polls that meet specific thresholds were included - for more details on the data cleaning process, refer to @sec-data-cleaning.
+To ensure the quality of the analysis, only those polls that meet specific thresholds were included - for more details on the data cleaning process, refer to [Appendix -@sec-data-cleaning].
 

### 2024-11-04 paper/paper.qmd
committing changes to appendix for paper.qmd and testing scripts

@@
 title: "Forecasting the 2024 U.S. Presidential Election: Kamala Harris' Projected Victory"
-subtitle: A Bayesian Analysis of Polling Data from Key Battleground States"
+subtitle: "A Bayesian Analysis of Polling Data from Key Battleground States - Estimating Voter Support Dynamics for Donald Trump and Kamala Harris in the 2024 U.S. Presidential Election"
 author: 
@@
 Utilizing a Bayesian framework and state-level polling data, we estimate the probabilities of victory for candidates Donald Trump and Kamala Harris.
-Our findings indicate that Harris is projected to secure a win in six out of seven key battleground states, such as Michigan, Nevada, and North Carolina, with predicted support exceeding 50%.
+Our findings indicate that Harris is projected to secure a win in six out of seven key battleground states, with predicted support exceeding 50%.
 This analysis not only forecasts the likely winner of the U.S.
@@
 As candidates Donald Trump and Kamala Harris intensify their campaigns, understanding voter preferences in these key regions is essential for accurate election forecasting.
-This paper utilizes state-level polling data from FiveThirtyEight [@FiveThirtyEight] and the statistical programming language R [@citeR] to develop a predictive model estimating each candidate's likelihood of victory in critical battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+This paper utilizes state-level polling data from Project 538 [@FiveThirtyEight] and the statistical programming language R [@citeR] to develop a predictive model estimating each candidate's likelihood of victory in critical battleground states, including Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
 
@@
 The implications of our findings extend beyond mere predictions, offering a framework for analyzing potential shifts in voter sentiment as Election Day approaches.
-The remainder of this paper is organized as follows: Section @sec-data details the data and variables utilized in the analysis; Section @sec-model outlines the model setup and estimation strategies employed; Section @sec-results presents the results, including visualizations of the predicted election outcomes; and Section @sec-discussion discusses the implications and limitations of our findings.
-
-All programming for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
+The remainder of this paper is organized as follows: @sec-data details the data and variables utilized in the analysis; @sec-model outlines the model setup and estimation strategies employed; @sec-results presents the results, including visualizations of the predicted election outcomes; and @sec-discussion discusses the implications and limitations of our findings.
+
+Data cleaning, analysis, modeling, and visualization for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
 states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.
@@
 ## Overview
-The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538's polling project [@FiveThirtyEight] on October 19th, 2024.
+The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538 [@FiveThirtyEight] on October 19th, 2024.
 The data is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
@@
 We collect observations from the data where the **‘candidate_name’** is Trump and Harris.
-We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’**, and **‘state’**, as well as the electorate support for each candidate is denoted as **‘pct’**.
+We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’** and **‘state’**, as well as the electorate support for each candidate, which is denoted as **‘pct’**.
 Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high **‘numeric_grade’** and **‘transparency_score’**, details of data cleaning are further explained in [Appendix -@sec-data-details].
@@
 - **'State'**: Indicates the state where the polling took place, critical for understanding regional support dynamics.
-- **'Polling Date'**: The date when the poll was conducted, which helps in analyzing trends over time.
+- **'End Date'**: The date when the poll ended, which helps in analyzing trends over time.
+- **'Start Date'**: The date when the poll started, to filter for Kamala Harris's late entry into the presidential race 
 - **'Pollster'**: The organization conducting the poll, providing insight into the reliability and methodology of the polling data.
@@
 - **'Transparency Score'**: A measure indicating how transparent the pollster is regarding their methodology, further enhancing data credibility.
-- **'Candidate Name'**: Identifies the candidate being polled, specifically focusing on Donald Trump and Kamala Harris.
+- **'Candidate Name'**: Identifies the running candidate, specifically focusing on Donald Trump and Kamala Harris.
 - **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.
@@
 ### Outcome Variable
-As our focus is to analyze and predict voter support for Kamala Harris and Donald Trump during the current presidential election cycle, the primary outcome variables of interest are the percentages of support for each candidate, denoted as ‘pct’.The pct is categorized by candidate and scaled to a proportion by diving by 100.
+As our focus is to analyze and predict voter support for Kamala Harris and Donald Trump during the current presidential election cycle, the primary outcome variables of interest are the percentages of support for each candidate, denoted as **‘pct’**.The pct is categorized by candidate and scaled to a proportion by diving by 100.
 This variable represents the proportion of voter support for each candidate based on polling data.
-Through understanding the percentage of support (pct) for each candidate, we can observe and infer the candidate that has a leading advantage in winning the presidential election race.
+Through understanding the percentage of support (**'pct'**) for each candidate, we can observe and infer the candidate that has a leading advantage in winning the presidential election race.
 
@@
 Therefore, any polls that has an end date before July would not contain polling results for Harris.
-The data from [@FiveThirtyEight] is last updated on October 15, 2024, thus the latest end date being October 21, 2024.
+The data from Project 538 [@FiveThirtyEight] is last updated on October 15th, 2024, thus the latest end date being October 14th, 2024.
 Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time.
@@
 #### Numeric Grade
-Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
+Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
 In the context of our paper, numeric grade allows us to weigh the reliability of different polls.
@@
 #### Transparency Score
-Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by [@FiveThirtyEight].
+Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by Project 538 [@FiveThirtyEight].
 This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error.
@@
 Notably, Emerson is the most frequent pollster, conducting the highest number of polls, followed by Quinnipiac and Siena/NYT.
-Other pollsters, such as YouGov, CNN/SSRS, and Marist, also contribute a significant number of polls, whereas some organizations like The Washington Post and SurveyUSA have comparatively fewer entries in the dataset.This distribution is crucial for evaluating the consistency and breadth of the polling data.
+Other pollsters, such as YouGov, CNN/SSRS, and Marist, also contribute a significant number of polls, whereas some organizations like The Washington Post and SurveyUSA have comparatively fewer entries in the dataset.
+This distribution is crucial for evaluating the consistency and breadth of the polling data.
 Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others.
@@
 This approach captures how support fluctuates leading up to Election Day and predicts potential electoral outcomes.
-To analyze changes in voter support over time, the models use natural splines with the **'end_date_num'** (days from the earliest poll date) as a predictor, allowing for the identification of non-linear trends influenced by campaign events and news cycles.
+To analyze changes in voter support over time, the models use natural splines with the **'end_date_num'** (days from the earliest **'end_date'** to the latest **'end_date'**) as a predictor, allowing for the identification of non-linear trends influenced by campaign events and news cycles.
 Additionally, state-level effects are incorporated as random factors, enabling the models to account for unique political landscapes across the battleground states.
@@
 The Beta distribution is ideal for modeling this proportion data, as it allows us to avoid unrealistic predictions outside the (0, 1) range.
-To capture non-linear trends in voter support as Election Day approaches, we employe natural splines with **end_date_num**, which represents the number of days since the earliest poll date.
+To capture non-linear trends in voter support as Election Day approaches, we employe natural splines with **'end_date_num'**, which represents the number of days from earliest **'end_date'** to latest **'end_date'**.
 Additionally, incorporating state-specific random effects acknowledges the variations in voter preferences across different regions, allowing for more accurate modeling of baseline support levels.
@@
 It provides us with a high level understanding of the current voter dynamic.
-The pct results are aggregated averages of all pct data points for Trump and Harris in each state.
-By comparing the percentage of support for T.rump and Harris, we can determine the candidate that has a leading advantage in winning each state.
-For example,Trump has 49.2% support in Arizona while Harris has 46.8% support, which means Trump has won the state of Arizona.
+The percentages of support for Trump and Harris are aggregated averages of all **'pct'** data points for Trump and Harris in each state.
+By comparing the percentage of support for Trump and Harris, we can determine the candidate that has a leading advantage in winning each state.
+For example, Trump has 49.2% support in Arizona while Harris has 46.8% support, which means Trump has won the state of Arizona.
 ```{r}
@@
 ```
-This distribution is crucial for evaluating the consistency and breadth of the polling data.
-Pollsters with a higher count of polls, like Emerson, contribute substantially to the dataset and can have a greater influence on the model's predictions, particularly if their methodology or sample selection differs from others.
-On the other hand, pollsters with fewer entries may provide valuable but limited data points, potentially adding unique perspectives but with less influence on overall trends.
 
@@
 Across several states, there are noticeably increased fluctuations in polling averages for both candidates closer to the election particularly from late September to October.
-This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 4th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
+This pattern may indicate heightened voter interest and engagement as the election date approaches which is November 5th, 2024, it could also reflect changes in public opinion due to campaign events, political debates, or other external factors.
 ```{r}
@@
   kable_styling("striped", full_width = F)
-
 ```
@@
 ## US Map Showing Predicted Winner by State on Election Day
-To visualize the colour distribution of the predicted results, we emloy the US map from [@usmap] to display the results from @tbl-state-probabilities.
+To visualize the colour distribution of the predicted results, we emloy the US map [@usmap] to display the results from @tbl-state-probabilities.
 
@@
 This bias highlights a tendency to focus on recent events while overlooking broader trends.
-For example, in 2012, Barack Obama outperformed polls, a fact often overshadowed by recent election narratives[@discussion-article3].
+For example, in 2012, Barack Obama outperformed polls, a fact often overshadowed by recent election narratives [@discussion-article3].
 This can skew how people respond to polls and how pollsters interpret results, introducing potential polling errors that our model cannot fully account for.
@@
 Social stigma may deter some voters from expressing conservative support openly, leading to under-reported Trump support in polls.
-Additionally, Trump’s supporters often have lower social trust, making them less likely to participate in polls altogether[@discussion-article3].
+Additionally, Trump’s supporters often have lower social trust, making them less likely to participate in polls altogether [@discussion-article3].
 If certain groups systematically avoid participation, polling data may under-represent these voters, resulting in an underestimation of Trump’s support in our model.
 
-Harris faces potential challenges such as the Bradley effect and gender-related bias.
+On the other hand, Harris faces potential challenges such as the Bradley effect and gender-related bias.
 As a Black female candidate, Harris may encounter undecided voters leaning against her due to underlying biases, similar to what Hillary Clinton faced in 2016 [@discussion-article3].
@@
 In contrast, significant deviations may reveal areas for model improvement or highlight limitations in our approach.
-This diagnostic process aims to enhance the credibility of our findings and deepen our understanding of the electoral landscape as the 2024 Presidential Election approaches
+This diagnostic process aims to enhance the credibility of our findings and deepen our understanding of the electoral landscape as the 2024 Presidential Election approaches.
+
+In addition to the Bayesian Beta regression model, other models such as generalized linear models (GLMs) with logistic regression could be employed to predict voter support; however, these may not adequately capture the non-linear trends inherent in polling data.
+Alternatively, machine learning approaches like random forests or gradient boosting could provide robust predictions by leveraging complex interactions in the data, but they may sacrifice interpretability compared to the Bayesian framework.
+Ultimately, the choice of model depends on the trade-off between flexibility, interpretability, and the specific characteristics of the data at hand.
 
@@
 
-### Strengths of the NYT/Siena College questionnaires
-First, they have a comprehensive question design.
+### Strengths of the NYT/Siena College Questionnaires
+
+- They have a comprehensive question design.
 NYT/Sienna College polls ask a wide range of questions that provides a detailed view of the respondent’s preferences and opinions.
 It helps produce multi-dimensional data that allows for in-depth analysis and segmentation of voter groups.
-Second, the questionnaires are tailored to likely voters, as identified through the L-2 voter file.
+- The questionnaires are tailored to likely voters, as identified through the L-2 voter file.
 This means that the voters are more likely to be up-to-date with relevant policies proposed by different parties, and have formed their own opinions about them, this improves the predictive reliability of the poll.
-Third, the questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.
-
-### Weaknesses of the NYT/Siena College questionnaires
-First, the length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls.
+- The questionnaires contain detailed demographic questions about the respondents which allow NYT/Siena College to stratify responses by voter subgroups, identifying varied voter bases and diverse issues.
+
+### Weaknesses of the NYT/Siena College Questionnaires
+
+- The length and complexity of the questionnaires might cause the respondents to be less likely to respond or finish the polls.
 Respondents may rush through questions or drop out before completion.
 Respondents may also be prone to moderacy response bias where they have the tendency to choose middle options regardless of question content, as well as response order bias where respondents choose answers based on their order.
-Second, questionnaires have a dependence on self-reported likelihood to vote.
+- Questionnaires have a dependence on self-reported likelihood to vote.
 Self-reported likelihood can be unreliable, especially for individuals whose intentions to vote may change close to the actual election date.
 This can reduce the accuracy of predictions based on likely voter models.
-Third, there are limited open-ended questions.
+- There are limited open-ended questions.
 Many voters may have mixed opinions on specific issues or policies that cannot be explained in a single-choice answer.
@@
 presidential election.
-To demonstrate our idealized methodology, we generated a survey on the 2024 US Presidential election, which can be accessed via the the URL provided in [@US-election-survey] under [References -@sec-references].
+We aim to align these methods closely with real-world considerations to ensure it is efficient and represents the views of a broad population well, within a budget of $100,000.
+To demonstrate our idealized methodology, the following survey is generated, which can be accessed via the the URL provided in [@US-election-survey] under [References -@sec-references].
+
+## Target Population, Sampling Frame, and Sample
+
+- **Target population**: The target population consists of all registered voters across the United States who are eligible to vote in the upcoming election.
+Since the focus is on achieving state-level accuracy, this population includes individuals across all states, spanning various demographic groups, household statuses, and voting behaviors.
+- **Sampling frame**: The sampling frame is the voter file containing demographic and historical voting information about registered voters.
+This voter file is likely provided by a third-party data vendor (such as L-2 or a similar organization) and includes essential information like age, gender, race/ethnicity, household status, party registration, and voting history.
+The voter file ensures access to a comprehensive, up-to-date list that reflects the characteristics of the registered voting population.
+- **Sample**: The sample consists of individuals selected from each stratum within each state.
+We will select 100 respondents from each stratum identified in the sampling frame, ensuring representation across key categories such as age groups, gender, race/ethnicity, household status, and party affiliation.
+By carefully selecting individuals from each stratum, we ensure that the sample reflects the voting intention of the target population, enabling us to make accurate predictions about voter behavior.
+
+## Sampling Methodology & Justification
+Using data from voter files, which contain demographic information about registered voters, we aim to ensure that our sample accurately represents the population.
+To achieve this, we will use stratified random sampling.
+Given the importance of state-level forecasts due to the Electoral College system, we will first stratify by state.
+Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographics critical for predicting voter behavior.
+The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).
+Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth.
+Voting history will be classified into those who voted in the previous election and non-voters, while party registration will include Democrats, Republicans, and Independents.
+
+This stratified sampling approach ensures that our sample accurately reflects key demographics, essential for making reliable state-level predictions.
+By segmenting the sample by state and further by demographic groups—such as age, gender, race/ethnicity, household status, voting history, and party registration—we ensure that our sample reflects the key characteristics of the population.
+This method not only enhances the reliability of our state-level predictions, which is crucial due to the Electoral College system but also optimizes resources to be cost-effective to capture all significant voter demographics within budget.
+
+## Sample Recruitment Method
+
+## Survey Design
+The survey will be designed to capture critical insights into voter preferences, opinions, and behaviors relevant to the 2024 election.
+It contains questions on candidates, key political issues, and demographic information, enabling a detailed analysis of trends within each stratum.
+To reduce survey fatigue and encourage high-quality responses, we prioritized concise, clear questions.
+To ensure inclusivity and ease of response, the survey will be accessible online and formatted for both desktop and mobile use.
+
+The survey questions capture both quantitative and qualitative data that reflect voter intentions and behavior.
+By including a range of topics—from candidate preference to issue prioritization and engagement with political news.
+Each question is intended to collect actionable information that contributes to understanding voter stability and sentiment, which are all key in forming reliable predictions for the election.
+The scale-based approach allows respondents to express the intensity of their preferences, providing more nuanced data than simple yes-or-no answers.
+The questions are direct and structured to maintain respondent engagement, minimizing the risk of survey fatigue.
+Additionally, by setting a hypothetical match-up between prominent candidates- Kamala Harris and Donald Trump, we introduce a realistic context that grounds responses in potential voting scenarios, which adds to the predictive power of our data.
+Also, a validation question is asked to ensure our data remains reliable and accurate.
+It measures if the responder maintains integrity and attentiveness.
+
+## Data Validation
+After completing the survey, we will adjust the data through weighting to ensure that our samples accurately reflect the broader population for predictive purposes.
+Each stratum will be assigned a weight proportionally based on its size.
+For each group, we will calculate the weight using the formula:
+
+$$
+\text{Weight} = \frac{\text{Population Proportion}}{\text{Sample Proportion}}
+$$
+This approach assigns higher weights to underrepresented groups and lower weights to overrepresented groups, ensuring that each group's influence in the sample aligns with its actual population [@idealized].
+For instance, if young voters (ages 18-29) constitute 30% of the population but only make up 20% of our survey sample, this demographic is considered underrepresented.
+In this case, we would assign a larger weight to young voters, increasing their influence to better reflect their true proportion in the population.
+Additionally, extra weight will be given to respondents from strata who are less likely to participate in surveys.
+To avoid duplicate responses, each voter will be assigned a unique ID, retaining only the first response from each ID.
+To ensure that selected voters are completing the survey correctly, we will track responses in real time using a centralized system.
+Additionally, we will include validation questions in the survey to catch careless or fraudulent responses.
 
@@
 
-### Sampling Approach
-Using data from voter files, which contain demographic information about registered voters, we aim to ensure that our sample accurately represents the population.
-To achieve this, we will use stratified random sampling.
-Given the importance of state-level forecasts due to the Electoral College system, we will first stratify by state.
-
-Within each state, we will apply stratified sampling again, dividing the population into 6 groups and selecting 100 random samples from each stratum within those groups, ensuring our sample captures key demographic critical for predicting voter behavior.
-The groups and their respective strata include age groups (18-29, 30-44, 45-64, and 65+), gender (Male, Female, and Other), and race/ethnicity (White, Black, Latino, Asian, and Other).
-Additionally, household status will be categorized as either renting or owning, with home ownership serving as a proxy for wealth.
-Voting history will be classified into those who voted in the previous election and non-voters, while party registration will include Democrats, Republicans, and Independents.
-
-For the recruitment of participants, we will begin by sending survey invitations to our selected voters via email, offering a $20 incentive to encourage participation.
-To improve response rates and reduce non-response bias, we will send a reminder email three days later.
-This approach will help minimize expenses to some extent.
-If we do not receive a response within a week (7 days), we will follow up by phone to reach those who did not respond to the email invitation.
-
-## Data Validation
-
-After completing the survey, we will adjust the data through weighting to ensure the sample accurately reflects the broader population for predictive purposes.
-More weight will be assigned proportionally to each stratum based on its size, and additional weight will be given to respondents from strata that are less likely to participate in surveys.
-And to avoid duplicate responses, each voter will be assigned a unique ID with only the first response from each ID being retained.
-To ensure the selected voters are completing the survey correctly, we will  track responses in real-time using a centralized system.
-Additionally, we will include validation questions in the survey to catch careless or fraudulent responses.
-
 \newpage

### 2024-11-04 scripts/01-test_simulated_data.R

@@ -1,57 +1,59 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Conducts data quality tests on simulated_data to ensure the quality and validity of simulated data
+# Authors: Jimin Lee, Sarah Ding, Xiyann Chen
+# Date: 4 November 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca, sarah.ding@mail.utoronto.ca, xiyan.chen@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: R Packages (tidyverse, janitor, lubridate, arrow, testthat), simulated_data
+# Additional Information: N/A
 
 
-#### Workspace setup ####
+# Load necessary libraries
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-# Test 1: Check Support Percentage Consistency
-test_support_percentage <- function(data) {
-  support_summary <- data %>%
-    group_by(candidate) %>%
-    summarise(avg_support = mean(support_percent), .groups = 'drop')
+# Load the simulated data
+simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+#### Testing script ####
+
+# Test 1: Check the number of rows and columns
+test_that("Simulated data has the correct number of rows and columns", {
+  expect_equal(nrow(simulated_data), 1000)
+  expect_equal(ncol(simulated_data), 6)  # poll_id, pollster, state, sample_size, candidate, support_percent
+})
+
+# Test 2: Check for unique poll IDs
+test_that("Poll IDs are unique", {
+  expect_equal(length(unique(simulated_data$poll_id)), 1000)
+})
+
+# Test 3: Verify sample sizes are within specified range
+test_that("Sample sizes are within the range of 500 to 3000", {
+  expect_true(all(simulated_data$sample_size >= 500 & simulated_data$sample_size <= 3000))
+})
+
+# Test 4: Verify support percentages based on candidate
+test_that("Support percentages are correctly assigned", {
+  # Check support percentages for Donald Trump
[…]

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Conducts data quality tests on claen_poll_data to insure quality and validity of data being used for trump_model.rds and harris_model.rds
+# Author: Jimin Lee, Sarah Ding, Xiyann Chen
+# Date: 26 September 2024
+# Contact: jamiejimin.lee@mail.utoronto.ca, sarah.ding@mail.utoronto.ca, xiyan.chen@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: R Packages (tidyverse, janitor, lubridate, arrow, testthat), clean_poll_data
+# Additional Information: N/A
 
 
 #### Workspace setup ####

### 2024-11-04 paper/paper.qmd
committing text changes to paper.qmd

@@
 
-Data cleaning, analysis, modeling, and visualization for this paper were conducted using the statistical programming language `R` [@citeR] and the following packages: `tidyverse` package of @Tidyverse , `janitor` package of @Janitor for data cleaning, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models, `usmap` package of @usmap for mapping U.S.
-states, `arrow` package of @arrow for data integration with Apache Arrow, `lubridate` package of @Lubridate for date handling, `rstanarm` package of @rstanarm and `brms` package of @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret`package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.
-
 # Data {#sec-data}
@@
 
+Data cleaning and analysis are conducted by employing the following packages: `tidyverse` package of @Tidyverse, `janitor` package of @Janitor, `lubridate` package of @Lubridate,  `arrow` package of @arrow
+
 ## Measurement
@@
 - **'Candidate Name'**: Identifies the running candidate, specifically focusing on Donald Trump and Kamala Harris.
-- **Percentage of Support (pct)**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.
+- **Percentage of Support ('pct')**: Represents the percentage of respondents supporting each candidate, which is the primary outcome variable of interest.
 
@@
 ```
-@tbl-outcomevar reveals a total of 107 polls conducted, calculated as the total count of rows in our cleaned dataset.
-‘Total polls’ reflects the total number of unique poll entries after filtering for only battleground states, polls with sufficient data quality, and polls that report support percentages for both Trump and Harris.
-The 107 polls conducted display an average support level of 47.801% for Trump and 47.999% for Harris.
-This suggests that based on the dataset, Trump has a leading advantage in winning the presidential election race by having a slighty higher average level of support than Harris.
-The highest percentage of support for Trump in the 107 polls was 51.700% while the lowest recorded support was 43%.
-For Harris, the highest percentage of support also reached 51.200%, with a minimum of 43.600%.
-These statistics illustrate the competitive landscape of the presidential election race, without a clear leading candidate in the current race.
-
-It is important to note that since **'pct'** stands for percentage, there may be a misconception that the values for Trump and Harris will always add up to 100%.
-For instance, in the summary statistics presented in Table 1, the average support for Donald Trump is 47.801%, while the average support for Kamala Harris is 47.999%.
-This results in a combined average support of 95.800%, indicating that there is a significant portion of respondents who are either undecided or supporting other candidates.
+@tbl-outcomevar reveals a total of 107 polls conducted, representing unique poll entries filtered for battleground states, sufficient data quality, and reported support percentages for both Trump and Harris.
+Among these, the average support level is 47.801% for Trump and 47.999% for Harris, indicating a slight advantage for Trump in the presidential race.
+The highest recorded support for Trump is 51.700%, while the lowest is 43%; for Harris, the highest is 51.200% and the lowest is 43.600%.
+These figures illustrate a competitive landscape without a clear leading candidate.
+It is essential to note that the percentages for Trump and Harris do not sum to 100%, as the combined average of 95.800% highlights a significant portion of respondents who are undecided or supporting other candidates.
 This highlights the competitive nature of the election and emphasizes the necessity of examining not only the percentages attributed to the main candidates but also the context of voter preferences that could influence the final outcome.
@@
 #### State
+The state (**'state'**) variable indicates the U.S.
+state where the poll was conducted or targeted.
+In the context of the U.S.
+presidential election, the outcome is frequently determined by several key battlegroud states.
+In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
+To enhance our model and refine our predictions, we focused exclusively on these seven battleground states.
+The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
 ```{r}
@@
 ```
-The state (**'state'**) variable indicates the U.S.
-state where the poll was conducted or targeted.
-In the context of the U.S.
-presidential election, the outcome is frequently determined by several key battlegroud states.
-In the current 2024 election cycle, there are seven battleground states that are pivotal in the contest between Kamala Harris, representing the Democratic Party, and Donald Trump, representing the Republican Party.
-To enhance our model and refine our predictions, we focused exclusively on these seven battleground states.
-The polling results from each of these states play a crucial role in predicting electoral outcomes, as the percentage of support is evaluated on a state-by-state basis.
 The distribution of the state variable is visualized in @fig-state-distribution and indicates that some states, such as Pennsylvania, North Carolina and Wisconsin have more polling observations compared to others, like Nevada.
@@
 #### End Date {#sec-end-date}
-End date (**'end_date'**) represents the date the poll ended, essentially the date the poll stopped collecting data from electorates.
-In the context of our paper, the end date of each poll serves as one of the predictive variables and is used to assess changes in voter support over time.
-Including end date as a variable in our model can help us identify patterns in voter support as Election Day approaches, potentially reflecting the impact of campaign events, debates, or other political developments.
-@fig-end-date-distribution displays the distribution of polling end dates from July 2024 to October 2024.
+End date (**'end_date'**) marks the conclusion of data collection for each poll, serving as a predictive variable in our analysis of changes in voter support over time.
+This inclusion helps identify patterns as Election Day approaches, potentially reflecting the impact of campaign events, debates, or political developments.
 ```{r}
@@
 ```
-We filtered for the earliest end date as July 2024 since Harris officially announced her presidential campaign on July 21, 2024.
-Therefore, any polls that has an end date before July would not contain polling results for Harris.
-The data from Project 538 [@FiveThirtyEight] is last updated on October 15th, 2024, thus the latest end date being October 14th, 2024.
-Furthermore, @fig-end-date-distribution also reveals that polling activity is not evenly distributed over time.
-There are clear peaks where the number of polls conducted is higher, notably in early September, late September, as well as early October.
-This concentration of polls likely corresponds to period of heightened political interest or significant campaign events that prompt increased polling.
+@fig-end-date-distribution shows polling end dates from July 2024 to October 2024, with the earliest date set to July 2024, coinciding with Kamala Harris's official presidential campaign announcement on July 21, 2024.
+The latest end date, October 14, 2024, reflects the last data update from Project 538 [@FiveThirtyEight] on October 15, 2024.
+The distribution reveals that polling activity is not evenly spread, with notable peaks in early September, late September, and early October, likely corresponding to periods of heightened political interest or significant campaign events that drive increased polling activity.
 
@@
 #### Numeric Grade
-Numeric Grade (**'numeric_grade'**) represents the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors such as methodology, transparency, and historical accuracy.
-In the context of our paper, numeric grade allows us to weigh the reliability of different polls.
-By including numeric grade as a variable in our model, we aim to account for variations in poll quality, ensuring that polls with higher ratings (indicating higher reliability) contribute more significantly to our predictions of voter support.
+Numeric Grade (**'numeric_grade'**) assesses the quality of each poll, as rated by Project 538 [@FiveThirtyEight], based on factors like methodology, transparency, and historical accuracy.
+In our paper, this variable allows us to weigh the reliability of different polls, ensuring that those with higher ratings contribute more significantly to our voter support predictions.
 ```{r}
@@
 ```
-@fig-numgrade-distribution displays the distribution of numeric grades for the polls included in our analysis, ranging from 2.5 to 3.0.
-Polls with higher grades are generally seen as more methodologically rigorous and transparent, which helps improve the reliability of our model’s predictions.
-The distribution shows several peaks, indicating that certain grade levels (around 2.7, 2.8, and 3.0) are more common among the polling data we collected.
-This variability in numeric grade highlights the diversity in polling quality within our dataset.
-The presence of peaks suggests that some pollsters consistently receive higher ratings, likely due to their use of more rigorous methods or transparent reporting standards.
-Conversely, the valleys in the distribution may indicate fewer polls from sources with lower numeric grades, reflecting a potential tendency for lower-quality polls to be less prevalent or less frequently conducted.
-By accounting for numeric grade, we can improve the accuracy of predictions by emphasizing data from more reliable sources.
-This consideration is crucial, as polling quality can impact the stability of the model’s forecasts, especially in closely contested states where polling accuracy is essential for reliable predictions.
+@fig-numgrade-distribution illustrates the distribution of numeric grades in our analysis, which ranges from 2.5 to 3.0.
+Higher grades indicate more methodologically rigorous and transparent polls, enhancing the reliability of our model's predictions.
+The distribution shows several peaks around grades 2.7, 2.8, and 3.0, suggesting that certain pollsters consistently achieve higher ratings due to better methodologies or transparency.
+Conversely, valleys in the distribution may reflect a scarcity of polls from lower-rated sources.
+By incorporating numeric grade, we enhance prediction accuracy by prioritizing data from more reliable sources, which is particularly crucial in closely contested states where polling accuracy is essential for reliable forecasts.
 
 #### Transparency Score
-Transparency Score (**'transparency_score'**) represents the degree of methodological transparency provided by each pollster, rated on a scale from 0 to 10 by Project 538 [@FiveThirtyEight].
-This score reflects the extent to which pollsters disclose details about their data collection and analysis methods, such as sample size, weighting, and margin of error.
-In the context of our paper, transparency score is an essential variable for evaluating the reliability of polling data, as higher transparency scores generally indicate a greater level of methodological rigor and openness.
-@fig-transparency-score-distribution displays the distribution of transparency scores across the polling data for key battleground states.
-The plot shows that transparency scores cluster heavily around ratings of  7 and 9, suggesting that a majority of polls in our dataset come from sources that maintain a relatively high level of transparency.
-There are smaller peaks at lower scores (5 and 10), indicating that while some pollsters offer limited transparency, a few polls also achieve perfect transparency scores.
+Transparency Score (**'transparency_score'**) quantifies the methodological transparency of each pollster on a scale from 0 to 10, as rated by Project 538 [@FiveThirtyEight].
+This score reflects the extent of disclosure regarding data collection and analysis methods, including sample size, weighting, and margin of error.
+In our paper, the transparency score is crucial for assessing the reliability of polling data; higher scores typically indicate greater methodological rigor.
 ```{r}
@@
 ```
-This distribution is important for understanding the credibility of our data sources.
-Polls with higher transparency scores are more likely to provide consistent and reliable results, as they openly disclose their methodologies, allowing us to assess potential biases or limitations in their data.
-In contrast, polls with lower transparency scores may lack sufficient methodological detail, making it challenging to gauge their accuracy.
-Incorporating transparency scores into our model helps differentiate between high- and low-transparency polls.
-This adjustment is crucial as transparency can influence the stability of the model’s predictions, especially in closely contested states where the reliability of polling data may impact the accuracy of our forecasted election outcomes.
+@fig-transparency-score-distribution illustrates the distribution of transparency scores in our dataset for key battleground states, showing a concentration around scores of 7 and 9, indicating that most polls come from sources with relatively high transparency.
+Smaller peaks at lower scores (5) and a perfect score (10) suggest that while some pollsters provide limited transparency, a few achieve full disclosure.
+This distribution is vital for evaluating data credibility, as polls with higher transparency scores are more likely to yield consistent and reliable results, enabling us to identify potential biases.
+Incorporating transparency scores into our model helps differentiate between high- and low-transparency polls, which is essential for ensuring the stability of our predictions, particularly in closely contested states where polling accuracy is critical.
 
@@
 ## Model Setup
+Modeling processes are conducted by employing the following packages: `rstanarm` package of @rstanarm, `brms` package of  @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret` package of @caret for machine learning support, and `broom.mixed` package of @broom.mixed for tidying mixed models.
 
@@
 Furthermore, the model's structure allows for clear interpretation in terms of victory probabilities, aligning well with our goals of forecasting electoral outcomes.
-Further details are discussed in [Appendix -@sec-model-details].
+Details are further discussed in [Appendix -@sec-model-details].
+
+In addition to the Bayesian Beta regression model, other models such as generalized linear models (GLMs) with logistic regression could be employed to predict voter support; however, these may not adequately capture the non-linear trends inherent in polling data.
+Alternatively, machine learning approaches like random forests or gradient boosting could provide robust predictions by leveraging complex interactions in the data, but they may sacrifice interpretability compared to the Bayesian framework.
+Ultimately, the choice of model depends on the trade-off between flexibility, interpretability, and the specific characteristics of the data at hand.
 
@@
 ### Trump Model Summary
-@tbl-trump-model-summary displays fixed effects estimates for several key terms that capture the influence of time on his support in the battleground states.
-The intercept, with an estimate of -0.0923, represents the baseline log-odds of Trump’s support when other predictors are held constant.
-This negative intercept suggests a slightly lower baseline level of support.
-The term nsenddatenumdfEQ42, with an estimate of 0.0803 and a statistically significant positive confidence interval (0.0179 to 0.1426), indicates an increase in support around this time period.
-Similarly, nsenddatenumdfEQ44 has an estimate of 0.0789 and a confidence interval above zero (0.0014 to 0.1589), also showing a positive effect on support.
-Other terms such as nsenddatenumdfEQ41 and nsenddatenumdfEQ43 have less significant impacts, suggesting that certain time periods were more influential than others in shaping Trump’s support.
-The standard errors for the fixed effects terms range from 0.0306 to 0.0676 for nsenddatenumdfEQ43.
-The relatively low standard error of 0.0306 on the intercept suggests that the baseline estimate is quite stable, meaning we can be more confident about this baseline support level.
-However, the standard error for nsenddatenumdfEQ43 (0.0676) is higher than for other terms, indicating that this estimate is less precise.
+The summary in @tbl-trump-model-summary presents fixed effects estimates reflecting the influence of time on Trump's support in battleground states.
+The intercept, estimated at -0.0923, indicates a slightly lower baseline log-odds of support when other predictors are held constant.
+The term nsenddatenumdfEQ42 shows an estimate of 0.0803 with a statistically significant confidence interval (0.0179 to 0.1426), suggesting an increase in support during this period.
+Similarly, nsenddatenumdfEQ44 has an estimate of 0.0789 and a confidence interval above zero (0.0014 to 0.1589), also indicating a positive effect.
+In contrast, nsenddatenumdfEQ41 and nsenddatenumdfEQ43 exhibit less significant impacts, suggesting variability in time periods' influence on support.
+Standard errors for the fixed effects range from 0.0306 to 0.0676, with the lower standard error of 0.0306 for the intercept indicating a stable baseline estimate.
+However, the higher standard error of 0.0676 for nsenddatenumdfEQ43 reflects less precision in that estimate.
+Overall, these results imply that Trump's support fluctuated over time, likely in response to specific campaign events, with significant terms indicating changes in voter preference effectively captured by the model.
+Standard errors further inform the reliability of these estimates.
 ```{r}
@@
 ```
-These results imply that support for Trump varied over time in ways that may correspond to specific campaign events or other influences affecting voter sentiment.
-The significance of terms like nsenddatenumdfEQ42 and nsenddatenumdfEQ44 suggests that changes in voter preference could be captured effectively by this model.
-And by checking standard errors, helping us determine how much we can rely on those data.
+Overall, these results imply that Trump's support fluctuated over time, likely in response to specific campaign events, with significant terms indicating changes in voter preference effectively captured by the model.
+Standard errors further inform the reliability of these estimates.
 
@@
 @tbl-harris-model-summary similarly shows the impact of time on her support in the battleground states.
-The intercept for the Harris model is estimated at -0.1508, suggesting a slightly lower baseline support level compared to Trump(-0.0923).
-The term nsenddatenumdfEQ41 has an estimate of 0.1225 and a confidence interval 0.0553 to 0.1903, indicating a significant increase in support during this period.
-Another positive and significant effect is seen in nsenddatenumdfEQ43, with an estimate of 0.1191 and confidence interval from -0.0071 to 0.2417, which suggests a trend toward increased support.
-While other terms stay positive, they have less impacts on support.
-Additionally, the standard errors for fixed effects are generally smaller, ranging from 0.0279 for the intercept to 0.0631 for nsenddatenumdfEQ43.
-The larger standard error for nsenddatenumdfEQ43 (0.0631) suggests that there would be some uncertainty in how this time point influences support, though the positive estimate still points toward an upward trend.
-In comparison to the Trump model, the standard errors in the Harris model are slightly smaller for most terms, indicating potentially more consistent support trends over time in the data analyzed.
+The intercept is estimated at -0.1508, indicating a lower baseline support level compared to Trump's intercept of -0.0923.
+The term nsenddatenumdfEQ41 shows an estimate of 0.1225 with a confidence interval of 0.0553 to 0.1903, reflecting a significant increase in support during this period.
+Another notable effect is observed in nsenddatenumdfEQ43, which has an estimate of 0.1191 and a confidence interval from -0.0071 to 0.2417, suggesting a trend toward increased support.
+While other terms remain positive, they have less impact.
+The standard errors for fixed effects range from 0.0279 for the intercept to 0.0631 for nsenddatenumdfEQ43, with the larger standard error for nsenddatenumdfEQ43 indicating some uncertainty about its influence on support, despite the positive estimate suggesting an upward trend.
+Compared to the Trump model, the Harris model's standard errors are generally smaller, indicating more consistent support trends over time.
 ```{r}
@@
 ```
-These results highlight that Harris's support also fluctuated over time, with nsenddatenumdfEQ41 and nsenddatenumdfEQ43 marking periods of notable increases.
-Helping us understand the pattern, this model predicts Harris's probability of victory and provides insights for future changes in support as the election nears.
+These results reveal fluctuations in Harris's support, particularly during the significant periods marked by nsenddatenumdfEQ41 and nsenddatenumdfEQ43, which aid in predicting her probability of victory and potential shifts in support as the election approaches.
 
 # Results {#sec-results}
-In this section, we present the results of our analysis, focusing on the electoral support for Donald Trump and Kamala Harris across key battleground states during the 2024 Presidential Election cycle.
-We begin by examining the state-level polling averages for both candidates, which provide a snapshot of voter preferences as captured by various pollsters.
-Next, we explore how polling averages have evolved over time, highlighting trends that may influence election outcomes.
-We then assess the predicted probabilities of Trump winning on Election Day, followed by a detailed look at the likelihood of winning by state.
+In this section, we present the analysis of electoral support for Donald Trump and Kamala Harris in key battleground states during the 2024 Presidential Election cycle.
+We start by examining state-level polling averages for both candidates, providing insights into voter preferences as reported by various pollsters.
+We then analyze the evolution of these polling averages over time, identifying trends that could affect election outcomes.
+Following this, we assess the predicted probabilities of Trump winning on Election Day and examine the likelihood of winning by state.
 Finally, we visualize the predicted winner by state on a US map, summarizing our findings and their implications for the electoral landscape.
+Visualizations of results are generated by employing the following packages: `usmap` package of @usmap for mapping U.S.
+states, `kableExtra` package of @kableExtra for dynamic report generation and table formatting, `modelsummary` package of @Modelsummary for summarizing models.
 
@@
 If certain groups systematically avoid participation, polling data may under-represent these voters, resulting in an underestimation of Trump’s support in our model.
-
 On the other hand, Harris faces potential challenges such as the Bradley effect and gender-related bias.
@@
 The data cleaning process for our analysis focused on refining the raw polling data obtained from Project 538's polling project [@FiveThirtyEight].
-The raw data contains polling information for various U.S.
-states, but for our specific analysis, we concentrated on seven designated battleground states: Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-This is due to the fact that these battleground states have been recognized as pivotal in determining this year's election outcome.
-
-To prepare the data for modeling, we performed several key steps.
-First, we filtered the data to retain only relevant columns, including state, polling dates, pollster information, numeric grades, transparency scores, candidate names, and their respective support percentages.
-This filtering process ensured that we excluded any unnecessary data that could complicate our analysis.
-Next, we implemented thresholds for the numeric grade (minimum of 2.5) and transparency score (greater than 5) to ensure that only high-quality polling data was considered.
-We also converted polling dates into a standardized date format and created a new numeric variable representing the number of days since the earliest polling date.
-Additionally, we handled specific data inconsistencies, such as replacing "Nebraska CD-2" with "Nebraska." The polling percentages were then aggregated by averaging the values for each candidate while preserving other important attributes such as pollster and state.
-This aggregation allowed us to focus on polling data that represented a balanced view of voter support across the selected battleground states.
-The cleaned dataset was then saved in a parquet format for efficient processing in subsequent analyses.
-
-To ensure the integrity of our cleaned polling dataset, we implemented a comprehensive data testing process.
-This included verifying the presence of all required columns, checking data types for consistency, and ensuring there were no missing values in critical fields.
-We validated that candidate support percentages fell within the logical range of 0 to 100 and confirmed that numeric grades and transparency scores met predefined thresholds.
-By rigorously applying these tests, we ensured that the dataset is reliable and suitable for subsequent modeling and analysis, providing a solid foundation for our exploration of voter behavior in the battleground states.
+We concentrated on seven key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—recognized as pivotal for this year's election outcome.
+Key steps included filtering relevant columns such as state, polling dates, pollster information, numeric grades, transparency scores, candidate names, and support percentages to eliminate unnecessary data.
+We set thresholds for numeric grades (minimum of 2.5) and transparency scores (greater than 5) to ensure the inclusion of high-quality polling data.
+Polling dates were standardized, and we created a numeric variable representing the number of days since the earliest polling date.
+Additionally, we addressed data inconsistencies, like replacing "Nebraska CD-2" with "Nebraska," and aggregated polling percentages by averaging candidate support while preserving essential attributes.
+The cleaned dataset was saved in parquet format for efficient processing in subsequent analyses.
+
+To ensure the integrity of our cleaned data, we implemented a comprehensive data testing process.
+This involved verifying the presence of required columns, checking data types for consistency, and ensuring no missing values in critical fields.
+We validated that candidate support percentages fell within the logical range of 0 to 100 and confirmed that numeric grades and transparency scores met our predefined thresholds.
+By rigorously applying these tests, we established that the data is reliable and suitable for modeling and analysis, providing a solid foundation for exploring voter behavior in the selected battleground states.
 
@@
 
-In addition to the Bayesian Beta regression model, other models such as generalized linear models (GLMs) with logistic regression could be employed to predict voter support; however, these may not adequately capture the non-linear trends inherent in polling data.
-Alternatively, machine learning approaches like random forests or gradient boosting could provide robust predictions by leveraging complex interactions in the data, but they may sacrifice interpretability compared to the Bayesian framework.
-Ultimately, the choice of model depends on the trade-off between flexibility, interpretability, and the specific characteristics of the data at hand.
-
 ## Posterior Predictive Check for the Trump Model
@@
 
-### Model Diagnostics - Harris Model
+### Posterior Predictive Check for the Harris Model
 The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.

### 2024-11-04 paper/paper.qmd
committing text changes to paper.qmd in appendix

@@
 ## Overview
-The data for this paper contains the polling information of the 2024 Presidential Election cycle and was obtained from Project 538 [@FiveThirtyEight] on October 19th, 2024.
-The data is periodically updated throughout the presidential election campaign to reflect current and most up-to-date polling results.
-It consists of polling information from various pollsters such as the New York Times, ActiVote, Morning Consult, etc.
-For every poll conducted, relevant variables such as **'state'**, **'start_date'**, **'end_date'**, **'transparency_score'**, **'candidate_name'**, and percentage of support (**'pct'**) for the appropriate candidates are included.
-The data compiles information from all major pollsters in the US and provides us with an understanding of the current state of the presidential election.
-
-The present analysis focuses on Donald Trump and Kamala Harris, the two leading candidates in the current presidential election race.
-For all the variables mentioned (see @sec-variables), we collected observations for both of the candidates.
-We collect observations from the data where the **‘candidate_name’** is Trump and Harris.
-We are interested in the support for both candidates over time, for each state, therefore, we collect observations of **‘end_date’** and **‘state’**, as well as the electorate support for each candidate, which is denoted as **‘pct’**.
-Furthermore, since there are many pollsters present in the data, we extracted pollsters with high-quality data that is reliable, which is denoted as having a high **‘numeric_grade’** and **‘transparency_score’**, details of data cleaning are further explained in [Appendix -@sec-data-details].
+The data for this paper, obtained from Project 538 [@FiveThirtyEight] on October 19, 2024, encompasses polling information for the 2024 Presidential Election cycle, reflecting the most current polling results from various pollsters, including the New York Times, ActiVote, and Morning Consult.
+Key variables include **'state'**, **'start_date'**, **'end_date'**, **'transparency_score'**, **'candidate_name'**, and 'percentage of support (**'pct'**) for each candidate.
+Our analysis focuses on Donald Trump and Kamala Harris, the leading candidates, collecting observations where ‘candidate_name’ matches either Trump or Harris.
+We aim to track support over time for each candidate by extracting relevant data on **‘end_date’**, **‘state’**, and **‘pct’**.
+To ensure reliability, we included only polls from pollsters with high **‘numeric_grade’** and **‘transparency_score’**.
+Details on the data cleaning process are provided in [Appendix -@sec-data-details].
 
@@
 This initial conversion may not capture voters’ internal conflicts or reservations, as respondents must choose a definitive option.
-
 These individual responses are aggregated into percentages, treating each “Trump” or “Harris” response equally, regardless of confidence level, further smoothing out complexities.
@@
 #| label: tbl-outcomevar
-#| tbl-cap: "Summary Statistics for Trump and Harris Polling Data - Presents key summary statistics, including total polls, average support percentages, and the maximum and minimum polling levels for both candidates across the specified battleground states.
-All values are rounded to three decimal places, providing an overview of the polling landscape as of the selected election date."
+#| tbl-cap: "Summary statistics of polling data, showing a total of 107 polls conducted.
+Average support is 47.801% for Donald Trump and 47.999% for Kamala Harris, with similar maximum and minimum support levels, indicating a competitive electoral landscape."
 
@@
 ## Model Setup
-Modeling processes are conducted by employing the following packages: `rstanarm` package of @rstanarm, `brms` package of  @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret` package of @caret for machine learning support, and `broom.mixed` package of @broom.mixed for tidying mixed models.
+Modeling processes are conducted by employing the following packages: `rstanarm` package of @rstanarm, `brms` package of  @brms for Bayesian regression modeling, `Matrix` package of @Matrix for handling sparse and dense matrices, `splines` package of @splines for regression splines, `caret` package of @caret for machine learning support, and `broom.mixed` package of @broommixed for tidying mixed models.
 
@@
 ## Sample Recruitment Methods
-Each poll is conducted by phone using live interviewers at call centers based in Florida, New York, South Carolina, Texas, and Virginia.
-Most people that the live interviewers call do not answer the call, and some don’t finish the survey, leaving only about two percent both answer their phone and complete the poll.
-To mitigate this:
-First, the pollster keeps calling numbers until someone picks up, even though it can take hours–a reason why a high-quality poll is time-consuming and expensive to conduct. 
-Second, voter files are used to help make sure the pollster is reaching a representative sample; for example, they may reach out to a higher percentage of people who are less likely to respond. 
-Third, pollster mathematically adjusts the responses they do receive so they are representative of the wider population, it includes adjusting by age, race, gender, voting history and etc.
+We will initiate our survey by sending invitations to the 2,000 selected voters via email.
+To enhance response rates and reduce non-response bias, we will send a reminder email three days later.
+If we do not receive a response within a week (7 days), we will follow up by phone to reach those who did not respond to the email invitation.
+To effectively manage the recruitment process, we will allocate `$100,000` of the budget for staffing.
+We will employ a small team of around 30 people responsible for sending invitations during the first three days and following up with non-respondents over the entire two-week process.
+Each employee will be paid `$24` per hour for 40 hours per week for two weeks, resulting in a total of 40 hours worked per employee and `$960` for each, `$28,800` in total.
+Additionally, we will allocate `$86,000` for the phone follow-up process as well as the data cleaning and analysis process.
+We will employ a team of 20 data scientists to perform data engineering and analysis, with them being paid $3560 each for the duration of their work term, which will be around three to five weeks.
+
+Overall, this recruitment method is designed to be both efficient and effective.
+By reaching out to a diverse group of voters and encouraging participation through personalized follow-ups, we can collect high-quality data that accurately reflects the opinions and behaviors of the broader population.
+Additionally, this recruitment method minimizes early-stage labor costs by using email as the initial contact method, reserving more follow-up efforts for those who have not yet responded after a week.
+This approach not only helps a higher participation rate but also improves the reliability of the data collected.
 

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd for formatting purposes

@@
 This inclusion helps identify patterns as Election Day approaches, potentially reflecting the impact of campaign events, debates, or political developments.
+@fig-end-date-distribution shows polling end dates from July 2024 to October 2024, with the earliest date set to July 2024, coinciding with Kamala Harris's official presidential campaign announcement on July 21, 2024.
+The latest end date, October 14, 2024, reflects the last data update from Project 538 [@FiveThirtyEight] on October 15, 2024.
+The distribution reveals that polling activity is not evenly spread, with notable peaks in early September, late September, and early October, likely corresponding to periods of heightened political interest or significant campaign events that drive increased polling activity.
 ```{r}
@@
 ```
-@fig-end-date-distribution shows polling end dates from July 2024 to October 2024, with the earliest date set to July 2024, coinciding with Kamala Harris's official presidential campaign announcement on July 21, 2024.
-The latest end date, October 14, 2024, reflects the last data update from Project 538 [@FiveThirtyEight] on October 15, 2024.
-The distribution reveals that polling activity is not evenly spread, with notable peaks in early September, late September, and early October, likely corresponding to periods of heightened political interest or significant campaign events that drive increased polling activity.
 

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd for formatting purposes

@@
 In our paper, this variable allows us to weigh the reliability of different polls, ensuring that those with higher ratings contribute more significantly to our voter support predictions.
+@fig-numgrade-distribution illustrates the distribution of numeric grades in our analysis, which ranges from 2.5 to 3.0.
+Higher grades indicate more methodologically rigorous and transparent polls, enhancing the reliability of our model's predictions.
+The distribution shows several peaks around grades 2.7, 2.8, and 3.0, suggesting that certain pollsters consistently achieve higher ratings due to better methodologies or transparency.
+Conversely, valleys in the distribution may reflect a scarcity of polls from lower-rated sources.
+By incorporating numeric grade, we enhance prediction accuracy by prioritizing data from more reliable sources, which is particularly crucial in closely contested states where polling accuracy is essential for reliable forecasts.
 ```{r}
@@
 ```
-@fig-numgrade-distribution illustrates the distribution of numeric grades in our analysis, which ranges from 2.5 to 3.0.
-Higher grades indicate more methodologically rigorous and transparent polls, enhancing the reliability of our model's predictions.
-The distribution shows several peaks around grades 2.7, 2.8, and 3.0, suggesting that certain pollsters consistently achieve higher ratings due to better methodologies or transparency.
-Conversely, valleys in the distribution may reflect a scarcity of polls from lower-rated sources.
-By incorporating numeric grade, we enhance prediction accuracy by prioritizing data from more reliable sources, which is particularly crucial in closely contested states where polling accuracy is essential for reliable forecasts.
 
@@
 ```
-@fig-start-date-distribution displays the distribution of polling start dates from late July to mid-October 2024, showing when polling efforts began across key battleground states.
-The histogram reveals peaks in polling activity, particularly in mid-August, early September, and late September, indicating periods of heightened interest or campaign events that likely prompted increased polling.
-This distribution provides insights into the timing and intensity of polling efforts throughout the campaign season.
-By including only polls initiated after Harris’s campaign launch, we capture voter preferences in a context where both major candidates are actively campaigning, thus making the polling data more relevant for direct comparisons.
-This filtered timeline ensures our model focuses on data that reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as we approach Election Day.
+@fig-start-date-distribution illustrates the distribution of polling start dates from late July to mid-October 2024, highlighting when polling efforts commenced in key battleground states.
+The histogram shows peaks in activity during mid-August, early September, and late September, suggesting increased polling during significant campaign events.
+By including only polls initiated after Harris's campaign launch, we capture voter preferences in a context where both major candidates are actively campaigning.
+This approach ensures our model focuses on data that accurately reflects the competitive dynamics between Harris and Trump, enhancing the validity of our predictive analysis as Election Day approaches.
 

### 2024-11-04 paper/paper.qmd
commit changes to paper.qmd for formatting

@@
 ## Posterior Predictive Check for the Trump Model
-The posterior predictive check for the Trump model is illustrated @fig-trump-check, showcasing the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
-The plot displays a density estimate for both the actual observed data (in dark blue) and the predicted values generated by the model (in light blue).
-The primary aim of this diagnostic check is to assess how well the model captures the distribution of observed support for Donald Trump.
 ```{r}
@@
 ```
-From the visualization, we can observe that the predicted values closely align with the actual observed support percentages.
+The posterior predictive check for the Trump model is illustrated @fig-trump-check, showcasing the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+The plot displays a density estimate for both the actual observed data (in dark blue) and the predicted values generated by the model (in light blue).
+The primary aim of this diagnostic check is to assess how well the model captures the distribution of observed support for Donald Trump.
+@fig-trump-check reveals that the predicted values closely align with the actual observed support percentages.
 The dark blue curve representing the observed data indicates a concentrated support range around 48% to 52%, suggesting that the model effectively captures the central tendency of voter support for Trump.

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd for formatting

@@
 ### Posterior Predictive Check for the Harris Model
-The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
-@fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
-The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
 ```{r}
@@
 #| label: fig-harris-check
-#| fig-cap: "Posterior predictive check for Kamala Harris's model, illustrating the distribution of observed support percentages $y$ against the predicted support values $y_{\text{rep}}$.
-The graph shows a strong correspondence between the observed and predicted distributions, indicating that the model accurately captures the dynamics of voter support for Harris in the given dataset."
+#| fig-cap: "Posterior predictive check for Kamala Harris's model.
+Comparison of observed support percentages ($y$) and predicted values ($y_{\text{rep}}$) shows strong alignment, confirming the model's accuracy in capturing voter support dynamics for Harris."
 
@@
 ```
-In the plot, we observe a strong overlap between the $y$ and $y_{\text{rep}}$ distributions, with the dark blue curve representing the observed support percentages primarily concentrated around 48% to 52%.
+The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+@fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
+The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
+We observe a strong overlap between the $y$ and $y_{\text{rep}}$ distributions, with the dark blue curve representing the observed support percentages primarily concentrated around 48% to 52%.
 This indicates that the model successfully approximates the actual voting support dynamics for Harris.

### 2024-11-04 paper/paper.qmd
committing changes to paper.qmd for formatting changes

@@
 
-## Posterior Predictive Check for the Trump Model
+### Posterior Predictive Check for the Trump Model
 ```{r}
@@
 ### Posterior Predictive Check for the Harris Model
+The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
+@fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
+The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
 ```{r}
@@
 ```
-The posterior predictive check for the Harris model illustrates the relationship between the observed support percentages $y$ and the predicted support values $y_{\text{rep}}$.
-@fig-harris-check displays a density estimate for both the actual observed data (depicted in dark blue) and the predicted values generated by the model (shown in light blue).
-The purpose of this diagnostic check is to evaluate how well the model captures the distribution of observed support for Kamala Harris.
 We observe a strong overlap between the $y$ and $y_{\text{rep}}$ distributions, with the dark blue curve representing the observed support percentages primarily concentrated around 48% to 52%.
